### 2020-05-22 documents/ovando-etal-assessing-global-fisheries.Rmd
initial commit

---
title: "The Future of Global Fishery Assessments"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2: default
  bookdown::pdf_document2: default
abstract: Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals. Numerous high-profile publications have produced estimates of the global status of fisheries, but limited quantity and quality of data, conflicting results, and methodological concerns mean that our understanding of global fisheries is still highly uncertain, especially in critical areas such as Southeast Asia and Africa. We present an evaluation of the past and future of global fishery assessments. Global assessments have typically involved aggregating best-available stock assessments where possible with various data-limited assessments for regions lacking formal evaluations of stock status. The data-limited methods used have ranged from simple catch heuristics to super-ensemble modeling. Despite attempts at model improvement, there have been discrepancies both among global fishery assessments and between the data-limited components of these methods and more data-rich models,demonstrating that our understanding of global data-limited fisheries remains limited. How then can we most efficiently move forward? We utilize a flexible assessment model together with the RAM Legacy Stock Assessment Database not strictly as a stock assessment tool, but to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion and increasingly available but historically underutilized data such as trawl footprints. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world. We close by discussing priority areas for future data collection. Our results make the case that improving our understanding of global fisheries depends not on better models, but on better data.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
library(tidyverse)
library(sf)
library(hrbrthemes)
library(extrafont)
library(scales)
library(here)
extrafont::loadfonts()

results_name <- "v0.5"

results_dir <- here("results",results_name)

load(file = file.path(results_dir, "paper_plots.Rdata"))

pub_theme <- theme_ipsum(base_size = 12, 
                         axis_text_size = 12)

theme_set(pub_theme)

```

Target journal: Nature Sustainability

https://www.nature.com/natsustain/about/content

3500 words main text

50 references (main text)

3000 words (max) methods

# Introduction (500 words)

*Insert mandatory opening sentence about the importance of global fisheries*. On-the ground fisheries management decisions are ideally based on assessment models, designed to provide robust estimates of individual stock status that can inform effective management policies. However, data and resource limitations prohibit our ability to create bespoke tactical assessments for every fishery resource around the globe. As such, we also depend on widely applicable methods intended to provide a general sense of the status of fisheries around the globe. While not suitable for management of individual stocks, these broad strategic models play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals.

The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the most widely followed estimate of global stock status. As of the most recent report, the FAO estimates that 59.9% of marine fish stocks are maximally sustainably fishery, 7% are underfished ,and 33.1% are overfished [@fao2018]. While foundational, the SOFIA reports depend on a complex process of aggregating results from tactical models, expert judgement, etc. **make clear argument why other/new global assessment methods are needed**.

Several publications have explored alternative methods for estimate global stock status, including Costello et al. 2016 [@costello2016a]  and Rosenberg et al. 2018 [@rosenberg2018], and Pauly 2007 [@pauly2007]. Each of these methods sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses. However, comparing estimates from each of these methods against best-available knowledge of stock status from the RAM Legacy Stock Assessment Database [@ricard2012], as well as broader intuition. reveals some systemic problems with these assessments.

We have stock assessment data from RAM from the NE Pacific,the NE Atlantic and the Mediterranean, and of these the NE Pacific is in the best state (mean RAM B/B~MSY~ =1.8), NE Atlantic (mean RAM B/B~MSY~ =1.3) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.5).  We do not have formal stock assessment data from S. and SE Asia (Western Central Pacific and Eastern Indian Ocean), but based on surveys and expert opinion we believe it should be worse than Northeast Pacific. The Costello et al. 2016 [@costello2016a] model shows the correct general pattern, but we would expect the state of fisheries in the Mediterranean/Black sea regions and SE Asia to be much worse, and the NE Pacific better by comparison. The Rosenberg super-ensemble method[@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)) *probably want to make this a figure. Also, Ray, do you remember if these estimates are unassessed only, or unassessed + RAM?*

```{r status-tab}

readr::read_csv(here("data","status-table.csv")) %>% 
  knitr::kable(caption = "Status by FAO Statistical area from four different assessment methods",
               booktabs = TRUE) %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

Despite numerous advances in statistical methods, we are currently left with strong questions about our understanding of the state of global fisheries, particularly in the regions of the world lacking formal stock assessments. We build off of these prior works to demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection effort.

# Results (2000 words)

The FAO's catch statistics remain the most widely accessible and used source of information on global fisheries. Several methods have attempted to infer stock status (generally biomass relative to biomass at maximum sustainable yield, B/B_msy_) based on a fishery's catch history alone [summarized by @free2020]. However, these "catch-only" methods have serious shortcomings that limit their performance, as evidenced by the status discrepancies by region they produce (Table.\@ref(tab:status-tab)), as well as results from several studies examining the performance of such methods [@free2020; @pauly2013].

However, catches are not the only broadly available source of fishery data. We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@anderson2015], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c]. We use a flexible assessment package ([sraplus](https://github.com/DanOvando/sraplus#sraplus)) to compare estimates of stock status around the globe using different combination of these data. We find that our perception of global stock status varies wildly depending on which data source are used, without any clear evidence of a "best" source of data to use *this needs to be much clearer*.

## Case Study: Europe vs. Northeast Asia or East India

Don't know if this section is needed? Key messages if so (and might want to pick just one case study)

- Europe vs North East Asia
  - sort of a more positive story: adding in the Rousseau effort data at least produces a more plausible split between Europe and Northeast Asia
  - Nicely illustrates bi-modal heuristics behavior. 
  - Also lends to shortcomings though: assuming one effort series for all stocks clearly produces some weird results, e.g. all those basically unfished EU stocks

```{r eu-v-na, fig.cap="Estimates of stock status in Europe and Northeast Asia using either a simple catch heuristic as presented in @martell2013 or an abundance index based on continental-scale effort trends reported in @rousseau2019. Each point is a stock, point size is a function of stock size."}
continent_kobe_plot
```

- East India
  - Local data can help: heuristic alone bi-modal and optimistic, SAR alone really pessimistic, CPUE + SAR a bit more nuanced?
  - A little bit a of a conflict here, since paper so far has been about global datasets, and here we're introducing a more local study (trawl survey abundance from eastern bay of Bengal)

```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio. Each point is a stock."}
ei_kobe_plot
```

## Conflicting Estimates of Global Status

This is the main plot of the paper Percent of stocks classified as overfished depending on what data you use. CPUE data are at the continental scale (e.g. use North America Rousseau effort for USA, Mexico, Canada).

key message: Different data sources give you different estimates of stock status, and not in a way that makes it clear that any one is inherently superior to another

A few key questions here

- Should we use a more recent FAO report? Don't remember why we chose the 2011 report for the original report

- Currently this runs sraplus on all stocks in the region. I think this is informative since it says, suppose we only used sraplus, what would we think, and can compare that to RAM etc. in our heads (or could add a map panel). An alternative though would be to stitch together status, i.e. use RAM where there is RAM, use sraplus otherwise. That would be a better "estimate" of the actual stock status in an area, but would make the contrast between the methods less clear

- This used the continent level Rousseau data - can we get the finer scale version for this paper? Would make things a lot clearer I think

*will add in RAM panel once I match up RAM stocks to FAO region*

```{r po-map, fig.width=8, fig.height=8, fig.cap="Estimated percent of fisheries 'overfished' (B/B_msy_ less than 0.8) using different dat sources. CPUE refers to an index of abundance based on the Rousseau effort data. Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data. The FAO Report 569 is our reference status estimate provided by the FAO. ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region. fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
po_map_plot
```

## Value of Information

Provide some guidance on where to prioritize data collection. e.g. how long of a CPUE series is needed, expansion of f based priors / value of f, advocate for merging together of the "Catch only" world with the "length-based world". E.g. we talk about catch being the only widely available statistic, but in many data-limited communities all they have are lengths. A global effort to create a centralized database of length distributions could allow for length-based priors on spr/f that could be paired with catches?

# Discussion (500 words)

*these are just rough outline paragraphs*

Global-level assessment are critical for guiding management agendas for the world's oceans. Despite this need, and despite advances in stock assessment methods, our understanding of many of the world's fisheries remains poor. We show that while new sources of widely available data are becoming increasingly available, none yet is capable of providing a clear improvement in our estimates of global stock status.

Our estimate of global stock status varies wildly depending on which source of data we use. Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status? Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:po-map)). Globally available effort data from Rousseau allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region. These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability. The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations. Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks.

We have stated that improving global estimates of stock status depends on improved data, not improved models *per say*. This is not to suggest that improvements in stock assessment methods, even for relatively simple models such as the biomass dynamics types used here, are not of use. At the tactical level, such improvements can make meaningful changes in management outcomes. However, at the global strategic levels, the improvements to be made from model development are marginal compared to the fundamental uncertainties stemming from currently available data.

What does the future of global fishery assessment look like? Investment must be made in improving data collection at the global scale. Our analysis suggests several possible paths for this process, in particular creation of a global repository or length-composition data, and expansion of FMI surveys and SAR data. Global Fishing Watch provides another avenue, but requires further research on the best ways of translating raw effort data into indices of stock status. Interpreting these data responsibly will require support from regional experts, for example in the form of expert panels charged with determine the most representative broadly available data for different portions of the oceans.

We have accumulated lots of evidence for status of global fisheries, each of which has strengths and weaknesses.These were important foundations. But, if we want to move forwards, the answer is better data.

# Methods (3000 words)

## `sraplus`

Description of the package itself focused on key differences to other packages

## Swept Area Ratio Priors

Basic description and out-of-sample predictive power

## Fisheries Management Index Priors

Basic description and out-of-sample predictive power

## Catch-Only Priors?

At the moment only using Heuristics. I've been exploring an update to the regression based catch-only models, can get up to an out of sample r^2 of about 0.3 and that's about it. Probably worth putting in just to make clear that we're making about as much use out of the catch data alone as we possible can.

## Effort Data

Description of Rousseau data, steps taken to use it (Baranov equation, choice about tech. creep etc.)

## Value of Information Calculations

Details of value of information calculates (fitting to messed up versions of RAM)

## Estimation model

Important details of sraplus, basic equations, etc. Focus on prior-predictive tuning process to validate use of priors on terminal status.

# References

### 2020-05-22 functions/calc-error.R

@@ -1,0 +1,43 @@
+calc_error <- function(data, priors,index_type, q = 1e-6) {
+  #
+  # data <- ram_fits$data[[1]]
+  #
+  # priors <- ram_fits$priors[[1]]
+
+  if (index_type == "survey") {
+    true <-  data$index[priors$index_years]
+  } else {
+    true <- data$approx_cpue[priors$index_years]
+
+  }
+
+  observed <- priors$index
+
+  index_rmse <- sqrt(mean((true - observed) ^ 2))
+
+  true <-  data$u_v_umsy[priors$u_years]
+
+  observed <- priors$u_v_umsy
+
+  u_rmse <- sqrt(mean((true - observed) ^ 2))
+
+  ref_b <- dplyr::case_when(priors$b_ref_type == "k" ~ last(data$b_rel),
+                            TRUE ~ last(data$b_v_bmsy))
+
+  terminal_b_rmse = sqrt((ref_b - priors$terminal_state) ^ 2)
+
+
+  ref_b <-
+    dplyr::case_when(priors$b_ref_type == "k" ~ first(data$b_rel),
+                     TRUE ~ first(data$b_v_bmsy))
+
+  initial_b_rmse = sqrt((ref_b - priors$initial_state) ^ 2)
+
+  outs <- ls()[str_detect(ls(), "rmse")]
+
+  out <- map_dfc(outs, ~ data_frame(.x = get(.x))) %>%
+    set_names(outs)
+
+  return(out)
+
+}

### 2020-05-22 functions/fit-prior-regressions.R

@@ -1,0 +1,128 @@
+fit_prior_regressions <-
+  function(split,
+           model_structure,
+           use_splits = TRUE,
+           chains = 4,
+           cores = 4,
+           iter = 2000,
+           adapt_delta = adapt_delta,
+           max_treedepth = max_treedepth,
+           produce = "summary",
+           refresh = 0,
+           post_draws = 10,
+           ...) {
+    args <- list(...)
+
+
+    # chains = 4
+    #
+    # cores = 4
+
+    # ram_v_fmi <- ram_v_fmi %>%
+    #   mutate(log_value = log(value))
+    #
+    #
+    # training <- ram_v_fmi %>%
+    #   filter(metric == "mean_b") %>%
+    #   slice(1:100)
+    #
+    #
+    # testing <- ram_v_fmi %>%
+    #   filter(metric == "mean_b") %>%
+    #   slice(101:173)
+    #
+    # model_structure <-
+    #   "log_value ~ research + management + enforcement + socioeconomics +(1|country_rfmo)"
+
+    if (use_splits == TRUE){
+
+    training <- rsample::training(split)
+
+    testing <- rsample::testing(split)
+
+    } else {
+
+      training <- split
+
+      testing <- split
+    }
+
+    # has_nas <- map_lgl(training, ~any(is.na(.x))) | map_lgl(testing, ~any(is.na(.x)))
+
+
+
+    model <- as.formula(model_structure)
+
+    if (str_detect(model_structure, "\\|")) {
+      fit <-
+        stan_glmer(model,
+                             data = training,
+                             chains = chains,
+                             cores = cores,
+                             refresh = refresh)
+    } else {
+      fit <-
+        stan_glm(model,
+                           data = training,
+                           chains = chains,
+                           cores = cores,
+                           refresh = refresh)
+
+    }
+
+    training_draws <-
+      tidybayes::add_fitted_draws(model = fit,
+                                  newdata = training,
+                                  value = "fit_pred",
+                                  n = post_draws) %>%
+      mutate(type = "fitted") %>%
+      ungroup()
+
+    pp_training_draws <-
+      tidybayes::add_predicted_draws(model = fit, newdata = training,n = post_draws) %>%
+      ungroup()
+
+    training_draws <- training_draws %>%
+      mutate(pp_pred = pp_training_draws$.prediction) %>%
+      gather(metric, prediction, fit_pred, pp_pred)
+
[…]

### 2020-05-22 functions/fit_continent_examples.R

@@ -1,0 +1,176 @@
+fit_continent_examples <-
+  function(scientific_name,
+           data,
+           model = "sraplus_tmb",
+           engine = "tmb",
+           fmi_f_reg = NA,
+          estimate_shape = TRUE,
+          estimate_proc_error = TRUE,
+          estimate_qslope = FALSE,
+          use_baranov = TRUE,
+          first_effort_year = 1950,
+          chains = 1,
+          cores = 1,
+          q_slope_prior = 0.025) {
+    # heuristic example
+    # i = 2
+    # engine = "tmb"
+    # model = "sraplus_tmb"
+    # data <- huh$data[[1]]
+    # scientific_name <- huh$scientific_name[[1]]
+    # catches <- data$total_catch
+
+    # estimate_proc_error = FALSE
+    # estimate_qslope = TRUE
+    # estimate_shape = TRUE
+    # use_baranov = TRUE
+    # first_effort_year = 1950
+
+    catches <- data$total_catch
+
+    years <- data$year
+
+    basic_driors <-
+      format_driors(
+        taxa = scientific_name,
+        catch = catches,
+        years = seq_along(years),
+        initial_state = NA,
+        initial_state_cv = 0.2,
+        terminal_state = NA,
+        terminal_state_cv =  0.05,
+        use_heuristics = TRUE,
+        b_ref_type = "k"
+      )
+
+
+    basic_fit <-
+      fit_sraplus(
+        driors = basic_driors,
+        include_fit = TRUE,
+        model = model,
+        engine = "sir",
+        estimate_shape = estimate_shape,
+        estimate_qslope = estimate_qslope,
+        estimate_proc_error = estimate_proc_error
+      )
+
+    # basic_fit_results <- basic_fit$results %>%
+    #   filter(variable != "b_t", variable != "index_hat_t") %>%
+    #   mutate(variable = case_when(
+    #     variable == "b_bmsy_t" ~ "b",
+    #     variable == "c_msy_t" ~ "c",
+    #     variable == "dep_t" ~ "dep",
+    #     variable == "u_umsy_t" ~ "u",
+    #     TRUE ~ variable
+    #   )) %>%
+    #   mutate(year = min(years) + year - 1,
+    #          data = "heuristic")
+
+    basic_fit_results <- basic_fit$results %>%
[…]

### 2020-05-22 functions/prepare-sofia-data.R

@@ -1,0 +1,644 @@
+prepare_sofia_data <- function(min_years_catch = 20,
+                               crazy_b = 4,
+                               crazy_u = 5,
+                               lookup_fmi_names = FALSE,
+                               q = 1e-6) {
+  # min_years_catch = 20
+  # crazy_b = 4
+  # crazy_u = 5
+  # lookup_fmi_names = FALSE
+  
+  # load(here::here("data", "Return.Rdata"))
+  
+  
+  if (file.exists(here("data", "ram.zip")) == FALSE) {
+    # for now storing in my google drive... will need to put this in a better and public location
+    
+    download.file("https://www.dropbox.com/s/jpgz0a5s5of3qev/RAM%20v4.491%20Files%20(1-14-20).zip?dl=1", destfile = here("data","ram.zip"))
+    
+    unzip(here("data","ram.zip"), exdir = here("data"))
+  }
+  
+  
+  ram_dirs <- list.files(here("data","ram"))
+  
+  ram_dirs <- ram_dirs[str_detect(ram_dirs,"RAM")]
+  
+  ram_files <- list.files(here("data","ram",ram_dirs), recursive = TRUE)
+  
+  ram_files <- ram_files[str_detect(ram_files,".RData")]
+  
+  ram_files <- ram_files[str_detect(ram_files,"Model Fit")]
+  
+  load(here("data","ram",ram_dirs,ram_files[1]))
+  
+  # process data ------------------------------------------------------------
+  
+  stock <- stock %>%
+    left_join(area, by = "areaid")
+  # catches
+  ram_catches <- tcbest.data %>%
+    mutate(year = rownames(.) %>% as.integer()) %>%
+    as_data_frame() %>%
+    gather(stockid, catch,-year)
+  
+  # B/Bmsy
+  ram_b_v_bmsy <- divbpref.data %>%
+    mutate(year = rownames(.) %>% as.integer()) %>%
+    tibble() %>%
+    gather(stockid, b_v_bmsy,-year)
+  
+  
+  # U/Umsy
+  ram_u_v_umsy <- divupref.data %>%
+    mutate(year = rownames(.) %>% as.integer()) %>%
+    as_data_frame() %>%
+    gather(stockid, u_v_umsy,-year)
+  
+  # Effort
+  ram_effort <- effort.data %>%
+    mutate(year = rownames(.) %>% as.integer()) %>%
+    as_data_frame() %>%
+    gather(stockid, effort,-year)
+  
+  # biomass
+  
+  
[…]

### 2020-05-22 functions/run-biosraplus.R

@@ -1,0 +1,200 @@
+run_biosraplus <- function(iter = 1000,
+                           dat,
+                           priors,
+                           refresh = 0,
+                           show_messages = FALSE,
+                           sra_model,
+                           q = 1e-6) {
+
+
+  genus_species <-
+    unique(dat$scientificname) %>% str_split(" ", simplify = TRUE)
+
+  shh <- quietly(FishLife::Search_species)
+
+  fish_search <-
+    shh(Genus = genus_species[1], Species = genus_species[2])
+
+  if (is.null(fish_search$error)) {
+    taxon <- fish_search$result$match_taxonomy[1] %>%
+      str_split("_") %>%
+      unlist()
+
+  } else{
+    shh <- quietly(FishLife::Search_species)
+
+    taxon <-
+      shhh(Genus = genus_species[1])$result$match_taxonomy[1] %>%
+      str_split("_") %>%
+      unlist()
+
+  }
+
+
+
+
+  taxon <-
+    data_frame(
+      Class = taxon[[1]],
+      Order = taxon[[2]],
+      Family = taxon[[3]],
+      Genus = taxon[[4]],
+      Species = taxon[[5]]
+    )
+
+  sp <- shh(
+    Class = taxon["Class"],
+    Order = taxon["Order"],
+    Family = taxon["Family"],
+    Genus = taxon["Genus"],
+    Species = taxon["Species"],
+    ParentChild_gz = Return$ParentChild_gz
+  )$result$match_taxonomy[1]
+
+  taxa_location <- grep(sp, Return$ParentChild_gz[, "ChildName"])[1]
+
+  mean_lh <- Return$beta_gv[taxa_location,]
+
+  cov_lh <- Return$Cov_gvv[taxa_location, ,]
+  ##------------ deviates from multivariate normal -------------##
+  ## parameters for multvariate normal distribution
+  ##  params_mvn <- c("tm", "M", "tmax", "Loo", "K", "Winfinity") #"ln_var", "Loo", "h", "logitbound_h")
+  params_mvn <-
+    c("intrinsic_growth_rate", "ln_var")
+  mean_lh <- mean_lh[which(names(mean_lh) %in% params_mvn)]
+
+  cov_lh <-
[…]

### 2020-05-22 scripts/fit-ram.R

@@ -1,0 +1,264 @@
+  library(tidyverse)
+  library(here)
+  library(sraplus)
+  library(furrr)
+  
+  # options -----------------------------------------------------------------
+  
+  min_years_catch <- 20
+  
+  crazy_b <- 4 # threshold for suspect B/Bmsy value
+  
+  crazy_u <- 5 # threshold for suspect U/Umsy value
+  
+  
+  # load data ---------------------------------------------------------------
+  
+  
+  if(!dir.exists("data-raw")){
+    dir.create("data-raw")
+  }
+  
+  if (!file.exists(here("data-raw","ram.zip"))) {
+    
+    download.file("https://www.dropbox.com/s/jpgz0a5s5of3qev/RAM%20v4.491%20Files%20(1-14-20).zip?dl=1", destfile = here::here("data-raw","ram.zip"))
+    
+    unzip(here::here("data-raw","ram.zip"), exdir = "data-raw")
+    
+  }
+  
+  
+  
+  # process RAM data --------------------------------------------------------
+  
+  ram_dirs <- list.files("data-raw")
+  
+  ram_dirs <- ram_dirs[str_detect(ram_dirs,"RAM v\\d")]
+  
+  ram_files <- list.files(here("data-raw",ram_dirs), recursive = TRUE)
+  
+  ram_files <- ram_files[str_detect(ram_files,".RData")]
+  
+  ram_files <- ram_files[str_detect(ram_files,"Assessment Data")]
+  
+  load(here("data-raw",ram_dirs,ram_files[1]))
+  
+  stock <- stock %>%
+    left_join(area, by = "areaid")
+  # catches
+  ram_catches <- tcbest.data %>%
+    mutate(year = rownames(.) %>% as.integer()) %>%
+    as_data_frame() %>%
+    gather(stockid, catch,-year)
+  
+  # B/Bmsy
+  ram_b_v_bmsy <- divbpref.data %>%
+    mutate(year = rownames(.) %>% as.integer()) %>%
+    as_data_frame() %>%
+    gather(stockid, b_v_bmsy,-year)
+  
+  # U/Umsy
+  ram_u_v_umsy <- divupref.data %>%
+    mutate(year = rownames(.) %>% as.integer()) %>%
+    as_data_frame() %>%
+    gather(stockid, u_v_umsy,-year)
+  
+  # Effort
+  ram_effort <- effort.data %>%
[…]

### 2020-06-14 functions/fit-fao.R
updated to FAO query instead of stored data

@@ -36,7 +36,7 @@ fit_fao <-
         years = seq_along(years),
         use_heuristics = TRUE,
         b_ref_type = "k",
-        isscaap_group = unique(data$isscaap_group)
+        isscaap_group = unique(data$isscaap_group)[1]
       )
 
     basic_fit <-
@@ -69,7 +69,7 @@ fit_fao <-
           initial_state_cv = basic_driors$initial_state_cv,
           sar = last(data$sar),
           sar_cv = reg_cv,
-          isscaap_group = unique(data$isscaap_group)
+          isscaap_group = unique(data$isscaap_group)[1]
         )
 
       sar_fit <-
@@ -87,6 +87,8 @@ fit_fao <-
         filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy","depletion")) %>%
         mutate(year = rep(years, each = 4)) %>%
         mutate(data = "sar")
+      
+      rm(sar_fit)
 
     } else {
       used_sar <- FALSE
@@ -94,10 +96,9 @@ fit_fao <-
 
 
     # fmi fit
-
     fmi_dat <- support_data$mean_regional_isscaap_fmi %>%
       filter(fao_area_code == unique(as.numeric(data$fao_area)),
-             isscaap_group == unique(data$isscaap_group))
+             isscaap_group == unique(data$isscaap_group)[1])
 
 
     fmi <- c(research = fmi_dat$research, management = fmi_dat$management, socioeconomics = fmi_dat$socioeconomics, enforcement = fmi_dat$enforcement)
@@ -137,7 +138,7 @@ fit_fao <-
           f_ref_type = "fmsy",
           fmi_cv = reg_cv,
           use_b_reg = FALSE,
-          isscaap_group = unique(data$isscaap_group)
+          isscaap_group = unique(data$isscaap_group)[1]
         )
       fmi_fit <-
         fit_sraplus(
@@ -154,6 +155,8 @@ fit_fao <-
         filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy","depletion")) %>%
         mutate(year = min(years) + year - 1) %>%
         mutate(data = "fmi")
+      
+      rm(fmi_fit)
 
       used_fmi <- TRUE
     } else {
@@ -175,7 +178,7 @@ fit_fao <-
           u_years = u_years,
           initial_state = basic_driors$initial_state,
           initial_state_cv = 0.2,
-          isscaap_group = unique(data$isscaap_group)
[…]

### 2020-06-14 functions/prepare-sofia-data.R

@@ -193,6 +193,200 @@ prepare_sofia_data <- function(min_years_catch = 20,
   
   # fao and effort data
   
+  # this is going to take some done, you are here
+  
+  # fishstat_dat <-
+  #   readr::read_csv(
+  #     here::here("data", "fao_capture_1950-2016.csv")
+  #   ) %>%
+  #   janitor::clean_names() %>% # clean up the names to something usable
+  #   tibble::rownames_to_column(var = "id") %>% # add an id column
+  #   dplyr::filter(
+  #     !stringr::str_detect(tolower(country_country), "total"),!stringr::str_detect(country_country, "\\d"),!stringr::str_detect(
+  #       tolower(production_source_detailed_production_source),
+  #       "aquaculture"
+  #     )
+  #   ) # get rid of totals rows, numbered rows, and aquaculture datta
+  # 
+  # 
+  # fishstat_dat <- fishstat_dat %>% # convert the individual year columns to a tidy format
+  #   tidyr::pivot_longer(
+  #     cols = starts_with("x"), # noticed that all year columns started with x
+  #     names_prefix = "x", # gets rid of that x in front of the years
+  #     # names_ptypes = list(year = integer()), # convert each column years to an integer after stripping off the x
+  #     names_to = "year",# the variable the former column names will go into
+  #     values_to = "capture" # the variable the capture values will go ti
+  #   ) %>%
+  #   mutate(
+  #     year = as.integer(year),
+  #     capture_notes = str_replace_all(capture, "\\d", ''), #pull out the text
+  #     capture_numbers = str_replace_all(capture, "\\D", "") %>% as.numeric(), #pull out the numbers and convert the numerics. Note the use of a nested pipe! R is cool.
+  #     clean_capture = as.numeric(capture) # SERIOUSLY awful step to accounnt for the like 10 stocks entered as 70000000e-7.... as.numeric will keep any clean numbers and convert any numbers + text to NA
+  #   ) %>%
+  #   mutate(capture = ifelse(test = is.na(clean_capture), yes = capture_numbers, no = clean_capture)) %>%  # keep any stock that was a clean number (which catches the 7e-7, and add in anything else that was like 500 F
[…]

### 2020-06-14 make-assessing-global-fisheries.R

@@ -26,7 +26,7 @@ Sys.unsetenv("PKG_CXXFLAGS")
 
 rstan::rstan_options(auto_write = TRUE)
 
-
+sraplus::get_tmb_model()
 # options -----------------------------------------------------------------
 
 min_years_catch <- 20
@@ -67,7 +67,7 @@ run_continent_examples <- FALSE
 
 run_ei_example <- FALSE
 
-run_sofia_comparison <- TRUE
+run_sofia_comparison <- FALSE
 
 engine = "stan"
 
@@ -752,12 +752,19 @@ total_stocks <- fao %>%
   mutate(continent = tolower(continent)) %>%
   left_join(total_nominal_effort, by = c("year", "continent" = "region"))
 
+fao %>% 
+  group_by(year) %>% 
+  summarise(catch = sum(capture, na.rm = TRUE)) %>% 
+  ggplot(aes(year, catch)) + 
+  geom_line() + 
+  scale_y_continuous(labels = comma)
 
-# total_stocks %>%
-#   ggplot(aes(year, total_catch, color = scientific_name)) +
-#   geom_line(show.legend = FALSE) +
-#   facet_wrap(~continent) +
-#   scale_y_log10()
+
+total_stocks %>%
+  ggplot(aes(year, total_catch, color = scientific_name)) +
+  geom_line(show.legend = FALSE) +
+  facet_wrap(~continent) +
+  scale_y_log10()
 
 # future::plan(future::multiprocess, workers = 4)
 
@@ -1281,12 +1288,12 @@ assign_effort <-
            fao_catch,
            scalar = 1000) {
     # data <- temp_fao$data[[1]]
-    #
+    # 
     # fao_catch <- fao
-    #
+    # 
     # effort <- rous_data
-    #
-    # fao_stock <- "33-Miscellaneous coastal fishes_77_33"
+    # 
+    # fao_stock <- temp_fao$stockid[1]
     
     comm_name <-
       str_split(fao_stock, pattern = '_')[[1]][1] %>% str_remove_all("(\\d)|(-)")
@@ -1294,10 +1301,10 @@ assign_effort <-
     isscp_number <-
       as.numeric(str_split(fao_stock, pattern = '_')[[1]][3])
     
-    fao_code <- str_split(fao_stock, pattern = '_')[[1]][2]
+    fao_code <- as.numeric(str_split(fao_stock, pattern = '_')[[1]][2])
     
     fao_matches <- fao %>% {
[…]

### 2020-06-15 functions/fit-fao.R
fixed a bunch of issues stemming from the new FAO Data

@@ -9,25 +9,27 @@ fit_fao <-
            q_slope_prior = 0.025,
            estimate_proc_error = TRUE,
            estimate_qslope = FALSE,
+           default_initial_state = 1,
+           default_initial_state_cv = 0.2,
            chains = 1,
            cores = 1,
            reg_cv = NA,
            draws = 2e6,
            refresh = 0,
            thin_draws = TRUE) {
-#
-#     data <- test$data[[1]]
-
+    #
+    #     data <- test$data[[1]]
+    
     ogengine <- engine
-
+    
     scientific_name <- unique(data$species)
-
+    
     catches <- data$catch
-
+    
     years <- data$year
-
+    
     u_years <- seq_along(years)[!is.na(data$mean_u_umsy)]
-
+    
     u_umsy <- data$mean_u_umsy[u_years]
     basic_driors <-
       format_driors(
@@ -36,9 +38,9 @@ fit_fao <-
         years = seq_along(years),
         use_heuristics = TRUE,
         b_ref_type = "k",
-        isscaap_group = unique(data$isscaap_group)[1]
+        isscaap_group = unique(data$isscaap_group)
       )
-
+    
     basic_fit <-
       fit_sraplus(
         driors = basic_driors,
@@ -48,30 +50,30 @@ fit_fao <-
         draws = draws,
         thin_draws = thin_draws
       )
-
+    
     basic_fit_results <- basic_fit$results %>%
-      filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy","depletion")) %>%
+      filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy", "depletion")) %>%
       mutate(year = rep(years, each = 4)) %>%
       mutate(data = "heuristic")
-
-
+    
+    
     # sar fit
-
+    
     if (!is.na(last(data$sar))) {
       used_sar <- TRUE
-
+      
       sar_driors <-
         sraplus::format_driors(
           taxa = scientific_name,
           catch = catches,
           years = seq_along(catches),
-          initial_state = basic_driors$initial_state,
-          initial_state_cv = basic_driors$initial_state_cv,
+          initial_state = default_initial_state,
+          initial_state_cv = default_initial_state,
           sar = last(data$sar),
           sar_cv = reg_cv,
-          isscaap_group = unique(data$isscaap_group)[1]
+          isscaap_group = unique(data$isscaap_group)
         )
-
+      
       sar_fit <-
         fit_sraplus(
[…]

### 2020-06-15 functions/prepare-sofia-data.R

@@ -357,18 +357,27 @@ prepare_sofia_data <- function(min_years_catch = 20,
   asfis <- read_delim(here("data","fao","ASFIS_sp_2020.txt"), delim = ",") %>% 
     janitor::clean_names() %>% 
     rename(isscaap_code = isscaap) %>% 
-    select(isscaap_code, scientific_name) %>% 
+    select(isscaap_code, scientific_name, taxocode) %>% 
     unique()
+  
+  # major issue with NEIs here. There is no database that has both isscaap group and isscaap code, so you need
+  # to do a complicated merge based on scientific name. 
+  # the problem there is that some "scientific names" 
   
   fao_capture <- read_csv(here("data","fao","TS_FI_CAPTURE.csv")) %>% 
     janitor::clean_names()
   
   sp_groups <- read_csv(here("data","fao","CL_FI_SPECIES_GROUPS.csv")) %>% 
     janitor::clean_names() %>% 
-    select(x3alpha_code:identifier, contains("_en"), scientific_name:cpc_group) %>% 
+    select(x3alpha_code:identifier, contains("_en"), author:cpc_group) %>% 
     rename(species_name_en = name_en) %>% 
-    left_join(asfis, by = "scientific_name")
+    left_join(asfis, by = c("taxonomic_code" = "taxocode"))
   
+  # sp_groups %>%
+  #   group_by(x3alpha_code) %>%
+  #   summarise(ni = n_distinct(isscaap_group)) %>%
+  #   arrange(desc(ni))
+
   country_groups <- read_csv(here("data","fao","CL_FI_COUNTRY_GROUPS.csv")) %>% 
     janitor::clean_names() %>% 
     mutate(un_code = as.numeric(un_code)) %>% 
@@ -386,6 +395,12 @@ prepare_sofia_data <- function(min_years_catch = 20,
   fao_capture <- fao_capture %>% 
     left_join(country_groups, by = c("country" = "un_code")) %>% 
     left_join(fao_areas, by = "fishing_area")
+  
+  # fao_capture %>%
+  #   group_by(species) %>%
+  #   summarise(ni = n_distinct(isscaap_group)) %>%
+  #   arrange(desc(ni))
+
   
   fao_to_effort <-
     read_csv(here::here("data", "fao-to-bell-region.csv")) %>%
@@ -454,6 +469,14 @@ prepare_sofia_data <- function(min_years_catch = 20,
   fao_species <- fao %>%
[…]

### 2020-06-15 make-assessing-global-fisheries.R

@@ -39,7 +39,7 @@ draws <- 2500
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
-n_cores <- 4
+n_cores <- 6
 # number of cores for parallel processing
 
 future::plan(multiprocess, workers = n_cores)
@@ -1095,7 +1095,8 @@ fao2011 <-
     year = as.numeric(year),
     area = as.numeric(area)
   ) %>%
-  filter(!is.na(area))
+  filter(!is.na(area),
+         !is.na(name))
 
 
 n_status_plot <- fao2011 %>%
@@ -1287,14 +1288,14 @@ assign_effort <-
            effort,
            fao_catch,
            scalar = 1000) {
-    # data <- temp_fao$data[[1]]
-    # 
+    # data <- temp_fao$data[[which(temp_fao$stockid == huh)]]
+
     # fao_catch <- fao
-    # 
+
     # effort <- rous_data
-    # 
-    # fao_stock <- temp_fao$stockid[1]
-    
+
+    # fao_stock <- temp_fao$stockid[which(temp_fao$stockid == huh)]
+
     comm_name <-
       str_split(fao_stock, pattern = '_')[[1]][1] %>% str_remove_all("(\\d)|(-)")
     
@@ -1352,6 +1353,17 @@ fao2011 <- temp_fao %>%
 #   geom_line() + 
 #   facet_wrap(~area)
 
+# temp <- fao2011 %>%
+#   # filter(area %in% 67) %>%
+#   group_by(stockid) %>%
+#   nest() %>%
+#   ungroup()
+# 
+# huh <- temp$stockid[4]
+# 
+# wtf <- fao2011 %>% 
+#   filter(stockid == huh)
+
 
 fao2011 <- fao2011 %>%
   group_by(stockid) %>%
@@ -1387,7 +1399,7 @@ if (run_sofia_comparison == TRUE) {
     group_by(stockid) %>%
     nest() %>%
     ungroup() %>%
-    # sample_n(4) %>%
+    # sample_n(25) %>%
     mutate(
       fits = future_map(
         data,
@@ -1401,15 +1413,16 @@ if (run_sofia_comparison == TRUE) {
       )
     )
   
+
   # "Pink(=Humpback)salmon_67_23"
-  
-  # fao2011_fits <- fao2011 %>%
-  #   filter(area %in% 67) %>%
+  # 
+  # temp <- fao2011 %>%
[…]

### 2020-06-15 documents/ovando-etal-assessing-global-fisheries.Rmd
Finer scale effort (#2)
* incorporated the finer-scale rousseau data

* adding nominal CPUE to sofia comparisons

* updated to FAO query instead of stored data

* fixed a bunch of issues stemming from the new FAO Data

@@
 date: "`r Sys.Date()`"
-bibliography: ["references.bib"]
+bibliography: ["../references.bib"]
 csl: nature.csl
 output: 
+  bookdown::pdf_document2: default
   bookdown::word_document2: default
-  bookdown::pdf_document2: default
 abstract: Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals.

### 2020-06-15 functions/fit-fao.R

@@ -9,24 +9,27 @@ fit_fao <-
            q_slope_prior = 0.025,
            estimate_proc_error = TRUE,
            estimate_qslope = FALSE,
+           default_initial_state = 1,
+           default_initial_state_cv = 0.2,
            chains = 1,
            cores = 1,
            reg_cv = NA,
            draws = 2e6,
-           refresh = 0) {
-#
-#     data <- test$data[[1]]
-
+           refresh = 0,
+           thin_draws = TRUE) {
+    #
+    #     data <- test$data[[1]]
+    
     ogengine <- engine
-
+    
     scientific_name <- unique(data$species)
-
+    
     catches <- data$catch
-
+    
     years <- data$year
-
+    
     u_years <- seq_along(years)[!is.na(data$mean_u_umsy)]
-
+    
     u_umsy <- data$mean_u_umsy[u_years]
     basic_driors <-
       format_driors(
@@ -37,98 +40,117 @@ fit_fao <-
         b_ref_type = "k",
         isscaap_group = unique(data$isscaap_group)
       )
-
+    
     basic_fit <-
       fit_sraplus(
         driors = basic_driors,
         include_fit = include_fit,
         model = model,
         engine = "sir",
-        draws = draws
+        draws = draws,
+        thin_draws = thin_draws
       )
-
+    
     basic_fit_results <- basic_fit$results %>%
-      filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy","depletion")) %>%
+      filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy", "depletion")) %>%
       mutate(year = rep(years, each = 4)) %>%
       mutate(data = "heuristic")
-
-
+    
+    
     # sar fit
-
+    
     if (!is.na(last(data$sar))) {
       used_sar <- TRUE
-
+      
       sar_driors <-
         sraplus::format_driors(
           taxa = scientific_name,
           catch = catches,
           years = seq_along(catches),
-          initial_state = basic_driors$initial_state,
-          initial_state_cv = basic_driors$initial_state_cv,
+          initial_state = default_initial_state,
+          initial_state_cv = default_initial_state,
           sar = last(data$sar),
           sar_cv = reg_cv,
           isscaap_group = unique(data$isscaap_group)
         )
-
+      
       sar_fit <-
         fit_sraplus(
           driors = sar_driors,
           include_fit = include_fit,
           model = model,
[…]

### 2020-06-15 functions/prepare-sofia-data.R

@@ -16,7 +16,7 @@ prepare_sofia_data <- function(min_years_catch = 20,
     
     download.file("https://www.dropbox.com/s/jpgz0a5s5of3qev/RAM%20v4.491%20Files%20(1-14-20).zip?dl=1", destfile = here("data","ram.zip"))
     
-    unzip(here("data","ram.zip"), exdir = here("data"))
+    unzip(here("data","ram.zip"), exdir = here("data","ram"))
   }
   
   
@@ -173,7 +173,7 @@ prepare_sofia_data <- function(min_years_catch = 20,
   
   # load prices
   
-  prices <<-
+  prices <-
     readr::read_csv(here::here("data", "Exvessel Price Database.csv")) %>%
     janitor::clean_names() %>%
     rename(scientificname = scientific_name) %>%
@@ -186,10 +186,221 @@ prepare_sofia_data <- function(min_years_catch = 20,
     group_by(scientificname) %>%
     mutate(lag_exvessel = lag(exvessel))
   
+  assign("prices", prices, envir = .GlobalEnv)
+  
   ram_data <- ram_data %>%
     left_join(prices, by = c("scientificname", "year"))
   
   # fao and effort data
+  
+  # this is going to take some done, you are here
+  
+  # fishstat_dat <-
+  #   readr::read_csv(
+  #     here::here("data", "fao_capture_1950-2016.csv")
+  #   ) %>%
+  #   janitor::clean_names() %>% # clean up the names to something usable
+  #   tibble::rownames_to_column(var = "id") %>% # add an id column
+  #   dplyr::filter(
+  #     !stringr::str_detect(tolower(country_country), "total"),!stringr::str_detect(country_country, "\\d"),!stringr::str_detect(
+  #       tolower(production_source_detailed_production_source),
+  #       "aquaculture"
+  #     )
+  #   ) # get rid of totals rows, numbered rows, and aquaculture datta
+  # 
+  # 
+  # fishstat_dat <- fishstat_dat %>% # convert the individual year columns to a tidy format
+  #   tidyr::pivot_longer(
+  #     cols = starts_with("x"), # noticed that all year columns started with x
+  #     names_prefix = "x", # gets rid of that x in front of the years
+  #     # names_ptypes = list(year = integer()), # convert each column years to an integer after stripping off the x
+  #     names_to = "year",# the variable the former column names will go into
[…]

### 2020-06-15 make-assessing-global-fisheries.R

@@ -26,7 +26,7 @@ Sys.unsetenv("PKG_CXXFLAGS")
 
 rstan::rstan_options(auto_write = TRUE)
 
-
+sraplus::get_tmb_model()
 # options -----------------------------------------------------------------
 
 min_years_catch <- 20
@@ -39,7 +39,8 @@ draws <- 2500
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
-n_cores <- 4 # number of cores for parallel processing
+n_cores <- 6
+# number of cores for parallel processing
 
 future::plan(multiprocess, workers = n_cores)
 
@@ -751,12 +752,19 @@ total_stocks <- fao %>%
   mutate(continent = tolower(continent)) %>%
   left_join(total_nominal_effort, by = c("year", "continent" = "region"))
 
+fao %>% 
+  group_by(year) %>% 
+  summarise(catch = sum(capture, na.rm = TRUE)) %>% 
+  ggplot(aes(year, catch)) + 
+  geom_line() + 
+  scale_y_continuous(labels = comma)
 
-# total_stocks %>%
-#   ggplot(aes(year, total_catch, color = scientific_name)) +
-#   geom_line(show.legend = FALSE) +
-#   facet_wrap(~continent) +
-#   scale_y_log10()
+
+total_stocks %>%
+  ggplot(aes(year, total_catch, color = scientific_name)) +
+  geom_line(show.legend = FALSE) +
+  facet_wrap(~continent) +
+  scale_y_log10()
 
 # future::plan(future::multiprocess, workers = 4)
 
@@ -1028,7 +1036,7 @@ fao_country_region_key <- fao %>%
   filter(ptc > 0.25) %>%
   ungroup()
 
-fmi$country <- countrycode::countrycode(fmi$country_rfmo, "country.name","fao.name")
+fmi$country <- countrycode::countrycode(fmi$country_rfmo, "country.name","un.name.en")
 
 
 
@@ -1087,7 +1095,8 @@ fao2011 <-
     year = as.numeric(year),
     area = as.numeric(area)
   ) %>%
-  filter(!is.na(area))
+  filter(!is.na(area),
+         !is.na(name))
 
 
 n_status_plot <- fao2011 %>%
@@ -1129,7 +1138,7 @@ catch_status_plot <- fao2011 %>%
 load(here::here("data","PNAS-data","TBPdata.Rdata"))
 
[…]

### 2020-06-15 functions/prepare-sofia-data.R
removing data you don't need anymore

@@ -172,25 +172,25 @@ prepare_sofia_data <- function(min_years_catch = 20,
   # load other data ---------------------------------------------------------
   
   # load prices
+  # 
+  # prices <-
+  #   readr::read_csv(here::here("data", "Exvessel Price Database.csv")) %>%
+  #   janitor::clean_names() %>%
+  #   rename(scientificname = scientific_name) %>%
+  #   mutate(log_exvessel = log(exvessel)) %>%
+  #   group_by(asfis_species, pooled_commodity, group_for_pairing) %>%
+  #   mutate(lag_exvessel = lag(exvessel)) %>%
+  #   ungroup() %>%
+  #   group_by(scientificname, year) %>%
+  #   summarise(exvessel = mean(exvessel, na.rm = TRUE)) %>%
+  #   group_by(scientificname) %>%
+  #   mutate(lag_exvessel = lag(exvessel))
+  # 
+  # assign("prices", prices, envir = .GlobalEnv)
   
-  prices <-
-    readr::read_csv(here::here("data", "Exvessel Price Database.csv")) %>%
-    janitor::clean_names() %>%
-    rename(scientificname = scientific_name) %>%
-    mutate(log_exvessel = log(exvessel)) %>%
-    group_by(asfis_species, pooled_commodity, group_for_pairing) %>%
-    mutate(lag_exvessel = lag(exvessel)) %>%
-    ungroup() %>%
-    group_by(scientificname, year) %>%
-    summarise(exvessel = mean(exvessel, na.rm = TRUE)) %>%
-    group_by(scientificname) %>%
-    mutate(lag_exvessel = lag(exvessel))
-  
-  assign("prices", prices, envir = .GlobalEnv)
-  
-  ram_data <- ram_data %>%
-    left_join(prices, by = c("scientificname", "year"))
-  
+  # ram_data <- ram_data %>%
+  #   left_join(prices, by = c("scientificname", "year"))
+  # 
   # fao and effort data
   
   # this is going to take some done, you are here

### 2020-06-16 functions/fit-fao.R
using initial state priors by defalut

@@ -9,7 +9,7 @@ fit_fao <-
            q_slope_prior = 0.025,
            estimate_proc_error = TRUE,
            estimate_qslope = FALSE,
-           default_initial_state = 1,
+           default_initial_state = NA,
            default_initial_state_cv = 0.2,
            chains = 1,
            cores = 1,
@@ -249,7 +249,6 @@ fit_fao <-
           isscaap_group = unique(data$isscaap_group),
           q_slope_prior = q_slope_prior
         )
-      
       sfs <- safely(fit_sraplus)
       
       cpue_fit <-

### 2020-06-16 make-assessing-global-fisheries.R

@@ -39,7 +39,7 @@ draws <- 2500
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
-n_cores <- 6
+n_cores <- 3
 # number of cores for parallel processing
 
 future::plan(multiprocess, workers = n_cores)
@@ -67,7 +67,7 @@ run_continent_examples <- FALSE
 
 run_ei_example <- FALSE
 
-run_sofia_comparison <- FALSE
+run_sofia_comparison <- TRUE
 
 engine = "stan"
 
@@ -1422,7 +1422,7 @@ if (run_sofia_comparison == TRUE) {
   #   ungroup() %>%
   #   # sample_n(1) %>%
   #   # filter(stockid ==  "Pink(=Humpback)salmon_67_23") %>%
-  #   slice(141) %>% 
+  #   slice(141) %>%
   #   mutate(
   #     fits = map(
   #       data,
@@ -1433,7 +1433,7 @@ if (run_sofia_comparison == TRUE) {
   #       cores = 1
   #     )
   #   )
-  # 
+
   
   write_rds(fao2011_fits, path = file.path(results_path, "fao2011-fits.rds"))
   

### 2020-06-17 functions/generate-priors.R
passing to squid

@@ -2,13 +2,10 @@ generate_priors <-
   function(dat,
            use_index = FALSE,
            index_type = "survey",
-           use_initial_state = TRUE,
+           initial_state_type,
            initial_state_cv = 0.1,
            use_terminal_state = TRUE,
            terminal_state_cv = 0.1,
-           use_terminal_u = FALSE,
-           terminal_u = NA,
-           terminal_u_cv = 0.1,
            known_carry  = FALSE,
            carry_cv = 0.25,
            use_u_priors = FALSE,
@@ -20,18 +17,21 @@ generate_priors <-
            index_window = 1,
            effort_window = 1,
            u_window = 1,
-           u_freq = 1,
            u_cv = 0.1,
            index_freq = 1,
-           effort_freq = 1,
-           use_heuristics = FALSE) {
-    if (use_initial_state == TRUE) {
+           effort_freq = 1) {
+    if (initial_state_type == "known") {
       initial_state = dplyr::case_when(b_ref_type == "k" ~ dat$b_rel[1],
                                        TRUE ~ dat$b_v_bmsy[1])
 
+    } else if (initial_state_type == "unfished"){
+      initial_state <- dplyr::case_when(b_ref_type == "k" ~ 1,
+                                        TRUE ~ 2.5)
     } else {
+      
       initial_state <- NA
-    }
+    
+      }
 
     if (use_terminal_state == TRUE) {
       terminal_state = dplyr::case_when(b_ref_type == "k" ~ last(dat$b_rel),
@@ -40,27 +40,6 @@ generate_priors <-
     } else {
       terminal_state <- NA
     }
-
-    if (use_heuristics == TRUE) {
-      temp <-
-        if (dat$catch[1] / max(dat$catch, na.rm = TRUE) < 0.2) {
-          0.7
-        } else {
-          0.4
-        }
-
-      initial_state = dplyr::case_when(b_ref_type == "k" ~ temp,
-                                       TRUE ~ temp * 2)
-
-
-      temp_terminal <-
-        ifelse((last(dat$catch) / max(dat$catch)) > 0.5, 0.6, 0.2)
-
-      terminal_state = dplyr::case_when(b_ref_type == "k" ~ temp_terminal,
-                                        TRUE ~ temp_terminal * 2)
-
-
-    } # close if use heuristics
 
 
     if (known_carry == TRUE) {
[…]

### 2020-06-17 functions/prepare-sofia-data.R

@@ -109,11 +109,12 @@ prepare_sofia_data <- function(min_years_catch = 20,
   
   # filter data -------------------------------------------------------------
   
-  
+  #ACMACKSARG	
   # for now, only include continuous catch series
   
   ram_data <- ram_data %>%
     filter(is.na(catch) == FALSE) %>%
+    # filter(stockid == "ATBTUNAEATL") %>% 
     group_by(stockid) %>%
     mutate(delta_year = year - lag(year)) %>%
     mutate(delta_year = case_when(year == min(year) ~ as.integer(1),
@@ -128,7 +129,7 @@ prepare_sofia_data <- function(min_years_catch = 20,
     group_by(stockid) %>%
     mutate(
       has_tb0 = !all(is.na(TB0)),
-      has_tb = !all(is.na(total_biomass)),
+      has_tb = all(!is.na(total_biomass)),
       first_catch_year = year[which(catch > 0)[1]]
     ) %>%
     filter(year >= first_catch_year) %>%
@@ -140,11 +141,11 @@ prepare_sofia_data <- function(min_years_catch = 20,
       b_rel = dplyr::case_when(
         has_tb0 ~ total_biomass / max(TB0),
         has_tb ~ total_biomass / max(total_biomass),
-        TRUE ~ b_v_bmsy / 2
+        TRUE ~ b_v_bmsy / 2.5
       )
     ) %>%
     mutate(approx_cpue = pmin(quantile(approx_cpue, 0.9, na.rm = TRUE), approx_cpue)) %>%
-    ungroup()
+    ungroup() 
   
   ram_b_plot <- ram_data %>%
     ggplot(aes(x = year, y = b_v_bmsy)) +

### 2020-06-17 functions/run-assessments.R

@@ -19,6 +19,7 @@ run_assessments <-
            Kscale = NA,
            estimate_shape = FALSE,
            estimate_proc_error = TRUE,
+           estimate_initial_state = TRUE,
            max_time = Inf,
            refresh = 0) {
     # fits <- vector("list", length(assessments))
@@ -26,7 +27,7 @@ run_assessments <-
     # names(fits) <- assessments
 
     fits <- list()
-
+    
     if ("sraplus" %in% assessments) {
       driors <-
         sraplus::format_driors(
@@ -63,6 +64,7 @@ run_assessments <-
           n_keep = keep,
           estimate_proc_error = estimate_proc_error,
           estimate_shape = estimate_shape,
+          estimate_initial_state = estimate_initial_state,
           refresh = refresh
         )
 
@@ -76,9 +78,10 @@ run_assessments <-
           draws = draws,
           n_keep = keep,
           estimate_proc_error = estimate_proc_error,
-          estimate_shape = estimate_shape
+          estimate_shape = estimate_shape,
+          estimate_initial_state = estimate_initial_state
         )
-
+        
 
       }
 

### 2020-06-17 make-assessing-global-fisheries.R

@@ -24,6 +24,8 @@ library(rstanarm)
 library(sf)
 Sys.unsetenv("PKG_CXXFLAGS")
 
+options(dplyr.summarise.inform = FALSE)
+
 rstan::rstan_options(auto_write = TRUE)
 
 sraplus::get_tmb_model()
@@ -39,7 +41,7 @@ draws <- 2500
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
-n_cores <- 3
+n_cores <- 2
 # number of cores for parallel processing
 
 future::plan(multiprocess, workers = n_cores)
@@ -50,24 +52,24 @@ future::plan(multiprocess, workers = n_cores)
 results_name <- "v0.5"
 
 
-results_description <- "publication version of results"
+results_description <- "publication version of results updated VOI from version 0.5 "
 
-run_voi_models <- FALSE
+run_voi_models <- TRUE
 # sub options for run_voi_models
 fit_models <- FALSE
 
-write_results <- FALSE
+write_results <- TRUE
 
 process_fits <- FALSE
 
 # other things
-run_case_studies <- FALSE
+run_case_studies <- TRUE
 
-run_continent_examples <- FALSE
+run_continent_examples <- TRUE
 
-run_ei_example <- FALSE
+run_ei_example <- TRUE
 
-run_sofia_comparison <- TRUE
+run_sofia_comparison <- FALSE
 
 engine = "stan"
 
@@ -109,8 +111,9 @@ prepare_sofia_data(lookup_fmi_names = FALSE)
 
 has_total_biomass = ram_data %>%
   group_by(stockid) %>%
-  summarise(hastb = any(!is.na(total_biomass))) %>%
-  filter(hastb)
+  summarise(hastb = !any(is.na(total_biomass)),
+            hasb = !any(is.na(b_v_bmsy))) %>%
+  filter(hastb & hasb)
 
 ram_data <- ram_data %>%
   filter(stockid %in% has_total_biomass$stockid) %>%
@@ -126,8 +129,7 @@ ram_data <- ram_data %>%
   group_by(stockid) %>%
   mutate(n = length(catch)) %>%
   filter(n >= min_years_catch) %>%
-  ungroup()
-
+  ungroup() 
 
 # fit voi models --------------------------------------------------------------
 
@@ -136,30 +138,60 @@ if (run_voi_models == TRUE){
   
   
   stress_tests <- data_frame(
[…]

### 2020-06-18 functions/run-assessments.R
back to abalone

@@ -81,7 +81,6 @@ run_assessments <-
           estimate_shape = estimate_shape,
           estimate_initial_state = estimate_initial_state
         )
-        
 
       }
 

### 2020-06-18 make-assessing-global-fisheries.R

@@ -41,7 +41,7 @@ draws <- 2500
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
-n_cores <- 2
+n_cores <- 3
 # number of cores for parallel processing
 
 future::plan(multiprocess, workers = n_cores)
@@ -56,11 +56,11 @@ results_description <- "publication version of results updated VOI from version 
 
 run_voi_models <- TRUE
 # sub options for run_voi_models
-fit_models <- FALSE
+fit_models <- TRUE
 
 write_results <- TRUE
 
-process_fits <- FALSE
+process_fits <- TRUE
 
 # other things
 run_case_studies <- TRUE
@@ -156,7 +156,7 @@ if (run_voi_models == TRUE){
     estimate_shape = sample(c(FALSE, TRUE), draws, replace = TRUE),
     estimate_proc_error = sample(c(FALSE, TRUE), draws, replace = TRUE)
   ) %>%
-    mutate(error_cv = map_dbl(error_cv, ~ ifelse(.x == 0, 1e-6, runif(1, 0.01, .25)))) %>%
+    mutate(error_cv = map_dbl(error_cv, ~ ifelse(.x == 0, 1e-6, runif(1, 0.05, 0.2)))) %>%
     mutate(
       b_ref_type = ifelse(
         initial_state_type == "prior",
@@ -179,12 +179,12 @@ if (run_voi_models == TRUE){
   #   index_window = sample(c(1, 0.25, .5), draws, replace = TRUE),
   #   index_freq = sample(c(1, 2), draws, replace = TRUE),
   #   u_window = sample(c("snapshot", "recent", "complete"), draws, replace = TRUE),
-  #   b_ref_type = sample(c("k", "b"),  draws, replace = TRUE),
+  #   b_ref_type = sample(c("b"),  draws, replace = TRUE),
   #   f_ref_type = sample(c("f", "fmsy"),  draws, replace = TRUE),
   #   estimate_shape = sample(c(FALSE, TRUE), draws, replace = TRUE),
   #   estimate_proc_error = sample(c(FALSE, TRUE), draws, replace = TRUE)
   # ) %>%
-  #   mutate(error_cv = map_dbl(error_cv, ~ ifelse(.x == 0, 1e-6, runif(1, 0.01, .25)))) %>%
+  #   mutate(error_cv = map_dbl(error_cv, ~ ifelse(.x == 0, 1e-6, runif(1, 0.05, 0.2)))) %>%
[…]

### 2020-06-18 make-assessing-global-fisheries.R
VOI analysis up and running again

@@ -56,11 +56,11 @@ results_description <- "publication version of results updated VOI from version 
 
 run_voi_models <- TRUE
 # sub options for run_voi_models
-fit_models <- TRUE
+fit_models <- FALSE
 
 write_results <- TRUE
 
-process_fits <- TRUE
+process_fits <- FALSE
 
 # other things
 run_case_studies <- TRUE
@@ -415,7 +415,9 @@ if (run_voi_models == TRUE){
   wtf <- fits %>% 
     filter(use_terminal_state == TRUE,
            metric == "b_v_bmsy",
-           b_ref_type == "b") %>% 
+           b_ref_type == "b",
+           error_cv == min(error_cv),
+           use_index == FALSE) %>% 
     arrange(desc(rmse))
   
   View(wtf)
@@ -603,7 +605,7 @@ if (run_voi_models == TRUE){
   
   u_voi_fit <-
     rstanarm::stan_glmer(
-      rmse ~  initial_state_type  + use_index  + use_u_priors+ u_window + estimate_shape + use_terminal_state + (1|stockid),
+      rmse ~  initial_state_type  + use_index*u_window + estimate_shape + estimate_proc_error + use_terminal_state + (1|stockid),
       data = voi_data %>% filter(metric == "u_v_umsy", is.finite(rmse)),
       cores = 4,
       family = Gamma(link = "log")
@@ -611,7 +613,7 @@ if (run_voi_models == TRUE){
   
   u_voi_plot <- bayesplot::mcmc_areas(
     as.array(u_voi_fit),
-    regex_pars = c("use", "state", "estimate"),
+    regex_pars = c("use", "state", "estimate","index","u_"),
     prob = 0.8,
     # 80% intervals
     prob_outer = 0.95,
@@ -655,7 +657,7 @@ if (run_voi_models == TRUE){
   
   acc_voi_fit <-
     rstanarm::stan_glmer(
-      bin_acc ~  use_initial_state + use_index + use_u_priors+ use_terminal_u + estimate_shape + (1|stockid) ,
+      bin_acc ~  initial_state_type  + use_index*u_window + estimate_shape + estimate_proc_error + use_terminal_state + (1|stockid),
       data = acc_data,
       cores = 4,
       family = binomial()
@@ -664,7 +666,7 @@ if (run_voi_models == TRUE){
   
   acc_voi_plot <- bayesplot::mcmc_areas(
     as.array(acc_voi_fit),
[…]

### 2020-07-21 functions/fit-fao.R
merging back to master then reorganizing a bit

@@ -57,6 +57,37 @@ fit_fao <-
       mutate(data = "heuristic")
     
     
+    # catch only fit
+    
+    com_driors <-
+      format_driors(
+        taxa = scientific_name,
+        catch = catches,
+        years = seq_along(years),
+        use_heuristics = FALSE,
+        initial_state = NA,
+        initial_state_cv = NA,
+        b_ref_type = "k",
+        isscaap_group = unique(data$isscaap_group),
+        use_catch_prior = TRUE
+      )
+    
+    com_fit <-
+      fit_sraplus(
+        driors = com_driors,
+        include_fit = include_fit,
+        model = model,
+        engine = "sir",
+        draws = draws,
+        thin_draws = thin_draws
+      )
+    
+    com_fit_results <- com_fit$results %>%
+      filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy", "depletion")) %>%
+      mutate(year = rep(years, each = 4)) %>%
+      mutate(data = "catch_only")
+    
+    
     # sar fit
     
     if (!is.na(last(data$sar))) {
@@ -528,7 +559,8 @@ fit_fao <-
       used_cpue_plus <- FALSE
     }
     
-    results <- basic_fit_results %>% {
+    results <- basic_fit_results %>% 
+      bind_rows(com_fit_results) %>% {
       if (used_sar == TRUE) {
         bind_rows(., sar_fit_results)
       }

### 2020-07-21 make-assessing-global-fisheries.R

@@ -1511,11 +1511,21 @@ fao_status <- fao_status %>%
 
 support_data <- list(mean_regional_isscaap_fmi = mean_regional_isscaap_fmi)
 
-areas <- c(67, 57, 37,71)
+# areas <- c(67, 57, 37,71)
 
 areas <- unique(fao_status$fao_area_code)
-# annnnnd try and run assessments
 
+# fao_status %>% 
+#   filter(year == max(year)) %>% 
+#   summarise(catch = sum(catch))
+# 
+# fao %>% 
+#   filter(year == max(year)) %>% 
+#   summarise(catch = sum(capture))
+
+
+# annnnnd try and run assessments
+# browser()
 if (run_sofia_comparison == TRUE) {
   # future::plan(future::multiprocess, workers = 4)
   
@@ -1528,14 +1538,14 @@ if (run_sofia_comparison == TRUE) {
     group_by(stockid) %>%
     nest() %>%
     ungroup() %>%
-    # sample_n(3) %>% 
+    # sample_n(3) %>%
     mutate(
       fits = future_map(
         data,
         safely(fit_fao),
         support_data = support_data,
         default_initial_state = 1,
-        min_effort_year = 1960,
+        min_effort_year = 1975,
         engine = "stan",
         cores = 1,
         .progress = TRUE,
@@ -1572,10 +1582,11 @@ if (run_sofia_comparison == TRUE) {
   
 
 }
+
 reserve <- fao_status_fits
 
 ugh <- map_dbl(fao_status$data, ~unique(.x$area)) == 67
-#
+
 fao_worked <- map(fao_status_fits$fits, "error")
 
 

### 2020-07-21 functions/fit-fao.R
Update voi (#3)
* VOI analysis up and running again

@@ -9,7 +9,7 @@ fit_fao <-
            q_slope_prior = 0.025,
            estimate_proc_error = TRUE,
            estimate_qslope = FALSE,
-           default_initial_state = 1,
+           default_initial_state = NA,
            default_initial_state_cv = 0.2,
            chains = 1,
            cores = 1,
@@ -57,6 +57,37 @@ fit_fao <-
       mutate(data = "heuristic")
     
     
+    # catch only fit
+    
+    com_driors <-
+      format_driors(
+        taxa = scientific_name,
+        catch = catches,
+        years = seq_along(years),
+        use_heuristics = FALSE,
+        initial_state = NA,
+        initial_state_cv = NA,
+        b_ref_type = "k",
+        isscaap_group = unique(data$isscaap_group),
+        use_catch_prior = TRUE
+      )
+    
+    com_fit <-
+      fit_sraplus(
+        driors = com_driors,
+        include_fit = include_fit,
+        model = model,
+        engine = "sir",
+        draws = draws,
+        thin_draws = thin_draws
+      )
+    
+    com_fit_results <- com_fit$results %>%
+      filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy", "depletion")) %>%
+      mutate(year = rep(years, each = 4)) %>%
+      mutate(data = "catch_only")
+    
+    
     # sar fit
     
     if (!is.na(last(data$sar))) {
@@ -100,7 +131,7 @@ fit_fao <-
     # fmi fit
     fmi_dat <- support_data$mean_regional_isscaap_fmi %>%
       filter(
-        fao_area_code == unique(as.numeric(data$fao_area)),
+        fao_area_code == unique(as.numeric(data$fao_area_code)),
         isscaap_group == unique(data$isscaap_group)
       )
     
@@ -249,7 +280,6 @@ fit_fao <-
           isscaap_group = unique(data$isscaap_group),
           q_slope_prior = q_slope_prior
         )
-      
       sfs <- safely(fit_sraplus)
       
       cpue_fit <-
@@ -529,7 +559,8 @@ fit_fao <-
       used_cpue_plus <- FALSE
     }
     
-    results <- basic_fit_results %>% {
+    results <- basic_fit_results %>% 
+      bind_rows(com_fit_results) %>% {
       if (used_sar == TRUE) {
         bind_rows(., sar_fit_results)
       }

### 2020-07-21 functions/run-assessments.R

@@ -19,6 +19,7 @@ run_assessments <-
            Kscale = NA,
            estimate_shape = FALSE,
            estimate_proc_error = TRUE,
+           estimate_initial_state = TRUE,
            max_time = Inf,
            refresh = 0) {
     # fits <- vector("list", length(assessments))
@@ -26,7 +27,7 @@ run_assessments <-
     # names(fits) <- assessments
 
     fits <- list()
-
+    
     if ("sraplus" %in% assessments) {
       driors <-
         sraplus::format_driors(
@@ -63,6 +64,7 @@ run_assessments <-
           n_keep = keep,
           estimate_proc_error = estimate_proc_error,
           estimate_shape = estimate_shape,
+          estimate_initial_state = estimate_initial_state,
           refresh = refresh
         )
 
@@ -76,9 +78,9 @@ run_assessments <-
           draws = draws,
           n_keep = keep,
           estimate_proc_error = estimate_proc_error,
-          estimate_shape = estimate_shape
+          estimate_shape = estimate_shape,
+          estimate_initial_state = estimate_initial_state
         )
-
 
       }
 

### 2020-07-21 make-assessing-global-fisheries.R

@@ -24,6 +24,8 @@ library(rstanarm)
 library(sf)
 Sys.unsetenv("PKG_CXXFLAGS")
 
+options(dplyr.summarise.inform = FALSE)
+
 rstan::rstan_options(auto_write = TRUE)
 
 sraplus::get_tmb_model()
@@ -39,7 +41,7 @@ draws <- 2500
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
-n_cores <- 6
+n_cores <- 3
 # number of cores for parallel processing
 
 future::plan(multiprocess, workers = n_cores)
@@ -50,13 +52,13 @@ future::plan(multiprocess, workers = n_cores)
 results_name <- "v0.5"
 
 
-results_description <- "publication version of results"
+results_description <- "publication version of results updated VOI from version 0.5 "
 
 run_voi_models <- FALSE
 # sub options for run_voi_models
 fit_models <- FALSE
 
-write_results <- FALSE
+write_results <- TRUE
 
 process_fits <- FALSE
 
@@ -109,8 +111,9 @@ prepare_sofia_data(lookup_fmi_names = FALSE)
 
 has_total_biomass = ram_data %>%
   group_by(stockid) %>%
-  summarise(hastb = any(!is.na(total_biomass))) %>%
-  filter(hastb)
+  summarise(hastb = !any(is.na(total_biomass)),
+            hasb = !any(is.na(b_v_bmsy))) %>%
+  filter(hastb & hasb)
 
 ram_data <- ram_data %>%
   filter(stockid %in% has_total_biomass$stockid) %>%
@@ -126,8 +129,7 @@ ram_data <- ram_data %>%
   group_by(stockid) %>%
   mutate(n = length(catch)) %>%
   filter(n >= min_years_catch) %>%
-  ungroup()
-
+  ungroup() 
 
 # fit voi models --------------------------------------------------------------
 
@@ -136,31 +138,61 @@ if (run_voi_models == TRUE){
   
   
   stress_tests <- data_frame(
-    use_index = sample(c(TRUE,FALSE),  draws, replace = TRUE),
+    use_index = sample(c(TRUE, FALSE),  draws, replace = TRUE),
     index_type = sample(c("survey"), draws, replace = TRUE),
-    use_initial_state = sample(c(TRUE, FALSE),  draws, replace = TRUE),
-    use_heuristics = sample(c(TRUE, FALSE),  draws, replace = TRUE),
[…]

### 2020-07-21 documents/ovando-etal-assessing-global-fisheries.Rmd

--- 2020-05-22 documents/ovando-etal-assessing-global-fisheries.Rmd
@@
 date: "`r Sys.Date()`"
-bibliography: ["references.bib"]
+bibliography: ["../references.bib"]
 csl: nature.csl
@@
 pub_theme <- theme_ipsum(base_size = 12, 
-                         axis_text_size = 12)
+                         axis_text_size = 12) + 
+    theme(panel.spacing = unit(1,"lines"))
 
@@
 *Insert mandatory opening sentence about the importance of global fisheries*.
-On-the ground fisheries management decisions are ideally based on assessment models, designed to provide robust estimates of individual stock status that can inform effective management policies.
-However, data and resource limitations prohibit our ability to create bespoke tactical assessments for every fishery resource around the globe.
-As such, we also depend on widely applicable methods intended to provide a general sense of the status of fisheries around the globe.
-While not suitable for management of individual stocks, these broad strategic models play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals.
+The United Nations Sustainable Development Goal 14 calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
+A substantial part of achieving this goal rests on ensuring that marine fisheries are managed levels that ensure the conservation and sustainable use of ocean ecosystems.
+Achieving this goal depends in part on our ability to effectively track the status of global marine fish stocks.
+The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking this important metric.
+As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fishery, 6.2% are underfished ,and 34.2% are overfished [@fao2020].
 
-The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the most widely followed estimate of global stock status.
-As of the most recent report, the FAO estimates that 59.9% of marine fish stocks are maximally sustainably fishery, 7% are underfished ,and 33.1% are overfished [@fao2018].
-While foundational, the SOFIA reports depend on a complex process of aggregating results from tactical models, expert judgement, etc.
-**make clear argument why other/new global assessment methods are needed**.
+However, there are two major areas of improvement needed to ensure that we are able to accurately track the status of global fisheries.
+XX% of the global landings covered by the SOFIA reports come from fisheries with formal statistical stock assessments, quantitative models that provide estimates of stock status based on the best available science.
+These stock assessments are then incorporated into the SOFIA report's estimates of global stock status.
+The remaining XX% of the stocks/landings covered by SOFIA lack formal stock assessments.
+For these important fisheries, FAO scientists must use a mixture of local expert knowledge, data synthesis, and expert opinion to infer stock status.
+While this process provides the best available estimate of stock status, ideally a more reproducible  XX something to communicate why we might want to improve things for these stocks XX.
 
-Several publications have explored alternative methods for estimate global stock status, including Costello et al.
-2016 [@costello2016a]  and Rosenberg et al.
-2018 [@rosenberg2018], and Pauly 2007 [@pauly2007].
+Second, there are many fisheries that while perhaps small on a global scale provide critical ecosystem services to their adjacent communities.
+These fisheries collectively, making up XX% of global landings, are not contained in the FAO assessment, but still fall under the SDG 14 goal XX to what extent is this true?
+XX rishi would be good to work on this section together.
+By my count the total catch form the SOFIA table sent to me is 42.68 MMT, so about half of global catch, but SOFIA 2020 says that "78.7 percent of current landings come from biologically sustainable stocks.", which would imply then that basically all landings are in SOFIA (since that doesn't leave much room for the assessed but overfished)?
+Confused by discrepancy hereXX.
+
+Ideally, both of these problems would be solved by simply running formal stock assessments on all significant fisheries.
+This is rendered impossible though both by lack of data and human resources.
+Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance such as catch-per-unit-effort (CPUE).
+While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect.
+Even when data are available though, running an individualized stock assessment is not a "just press play" process.
+Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment.
+This process takes considerable time and expertise.
+As such, effectively tracking the status of all global fisheries will require alternative methods.
+
+To resolve this problem, numerous papers have developed methods that while not intended for the management of individual stock are hoped to provide reasonable estimates of stock status in aggregate, particularly for those regions lacking the data or capacity for formal stock assessment.
+These include @costello2016a, @rosenberg2018, and @pauly2007.
 Each of these methods sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.
+While not suitable for management of individual stocks, these broad strategic models have played an important role in shaping the public narrative around ocean health.
+
 However, comparing estimates from each of these methods against best-available knowledge of stock status from the RAM Legacy Stock Assessment Database [@ricard2012], as well as broader intuition.
 reveals some systemic problems with these assessments.
-
 We have stock assessment data from RAM from the NE Pacific,the NE Atlantic and the Mediterranean, and of these the NE Pacific is in the best state (mean RAM B/B~MSY~ =1.8), NE Atlantic (mean RAM B/B~MSY~ =1.3) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.5).
@@
 
-Despite numerous advances in statistical methods, we are currently left with strong questions about our understanding of the state of global fisheries, particularly in the regions of the world lacking formal stock assessments.
-We build off of these prior works to demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection effort.
+The coming decades are a critical period for the future of global oceans.
+Climate change, population growth, and economic expansion will all place increasing pressure on fisheries.
+Mitigating these threats, as called for by SDG 14, will require an improved ability to track the status of fisheries around the world.
+At the same time, new sets of globally distributed data have recently become available, along with advances in statistical modeling methods.
+We leverage these opportunities to provide a global assessment of our perception of global stock status resulting from different use of different data sources.
+Our results demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection and interpretation effort.
 
@@
 We use a flexible assessment package ([sraplus](https://github.com/DanOvando/sraplus#sraplus)) to compare estimates of stock status around the globe using different combination of these data.
-We find that our perception of global stock status varies wildly depending on which data source are used, without any clear evidence of a "best" source of data to use *this needs to be much clearer*.
 
-## Case Study: Europe vs. Northeast Asia or East India
+## Case Studies
 
-Don't know if this section is needed? Key messages if so (and might want to pick just one case study)
+Evaluating the validity of data-limited XX need to introduce this idea earlier xx assessment methods can be a challenging task.
+Simulation testing provides one avenue, but can be difficult to create a simulation environment that adequately captures the problems of real-world ecosystems and data.
+We can also compare results between data-limited assessment and more data (or resource) rich gold-standard stock assessments.
+However, formally stock asssessed fisheries represent a very biased subset of global fisheries, and it is entirely possible that a method that performs well for the kinds of species and fisheries that are commonly stock assessed might struggle in the kinds of fisheries that typically are unassessed.
 
-- Europe vs North East Asia
-  - sort of a more positive story: adding in the Rousseau effort data at least produces a more plausible split between Europe and Northeast Asia
-  - Nicely illustrates bi-modal heuristics behavior. 
-  - Also lends to shortcomings though: assuming one effort series for all stocks clearly produces some weird results, e.g. all those basically unfished EU stocks
+Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region.
+As outlined in Table.1, while we often have no proof that results finding that for example fisheries in south-east Asia have similar stock status to those in Alaska, very few experts would agree that this is a plausible result XX tough statement to make in a paper xx.
+We use two case studies here to present the potential of additional, broadly applicable, data to potentially improve on "catch-only" methods.
+
+For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region.
+We can then compare the estimates of stock status that come from using the catch-data alone (with a heuristic algorithm in the manner of @martel2013), with the catch data along with priors on terminal depletion informed by the swept-area-ratio data, and using the swept-area ratio data together with a local abundance index.
+By it's nature the catch-only heuristic produces a bi-model distribution, with most fisheries reported to have a B/B~MSY~ of 1.5 and an F/F~MSY~ less than 1 (Fig.\@ref(fig:east-india)-heuristic).
+Local experts put the swept-area-ratio (SAR) value for the region at an incredibly high 17.
+Using this value to inform a prior on terminal stock status dramatically changes our picture of stock status in the Eastern Bay of Bengal, with now all stocks being overfished and experiencing high levels of overfishing (Fig.\@ref(fig:east-india)-sar).
+
+If all the stocks in question were demersal, shared similar life histories, and were all caught in the same multi-species fleet, this result might make sense.
+However, imposing the same prior on terminal stock status for all fisheries in a region is not ideal.
+Supplementing these data with a local index of abundance provides a more nuanced figure, with most stocks estimated as having B/B~MSY~ less than 1, but with a much wider range of fishing mortality rates reflecting difference in recent catch histories among individual stocks (Fig.\@ref(fig:east-india)-"cpue + sar").
+While we do not know the true state of stocks in the Eastern Bay of Bengal, inclusion of local data provide a picture more in line with local perception XX.
+
+```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio.
+Each point is a stock."}
+ei_kobe_plot 
+```
+
+It should be no surprise that incorporating locally tailored data produces more locally plausible results.
+Global assessment at the moment though requires depending on a few broadly available datasets.
+To give an illustration, we used `sraplus` to assess the status of in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from @rousseau2019 (xx technically this is the updated dataset xx).
+Interestingly, catch histories for Europe and Northeast Asia look somewhat similar xx plot of catch and effort histories xx.
+As a result, using a "catch-only" heuristic, both Europ and Northeast Asia are estimated to have very similar stock status, a similar finding as those illustrated in Table.XX.
+Adding in the effort data (as an effort index created for each stock as catch divided by the effective harvest rate based on the effort data), and the two regions begin to diverge.
+While according to @rousseau2019, Europe has seen a recent decrease in fishing effort from a peak in XX, effort in Northeast Asia has increased more or less continuosly.
+As a result, `sraplus` estimates that more stocks in Northeast Asia are overfished / experiencing overfishing than in Europe.
+However, the shared effort series across these stocks aslo produces a large number of effectively unfished stocks; a highly unlikely result for both regions.
+
+XX need to redo this analyis with the new effort data XX
 
@@
 
-- East India
-  - Local data can help: heuristic alone bi-modal and optimistic, SAR alone really pessimistic, CPUE + SAR a bit more nuanced?
-  - A little bit a of a conflict here, since paper so far has been about global datasets, and here we're introducing a more local study (trawl survey abundance from eastern bay of Bengal)
+## Conflicting Estimates of Global Status
 
-```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio.
-Each point is a stock."}
-ei_kobe_plot
-```
+This exercise reveals the potential and challenges using broadly available data for global fishery assessment.
+On the one hand, for both case studies moving beyond catch data alone produced a more plausible set of results.
+However, both results also raise a host of new issues.
+How realistic is it to share one index of abundance and SAR value for the entire eastern Bay of Bengal?
+Similarly, the sharing of a common effort series at the continental level does provide increased separation between European and northeast Asian stocks that may be plausible, but also estimates a suspicious amount of stocks to be essentially unfished (a dynamic which does not match assessments from that region XX plot RAM stocks for europe XX).
 
-## Conflicting Estimates of Global Status
+Extending these issues to the global level, we assessed the status of stocks assessed by the FAO's latest SOFIA report using various types of broadly available data.
+Given that the SOFIA assessments present the best estimates of stock status across both assessed and unassessed stocks, we would hope that any assessment method suitable would produce broadly match the values reported by the FAO.
+We focus here on the percent of stocks classified as "overfished" (B/B~MSY~ < 0.8) in each FAO statistical region.
+Following the SOFIA methodology, "Stocks" are classified by the sum catch histories across the same species within a major FAO statistical region.
+
+Our results show that our perception of global fishery depends completely on which data sources we utilize.
+That different data tell different stories should not be surprising.
+However, we also do not see a convergence towards a cohesive picutre of global status that closely matches the SOFIA report as we add more data.
+The most data-poor method employed, the catch heuristic alone, matches some patterns from the SOFIA report (little overfishing in the Northeast Pacific, some overfishing in the northwest Atlantic), it misses the mark badly in other regions, e.g.
+the waters around New Zealand.
+Our most data-rich scenario, an effort series from @rousseau2019, along with priors from swept-area-reatio and fisheries management index where available....
+
+to provide more plausible seeming results than catch data alone.
+A reasonable question then is, can we simply augment global catch statistics with the data presented here to provide an improved understanding of global fisheries?
 
@@
 A few key questions here
-
-- Should we use a more recent FAO report? Don't remember why we chose the 2011 report for the original report
 
@@
 
+** how do we know that these assessments themselves aren't actually the right ones"
+
+**address better than nothing**.
+Certainly not advocating for sitting on hands until better models come around, but important to point out that these models don't just produce noisy but unbiased estimates, they seem to be fundamentally biased in many places.
+But, no one method appears to be consistently biased (except for catch only)
+
 *these are just rough outline paragraphs*
+
+we do not claim that our model is the best model possible, or that others could not make improvements on our methods.
+Rather, we claim that any improvements that could be made, given the same data and broadness of scope, would likely be marginal relative to the inherent limitations of the data themselves.
 
@@
 
-Description of the package itself focused on key differences to other packages
+The core of our model is a Pella-Tomlinson [@pella1969] production model parameterzied in the manner of @winker2018.
+While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
+
+XX equations here XX.
+
+All of our estimates are Bayesian in nature.
+We can break the use of our model into two distinct categories: with data and without.
+By "data", we refer to measurements which are confronted with model estimates within a likelihood function.
+In our context, these include fishery-independent survey data, or a CPUE index.
+When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
+Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population, given the constraints of the model.
+
+We use a Bayesian methods as it suits the philosophy of our analysis: incorporating prior knowledge about stock status drawn from expert opinion or other similar stocks, and where possible confronting those priors with data.
+However, this process does introduce a problem, in that it a) certain parts of the prior probability distribution have zero support in the data (e.g.
+carrying capacity's so small that they crash the population) and b) they create priors on both the *input* (e.g.
+growth rates and carrying capacity) and *outputs* (recent stock status), effectively putting two prior distributions for the same outcome, a problem termed Borel's Paradox [@puntXX;@rafferty1997].
+This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
+However, particularly for the data-less version of our model, Borel's Paradox poses a particular problem.
+
+Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
+Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of r and k together with the model produces a deterministic stock status in each time step.
+Suppose then we wish to also provide a prior on stock status in the most recent year, for example to reflect local expert opinion based on sound reasoning.
+We have now placed two priors on stock status in the most recent year, one implicit and one explicit, violating Borel's Paradox.
+The net effect of this is that once these priors are run through the model, through for example a sample-importance-resampling (SIR) algorithm as generally applied in stock-reduction analysis, our post-model-pre-data "estimate" of stock status will appear to have "updated".
+This creates the illusion that we have learned something about stock status, when in fact we have simply used our model to reveal our true implied prior on stock status.
+This is technically incorrect, since we do not have a likelihood in the model at this point, and as such per Bayes Theorem have not updated our prior.
+Of more practical importance, this "implied prior" will always reflect a more optimistic stock status then our original prior on stock status, due to the constraints of the population model.
+
+The SRA algorithm works in two steps.
+First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep).
+From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status, e.g.
+if the bulk of the prior on terminal stock status was concentrated at 50% of *k*, combinations of *r* and *k* that produce terminal stock status near 50% of *k* are sampled proportionally more frequently.
+However, lower values of terminal stock status have fewer candidate values of *r* and *k*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
+Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *k* combinations: since under this model the population cannot be greater than carrying capacity, as for example *k* approaches infinity terminal stock status asymptotes at to 100% of *k*.
+The net result of this is that even though individual combinations of *r* and *k* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
+As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.
+
+To resolve this, @puntXX (either cite that "on assessment of the bering-chukchi-beufort sea stock of bowhead whales paper that I can't find a date for, or see if he has a new paper on thisXX) suggests a "backwards" fitting approach, in which the model estimates growth parameters and a terminal depletion, and given the model backs out a value of *k* that satisifes the priors (conditional on an assumption initial stock status).
+We do not adopt this approach here due to computational issues.
+Adopting the "backwards" approach requires numerically searching for the plausible *k* during each iteration of the model, which is time consuming.
+In addition, the backwards approach is not without its own complicaitons, for example the requirement of a tolerance around the value of terminal depletion produced by a given *k* and the actual value selected by the model in that time step.
+This discrepency can occur when a given terminal stock status is not technically possible given the model; to give an extreme example under this model no value of *k* can produce a terminal stock status of 0 if catches are 0 for the entire history of the fishery (assuming *k* > 0).
+
+We present a novel approximation to this problem here.
+Our solution amounts to a two-step SIR algorithm.
+We first run the standard SIR algorithm as described above.
+We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
+
+$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
+
+We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
+
+$$p(n_i) = \frac{p(bin_i)}{N_i}$$
+
+And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
+The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
+In effect, this answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
+
+XX insert figure if there's space illustrating this.
+
+@raftery1993
+
+## Priors and Data
+
+We briefly describe the construction of priors and nature of available data.
+All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [xx citation].
+These models are intended as a reasonably robust means of translating available data into usable prior for our model.
+Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with fishery status indicators of interest.
+Rather, each regression was tested to ensure that it is unlike that, given the same data, and alternative model would be unlikely to perform much better than those presented here XX this might be too much defenseXX
+
+For each statistical prior, we fit three different models, relating the data of interesting, e.g. swept area ratio (SAR), to each of *B/B~MSY~*, *F/F~MSY~*, and *F*
 
@@
 
-Basic description and out-of-sample predictive power
+@amoroso2018 provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
+This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
+As illustrated in @amoroso2018, there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
+Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
+if all trawl-able areas are trawled twice a year then the SAR will be 2.
+Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
+
+The final selected model relating SAR to to stock status metrics was
+
+$$log(value_i) \sim normal(log(SAR_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR}) $$
+
+XX testing training plot XX
 
@@
 
-Basic description and out-of-sample predictive power
+The Fisheries Management Index (FMI), as presented in @melnychuk2016, utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
+These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomics.
+The higher the score, the better the expert judges that a given metric is met in that fishery.
+Importantly, FMI surveys can be filled out in the absence of stock assessments.
+This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to @osio2015 and @cope2015).
 
-## Catch-Only Priors?
+The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
 
-At the moment only using Heuristics.
-I've been exploring an update to the regression based catch-only models, can get up to an out of sample r^2 of about 0.3 and that's about it.
-Probably worth putting in just to make clear that we're making about as much use out of the catch data alone as we possible can.
+$$log(value_i) \sim N(s(research_i) + s(management_i) +
+s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
+
+XX testing training plot XX
+
+## Catch-Only Priors
+
+Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012;@costello2016;@rosenberg2018;@pauly2007].
+While these catch-only methods have been shown to have series shortcomings [@free2020], as a point of reference we use two forms of catch-only priors: a statistical and a heuristic approach.
+
+For the statistical approach, we utilize data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics...
+
+For the heuristic approach, we borrow from the original methods of @martell2013 as presented in @andersno2017, and define a heuristic on initial and terminal depletion where
+
+```r
+        if (first(catch) / max(catch) < 0.2){
+
+        initial_depletion = 0.7
+
+        } else {
+
+        initial_depletion = 0.4
+
+        }
+
+        if (last(catch) / max(catch) > 0.5){
+
+        terminal_depletion <- 0.6
+
+        } else {
+
+        terminal_depletion <- 0.2
+
+        }
+
+```
+
+We also define two statistical regressions relating catch history to stock status.
+To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheres in RAM, in order identify four possible clusters of catch history types within the the data.
+Emergent clusters show for example one build around "one way trip" style catch histories, others with a boom and bust, etc.
+
+We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history.
+This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term).
+For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status
+
+$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma) $$
+
+For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data
+
+$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
+where *fyear* is the year of the fishery, starting from 0.
 
@@
 
-Description of Rousseau data, steps taken to use it (Baranov equation, choice about tech. creep etc.)
+While the FMI and SAR data show some potential for estimating U/U~MSY~ and B/B~MSY~ values, they are clearly not an ideal measure of stock status.
+An alternative is to actually develop indices of abundance that we can fit models to.
+@rousseau2019 provides a useful starting place for this exercise, presenting estimates of fishing effort throughout the globe from 1950 to 2015.
+Merging these data with capture statistics can provide a crude CPUE index that can then be used to fit a surplus production model.
+Developing an index of abundance in this manner is clearly a very rough analysis only intended to illustrate broad concepts.
+Any analysis of CPUE data can be severely affected by changes in technology, skill, hyperdepletion, hyperstability, changes in the spatial footprint of the fishery, and selectivity [@maunder2004a; @walters2003].
+In the case of this analysis, a kilowatt day of fishing with 1950s technology clearly represents a different amount of fishing power than a kilowatt day of fishing in 2012.
+To account for this somewhat, @rousseau2019 include a 2.6% annual "technological creep" factor, which we incorporate in our analysis.
+
+@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6\% technology rate).
+This assumes that every unit increase in effort is linearly greater than the unit of effort before it.
+When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
+This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
+To accomodate this, we generate a catch per effective harvest rate index of abundance, as
+
+$$I_t = \frac{catch_t}{(1 - e^{-q_tE_t})}$$
+
+Where $q_t$ can has a technology rate component $\tau$
+
+$$q_t = q_{t-1} \times (1  + \tau)$$
 
@@
 
-## Estimation model
-
-Important details of sraplus, basic equations, etc. Focus on prior-predictive tuning process to validate use of priors on terminal status.
-
 # References

### 2020-07-22 functions/fit_continent_examples.R
cleaning up

@@ -84,7 +84,7 @@ fit_continent_examples <-
         years = seq_along(years),
         effort = data$total_effort[effort_years],
         effort_years = effort_years,
-        q_slope_prior = 0.025,
+        q_slope_prior = q_slope_prior,
         q_slope_prior_cv = 0.2,
         b_ref_type = "k"
       )

### 2020-07-22 functions/prepare-sofia-data.R

@@ -556,6 +556,8 @@ prepare_sofia_data <- function(min_years_catch = 20,
   fmi$fao_country_name <-
     countrycode::countrycode(fmi$country_rfmo, "country.name", "un.name.en")
   
+  fmi$country <- countrycode::countrycode(fmi$country_rfmo, "country.name","un.name.en")
+  
   fmi$region <-
     countrycode::countrycode(fmi$country_rfmo, "country.name", "region")
   
@@ -768,6 +770,17 @@ prepare_sofia_data <- function(min_years_catch = 20,
     mutate(country =countrycode::countrycode(country, "country.name", "un.name.en"))
   
   # assign("effort_region_to_country",effort_region_to_country, envir = .GlobalEnv )
+  
+  rous_data <- read.csv(here("data", "MappedFAO.csv")) %>% 
+    na.omit() %>% 
+    as_tibble() %>% 
+    janitor::clean_names() %>% 
+    mutate(country = countrycode::countrycode(iso3, "iso3c", "un.name.en")) %>% 
+    filter(type2 == "I") %>% 
+    select(year,fao, effort_cell_reported_nom, country ) %>% 
+    rename(area = fao)
+  
+  assign("rous_data",rous_data, envir = .GlobalEnv )
   
   
   # ugh. manual refactoring of region codes

### 2020-07-22 make-assessing-global-fisheries.R

@@ -67,7 +67,7 @@ run_case_studies <- FALSE
 
 run_continent_examples <- FALSE
 
-run_ei_example <- FALSE
+run_ei_example <- TRUE
 
 run_sofia_comparison <- TRUE
 
@@ -371,7 +371,6 @@ if (run_voi_models == TRUE){
       summary_plot = map(fit, "summary_plot"),
       summary = map(fit, "fit_summary")
     )
-  browser()
   # test <- test %>%
   #   mutate(
   #     summary_plot = map_plot(fit_performance, "summary_plot"),
@@ -549,7 +548,6 @@ if (run_voi_models == TRUE){
   # value of information analyses
   
   # how does the comparison stack up when you have more data?
-  browser()
   rmse_bias_plot <- noerror_data %>%
     ggplot(aes(rmse, bias)) +
     geom_hex(aes(fill = ..density..), binwidth = c(0.1, 0.1)) +
@@ -782,7 +780,6 @@ if (run_voi_models == TRUE){
   
 } # close fit voi models
 
-
 # run asia-eu comparison ------------------------------------------------------
 
 
@@ -793,21 +790,20 @@ support_data <- list(
 
 
 
-total_nominal_effort <- effort_data %>%
-  filter(effort_type == "nominal") %>%
+total_nominal_effort <- rous_data %>%
+  left_join(effort_region_to_country, by = "country") %>% 
   group_by(region,year) %>%
-  summarise(total_effort = sum(effort)) %>%
+  summarise(total_effort = sum(effort_cell_reported_nom, na.rm = TRUE)) %>%
   ungroup() %>%
   mutate(region = tolower(region))
 
 
-
-total_effecive_effort <- effort_data %>%
-  filter(effort_type == "effective") %>%
-  group_by(region,year) %>%
-  summarise(total_effort = sum(effort)) %>%
-  ungroup() %>%
-  mutate(region = tolower(region))
+# total_effecive_effort <- effort_data %>%
+#   filter(effort_type == "effective") %>%
+#   group_by(region,year) %>%
+#   summarise(total_effort = sum(effort)) %>%
+#   ungroup() %>%
+#   mutate(region = tolower(region))
 
 
 total_nominal_effort %>%
@@ -836,22 +832,25 @@ total_stocks <- fao %>%
   mutate(continent = tolower(continent)) %>%
[…]

### 2020-07-22 functions/prepare-sofia-data.R
something weird happened to east india

@@ -831,6 +831,8 @@ prepare_sofia_data <- function(min_years_catch = 20,
   ei_capture <-
     readxl::read_xlsx(here::here("data", "India East Only.xlsx")) %>%
     janitor::clean_names() %>%
+    select(-total, -ratio, -x78) %>% 
+    filter(!is.na(country_country)) %>% 
     gather(year, capture, x1950:x2015) %>%
     mutate(year = as.numeric(str_replace_all(year, "\\D", "")))
  

### 2020-07-22 make-assessing-global-fisheries.R

@@ -1026,7 +1026,7 @@ if (run_ei_example == TRUE) {
   sfi <- safely(fit_india)
   
   ei_fits <- ei_data %>%
-    # s(4) %>%
+    # sample_n(4) %>%
     # filter(scientific_name == "Sphyraena spp") %>%
     mutate(ex_fits = map2(
       scientific_name,
@@ -1039,7 +1039,7 @@ if (run_ei_example == TRUE) {
         index_years = .y$year[!is.na(.y$cpue) & .y$year >= 1990],
         support_data = support_data,
         country = "India",
-        engine = "tmb",
+        engine = "stan",
         sar = 18.7,
         chains = 1,
         cores = 1,

### 2020-07-23 documents/ovando-etal-assessing-global-fisheries.Rmd
results with updatede priors, no real changes

@@
 output: 
-  bookdown::word_document2: default
+  bookdown::word_document2:
+    reference_docx: word-template.docx
   bookdown::pdf_document2: default
-abstract: Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals.
+abstract: |
+  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals.
 Numerous high-profile publications have produced estimates of the global status of fisheries, but limited quantity and quality of data, conflicting results, and methodological concerns mean that our understanding of global fisheries is still highly uncertain, especially in critical areas such as Southeast Asia and Africa.
@@
 
-*Insert mandatory opening sentence about the importance of global fisheries*.
-The United Nations Sustainable Development Goal 14 calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
-A substantial part of achieving this goal rests on ensuring that marine fisheries are managed levels that ensure the conservation and sustainable use of ocean ecosystems.
+The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
+A substantial part of achieving this goal rests on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems.
 Achieving this goal depends in part on our ability to effectively track the status of global marine fish stocks.
@@
 
-However, there are two major areas of improvement needed to ensure that we are able to accurately track the status of global fisheries.
-XX% of the global landings covered by the SOFIA reports come from fisheries with formal statistical stock assessments, quantitative models that provide estimates of stock status based on the best available science.
-These stock assessments are then incorporated into the SOFIA report's estimates of global stock status.
-The remaining XX% of the stocks/landings covered by SOFIA lack formal stock assessments.
-For these important fisheries, FAO scientists must use a mixture of local expert knowledge, data synthesis, and expert opinion to infer stock status.
-While this process provides the best available estimate of stock status, ideally a more reproducible  XX something to communicate why we might want to improve things for these stocks XX.
+However, there is room for substantial improvement in the methods used in estimating global stock status.
+49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020].
+The SOFIA process uses these formal assessments for their determinations of stock status wherever possible.
+However, that leaves 51% of global fisheries landings lacking in formal stock assessments.
+While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
 
-Second, there are many fisheries that while perhaps small on a global scale provide critical ecosystem services to their adjacent communities.
-These fisheries collectively, making up XX% of global landings, are not contained in the FAO assessment, but still fall under the SDG 14 goal XX to what extent is this true?
-XX rishi would be good to work on this section together.
-By my count the total catch form the SOFIA table sent to me is 42.68 MMT, so about half of global catch, but SOFIA 2020 says that "78.7 percent of current landings come from biologically sustainable stocks.", which would imply then that basically all landings are in SOFIA (since that doesn't leave much room for the assessed but overfished)?
-Confused by discrepancy hereXX.
-
-Ideally, both of these problems would be solved by simply running formal stock assessments on all significant fisheries.
-This is rendered impossible though both by lack of data and human resources.
-Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance such as catch-per-unit-effort (CPUE).
+The SOFIA report bases its estimates of included unassessed stocks mostly on expert elicitation.
+While local experts often are well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable.
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012;@costello2016a;@rosenberg2018;@pauly2007].Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance such as catch-per-unit-effort (CPUE).
 While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect.
@@
 This process takes considerable time and expertise.
-As such, effectively tracking the status of all global fisheries will require alternative methods.
 
-To resolve this problem, numerous papers have developed methods that while not intended for the management of individual stock are hoped to provide reasonable estimates of stock status in aggregate, particularly for those regions lacking the data or capacity for formal stock assessment.
-These include @costello2016a, @rosenberg2018, and @pauly2007.
-Each of these methods sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.
-While not suitable for management of individual stocks, these broad strategic models have played an important role in shaping the public narrative around ocean health.
+Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment.
+Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models [@free2020 and references therein].
+These models seek to infer stock status, for example in terms of biomass *b* relative to the biomass at maximum sustainable yield *b~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
+In addition, given that a global assessment by its nature requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time.
+As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks (high variance, low bias).Each of these methods sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.
 
-However, comparing estimates from each of these methods against best-available knowledge of stock status from the RAM Legacy Stock Assessment Database [@ricard2012], as well as broader intuition.
-reveals some systemic problems with these assessments.
-We have stock assessment data from RAM from the NE Pacific,the NE Atlantic and the Mediterranean, and of these the NE Pacific is in the best state (mean RAM B/B~MSY~ =1.8), NE Atlantic (mean RAM B/B~MSY~ =1.3) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.5).
+However, @free2020 demonstrated that these catch-only models can often produce both imprecise and biased estimates of B/B~MSY~.
+These issues become apparent when we consider some of the macro-level predictions made by these models.
+Examining stock assessment data from RAM from the NE Pacific, the NE Atlantic and the Mediterranean, the NE Pacific is in the best state (mean RAM b/b~MSY~ = 1.8), NE Atlantic (mean RAM b/b~MSY~ =1.3) and the Mediterranean the worst (mean RAM b/b~MSY~ =0.5).
 We do not have formal stock assessment data from S.
@@
 The Costello et al.
-2016 [@costello2016a] model shows the correct general pattern, but we would expect the state of fisheries in the Mediterranean/Black sea regions and SE Asia to be much worse, and the NE Pacific better by comparison.
-The Rosenberg super-ensemble method[@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific.
-The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)) *probably want to make this a figure.
-Also, Ray, do you remember if these estimates are unassessed only, or unassessed + RAM?*
+2016 [@costello2016a] model shows the correct general pattern in terms of relative ranking of regions, but we would expect the state of fisheries in the Mediterranean/Black sea regions and SE Asia to be much worse, and the NE Pacific better by comparison.
+The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific.
+The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
 
@@
 Climate change, population growth, and economic expansion will all place increasing pressure on fisheries.
-Mitigating these threats, as called for by SDG 14, will require an improved ability to track the status of fisheries around the world.
+While our understanding of global fisheries has improved greatly,  mitigating these threats, as called for by SDG 14, will require an improved ability to track the status of fisheries around the world, particularly those stocks that are currently unassessed.
 At the same time, new sets of globally distributed data have recently become available, along with advances in statistical modeling methods.
@@
 
+We use a new assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
 The FAO's catch statistics remain the most widely accessible and used source of information on global fisheries.
-Several methods have attempted to infer stock status (generally biomass relative to biomass at maximum sustainable yield, B/B_msy_) based on a fishery's catch history alone [summarized by @free2020].
-However, these "catch-only" methods have serious shortcomings that limit their performance, as evidenced by the status discrepancies by region they produce (Table.\@ref(tab:status-tab)), as well as results from several studies examining the performance of such methods [@free2020; @pauly2013].
-
 However, catches are not the only broadly available source of fishery data.
 We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@anderson2015], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c].
-We use a flexible assessment package ([sraplus](https://github.com/DanOvando/sraplus#sraplus)) to compare estimates of stock status around the globe using different combination of these data.
+`sraplus` provides a flexible framework for integrating these data into valid statistical model.
+`FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model.
+We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~.
+Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.
 
@@
 
-Evaluating the validity of data-limited XX need to introduce this idea earlier xx assessment methods can be a challenging task.
-Simulation testing provides one avenue, but can be difficult to create a simulation environment that adequately captures the problems of real-world ecosystems and data.
-We can also compare results between data-limited assessment and more data (or resource) rich gold-standard stock assessments.
-However, formally stock asssessed fisheries represent a very biased subset of global fisheries, and it is entirely possible that a method that performs well for the kinds of species and fisheries that are commonly stock assessed might struggle in the kinds of fisheries that typically are unassessed.
-
-Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region.
-As outlined in Table.1, while we often have no proof that results finding that for example fisheries in south-east Asia have similar stock status to those in Alaska, very few experts would agree that this is a plausible result XX tough statement to make in a paper xx.
-We use two case studies here to present the potential of additional, broadly applicable, data to potentially improve on "catch-only" methods.
-
+Two case studies provide examples of how augmenting catch data alone with local and globally available data can provide dramatically different estimates of local stock status.
 For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region.
@@
 Using this value to inform a prior on terminal stock status dramatically changes our picture of stock status in the Eastern Bay of Bengal, with now all stocks being overfished and experiencing high levels of overfishing (Fig.\@ref(fig:east-india)-sar).
+
+<!-- Evaluating the validity of data-limited assessments methods can be challenging.
+Simulation testing provides one avenue, but it can be difficult to create a simulation environment that adequately captures the problems of real-world ecosystems and data.
+We can also compare results between data-limited assessment and more data (or resource) rich gold-standard stock assessments.
+However, formally stock assessed fisheries represent a very biased subset of global fisheries, and it is entirely possible that a method that performs well for the kinds of species and fisheries that are commonly stock assessed might struggle in the kinds of fisheries that typically are unassessed.
+-->
+
+<!-- Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region.
+As outlined in Table.\@ref(tab:status-tab), while we often have no proof that results finding that for example fisheries in south-east Asia have similar stock status to those in Alaska, very few experts would agree that this is a plausible result XX tough statement to make in a paper xx.
+We use two case studies here to present the potential of additional, broadly applicable, data to potentially improve on "catch-only" methods.
+-->
 
@@
 To give an illustration, we used `sraplus` to assess the status of in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from @rousseau2019 (xx technically this is the updated dataset xx).
-Interestingly, catch histories for Europe and Northeast Asia look somewhat similar xx plot of catch and effort histories xx.
+Catch histories for Europe and Northeast Asia look somewhat similar xx plot of catch and effort histories xx.
 As a result, using a "catch-only" heuristic, both Europ and Northeast Asia are estimated to have very similar stock status, a similar finding as those illustrated in Table.XX.
@@
 
-Our results show that our perception of global fishery depends completely on which data sources we utilize.
+Our perception of global fishery depends completely on which data sources we utilize.
 That different data tell different stories should not be surprising.
-However, we also do not see a convergence towards a cohesive picutre of global status that closely matches the SOFIA report as we add more data.
+However, we also do not see a convergence towards a cohesive picture of global status that closely matches the SOFIA report as we add more data.
 The most data-poor method employed, the catch heuristic alone, matches some patterns from the SOFIA report (little overfishing in the Northeast Pacific, some overfishing in the northwest Atlantic), it misses the mark badly in other regions, e.g.
@@
 Global-level assessment are critical for guiding management agendas for the world's oceans.
-Despite this need, and despite advances in stock assessment methods, our understanding of many of the world's fisheries remains poor.
-We show that while new sources of widely available data are becoming increasingly available, none yet is capable of providing a clear improvement in our estimates of global stock status.
+Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world.
+Numerous efforts have sought to resolve this problem through development of new modeling methods.
+These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment.
+This project was motivated by the need to address this problem, and the increasing availability of broadly (if not globally) datasets that could be of use for fisheries assessment.
+While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.
 
-Our estimate of global stock status varies wildly depending on which source of data we use.
-Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status?
+Our estimates of global stock status vary wildly depending on which source of data we use.
+Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information?
 Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:po-map)).
@@
 Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks.
+None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RAM (XX RAM matches?
+need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).
+
+This is not to say that the data presented here are not valuable.
+The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e.
+those that sufficiently resemble the data on which the regressions were trained).
+Effort data such as those reconstructed by @rousseau2019a can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with relatively constant fishing technologies and a broadly selective fishing fleet.
+The challenge comes in attempting to indiscriminately apply these data to all areas: doing so results in substantially incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
 
@@
 This is not to suggest that improvements in stock assessment methods, even for relatively simple models such as the biomass dynamics types used here, are not of use.
-At the tactical level, such improvements can make meaningful changes in management outcomes.
-However, at the global strategic levels, the improvements to be made from model development are marginal compared to the fundamental uncertainties stemming from currently available data.
+We certainly do not claim that the modeling choices we have made here represent the best version possible.
+At the tactical level, modeling improvements can make meaningful changes in management outcomes.
+However, at the global strategic level, the evidence from multiple efforts at assessing global fisheries strongly suggests that any improvements to be made from model improvement are marginal compared to the fundamental uncertainties stemming from currently available data.
 
 What does the future of global fishery assessment look like?
-Investment must be made in improving data collection at the global scale.
+Investment must be made in improving  global data collection and training local fisheries scientists.
+While the data used in this study were not a silver bullet, they present a step forward, and local versions of these data could prove immensely helpful.
+Stock-specific, as opposed to the often broad regional estimates used here, estimates of swept-area-ratio and fisheries management index scores can provide improved priors on stock status for applicable fisheries.
+Informative priors on fishing mortality rates can be particularly useful in improving accuracy of data-limited assessment results (Fig.XX VOI XX)
+
 Our analysis suggests several possible paths for this process, in particular creation of a global repository or length-composition data, and expansion of FMI surveys and SAR data.
@@
 
-XX equations here XX.
+```{r}
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
+  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
+  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
+  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
+  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
+  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))
+
+knitr::kable(
+  cd,
+  "latex",
+  align = "l",
+  booktabs = TRUE,
+  escape = F,
+  row.names = F,
+  caption = "Estimated parameters, abbreviations, and default priors"
+) %>%
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+```
+
+The population growth equation is
+
+\begin{equation}
+  f(x)=\begin{cases}
+      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
+, & \text{if $B_t>0.25 \times K$}.\\
+     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
+  \end{cases}
+  (\#eq:hockey)
+\end{equation}
+
+Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of Bmsy/K to shift (when *m* is 2 the Bmsy / K = 0.5 and the model acts as a Schaefer model, lower values of m shift the production function left, higher values right), *c* is catch, and *p* is process error.
+The Pella-Tomlinson model does have an issue where growth rates can become unrealistically large when the population reaches low sizes.
+We deal with this problem by following the methods described in @winker2018 to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
+Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see @thorson2018).
+
+<!-- When the SIR method is used we use the supplied catch history.
+However, for the maximum likelihood approach, convergence is greatly facilitated by estimating catches ($\hat{c_t}$) rather than using the observed catches directly.
+$\hat{c_t}$ is given as -->
+
+<!-- \begin{equation} -->
+<!--   \hat{c_t} = u_{t}B_{t} -->
+<!--   (\#eq:chat) -->
+<!-- \end{equation} -->
+
+<!-- where $u_t$ is estimated through -->
+
+<!-- \begin{equation} -->
+<!--   u_{t} = \frac{1}{1 + e^{-u^*_t}} -->
+<!--   (\#eq:invf) -->
+<!-- \end{equation} -->
+
+<!-- Equation \@ref(eq:invf) ensures that $u_t$ is between 0 and 1, while the model estimates $u^*_t$ to improve performance.  -->
+
+<!-- $\hat{c_t}$ is then fit to the observed catches through the likelihood -->
+
+<!-- \begin{equation} -->
+<!--   log\left(c_t\right) \sim normal\left(log\left(\hat{c_t}\right),0.05\right) -->
+<!--   (\#eq:catch) -->
+<!-- \end{equation} -->
+
+<!-- Where we assume an observation coefficient of variation of 5\%.
+We estimate catch rather than directly using the observed catch in order to improve model performance.
+As priors on depletion, or indices of abundance, pull the estimated population closer to zero, there is a greater and greater chance that the observed catch history will result in a collapse of the population (catch in time *t* being greater than biomass in time *t*); It is much harder to find parameters that produce a final depletion of 1\% of carrying capacity while not crashing the population prior to then given a catch history, than it is to produce a final depletion of 99\% of carrying capacity.
+We can impose restrictions and penalties on population crashes, but this distorts the gradients of the model, causing the model to often become stuck once a few crashes have occurred.
+Since $u_t$ is constrained between 0 and 1, $\hat{c_t}$ will always be "viable", and by then fitting to the observed catches, the model has an easier time finding parameters that produce a near match to the observed catches while not crashing the population.
+As of now none of the data used in this exercise have missing years of catch data, but the model can handle those cases in the future due to the random walk priors imposed on $u_t$ detailed below.
+-->
+
+<!-- In order to help separate out observation error in the catch and process error in the population model, when catch is being estimated we assume that fishing mortality $u_t$ follows a random walk on average, assigning each $u_t$ a prior of -->
+
+<!-- \begin{equation} -->
+<!--   log(u_t) \sim normal(log(u_{t - 1}),0.1)) -->
+<!--   (\#eq:rwalk) -->
+<!-- \end{equation} -->
+
+We  allow for process error [in the manner of the stochastic stock reduction analysis error suggested by @walters2006].
+This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
+Process error *p* is assumed to be log-normally distributed, such that
+
+\begin{equation}
+  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
+  (\#eq:procerror)
+\end{equation}
 

### 2020-07-23 functions/fit-fao.R

@@ -40,7 +40,6 @@ fit_fao <-
         b_ref_type = "k",
         isscaap_group = unique(data$isscaap_group)
       )
-    
     basic_fit <-
       fit_sraplus(
         driors = basic_driors,
@@ -58,7 +57,6 @@ fit_fao <-
     
     
     # catch only fit
-    
     com_driors <-
       format_driors(
         taxa = scientific_name,
@@ -71,7 +69,6 @@ fit_fao <-
         isscaap_group = unique(data$isscaap_group),
         use_catch_prior = TRUE
       )
-    
     com_fit <-
       fit_sraplus(
         driors = com_driors,

### 2020-07-23 functions/fit_continent_examples.R

@@ -109,7 +109,8 @@ fit_continent_examples <-
         estimate_shape = estimate_shape,
         estimate_qslope = estimate_qslope,
         estimate_proc_error = estimate_proc_error,
-        use_baranov = use_baranov
+        use_baranov = use_baranov,
+        workers = 4
       )
     #
     # plot(cpue_driors$catch[cpue_driors$effort_years] / cpue_driors$effort)

### 2020-07-23 make-assessing-global-fisheries.R

@@ -63,9 +63,8 @@ write_results <- TRUE
 process_fits <- FALSE
 
 # other things
-run_case_studies <- FALSE
 
-run_continent_examples <- FALSE
+run_continent_examples <- TRUE
 
 run_ei_example <- TRUE
 
@@ -1516,9 +1515,7 @@ areas <- unique(fao_status$fao_area_code)
 #   filter(year == max(year)) %>% 
 #   summarise(catch = sum(capture))
 
-
 # annnnnd try and run assessments
-# browser()
 if (run_sofia_comparison == TRUE) {
   # future::plan(future::multiprocess, workers = 4)
   
@@ -1537,10 +1534,11 @@ if (run_sofia_comparison == TRUE) {
         data,
         safely(fit_fao),
         support_data = support_data,
-        default_initial_state = 1,
+        default_initial_state = NA,
+        default_initial_state_cv = NA,
         min_effort_year = 1975,
         engine = "stan",
-        cores = 1,
+        cores = 2,
         .progress = TRUE,
         .options = future_options(globals = support_data, packages = c("tidyverse","sraplus"))
       )

### 2020-07-26 make-assessing-global-fisheries.R
adding in RAM testing

@@ -1,4 +1,5 @@
 
+
 # setup -------------------------------------------------------------------
 
 library(tidyverse)
@@ -52,7 +53,8 @@ future::plan(multiprocess, workers = n_cores)
 results_name <- "v0.5"
 
 
-results_description <- "publication version of results updated VOI from version 0.5 "
+results_description <-
+  "publication version of results updated VOI from version 0.5 "
 
 run_voi_models <- FALSE
 # sub options for run_voi_models
@@ -64,11 +66,14 @@ process_fits <- FALSE
 
 # other things
 
-run_continent_examples <- TRUE
+run_continent_examples <- FALSE
 
-run_ei_example <- TRUE
+run_ei_example <- FALSE
 
-run_sofia_comparison <- TRUE
+run_sofia_comparison <- FALSE
+
+run_ram_tests <- TRUE
+
 
 engine = "stan"
 
@@ -108,6 +113,7 @@ if (dir.exists(here::here("results", results_name)) == FALSE) {
 
 prepare_sofia_data(lookup_fmi_names = FALSE)
 
+
 has_total_biomass = ram_data %>%
   group_by(stockid) %>%
   summarise(hastb = !any(is.na(total_biomass)),
@@ -117,7 +123,8 @@ has_total_biomass = ram_data %>%
 ram_data <- ram_data %>%
   filter(stockid %in% has_total_biomass$stockid) %>%
   mutate(has_things = !(is.na(u_v_umsy) |
-                          is.na(total_biomass) | is.na(catch) | catch == 0)) %>%
+                          is.na(total_biomass) |
+                          is.na(catch) | catch == 0)) %>%
   filter(has_things) %>%
   group_by(stockid) %>%
   mutate(delta_year = as.integer(year - lag(year))) %>%
@@ -128,18 +135,18 @@ ram_data <- ram_data %>%
   group_by(stockid) %>%
   mutate(n = length(catch)) %>%
   filter(n >= min_years_catch) %>%
-  ungroup() 
+  ungroup()
 
 # fit voi models --------------------------------------------------------------
 
 
-if (run_voi_models == TRUE){
-  
-  
+if (run_voi_models == TRUE) {
   stress_tests <- data_frame(
     use_index = sample(c(TRUE, FALSE),  draws, replace = TRUE),
     index_type = sample(c("survey"), draws, replace = TRUE),
-    initial_state_type = sample(c("unfished", "known", "prior","heuristic"), draws, replace = TRUE),
[…]

### 2020-07-30 documents/ovando-etal-assessing-global-fisheries.Rmd
1st real draft of paper done

@@
 ---
-title: "The Future of Global Fishery Assessments"
+title: "Global Assessments of Fishery Status need Better Data more than Better Models"
 author: 
@@
 abstract: |
-  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, and tracking large-scale objectives such as the United Nations Sustainable Development Goals.
-Numerous high-profile publications have produced estimates of the global status of fisheries, but limited quantity and quality of data, conflicting results, and methodological concerns mean that our understanding of global fisheries is still highly uncertain, especially in critical areas such as Southeast Asia and Africa.
-We present an evaluation of the past and future of global fishery assessments.
-Global assessments have typically involved aggregating best-available stock assessments where possible with various data-limited assessments for regions lacking formal evaluations of stock status.
-The data-limited methods used have ranged from simple catch heuristics to super-ensemble modeling.
-Despite attempts at model improvement, there have been discrepancies both among global fishery assessments and between the data-limited components of these methods and more data-rich models,demonstrating that our understanding of global data-limited fisheries remains limited.
-How then can we most efficiently move forward?
-We utilize a flexible assessment model together with the RAM Legacy Stock Assessment Database not strictly as a stock assessment tool, but to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status.
-We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion and increasingly available but historically underutilized data such as trawl footprints.
-These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world.
-We close by discussing priority areas for future data collection.
-Our results make the case that improving our understanding of global fisheries depends not on better models, but on better data.
+  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals.
+Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results and therefore, high uncertainties in our understanding of global fisheries.
+This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources.
+How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets?
+We developed a new flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status.
+We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data.
+These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection.
+We show that our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
 ---
@@
 library(tidyverse)
+library(viridis)
 library(sf)
@@
 library(scales)
+library(patchwork)
+library(rstanarm)
+library(bayesplot)
 library(here)
@@
 
+functions <- list.files(here::here("functions"))
+
+purrr::walk(functions, ~ source(here::here("functions", .x)))
+
 results_name <- "v0.5"
 
-results_dir <- here("results",results_name)
+results_path <- here("results",results_name)
 
-load(file = file.path(results_dir, "paper_plots.Rdata"))
+min_years_catch <- 25
 
-pub_theme <- theme_ipsum(base_size = 12, 
-                         axis_text_size = 12) + 
-    theme(panel.spacing = unit(1,"lines"))
+prepare_sofia_data(lookup_fmi_names = FALSE)
+
+ram_comp_data <- ram_data %>%
+  mutate(has_things = !(is.na(catch) | catch == 0)) %>%
+  filter(has_things) %>%
+  group_by(stockid) %>%
+  mutate(delta_year = as.integer(year - lag(year))) %>%
+  mutate(delta_year = case_when(year == min(year) ~ as.integer(1),
+                                TRUE ~ delta_year)) %>%
+  mutate(missing_gaps = any(delta_year > 1)) %>%
+  filter(missing_gaps == FALSE) %>%
+  group_by(stockid) %>%
+  mutate(n = length(catch)) %>%
+  filter(n >= min_years_catch) %>%
+  ungroup()
+
+ram_comp_data <- ram_comp_data %>%
+  select(
+    year,
+    stockid,
+    country,
+    primary_FAOarea,
+    scientificname,
+    commonname,
+    year,
+    catch,
+    b_v_bmsy,
+    u_v_umsy,
+    total_biomass
+  ) %>%
+  rename(scientific_name = scientificname) %>% 
+  mutate(fao_area_code = as.numeric(primary_FAOarea)) %>% 
+  left_join(fao_species, by = "scientific_name")
+
+assess_ram_fits <- read_rds(path = file.path(results_path,"assess_ram_fits.rds"))
+
+ram_v_sraplus_area <- read_rds(path = file.path(results_path,"ram_v_sraplus_area.rds"))
+
+ram_comp_data <- read_rds(path = file.path(results_path,"ram_comp_data.rds"))
+
+fao_sraplus_comp_data <- read_rds(path = file.path(results_path, "fao_sraplus_comp_data.rds"))
+
+load(file = file.path(results_path, "paper_plots.Rdata"))
+
+load(file = file.path(results_path, "voi_fits.RData"))
+
+pub_theme <- theme_ipsum(base_size = 10,
+                         axis_text_size = 10,
+                         axis_title_size = 12) +
+    theme(panel.spacing = unit(.5,"lines"))
 
@@
 As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fishery, 6.2% are underfished ,and 34.2% are overfished [@fao2020].
+While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment.
+Although revised periodically over time, the method has kept revisions to a minimally necessary levels for the purposes of temporal consistency and comparability.
+With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.
 
-However, there is room for substantial improvement in the methods used in estimating global stock status.
 49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020].
@@
 While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-
 The SOFIA report bases its estimates of included unassessed stocks mostly on expert elicitation.
 While local experts often are well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable.
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012;@costello2016a;@rosenberg2018;@pauly2007].Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance such as catch-per-unit-effort (CPUE).
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
+
+Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance such as catch-per-unit-effort (CPUE).
 While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect.
@@
 Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment.
-Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models [@free2020 and references therein].
+Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (@free2020 and references therein).
 These models seek to infer stock status, for example in terms of biomass *b* relative to the biomass at maximum sustainable yield *b~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
@@
 We do not have formal stock assessment data from S.
-and SE Asia (Western Central Pacific and Eastern Indian Ocean), but based on surveys and expert opinion we believe it should be worse than Northeast Pacific.
+and SE Asia (Western Central Pacific and Eastern Indian Ocean), but other data sources such as surveys and expert opinion suggest this region should have more depleted fisheries than Northeast Pacific.
 The Costello et al.
-2016 [@costello2016a] model shows the correct general pattern in terms of relative ranking of regions, but we would expect the state of fisheries in the Mediterranean/Black sea regions and SE Asia to be much worse, and the NE Pacific better by comparison.
+20162 model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and SE Asia seem to be overoptimistic, and the NE Pacific should be better by comparison.
 The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific.
-The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
+The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).
 
@@
   kableExtra::row_spec(0, bold = T)
+
 ```
@@
 
-We use a new assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
-The FAO's catch statistics remain the most widely accessible and used source of information on global fisheries.
+We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
+The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability.
 However, catches are not the only broadly available source of fishery data.
-We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@anderson2015], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c].
+We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@melnychuk2017], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c].
 `sraplus` provides a flexible framework for integrating these data into valid statistical model.
@@
 For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region.
-We can then compare the estimates of stock status that come from using the catch-data alone (with a heuristic algorithm in the manner of @martel2013), with the catch data along with priors on terminal depletion informed by the swept-area-ratio data, and using the swept-area ratio data together with a local abundance index.
+We can then compare the estimates of stock status that come from using the catch-data alone (with a heuristic algorithm in the manner of @martell2013), with the catch data along with priors on terminal depletion informed by the swept-area-ratio data, and using the swept-area ratio data together with a local abundance index.
 By it's nature the catch-only heuristic produces a bi-model distribution, with most fisheries reported to have a B/B~MSY~ of 1.5 and an F/F~MSY~ less than 1 (Fig.\@ref(fig:east-india)-heuristic).
@@
 However, imposing the same prior on terminal stock status for all fisheries in a region is not ideal.
-Supplementing these data with a local index of abundance provides a more nuanced figure, with most stocks estimated as having B/B~MSY~ less than 1, but with a much wider range of fishing mortality rates reflecting difference in recent catch histories among individual stocks (Fig.\@ref(fig:east-india)-"cpue + sar").
+Supplementing these data with a local index of abundance provides a more nuanced result, with most stocks estimated as having B/B~MSY~ less than 1, but with a much wider range of fishing mortality rates reflecting difference in recent catch histories among individual stocks (Fig.\@ref(fig:east-india)-"cpue + sar").
 While we do not know the true state of stocks in the Eastern Bay of Bengal, inclusion of local data provide a picture more in line with local perception XX.
@@
 It should be no surprise that incorporating locally tailored data produces more locally plausible results.
-Global assessment at the moment though requires depending on a few broadly available datasets.
-To give an illustration, we used `sraplus` to assess the status of in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from @rousseau2019 (xx technically this is the updated dataset xx).
-Catch histories for Europe and Northeast Asia look somewhat similar xx plot of catch and effort histories xx.
-As a result, using a "catch-only" heuristic, both Europ and Northeast Asia are estimated to have very similar stock status, a similar finding as those illustrated in Table.XX.
-Adding in the effort data (as an effort index created for each stock as catch divided by the effective harvest rate based on the effort data), and the two regions begin to diverge.
-While according to @rousseau2019, Europe has seen a recent decrease in fishing effort from a peak in XX, effort in Northeast Asia has increased more or less continuosly.
-As a result, `sraplus` estimates that more stocks in Northeast Asia are overfished / experiencing overfishing than in Europe.
-However, the shared effort series across these stocks aslo produces a large number of effectively unfished stocks; a highly unlikely result for both regions.
+Global assessment at the moment though depend on few broadly available datasets.
+We used `sraplus` to assess the status of stocks in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from [@rousseau2019] (xx technically this is the updated dataset xx).
+Catch histories for Europe show a steady increase reaching a peak in the 1990s, and a decline and stabilization since.
+Current catches in Europe are much higher than those at the start of the catch history.
+Catches in northeast Asia have steadily increased over time, with some stabilization as of late.
+For both regions, recent catches expressed as a fraction of maximum catch, they key metric from the perspective of a catch-only heuristic, are both high.
+However, the two regions exhibit markedly different effort histories according to @rousseau2019, with effort in Europe peaking with catches and subsequently declining, while effort in northeast Asia has increased steadily over time.
+As a result, using a "catch-only" heuristic, both Europe and Northeast Asia are estimated to have very similar stock status.
+Adding in an index of abundance, calculated for each stock as catch divided by the effective harvest rate based on the effort data for the region, produces a very different picture of stock status and introduces some contrast between the two regions.
+Based on the catch and effort data, `sraplus` estimates that more stocks in Northeast Asia are overfished / experiencing overfishing than in Europe.
+However, the shared effort series across these stocks also produces a large number of effectively unfished stocks; a highly unlikely result for both regions (Fig.\@ref(fig:eu-v-na)).
 
-XX need to redo this analyis with the new effort data XX
-
-```{r eu-v-na, fig.cap="Estimates of stock status in Europe and Northeast Asia using either a simple catch heuristic as presented in @martell2013 or an abundance index based on continental-scale effort trends reported in @rousseau2019.
-Each point is a stock, point size is a function of stock size."}
+```{r eu-v-na, fig.cap="Estimates of stock status in Europe and Northeast Asia using either a simple catch heuristic as presented in @martell2013 or an abundance index based on continental-scale effort trends reported in @rousseau2019 Each point is a stock, point size is a function of stock size."}
 continent_kobe_plot
@@
 
-This exercise reveals the potential and challenges using broadly available data for global fishery assessment.
-On the one hand, for both case studies moving beyond catch data alone produced a more plausible set of results.
-However, both results also raise a host of new issues.
+```{r perf-tab}
+breaks <- c(0.8, 1.2)
+
+breaks <- c(0, breaks, Inf)
+
+labels <- c("over", "fully", "under")
+
+rand_fits <- assess_ram_fits %>% 
+  filter(data == "ram-data") %>% 
+  mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
+         data = "guess")
+
+perf <- assess_ram_fits %>% 
+  bind_rows(rand_fits) %>% 
+  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
+  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
+  group_by(data) %>% 
+  summarise(
+    mpe = median((mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
+    mape = median(abs(mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
+    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
+  ) %>% 
+  arrange((mape))
+
+knitr::kable(perf, digits = 2,
+             caption = "Global summary statistics of models using different sources of data.
+mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).",
+              format = "pipe")
+
+```
+
+These case studies reveals the potential and challenges of using broadly available data for global fishery assessment.
+On the one hand, for both case studies moving beyond catch data alone produced a more plausible set of results in some ways.
+However, both case studies also raise a host of new issues.
 How realistic is it to share one index of abundance and SAR value for the entire eastern Bay of Bengal?
-Similarly, the sharing of a common effort series at the continental level does provide increased separation between European and northeast Asian stocks that may be plausible, but also estimates a suspicious amount of stocks to be essentially unfished (a dynamic which does not match assessments from that region XX plot RAM stocks for europe XX).
+Similarly, the sharing of a common effort series at the continental level does provide increased separation between European and northeast Asian stocks that may be plausible, but also estimates a suspicious amount of stocks to be essentially unfished.
 
-Extending these issues to the global level, we assessed the status of stocks assessed by the FAO's latest SOFIA report using various types of broadly available data.
-Given that the SOFIA assessments present the best estimates of stock status across both assessed and unassessed stocks, we would hope that any assessment method suitable would produce broadly match the values reported by the FAO.
-We focus here on the percent of stocks classified as "overfished" (B/B~MSY~ < 0.8) in each FAO statistical region.
-Following the SOFIA methodology, "Stocks" are classified by the sum catch histories across the same species within a major FAO statistical region.
+Extending these issues to the global level, we  used a variety of data sources to estimate stock status for fisheries listed in RAM.
+Our goal is to simulate a macro-level global assessment, where data are available at regional levels, but not for specific fisheries.
+As a proof of concept, we first fit models to abundance indices from each stock in RAM.
+There are two main reasons why an assessment model might produce poor results: model misspecification and poor data.
+Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries.
+Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`.
+Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality.
+Overall the `sraplus` estimates using the RAM data are reasonably good.
+The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias.
+The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time.
+Looking across space, we see clear geographic patterns in model performance.
+The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada).
+The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.
 
-Our perception of global fishery depends completely on which data sources we utilize.
-That different data tell different stories should not be surprising.
-However, we also do not see a convergence towards a cohesive picture of global status that closely matches the SOFIA report as we add more data.
-The most data-poor method employed, the catch heuristic alone, matches some patterns from the SOFIA report (little overfishing in the Northeast Pacific, some overfishing in the northwest Atlantic), it misses the mark badly in other regions, e.g.
-the waters around New Zealand.
-Our most data-rich scenario, an effort series from @rousseau2019, along with priors from swept-area-reatio and fisheries management index where available....
+Performance limitations then are likely to arise less from model misspecification but from the quality of the data themselves.
+These becomes even clearer once we consider the performance of `sraplus` models fit to globally available datasets.
+Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)).
+However, this is somewhat an artifact of the data.
+The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4.
+The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1.
+This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise.
+Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
+The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`.
+Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])`.
 
-to provide more plausible seeming results than catch data alone.
-A reasonable question then is, can we simply augment global catch statistics with the data presented here to provide an improved understanding of global fisheries?
+Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process.
+We also see that across the models performance (or lack thereof) is not consistent in space: different data to best/worst for different FAO regions.
+For example, the catch heuristic and models fit to nominal CPUE data both substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
+We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
+Overall, while some data sources performed slightly better than others by some metrics in some places, there was no consistent spatial pattern to this performance, and none of the datasets were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
 
-This is the main plot of the paper Percent of stocks classified as overfished depending on what data you use.
-CPUE data are at the continental scale (e.g.
-use North America Rousseau effort for USA, Mexico, Canada).
+```{r dat-description}
+dat_description <- tribble(~`Data Name`, ~Description,
+        "RAM-Data", "Fit to abundance index from RAM",
+        "sar", "Prior on terminal F/Fmsy set by regional swept area ratio",
+        "fmi", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
+        "cpue", "Fit to CPUE index created from RAM catch and regional effort index. 2.5% tech. creep",
+        "cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.5% tech. creep",
+        "nominal-cpue", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
+         "nominal-cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
+        "catch-only", "priors on terminal B/Bmsy informed by catch history",
+        "heuristic", "priors on initial and terminal B/Bmsy informed by Cmsy heuristc")
 
-key message: Different data sources give you different estimates of stock status, and not in a way that makes it clear that any one is inherently superior to another
+knitr::kable(dat_description,
+             caption = "Data sources used for terminal stock status estimate",
+             format = "pipe")
+# %>%
+#   kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+#   kableExtra::row_spec(0, bold = T)
+```
 
-A few key questions here
-
-- Currently this runs sraplus on all stocks in the region.
-I think this is informative since it says, suppose we only used sraplus, what would we think, and can compare that to RAM etc.
-in our heads (or could add a map panel).
-An alternative though would be to stitch together status, i.e.
-use RAM where there is RAM, use sraplus otherwise.
-That would be a better "estimate" of the actual stock status in an area, but would make the contrast between the methods less clear
-
-- This used the continent level Rousseau data - can we get the finer scale version for this paper? Would make things a lot clearer I think
-
-*will add in RAM panel once I match up RAM stocks to FAO region*
-
-```{r po-map, fig.width=8, fig.height=8, fig.cap="Estimated percent of fisheries 'overfished' (B/B_msy_ less than 0.8) using different dat sources.
+```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent B/B~MSY~ by FAO statistical area from different data sources.
+ram-data refers to catch and abundance index drawn from RAM.
 CPUE refers to an index of abundance based on the Rousseau effort data.
@@
 fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
-po_map_plot
+ram_mpe_map_plot
 ```
 
-## Value of Information
+```{r mape-map, fig.cap = "Median absolute percent error in most recent B/B~MSY~ by FAO statistical area from different data sources.
+ram-data refers to catch and abundance index drawn from RAM.
+CPUE refers to an index of abundance based on the Rousseau effort data.
+Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
+The FAO Report 569 is our reference status estimate provided by the FAO.
+ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
+fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
+ram_mape_map_plot
+```
 
-Provide some guidance on where to prioritize data collection.
-e.g.
-how long of a CPUE series is needed, expansion of f based priors / value of f, advocate for merging together of the "Catch only" world with the "length-based world".
-E.g.
-we talk about catch being the only widely available statistic, but in many data-limited communities all they have are lengths.
-A global effort to create a centralized database of length distributions could allow for length-based priors on spr/f that could be paired with catches?
+```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
+ram-data refers to catch and abundance index drawn from RAM.
+CPUE refers to an index of abundance based on the Rousseau effort data.
+Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
+The FAO Report 569 is our reference status estimate provided by the FAO.
+ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
+fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
+
+ram_acc_map_plot 
+```
 
 # Discussion (500 words)
-
-** how do we know that these assessments themselves aren't actually the right ones"
-
-**address better than nothing**.
-Certainly not advocating for sitting on hands until better models come around, but important to point out that these models don't just produce noisy but unbiased estimates, they seem to be fundamentally biased in many places.
-But, no one method appears to be consistently biased (except for catch only)
-
-*these are just rough outline paragraphs*
-
-we do not claim that our model is the best model possible, or that others could not make improvements on our methods.
-Rather, we claim that any improvements that could be made, given the same data and broadness of scope, would likely be marginal relative to the inherent limitations of the data themselves.
 
@@
 These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment.
-This project was motivated by the need to address this problem, and the increasing availability of broadly (if not globally) datasets that could be of use for fisheries assessment.
+However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes.
 While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.
+
+What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends of course on the needs of those using the assessment outputs.
+It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
+We have stated that addition of the globally available datasets compiled here did not substantially improve estimates of stock status, but in some instances using these data did provide some improvement over use of catch-only style methods.
+However, none of the data sources used here came close to the benchmark performance provided by our fits to the RAM data, either in overall performance or in capturing the geographic distribution of stock status.
+While it is unreasonable to expect any global-scale dataset to be able to perform as well as data pulled from RAM assessments themselves, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general trends in stock status in time and space.
+That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the data themselves.
+
+We have stated that improving global estimates of stock status depends on improved data, not improved models *per se*.
+This is not to suggest that improvements in stock assessment methods, even for relatively simple models such as the surplus production types used here, are not of use.
+We certainly do not claim that the modeling choices we have made here represent the best version possible.
+At the tactical level, modeling improvements can make meaningful changes in management outcomes.
+However, at the global strategic level, the evidence from multiple efforts at assessing global fisheries strongly suggests that any improvements to be made from model improvement are marginal compared to the fundamental uncertainties stemming from currently available data.
 
@@
 Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information?
-Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:po-map)).
-Globally available effort data from Rousseau allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region.
+Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)).
+Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used.
 These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability.
@@
 those that sufficiently resemble the data on which the regressions were trained).
-Effort data such as those reconstructed by @rousseau2019a can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with relatively constant fishing technologies and a broadly selective fishing fleet.
+Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with relatively constant fishing technologies and a broadly selective fishing fleet.
+For example, despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis.
+But, applying one SAR value to all stocks in a region causes innacurate estimates of stock status.
 The challenge comes in attempting to indiscriminately apply these data to all areas: doing so results in substantially incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
 
-We have stated that improving global estimates of stock status depends on improved data, not improved models *per say*.
-This is not to suggest that improvements in stock assessment methods, even for relatively simple models such as the biomass dynamics types used here, are not of use.
-We certainly do not claim that the modeling choices we have made here represent the best version possible.
-At the tactical level, modeling improvements can make meaningful changes in management outcomes.
-However, at the global strategic level, the evidence from multiple efforts at assessing global fisheries strongly suggests that any improvements to be made from model improvement are marginal compared to the fundamental uncertainties stemming from currently available data.
+It is easy enough to call for expanded data collection at a global scale.
+In world of finite resources though, it is important to consider what sources of data might provide the greatest value in improving our estimates of global stock status.
+We fit 3,000XX sraplus models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
+We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).
 
-What does the future of global fishery assessment look like?
-Investment must be made in improving  global data collection and training local fisheries scientists.
-While the data used in this study were not a silver bullet, they present a step forward, and local versions of these data could prove immensely helpful.
-Stock-specific, as opposed to the often broad regional estimates used here, estimates of swept-area-ratio and fisheries management index scores can provide improved priors on stock status for applicable fisheries.
-Informative priors on fishing mortality rates can be particularly useful in improving accuracy of data-limited assessment results (Fig.XX VOI XX)
+XX will simplify this plot down to key things and clean up the variable names XX
 
-Our analysis suggests several possible paths for this process, in particular creation of a global repository or length-composition data, and expansion of FMI surveys and SAR data.
-Global Fishing Watch provides another avenue, but requires further research on the best ways of translating raw effort data into indices of stock status.
-Interpreting these data responsibly will require support from regional experts, for example in the form of expert panels charged with determine the most representative broadly available data for different portions of the oceans.
+As we should expect having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 20%.
+In terms of data however, information on fishing mortality rates (F/F~MSY~) in the most recent year was almost as useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot).
+Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance.
+This is because we varied both the length and quality of the index of abundance, and on average an even noisy estimate of recent F/F~MSY~ may be more valuable than a short and noisy abundance index.
 
-We have accumulated lots of evidence for status of global fisheries, each of which has strengths and weaknesses.These were important foundations.
-But, if we want to move forwards, the answer is better data.
+```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types and model options on root mean squared error of B/Bmsy in the most recent 5 years of data available for each model fit.
+Dark blue line shows median, blue shaded area the 66% credible interval, full rea the 95% credible interval."}
+b_voi_plot + 
+  geom_vline(aes(xintercept = 0), alpha = 0.75)
+```
+
+Given the demonstrated value of F/Fmsy values in improving estimates of stock status, how might we go about obtaining estimates of fishing mortality rates at a global scale?
+We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/Fmsy for individual fisheries (Fig.SXX).
+While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
+
+Another avenue would be to prioritize the development of a global repository for length and age composition data.
+The data-limited assessment world has recently been developing along two almost parallel tracks.
+Catch-based methods have most commonly been employed in efforts to understand global fisheries, and on occasion for management of stocks mandated to manage catch but lacking sufficient data for a formal stock assessment [@berkson2015].
+Along with general performance questions around catch-only models, no matter what they require a complete, or at least near complete timeseries of captures.
+This rules catch-based methods out for many small scale fisheries around the world.
+As a result, a growing number of methods have focused on using length-composition data to provide estimates of stock status, including measures of fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
+These methods have a distinct advantage for many fisheries in that they can in theory function with as little as one year of data.
+As a result, many fisheries around the world are enacting programs to collect and utilize length composition data for fisheries assessment.
+
+However, while we have standardized databases for global marine catch [@fao2020] and life history estimates [@froese2000; @thorson2017c], no comparable repository for length composition data exists.
+Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which when paired with catch data and where possible indicies of abundance could meaningfully improve our understanding of global fisheries [@thorson2015f].
+Such a database could be constructed in a top-down manner in a manner similar to the catch records compiled by the FAO (xx maybe top down isn't best descriptor here xx,or a bottom-up platform, such as the [MERMAID](https://datamermaid.org/) platform for coral reef data, could be developed that would allow users around the world to self-contribute length composition studies.
+
+This is not to say that we should also invest in global strategies for calculating abundance indicies.
+Where possible, repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//).
+However for many areas of the world effective fishery independent estimates of stock abundance are unlikely to be feasible.
+Effort databases such as @rousseau2019, can be combined with catch histories to to construct fishery-dependent abundance indicies.
+While there are substantial difficulties and biases that can arise from improper use of fishery-dependent abundance data, improved understanding of the evolution of global fishing effort could be very useful in improving global estimates of fishery health.
+While we did include effort reconstructions in our analysis, and did not find it to provide substantial improvements in assessment results, we attribute this largely to two difficulties.
+Use of effort data requires making an assumption about the rate at which the fishing power of a unit of fishing effort has changed over time.
+@rousseau2019 used a fixed global rate of 2.6% to describe the increasing trend in fishing power, but as shown in their paper and in our analysis, any estimates of stock status using these effort data will be critically sensitive to the value of this parameter.
+Global studies of technology creep could help inform more realistic values and functional forms for the evolution of fishing skill in particular fisheries and regions.
+Secondly, the effort reconstructions as employed here are still at very coarse scales, requiring us to apply one effort series for all the stocks within a given region.
+Going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
+
+Expanded training of fisheries scientists around the globe is another critical need.
+Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and critically evaluate the results of any model based on local expertise.
+This is why stock assessments even in the data rich world are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
+Empowering a global network of fisheries scientists through training and peer-support would both improve the health and management of local stocks and provide a means for ensuring that global estimates of stock status are as accurate as they can be.
+
+The coming decades are a critical time for the future of fisheries and ocean health.
+Achieving the United Nations Sustainable Development Goal 14 for the conservation and sustainable use of the world's oceans depends on our ability to effectively assess the status of fish stocks around the world.
+The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
+However, this process still leaves a substantial number of fisheries and global catch unassessed xx still need to dial in the numbers here regarding what proportion of global catch is directly covered by SOFIA assessments xx.
+Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a clear solution to this problem.
+We demonstrate here that the bottleneck to improving estimates of global stock status depends on investing in an improved global network of fishery data and fisheries scientists.
 
@@
 
+All analysis were conducted in the R programming language [@rcoreteam2019].
+Unless otherwise stated all model fitting was conducted using stan [@carpenter2017], implemented through template model builder [@kristensen2016] by the tmbstan package [@monnahan2018].
+The `sraplus` package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
+
 ## `sraplus`
 
-The core of our model is a Pella-Tomlinson [@pella1969] production model parameterzied in the manner of @winker2018.
+The core of our model is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018.
 While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
@@
 The Pella-Tomlinson model does have an issue where growth rates can become unrealistically large when the population reaches low sizes.
-We deal with this problem by following the methods described in @winker2018 to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
+We deal with this problem by following the methods described in @winker2018.
+to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
 Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see @thorson2018).
@@
 
-We  allow for process error [in the manner of the stochastic stock reduction analysis error suggested by @walters2006].
+We  allow for process error (in the manner of the stochastic stock reduction analysis error suggested by @walters2006).
 This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
@@
 carrying capacity's so small that they crash the population) and b) they create priors on both the *input* (e.g.
-growth rates and carrying capacity) and *outputs* (recent stock status), effectively putting two prior distributions for the same outcome, a problem termed Borel's Paradox [@puntXX;@rafferty1997].
+growth rates and carrying capacity) and *outputs* (recent stock status), effectively putting two prior distributions for the same outcome, a problem termed Borel's Paradox XX punt citation.
 This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
@@
 However, lower values of terminal stock status have fewer candidate values of *r* and *k*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
-Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *k* combinations: since under this model the population cannot be greater than carrying capacity, as for example *k* approaches infinity terminal stock status asymptotes at to 100% of *k*.
+Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *k* combinations: since under this model the population cannot be greater than carrying capacity, as for example *k* approaches infinity terminal stock status asymptotes at close to 100% of *k*.
 The net result of this is that even though individual combinations of *r* and *k* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
@@
 
-To resolve this, @puntXX (either cite that "on assessment of the bering-chukchi-beufort sea stock of bowhead whales paper that I can't find a date for, or see if he has a new paper on thisXX) suggests a "backwards" fitting approach, in which the model estimates growth parameters and a terminal depletion, and given the model backs out a value of *k* that satisifes the priors (conditional on an assumption initial stock status).
+To resolve this, @puntXX (either cite that "on assessment of the bering-chukchi-beufort sea stock of bowhead whales paper that I can't find a date for, or see if he has a new paper on thisXX) suggests a "backwards" fitting approach, in which the model estimates growth parameters and a terminal depletion, and given the model backs out a value of *k* that satisfies the priors (conditional on an assumption initial stock status).
 We do not adopt this approach here due to computational issues.
 Adopting the "backwards" approach requires numerically searching for the plausible *k* during each iteration of the model, which is time consuming.
-In addition, the backwards approach is not without its own complicaitons, for example the requirement of a tolerance around the value of terminal depletion produced by a given *k* and the actual value selected by the model in that time step.
-This discrepency can occur when a given terminal stock status is not technically possible given the model; to give an extreme example under this model no value of *k* can produce a terminal stock status of 0 if catches are 0 for the entire history of the fishery (assuming *k* > 0).
+In addition, the backwards approach is not without its own complications, for example the requirement of a tolerance around the value of terminal depletion produced by a given *k* and the actual value selected by the model in that time step.
+This discrepancy can occur when a given terminal stock status is not technically possible given the model; to give an extreme example under this model no value of *k* can produce a terminal stock status of 0 if catches are 0 for the entire history of the fishery (assuming *k* > 0).
 
@@
 
-@raftery1993
+<!-- @raftery1993 -->
 
@@
 
-$$log(value_i) \sim normal(log(SAR_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR}) $$
+$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
 
-XX testing training plot XX
+```{r sar-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of swept area ratio (SAR) on F/F~MSY~"}
+fit <- sraplus::sar_models$fit[sraplus::sar_models$metric == "mean_uumsy"][[1]]
+
+fit_data <- fit$data
+
+pp <- posterior_predict(fit, draws = 1000)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
+  scale_x_continuous(name = "Observed Log F/Fmsy") + 
+  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+
+```
 
@@
 
-The Fisheries Management Index (FMI), as presented in @melnychuk2016, utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
-These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomics.
+The Fisheries Management Index (FMI), as presented in @melnychuk2017, utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
+These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
 The higher the score, the better the expert judges that a given metric is met in that fishery.
@@
 
-XX testing training plot XX
+```{r fmi-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of fisheries management index (FMI) on F/F~MSY~"}
+
+fit <- sraplus::fmi_models$fit[sraplus::fmi_models$metric == "mean_uumsy"][[1]]
+
+fit_data <- fit$data
+
+pp <- posterior_predict(fit, draws = 1000)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
+  scale_x_continuous(name = "Observed Log F/Fmsy") + 
+  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+```
 
@@
 
-Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012;@costello2016;@rosenberg2018;@pauly2007].
+Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
 While these catch-only methods have been shown to have series shortcomings [@free2020], as a point of reference we use two forms of catch-only priors: a statistical and a heuristic approach.
-
-For the statistical approach, we utilize data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics...
 
@@
 
-We also define two statistical regressions relating catch history to stock status.
-To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheres in RAM, in order identify four possible clusters of catch history types within the the data.
+For the statistical approach, we utilize data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
+To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
 Emergent clusters show for example one build around "one way trip" style catch histories, others with a boom and bust, etc.
@@
 where *fyear* is the year of the fishery, starting from 0.
+
+```{r catch-reg,fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) B/B~MSY~ for regression of catch on B/B~MSY~"}
+
+fit <- sraplus::catch_b_model
+
+fit_data <- fit$data
+
+pp <- posterior_predict(fit, draws = 1000)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_scatter_avg(y = fit_data$log_value, yrep = pp) + 
+  coord_flip() +
+  scale_y_continuous(name = "Observed Log B/Bmsy") + 
+  scale_x_continuous(name = "Posterior Predictive B/Bmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+```
 
@@
 This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
-To accomodate this, we generate a catch per effective harvest rate index of abundance, as
+To accommodate this, we generate a catch per effective harvest rate index of abundance, as
 
@@
 
-Details of value of information calculates (fitting to messed up versions of RAM)
+Details of value of information calculates (fitting to messed up versions of RAM) XX need to add this in XX
+
+# Supplemental Figures
+
+```{r fao-acc-map, fig.cap = "Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
+ram-data refers to catch and abundance index drawn from RAM.
+CPUE refers to an index of abundance based on the Rousseau effort data.
+Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
+The FAO Report 569 is our reference status estimate provided by the FAO.
+ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
+fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio"}
+
+fao_acc_map_plot
+
+```
+
+```{r}
+
+rand_fits <- fao_sraplus_comp_data %>% 
+  filter(data == "catch_only") %>% 
+  mutate(bin = sample(c("over", "fully", "under"), n(), replace = TRUE),
+         data = "guess") %>% 
+  mutate(correct = bin == fao_bin)
+
+fao_sraplus_acc <- fao_sraplus_comp_data %>% 
+  bind_rows(rand_fits) %>% 
+  group_by(data) %>% 
+  summarise(accuracy = mean(correct, na.rm = TRUE)) %>% 
+  arrange(desc(accuracy))
+
+knitr::kable(fao_sraplus_acc,
+             caption = "Accuracy of different data sources in classifying stocks reported in the FAO SOFIA report to FAO status categories.
+Guess corresponds to randomly assigning a stock to one of the FAO status categories.",
+             format = "pipe", 
+             digits = 2)
+
+```
+
+```{r}
+
+rand_fits <- fao_sraplus_comp_data %>% 
+  filter(data == "catch_only") %>% 
+  mutate(bin = sample(c("over", "fully", "under"), n(), replace = TRUE),
+         data = "guess") %>% 
+  mutate(correct = bin == fao_bin)
+
+fao_sraplus_area_acc <- fao_sraplus_comp_data %>% 
+  bind_rows(rand_fits) %>% 
+  group_by(data, fao_area_code) %>% 
+  summarise(accuracy = mean(correct, na.rm = TRUE)) %>% 
+  arrange(desc(accuracy)) %>% 
+  pivot_wider(names_from = data, values_from = accuracy) %>% 
+  arrange(fao_area_code)
+
+knitr::kable(fao_sraplus_area_acc,
+             caption = "Accuracy of different data sources in classifying stocks reported in the FAO SOFIA report to FAO status categories by FAO major statistical area.
+Guess corresponds to randomly assigning a stock to one of the FAO status categories.",
+             format = "pipe", 
+             digits = 2)
+
+```
 

### 2020-07-30 functions/run-assessments.R

@@ -46,7 +46,7 @@ run_assessments <-
           f_ref_type = priors$f_ref_type,
           u_years = priors$u_years,
           b_ref_type = priors$b_ref_type,
-          shape_prior = 1.01,
+          shape_prior_source = priors$shape_prior_source,
           u_cv = priors$u_cv)
 
       if (all(is.na(driors$index)) & (length(driors$u) == 1)) {

### 2020-07-30 make-assessing-global-fisheries.R

@@ -34,19 +34,16 @@ sraplus::get_tmb_model()
 
 min_years_catch <- 20
 
-crazy_b <- 4 # threshold for suspect B/Bmsy value
+crazy_b <- 5 # threshold for suspect B/Bmsy value
 
-crazy_u <- 5 # threshold for suspect U/Umsy value
+crazy_u <- 8 # threshold for suspect U/Umsy value
 
-draws <- 2500
+draws <- 3000
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
 n_cores <- 3
 # number of cores for parallel processing
-
-future::plan(multiprocess, workers = n_cores)
-
 
 # options(mc.cores = 1)
 
@@ -54,35 +51,41 @@ results_name <- "v0.5"
 
 
 results_description <-
-  "publication version of results updated VOI from version 0.5 "
+  "publication version of results"
 
-run_voi_models <- FALSE
+run_voi_models <- TRUE
 # sub options for run_voi_models
 fit_models <- FALSE
 
-write_results <- TRUE
+write_results <- FALSE
 
 process_fits <- FALSE
 
-# other things
+# other run options things
 
 run_continent_examples <- FALSE
 
 run_ei_example <- FALSE
 
-run_sofia_comparison <- TRUE
+run_sofia_comparison <- FALSE
 
-run_ram_tests <- TRUE
+run_ram_tests <- FALSE
 
-run_ram_comparison <- TRUE
-
-
+run_ram_comparison <- FALSE
 
 engine = "stan"
 
 catchability = 1e-2
 
-theme_set(theme_classic() + theme(strip.background = element_rect(color = "transparent")))
+pub_theme <- theme_ipsum(base_size = 10,
+                         axis_text_size = 10,
+                         axis_title_size = 12) +
+    theme(panel.spacing = unit(.5,"lines"))
+
+theme_set(pub_theme)
+
+
+# theme_set(theme_classic() + theme(strip.background = element_rect(color = "transparent")))
 # data(Return)
 
 # return is from here, WARNING, changes rapidly, things break check and make sure this isn't why
@@ -122,6 +125,8 @@ has_total_biomass = ram_data %>%
   summarise(hastb = !any(is.na(total_biomass)),
             hasb = !any(is.na(b_v_bmsy))) %>%
   filter(hastb & hasb)
+
+og_ram_data <- ram_data
 
 ram_data <- ram_data %>%
   filter(stockid %in% has_total_biomass$stockid) %>%
[…]

### 2020-07-30 documents/ovando-etal-assessing-global-fisheries.Rmd
off to github

@@
 
-# Introduction (500 words)
+# Introduction
 
@@
 
-# Results  (2000 words)
+# Results
 
@@
 
-## Case Studies
+<!-- ## Case Studies  -->
 
-Two case studies provide examples of how augmenting catch data alone with local and globally available data can provide dramatically different estimates of local stock status.
+<!-- Two case studies provide examples of how augmenting catch data alone with local and globally available data can provide dramatically different estimates of local stock status.
 For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region.
@@
 Using this value to inform a prior on terminal stock status dramatically changes our picture of stock status in the Eastern Bay of Bengal, with now all stocks being overfished and experiencing high levels of overfishing (Fig.\@ref(fig:east-india)-sar).
+-->
 
-<!-- Evaluating the validity of data-limited assessments methods can be challenging.
+<!-- <!-- Evaluating the validity of data-limited assessments methods can be challenging.
 Simulation testing provides one avenue, but it can be difficult to create a simulation environment that adequately captures the problems of real-world ecosystems and data.
@@
 However, formally stock assessed fisheries represent a very biased subset of global fisheries, and it is entirely possible that a method that performs well for the kinds of species and fisheries that are commonly stock assessed might struggle in the kinds of fisheries that typically are unassessed.
--->
+--> -->
 
-<!-- Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region.
+<!-- <!-- Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region.
 As outlined in Table.\@ref(tab:status-tab), while we often have no proof that results finding that for example fisheries in south-east Asia have similar stock status to those in Alaska, very few experts would agree that this is a plausible result XX tough statement to make in a paper xx.
 We use two case studies here to present the potential of additional, broadly applicable, data to potentially improve on "catch-only" methods.
--->
+--> -->
 
-If all the stocks in question were demersal, shared similar life histories, and were all caught in the same multi-species fleet, this result might make sense.
+<!-- If all the stocks in question were demersal, shared similar life histories, and were all caught in the same multi-species fleet, this result might make sense.
 However, imposing the same prior on terminal stock status for all fisheries in a region is not ideal.
@@
 While we do not know the true state of stocks in the Eastern Bay of Bengal, inclusion of local data provide a picture more in line with local perception XX.
+-->
 
-```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio.
-Each point is a stock."}
-ei_kobe_plot 
-```
+<!-- ```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio.
+Each point is a stock."} -->
+<!-- ei_kobe_plot  -->
+<!-- ``` -->
 
-It should be no surprise that incorporating locally tailored data produces more locally plausible results.
-Global assessment at the moment though depend on few broadly available datasets.
 We used `sraplus` to assess the status of stocks in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from [@rousseau2019] (xx technically this is the updated dataset xx).
@@
 
-These case studies reveals the potential and challenges of using broadly available data for global fishery assessment.
-On the one hand, for both case studies moving beyond catch data alone produced a more plausible set of results in some ways.
-However, both case studies also raise a host of new issues.
-How realistic is it to share one index of abundance and SAR value for the entire eastern Bay of Bengal?
-Similarly, the sharing of a common effort series at the continental level does provide increased separation between European and northeast Asian stocks that may be plausible, but also estimates a suspicious amount of stocks to be essentially unfished.
-
-Extending these issues to the global level, we  used a variety of data sources to estimate stock status for fisheries listed in RAM.
+Extending this idea to the global level, we  used a variety of data sources to estimate stock status for fisheries listed in RAM.
 Our goal is to simulate a macro-level global assessment, where data are available at regional levels, but not for specific fisheries.
 As a proof of concept, we first fit models to abundance indices from each stock in RAM.
+We assess performance by three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
+Classification accuracy is calcualted as the proportion of times that a given dataset results in a stock falling into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
 There are two main reasons why an assessment model might produce poor results: model misspecification and poor data.
@@
 Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality.
-Overall the `sraplus` estimates using the RAM data are reasonably good.
-The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias.
+Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(teb:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.
+
+<!-- The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias.
 The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time.
@@
 While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.
+-->
 
-Performance limitations then are likely to arise less from model misspecification but from the quality of the data themselves.
+Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves.
 These becomes even clearer once we consider the performance of `sraplus` models fit to globally available datasets.
@@
 
-# Discussion (500 words)
+# Discussion
 
@@
 
-Our estimates of global stock status vary wildly depending on which source of data we use.
+<!-- Our estimates of global stock status vary wildly depending on which source of data we use.
 Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information?
@@
 need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).
+-->
 
@@
 
-As we should expect having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 20%.
+Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 20%.
 In terms of data however, information on fishing mortality rates (F/F~MSY~) in the most recent year was almost as useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot).
@@
 
-# Methods (3000 words)
+# Methods
 

### 2020-07-30 documents/ovando-etal-assessing-global-fisheries.Rmd
a bit of cleanup

@@
   bookdown::pdf_document2: default
+params:
+  results_name: ["v0.5"]
+  min_years_catch: [25]
 abstract: |
@@
 We show that our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
+always_allow_html: true
 ---
@@
 
-results_name <- "v0.5"
+results_name <- params$results_name
 
@@
 
-min_years_catch <- 25
+min_years_catch <- params$min_years_catch
 
@@
 We assess performance by three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
-Classification accuracy is calcualted as the proportion of times that a given dataset results in a stock falling into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
+Classification accuracy is calculated as the proportion of times that a given dataset results in a stock falling into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
 There are two main reasons why an assessment model might produce poor results: model misspecification and poor data.
@@
 Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality.
-Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(teb:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
 While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.
@@
 For example, despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis.
-But, applying one SAR value to all stocks in a region causes innacurate estimates of stock status.
+But, applying one SAR value to all stocks in a region causes inaccurate estimates of stock status.
 The challenge comes in attempting to indiscriminately apply these data to all areas: doing so results in substantially incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.

### 2020-07-30 make-assessing-global-fisheries.R

@@ -73,7 +73,11 @@ run_ram_tests <- FALSE
 
 run_ram_comparison <- FALSE
 
-engine = "stan"
+knit_paper <- TRUE
+
+warning("Running full analysis takes upwards of 12 hours on 4 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 4 cores is not recommended.")
+
+engine <-  "stan"
 
 catchability = 1e-2
 
@@ -175,35 +179,7 @@ if (run_voi_models == TRUE) {
     mutate(b_ref_type = ifelse(initial_state_type == "prior",
                                "b",
                                b_ref_type))
-  
-  # stress_tests <- data_frame(
-  #   use_index = sample(c(FALSE),  draws, replace = TRUE),
-  #   index_type = sample(c("survey"), draws, replace = TRUE),
-  #   initial_state_type = sample(c("unfished", "known", "prior","heuristic"), draws, replace = TRUE),
-  #   initial_state_cv =  runif(draws, 0.05, 0.2),
-  #   use_terminal_state = sample(c(TRUE),  draws, replace = TRUE),
-  #   terminal_state_cv = runif(draws, 0.05, 0.2),
-  #   error_cv = sample(c(0, 1), draws, replace = TRUE),
-  #   stockid = sample(unique(ram_data$stockid), draws, replace = TRUE),
-  #   use_u_priors =  sample(c(FALSE),  draws, replace = TRUE),
-  #   u_cv = runif(draws, 0.05, 0.2),
-  #   index_window = sample(c(1, 0.25, .5), draws, replace = TRUE),
-  #   index_freq = sample(c(1, 2), draws, replace = TRUE),
-  #   u_window = sample(c("snapshot", "recent", "complete"), draws, replace = TRUE),
-  #   b_ref_type = sample(c("b"),  draws, replace = TRUE),
-  #   f_ref_type = sample(c("f", "fmsy"),  draws, replace = TRUE),
-  #   estimate_shape = sample(c(FALSE, TRUE), draws, replace = TRUE),
-  #   estimate_proc_error = sample(c(FALSE, TRUE), draws, replace = TRUE)
-  # ) %>%
-  #   mutate(error_cv = map_dbl(error_cv, ~ ifelse(.x == 0, 1e-6, runif(1, 0.05, 0.2)))) %>%
-  #   mutate(
-  #     b_ref_type = ifelse(
[…]

### 2020-07-30 make-assessing-global-fisheries.R
more cleaning

@@ -1116,7 +1116,7 @@ stats <-
 
 sofia_status$status <- stats
 
-# there are MANY mispellings of scientific names in SOFIA
+# there are many mispellings of scientific names in SOFIA that need correcting
 
 correct_sofia <- sofia_status %>% 
   select(common_name, scientific_name) %>% 
@@ -2200,7 +2200,7 @@ ram_mpe_map_plot <- fao_area_ram_status %>%
     limits = c(-1,3.5),
     breaks = seq(-1, 3.5, by = .5),
     guide = guide_colorbar(
-      barwidth = ggplot2::unit(15, "lines"),
+      barwidth = ggplot2::unit(10, "lines"),
       axis.linewidth = 1,
       ticks.colour = "black",
       frame.colour = "black"
@@ -2324,6 +2324,9 @@ save(file = here("results", results_name, "paper_plots.Rdata"),
 # knit paper --------------------------------------------------------------
 if (knit_paper == TRUE){
   
-  rmarkdown::render(here::here("documents","ovando-etal-assessing-global-fisheries.Rmd"), params = list(results_name = results_name,
-                                                                                                        min_years_catch = 25))
+  rmarkdown::render(
+    here::here("documents", "ovando-etal-assessing-global-fisheries.Rmd"),
+    params = list(results_name = results_name,
+                  min_years_catch = 25)
+  )
 }

### 2020-08-06 documents/ovando-etal-assessing-global-fisheries.Rmd
merging back to master

@@
   - James Thorson
+  - Yannick Rousseau
   - Yimin Ye
@@
 
-<!-- <!-- Evaluating the validity of data-limited assessments methods can be challenging.
+ <!-- Evaluating the validity of data-limited assessments methods can be challenging.
 Simulation testing provides one avenue, but it can be difficult to create a simulation environment that adequately captures the problems of real-world ecosystems and data.
@@
 However, formally stock assessed fisheries represent a very biased subset of global fisheries, and it is entirely possible that a method that performs well for the kinds of species and fisheries that are commonly stock assessed might struggle in the kinds of fisheries that typically are unassessed.
---> -->
+-->
 
-<!-- <!-- Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region.
+ <!-- Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region.
 As outlined in Table.\@ref(tab:status-tab), while we often have no proof that results finding that for example fisheries in south-east Asia have similar stock status to those in Alaska, very few experts would agree that this is a plausible result XX tough statement to make in a paper xx.
 We use two case studies here to present the potential of additional, broadly applicable, data to potentially improve on "catch-only" methods.
---> -->
+-->
 
@@
 However, the shared effort series across these stocks also produces a large number of effectively unfished stocks; a highly unlikely result for both regions (Fig.\@ref(fig:eu-v-na)).
+
+```{}
+
+```
 

### 2020-08-06 make-assessing-global-fisheries.R

@@ -1130,7 +1130,6 @@ correct_sofia <-  correct_sofia %>%
   mutate(good_scientific_name = map_chr(sciname, ~.x[1])) %>% 
   select(common_name,good_scientific_name )
 
-# correct_sofia$foo <- correct_sofia$bad_scientific_name != correct_sofia$good_scientific_name
 
 sofia_status <- sofia_status %>%
   left_join(correct_sofia, by = "common_name") %>%

### 2020-08-06 functions/fit-india.R
Cleanup (#4)
* major change making the analysis primarily RAM focused

@@ -102,7 +102,7 @@ sar_fit <-
       )
 
     sfs <- safely(fit_sraplus)
-
+    
     cpue_fit <-
       sfs(driors = cpue_driors,
           include_fit = include_fit,
@@ -115,7 +115,7 @@ sar_fit <-
           chains = chains,
           refresh = refresh,
           cores = cores,
-          adapt_delta = 0.9)
+          workers = 4)
 
     if (is.null(cpue_fit$error)){
       cpue_fit <- cpue_fit$result

### 2020-08-06 functions/fit_continent_examples.R

@@ -84,7 +84,7 @@ fit_continent_examples <-
         years = seq_along(years),
         effort = data$total_effort[effort_years],
         effort_years = effort_years,
-        q_slope_prior = 0.025,
+        q_slope_prior = q_slope_prior,
         q_slope_prior_cv = 0.2,
         b_ref_type = "k"
       )
@@ -109,7 +109,8 @@ fit_continent_examples <-
         estimate_shape = estimate_shape,
         estimate_qslope = estimate_qslope,
         estimate_proc_error = estimate_proc_error,
-        use_baranov = use_baranov
+        use_baranov = use_baranov,
+        workers = 4
       )
     #
     # plot(cpue_driors$catch[cpue_driors$effort_years] / cpue_driors$effort)

### 2020-08-06 functions/generate-priors.R

@@ -19,7 +19,8 @@ generate_priors <-
            u_window = 1,
            u_cv = 0.1,
            index_freq = 1,
-           effort_freq = 1) {
+           effort_freq = 1,
+           shape_prior_source = "thorson") {
     if (initial_state_type == "known") {
       initial_state = dplyr::case_when(b_ref_type == "k" ~ dat$b_rel[1],
                                        TRUE ~ dat$b_v_bmsy[1])
@@ -168,7 +169,6 @@ generate_priors <-
       
       
     } # close if use heuristics
-    
     priors <-
       list(
         carry = carry,
@@ -185,11 +185,12 @@ generate_priors <-
         index_years = index_years,
         effort_years = effort_years,
         u_years = u_years,
-        u_cv = u_cv
+        u_cv = u_cv,
+        shape_prior_source = shape_prior_source
       )
 
     # apply error to priors
-    error_prone <- !str_detect(names(priors), "cv|years|anchors")
+    error_prone <- !str_detect(names(priors), "cv|years|anchors|prior")
 
     priors <-
       map_if(priors, error_prone, ~ .x * exp(rnorm(length(.x), 0, error_cv) - error_cv ^

### 2020-08-06 functions/prepare-sofia-data.R

@@ -556,6 +556,8 @@ prepare_sofia_data <- function(min_years_catch = 20,
   fmi$fao_country_name <-
     countrycode::countrycode(fmi$country_rfmo, "country.name", "un.name.en")
   
+  fmi$country <- countrycode::countrycode(fmi$country_rfmo, "country.name","un.name.en")
+  
   fmi$region <-
     countrycode::countrycode(fmi$country_rfmo, "country.name", "region")
   
@@ -769,6 +771,17 @@ prepare_sofia_data <- function(min_years_catch = 20,
   
   # assign("effort_region_to_country",effort_region_to_country, envir = .GlobalEnv )
   
+  rous_data <- read.csv(here("data", "MappedFAO.csv")) %>% 
+    na.omit() %>% 
+    as_tibble() %>% 
+    janitor::clean_names() %>% 
+    mutate(country = countrycode::countrycode(iso3, "iso3c", "un.name.en")) %>% 
+    filter(type2 == "I") %>% 
+    select(year,fao, effort_cell_reported_nom, country ) %>% 
+    rename(area = fao)
+  
+  assign("rous_data",rous_data, envir = .GlobalEnv )
+  
   
   # ugh. manual refactoring of region codes
   
@@ -818,6 +831,8 @@ prepare_sofia_data <- function(min_years_catch = 20,
   ei_capture <-
     readxl::read_xlsx(here::here("data", "India East Only.xlsx")) %>%
     janitor::clean_names() %>%
+    select(-total, -ratio, -x78) %>% 
+    filter(!is.na(country_country)) %>% 
     gather(year, capture, x1950:x2015) %>%
     mutate(year = as.numeric(str_replace_all(year, "\\D", "")))
  

### 2020-08-06 make-assessing-global-fisheries.R

@@ -1,4 +1,5 @@
 
+
 # setup -------------------------------------------------------------------
 
 library(tidyverse)
@@ -33,49 +34,62 @@ sraplus::get_tmb_model()
 
 min_years_catch <- 20
 
-crazy_b <- 4 # threshold for suspect B/Bmsy value
+crazy_b <- 5 # threshold for suspect B/Bmsy value
 
-crazy_u <- 5 # threshold for suspect U/Umsy value
+crazy_u <- 8 # threshold for suspect U/Umsy value
 
-draws <- 2500
+draws <- 3000
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
 n_cores <- 3
 # number of cores for parallel processing
 
-future::plan(multiprocess, workers = n_cores)
-
-
 # options(mc.cores = 1)
 
 results_name <- "v0.5"
 
 
-results_description <- "publication version of results updated VOI from version 0.5 "
+results_description <-
+  "publication version of results"
 
-run_voi_models <- FALSE
+run_voi_models <- TRUE
 # sub options for run_voi_models
 fit_models <- FALSE
 
-write_results <- TRUE
+write_results <- FALSE
 
 process_fits <- FALSE
 
-# other things
-run_case_studies <- FALSE
+# other run options things
 
 run_continent_examples <- FALSE
 
 run_ei_example <- FALSE
 
-run_sofia_comparison <- TRUE
+run_sofia_comparison <- FALSE
 
-engine = "stan"
+run_ram_tests <- FALSE
+
+run_ram_comparison <- FALSE
+
+knit_paper <- TRUE
+
+warning("Running full analysis takes upwards of 12 hours on 4 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 4 cores is not recommended.")
+
+engine <-  "stan"
 
 catchability = 1e-2
 
-theme_set(theme_classic() + theme(strip.background = element_rect(color = "transparent")))
+pub_theme <- theme_ipsum(base_size = 10,
+                         axis_text_size = 10,
+                         axis_title_size = 12) +
+    theme(panel.spacing = unit(.5,"lines"))
+
+theme_set(pub_theme)
+
+
+# theme_set(theme_classic() + theme(strip.background = element_rect(color = "transparent")))
 # data(Return)
 
 # return is from here, WARNING, changes rapidly, things break check and make sure this isn't why
[…]

### 2020-08-06 functions/fit-ram.R
cmsy working for ram test

@@ -13,6 +13,7 @@ fit_ram <-
            default_initial_state_cv = 0.2,
            chains = 1,
            cores = 1,
+           cmsy_cores = 4,
            reg_cv = NA,
            draws = 2e6,
            refresh = 0,
@@ -56,7 +57,48 @@ fit_ram <-
       mutate(data = "heuristic")
     
     
+    
+    # updated cmsy fit
+    
+    sfc <- purrr::safely(funct_cmsy)
+    
+    ported_cmsy <- sfc(
+      catches = catches[years >= 1950],
+      catch_years = years[years >= 1950],
+      stock = scientific_name,
+      common_name = unique(data$common_name),
+      scientific_name = unique(data$scientific_name),
+      res = unique(data$resilience),
+      start.yr = years[years >= 1950][1],
+      end.yr = dplyr::last(years[years >= 1950]),
+      cores = cmsy_cores
+    )
+    
+    if (is.null(ported_cmsy$error)){
+      
+      cmsy_results <- ported_cmsy$result %>% 
+        mutate(c_div_msy  = ct / msy) %>% 
+        select(year, bt, b_bmsy, f_fmsy, c_div_msy) %>% 
+        rename(depletion = bt, b_div_bmsy = b_bmsy, u_div_umsy = f_fmsy) %>% 
+        pivot_longer(-year, names_to = "variable", values_to = "mean") %>% 
+        mutate(sd = NA, lower = NA, upper = NA, data = "cmsy")
+      
+      # tests <- cmsy_results %>% 
+      #   bind_rows(basic_fit_results)
+      # 
+      # tests %>% 
+      #   filter(variable == "c_div_msy") %>% 
+      #   select(year, mean, data) %>% 
+      #   ggplot(aes(year, mean, color = data)) + 
+      #   geom_line()
+    
+      cmsy_worked <- TRUE 
+    } else {
+      cmsy_worked <- FALSE
+    }
+
     # catch only fit
+    
     com_driors <-
       format_driors(
         taxa = scientific_name,
@@ -674,7 +716,14 @@ fit_ram <-
         } else{
           .
         }
+      } %>% {
+        if (cmsy_worked == TRUE){
+          bind_rows(., cmsy_results)
+        } else {
+          .
+        }
       }
+      
     
     # write(wtf,file = "wtf.txt", append = TRUE)
     

### 2020-08-06 make-assessing-global-fisheries.R

@@ -9,6 +9,7 @@ library(rstan)
 library(mvtnorm)
 library(FishLife)
 library(furrr)
+library(doParallel)
 library(future)
 library(viridis)
 library(sraplus)
@@ -53,7 +54,7 @@ results_name <- "v0.5"
 results_description <-
   "publication version of results"
 
-run_voi_models <- TRUE
+run_voi_models <- FALSE
 # sub options for run_voi_models
 fit_models <- FALSE
 
@@ -71,7 +72,7 @@ run_sofia_comparison <- FALSE
 
 run_ram_tests <- FALSE
 
-run_ram_comparison <- FALSE
+run_ram_comparison <- TRUE
 
 knit_paper <- TRUE
 
@@ -2006,20 +2007,38 @@ if (run_ram_comparison == TRUE) {
     geom_line() + 
     facet_wrap(~fao_area_code, scales = "free_y")
   
-  set.seed(24)
-  
   future::plan(multisession, workers = n_cores)
+  
+  comp_stocks <- ram_comp_data %>%
+    select(scientific_name) %>%
+    unique() %>%
+    mutate(fishbase_vuln = map_dbl(
+      scientific_name,
+      ~ rfishbase::species(.x, fields = "Vulnerability")$Vulnerability
+    )) %>%
+    mutate(
+      resilience = dplyr::case_when(
+        fishbase_vuln < 33 ~ "Low",
+        fishbase_vuln >= 66 ~ "High",
+        TRUE ~ "Medium"
+      )
+    ) %>% 
+    select(scientific_name, resilience)
+  
+  ram_comp_data <- ram_comp_data %>% 
+    left_join(comp_stocks, by = "scientific_name") %>% 
+    mutate(resilience = ifelse(is.na(resilience), "Medium",resilience))
   
   ram_status_fits <- ram_comp_data %>%
     # filter(fao_area_code %in% 67) %>%
     group_by(stockid) %>%
     nest() %>%
     ungroup() %>% 
-    # sample_n(1) %>%
+    # sample_n(10) %>%
     mutate(
       fits = future_map(
         data,
-       safely( fit_ram),
+       safely(fit_ram),
         support_data = support_data,
         default_initial_state = NA,
         default_initial_state_cv = NA,
@@ -2028,7 +2047,7 @@ if (run_ram_comparison == TRUE) {
         cores = 2,
         .progress = TRUE,
         .options = future_options(
[…]

### 2020-08-08 functions/SchaeferMC.R
off to squid

@@ -1,57 +1,0 @@
-SchaeferMC <-
-  function(ri,
-           ki,
-           startbio,
-           int.yr,
-           intbio,
-           endbio,
-           sigR,
-           pt,
-           duncert,
-           startbins,
-           ni,
-           yr,
-           ct,
-           end.yr) {
-    # create vector for initial biomasses
-    #
-
-    nyr <- length(yr)
-
-    startbt     <-
-      seq(
-        from = startbio[1],
-        to = startbio[2],
-        by = (startbio[2] - startbio[1]) / startbins
-      )
-    nstartbt    <- length(startbt)
-    npoints     <- length(ri)
-    # get index of intermediate year
-    int.yr.i     <- which(yr == int.yr)
-
-    #loop through r-k pairs with parallel search
-    mdat <-
-      SchaeferParallelSearch(
-        ni = ni,
-        nyr = nyr,
-        sigR = sigR,
-        duncert = duncert,
-        ct = ct,
-        int.yr = int.yr,
-        intbio = intbio,
-        startbt = startbt,
-        ki = ki,
-         i = i,
-        ri = ri,
-        int.yr.i = int.yr.i,
-        nstartbt = nstartbt,
-        yr = yr,
-        end.yr = end.yr,
-        endbio = endbio,
-        npoints = npoints,
-        pt = pt
-      )
-
-    cat("\n")
-    return(list(mdat))
-  } # end of SchaeferMC function

### 2020-08-08 functions/SchaeferParallelSearch.R

@@ -1,137 +1,0 @@
-#----------------------------------------------
-#  FUNCTIONS ----
-#----------------------------------------------
-# Monte Carlo filtering with Schaefer Function ----
-#----------------------------------------------
-SchaeferParallelSearch <-
-  function(ni,
-           nyr,
-           sigR,
-           duncert,
-           ct,
-           int.yr,
-           intbio,
-           startbt,
-           ki,
-           i,
-           ri,
-           int.yr.i,
-           nstartbt,
-           yr,
-           end.yr,
-           endbio,
-           npoints,
-           pt) {
-    ptm <- proc.time()
-    # create vectors for viable r, k and bt
-    inmemorytable <- vector()
-    # parallelised for the points in the r-k space
-    inmemorytable <-
-      foreach (
-        i = 1:npoints,
-        .combine = 'rbind',
-        .packages = 'foreach',
-        .inorder = TRUE
-      ) %dopar% {
-        nsbt = length(startbt)
-        VP   <- FALSE
-        for (nj in 1:nsbt) {
-          # create empty vector for annual biomasses
-          bt <- vector()
-          j <- startbt[nj]
-          # set initial biomass, including 0.1 process error to stay within bounds
-          bt[1] = j * ki[i] * exp(rnorm(1, 0, 0.1 * sigR))  ## set biomass in first year
-          # repeat test of r-k-startbt combination to allow for different random error
-          for (re in 1:ni)   {
-            #loop through years in catch time series
-            for (t in 1:nyr)  {
-              # for all years in the time series
-              xt = rnorm(1, 0, sigR) # set new process error for every year
-              zlog.sd = sqrt(log(1 + (duncert) ^ 2))
-              zt = rlnorm(1, meanlog = 0, sdlog = zlog.sd) # model the catch error as a log normal distribution.
-              # calculate biomass as function of previous year's biomass plus surplus production minus catch
-              bt[t + 1] <- ifelse(
-                bt[t] / ki[i] >= 0.25,
-                bt[t] + ri[i] * bt[t] * (1 - bt[t] / ki[i]) *
-                  exp(xt) - ct[t] * zt,
-                bt[t] + (4 * bt[t] / ki[i]) * ri[i] * bt[t] *
-                  (1 - bt[t] / ki[i]) * exp(xt) - ct[t] * zt
-              ) # assuming reduced r at B/k < 0.25
-
-              # if biomass < 0.01 k, discard r-k-startbt combination
[…]

### 2020-08-08 functions/fit-fao.R

@@ -16,7 +16,8 @@ fit_fao <-
            reg_cv = NA,
            draws = 2e6,
            refresh = 0,
-           thin_draws = TRUE) {
+           thin_draws = TRUE, 
+           cmsy_cores = 3) {
     #
     #     data <- test$data[[1]]
     
@@ -27,6 +28,44 @@ fit_fao <-
     catches <- data$catch
     
     years <- data$year
+    
+    
+    # fit cmsy
+    # 
+    sfc <- purrr::safely(portedcmsy::funct_cmsy)
+    
+    ported_cmsy <- sfc(
+      catches = catches[years >= 1950],
+      catch_years = years[years >= 1950],
+      stock = scientific_name,
+      common_name = NA,
+      scientific_name = scientific_name,
+      res = unique(data$resilience),
+      start.yr = years[years >= 1950][1],
+      end.yr = dplyr::last(years[years >= 1950]),
+      cores = cmsy_cores
+    )
+    if (is.null(ported_cmsy$error)) {
+      cmsy_results <- ported_cmsy$result %>%
+        mutate(c_div_msy  = ct / msy) %>%
+        select(year, bt, b_bmsy, f_fmsy, c_div_msy) %>%
+        rename(depletion = bt,
+               b_div_bmsy = b_bmsy,
+               u_div_umsy = f_fmsy) %>%
+        pivot_longer(-year, names_to = "variable", values_to = "mean") %>%
+        mutate(
+          sd = NA,
+          lower = NA,
+          upper = NA,
+          data = "cmsy"
+        )
+      
+      cmsy_worked <- TRUE
+    } else {
+      cmsy_worked <- FALSE
+    }
+    
+    # fit to u
     
     u_years <- seq_along(years)[!is.na(data$mean_u_umsy)]
     
@@ -595,15 +634,13 @@ fit_fao <-
         else{
           .
         }
+      }  %>% {
+        if (cmsy_worked) {
+          bind_rows(., cmsy_results)
+        } else {
+          .
+        }
       }
-    
-    # write(wtf,file = "wtf.txt", append = TRUE)
-    
-    # results %>%
-    #   filter(data == "sar") %>%
-    #   ggplot(aes(year, mean, color = data)) +
-    #   geom_line() +
-    #   facet_wrap(~variable, scales = "free_y")
     
     rm(list = ls()[!str_detect(ls(), "results")])
     gc()

### 2020-08-08 functions/fit-india.R

@@ -17,9 +17,53 @@ fit_india <-
            draws = 1e6,
            chains = 1,
            refresh = 500,
-           cores = 1){
+           cores = 1,
+           cmsy_cores = 8) {
+    sfc <- purrr::safely(portedcmsy::funct_cmsy)
+    
+    ported_cmsy <- sfc(
+      catches = catches[years >= 1950],
+      catch_years = years[years >= 1950],
+      stock = scientific_name,
+      common_name = NA,
+      scientific_name = scientific_name,
+      res = unique(data$resilience),
+      start.yr = years[years >= 1950][1],
+      end.yr = dplyr::last(years[years >= 1950]),
+      cores = cmsy_cores
+    )
+    if (is.null(ported_cmsy$error)) {
+      cmsy_results <- ported_cmsy$result %>%
+        mutate(c_div_msy  = ct / msy) %>%
+        select(year, bt, b_bmsy, f_fmsy, c_div_msy) %>%
+        rename(depletion = bt,
+               b_div_bmsy = b_bmsy,
+               u_div_umsy = f_fmsy) %>%
+        pivot_longer(-year, names_to = "variable", values_to = "mean") %>%
+        mutate(
+          sd = NA,
+          lower = NA,
+          upper = NA,
+          data = "cmsy"
+        )
+      
+      # tests <- cmsy_results %>%
+      #   bind_rows(basic_fit_results)
+      #
+      # tests %>%
+      #   filter(variable == "c_div_msy") %>%
+      #   select(year, mean, data) %>%
+      #   ggplot(aes(year, mean, color = data)) +
+      #   geom_line()
+      
+      cmsy_worked <- TRUE
+    } else {
+      cmsy_worked <- FALSE
+    }
+    
+    
     # heuristic example
-
+    
     basic_driors <-
       format_driors(
         taxa = scientific_name,
@@ -32,7 +76,7 @@ fit_india <-
         use_heuristics = TRUE,
         b_ref_type = "k"
       )
-
+    
     basic_fit <-
       fit_sraplus(
         driors = basic_driors,
@@ -41,50 +85,52 @@ fit_india <-
         engine = "sir",
         draws = draws
       )
-
-
+    
+    
     basic_fit_results <- basic_fit$results %>%
-      filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy","depletion")) %>%
[…]

### 2020-08-08 functions/fit-ram.R

@@ -9,11 +9,12 @@ fit_ram <-
            q_slope_prior = 0.025,
            estimate_proc_error = TRUE,
            estimate_qslope = FALSE,
+           estimate_shape = FALSE,
            default_initial_state = NA,
            default_initial_state_cv = 0.2,
            chains = 1,
            cores = 1,
-           cmsy_cores = 4,
+           cmsy_cores = 1,
            reg_cv = NA,
            draws = 2e6,
            refresh = 0,
@@ -21,6 +22,9 @@ fit_ram <-
     #
     #     data <- test$data[[1]]
     # browser()
+    
+    sfs <- safely(fit_sraplus)
+    
     ogengine <- engine
     
     scientific_name <- unique(data$scientific_name)
@@ -29,39 +33,8 @@ fit_ram <-
     
     years <- data$year
     
-    u_years <- seq_along(years)[!is.na(data$mean_u_umsy)]
-    
-    u_umsy <- data$mean_u_umsy[u_years]
-    basic_driors <-
-      format_driors(
-        taxa = scientific_name,
-        catch = catches,
-        years = seq_along(years),
-        use_heuristics = TRUE,
-        b_ref_type = "k",
-        isscaap_group = unique(data$isscaap_group)
-      )
-    basic_fit <-
-      fit_sraplus(
-        driors = basic_driors,
-        include_fit = include_fit,
-        model = model,
-        engine = "sir",
-        draws = draws,
-        thin_draws = thin_draws
-      )
-    
-    basic_fit_results <- basic_fit$results %>%
-      filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy", "depletion")) %>%
-      mutate(year = rep(years, each = 4)) %>%
-      mutate(data = "heuristic")
-    
-    
-    
-    # updated cmsy fit
-    
-    sfc <- purrr::safely(funct_cmsy)
-    
+    sfc <- purrr::safely(portedcmsy::funct_cmsy)
+    # browser()
     ported_cmsy <- sfc(
       catches = catches[years >= 1950],
       catch_years = years[years >= 1950],
@@ -73,6 +46,7 @@ fit_ram <-
       end.yr = dplyr::last(years[years >= 1950]),
       cores = cmsy_cores
     )
+    write_rds(ported_cmsy, "wtf.rds")
     
     if (is.null(ported_cmsy$error)){
       
@@ -91,11 +65,58 @@ fit_ram <-
[…]

### 2020-08-08 functions/fit_continent_examples.R

@@ -11,6 +11,7 @@ fit_continent_examples <-
           first_effort_year = 1950,
           chains = 1,
           cores = 1,
+          cmsy_cores = 4,
           q_slope_prior = 0.025) {
     # heuristic example
     # i = 2
@@ -26,10 +27,54 @@ fit_continent_examples <-
     # use_baranov = TRUE
     # first_effort_year = 1950
 
+    
     catches <- data$total_catch
-
+    
     years <- data$year
-
+    
+ 
+  
+    # fit cmsy
+  
+    sfc <- purrr::safely(portedcmsy::funct_cmsy)
+    
+    ported_cmsy <- sfc(
+      catches = catches[years >= 1950],
+      catch_years = years[years >= 1950],
+      stock = scientific_name,
+      common_name = NA,
+      scientific_name = scientific_name,
+      res = unique(data$resilience),
+      start.yr = years[years >= 1950][1],
+      end.yr = dplyr::last(years[years >= 1950]),
+      cores = cmsy_cores
+    )
+    if (is.null(ported_cmsy$error)){
+      
+      cmsy_results <- ported_cmsy$result %>% 
+        mutate(c_div_msy  = ct / msy) %>% 
+        select(year, bt, b_bmsy, f_fmsy, c_div_msy) %>% 
+        rename(depletion = bt, b_div_bmsy = b_bmsy, u_div_umsy = f_fmsy) %>% 
+        pivot_longer(-year, names_to = "variable", values_to = "mean") %>% 
+        mutate(sd = NA, lower = NA, upper = NA, data = "cmsy")
+      
+      # tests <- cmsy_results %>% 
+      #   bind_rows(basic_fit_results)
+      # 
+      # tests %>% 
+      #   filter(variable == "c_div_msy") %>% 
+      #   select(year, mean, data) %>% 
+      #   ggplot(aes(year, mean, color = data)) + 
+      #   geom_line()
+      
+      cmsy_worked <- TRUE 
+    } else {
+      cmsy_worked <- FALSE
+    }
+    
+    
+    # fit heuristic
+    
     basic_driors <-
       format_driors(
         taxa = scientific_name,
@@ -170,7 +215,13 @@ fit_continent_examples <-
     }
 
     results <- cpue_fit_results %>%
-      bind_rows(basic_fit_results)
+      bind_rows(basic_fit_results) %>% {
+        if (cmsy_worked){
+          bind_rows(., cmsy_results)
+        } else {
+          .
+        }
+      }
 
     return(results)
 

### 2020-08-08 functions/funct-cmsy.R

@@ -1,3517 +1,0 @@
-#' run functional cmsy
-#'
-#' ported from https://github.com/SISTA16/cmsy on August 5 2020
-#' No changes to any methods made
-#'
-#'
-#' @param catches a continuous vector of catches
-#' @param years a continuous vector of years corresponding to catches
-#' @param stock the name of the stock
-#' @param region the region the stock is in
-#' @param subregion the subregion the stock is in
-#' @param common_name the common name of the species
-#' @param ScientificName the scientific name of the species
-#' @param bio_index  a vector of an abundance index
-#' @param r.low  lower bounds of intrinsic growth rate r
-#' @param r.hi  upper bound of intrinsic growth rate r
-#' @param res  resilience category for the stock
-#' @param stb.low  lower end b status in starting year
-#' @param stb.hi upper end b status in starting year
-#' @param int.yr year for intermediate estimate of stock status
-#' @param intb.low lower end b status in intermediate year
-#' @param intb.hi upper end b status in intermediate year
-#' @param endb.low lower end b status in last year
-#' @param endb.hi upper end b status in last year
-#' @param q.start initial catchability
-#' @param q.end final catchability
-#' @param e.creep something to do with effort creep
-#' @param btype the type of biomass index, one of None, CPUE, biomass
-#' @param force.cmsy force use of cmsy instead of BSM, i think?
-#' @param comment any comment that goes with the stock
-#' @param source the source of any information, i think
-#' @param dataUncert set observation error as uncertainty in catch - default is SD=0.3
-#' @param sigmaR overall process error for CMSY; SD=0.1 is the default
-#' @param cor.log.rk empirical value of log r-k correlation in 140 stocks analyzed with BSM (without r-k correlation)
-#' @param rk.cor.beta beta.prior for rk cor+1
-#' @param nbk  Number of B/k priors to be used by BSM, with options 1 (first year), 2 (first & intermediate), 3 (first, intermediate & final bk priors)
-#' @param q.biomass.pr  if btype=="biomass" this is the prior range for q
-#' @param n  initial number of r-k pairs
-#' @param ni iterations for r-k-startbiomass combinations, to test different variability patterns; no improvement seen above 3
-#' @param nab  recommended=5; minimum number of years with abundance data to run BSM
[…]

### 2020-08-08 functions/mvn.R

@@ -1,13 +1,0 @@
-#-------------------------------------------------------------
-# Function to create multivariate-normal distribution for r-k
-#-------------------------------------------------------------
-mvn   <- function(n,mean.log.r,sd.log.r,mean.log.k,sd.log.k,cor.log.rk) {
-  cov.log.rk <- cor.log.rk*sd.log.r*sd.log.k # covariance with empirical correlation and prior variances  covar.log.rk = matrix(NA, ncol=2,nrow=2)   # contract covariance matrix
-  covar.log.rk      <- matrix(NA, ncol=2,nrow=2) # covariance matrix
-  covar.log.rk[1,1] <- sd.log.r^2                # position [1,1] is variance of log.r
-  covar.log.rk[2,2] <- sd.log.k^2               # position [2,2] is variance of log.k
-  covar.log.rk[1,2] = covar.log.rk[2,1] = cov.log.rk     # positions [1,2] and [2,1] are correlations
-  mu.log.rk  <- (c(mean.log.r,mean.log.k))      # vector of log.means
-  mvn.log.rk <- rmvnorm(n,mean=mu.log.rk,sigma=covar.log.rk,method="svd")
-  return(mvn.log.rk)
-}

### 2020-08-08 make-assessing-global-fisheries.R

@@ -24,6 +24,7 @@ library(tidybayes)
 library(rstan)
 library(rstanarm)
 library(sf)
+library(portedcmsy)
 Sys.unsetenv("PKG_CXXFLAGS")
 
 options(dplyr.summarise.inform = FALSE)
@@ -72,7 +73,7 @@ run_sofia_comparison <- FALSE
 
 run_ram_tests <- FALSE
 
-run_ram_comparison <- TRUE
+run_ram_comparison <- FALSE
 
 knit_paper <- TRUE
 
@@ -763,14 +764,35 @@ total_stocks %>%
 
 # future::plan(future::multiprocess, workers = 4)
 
+
+comp_stocks <- total_stocks %>%
+  select(scientific_name) %>%
+  unique() %>%
+  mutate(fishbase_vuln = map_dbl(
+    scientific_name,
+    ~ rfishbase::species(.x, fields = "Vulnerability")$Vulnerability
+  )) %>%
+  mutate(
+    resilience = dplyr::case_when(
+      fishbase_vuln < 33 ~ "Low",
+      fishbase_vuln >= 66 ~ "High",
+      TRUE ~ "Medium"
+    )
+  ) %>% 
+  select(scientific_name, resilience)
+
+total_stocks <- total_stocks %>% 
+  left_join(comp_stocks, by = "scientific_name") %>% 
+  mutate(resilience = ifelse(is.na(resilience), "Medium",resilience))
+
 if (run_continent_examples == TRUE) {
-  set.seed(42)
+  set.seed(24)
   
   continent_fits <- total_stocks %>%
     group_by(scientific_name, continent) %>%
     nest() %>%
     ungroup() #%>%
-  # sample_n(10)
+  # sample_n(3)
   
   sfe <- safely(fit_continent_examples)
   
@@ -786,7 +808,8 @@ if (run_continent_examples == TRUE) {
         estimate_proc_error = TRUE,
         estimate_qslope = FALSE,
         estimate_shape = FALSE,
-        first_effort_year = 1960
+        first_effort_year = 1960,
+        cmsy_cores = 8
       )
     )
   
@@ -931,11 +954,30 @@ ei_data <- ei_capture %>%
   left_join(ei_cpue, by = "year") %>%
   left_join(fao_names, by = c("species_asfis_species" = "common_name")) %>%
   ungroup() %>%
-  filter(!is.na(scientific_name)) %>%
+  filter(!is.na(scientific_name)) 
+
+comp_stocks <- ei_data %>%
[…]

### 2020-08-09 make-assessing-global-fisheries.R
write FAO fits to memory

@@ -25,13 +25,15 @@ library(rstan)
 library(rstanarm)
 library(sf)
 library(portedcmsy)
+
 Sys.unsetenv("PKG_CXXFLAGS")
+
+sraplus::get_tmb_model()
 
 options(dplyr.summarise.inform = FALSE)
 
 rstan::rstan_options(auto_write = TRUE)
 
-sraplus::get_tmb_model()
 # options -----------------------------------------------------------------
 
 min_years_catch <- 20
@@ -784,7 +786,6 @@ comp_stocks <- total_stocks %>%
 total_stocks <- total_stocks %>% 
   left_join(comp_stocks, by = "scientific_name") %>% 
   mutate(resilience = ifelse(is.na(resilience), "Medium",resilience))
-
 if (run_continent_examples == TRUE) {
   set.seed(24)
   
@@ -1586,7 +1587,6 @@ comp_stocks <- fao_status %>%
 fao_status <- fao_status %>% 
   left_join(comp_stocks, by = "scientific_name") %>% 
   mutate(resilience = ifelse(is.na(resilience), "Medium",resilience))
-
 if (run_sofia_comparison == TRUE) {
   # future::plan(future::multiprocess, workers = 4)
   
@@ -1602,8 +1602,9 @@ if (run_sofia_comparison == TRUE) {
     # sample_n(3) %>%
     # slice(10) %>% 
     mutate(
-      fits = future_map(
+      fits = future_map2(
         data,
+        stockid,
         safely(fit_fao),
         support_data = support_data,
         default_initial_state = NA,
@@ -1611,6 +1612,8 @@ if (run_sofia_comparison == TRUE) {
         min_effort_year = 1975,
         engine = "stan",
         cores = 2,
+        write_results = TRUE, 
+        results_path = results_path,
         .progress = TRUE,
         .options = future_options(
           globals = support_data,
@@ -1626,6 +1629,8 @@ if (run_sofia_comparison == TRUE) {
 } else {
   fao_status_fits <-
     read_rds(path = file.path(results_path, "fao_status_fits.rds"))
+  
+  
   
   
 }

### 2020-08-09 make-assessing-global-fisheries.R
mergin from abalone

@@ -44,7 +44,7 @@ draws <- 3000
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
-n_cores <- 3
+n_cores <- 2
 # number of cores for parallel processing
 
 # options(mc.cores = 1)
@@ -69,11 +69,11 @@ run_continent_examples <- FALSE
 
 run_ei_example <- FALSE
 
-run_sofia_comparison <- FALSE
+run_sofia_comparison <- TRUE
 
 run_ram_tests <- FALSE
 
-run_ram_comparison <- FALSE
+run_ram_comparison <- TRUE
 
 knit_paper <- TRUE
 
@@ -123,7 +123,7 @@ if (dir.exists(here::here("results", results_name)) == FALSE) {
 }
 
 
-prepare_sofia_data(lookup_fmi_names = FALSE)
+prepare_sofia_data(lookup_fmi_names = TRUE)
 
 
 has_total_biomass = ram_data %>%
@@ -1599,7 +1599,7 @@ if (run_sofia_comparison == TRUE) {
     group_by(stockid) %>%
     nest() %>%
     ungroup() %>%
-    # sample_n(3) %>%
+    sample_n(10) %>%
     # slice(10) %>% 
     mutate(
       fits = future_map(

### 2020-08-25 documents/ovando-etal-assessing-global-fisheries.Rmd
passing back to squid

@@
   Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals.
-Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results and therefore, high uncertainties in our understanding of global fisheries.
+While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor.
+Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results.
 This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources.
 How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets?
-We developed a new flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status.
+We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status.
 We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data.
 These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection.
-We show that our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
+Our results provide further evidence that catch data by themselves cannot provide reliable estimates of fishery health.
+Our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
 always_allow_html: true
@@
 library(here)
+library(sraplus)
 extrafont::loadfonts()
@@
 
+# rishi notes
+
+* more assertive on what the data needs really are
+  * where to apply globally 
+* For places with less coverage, what would we actually recommend there 
+* what do we know about catchabiliy
+- e.g. 51, fleets became industrial 
+  - hammer down on that's the issue with the effort 
+  - how do we translate effort into f bey area
+  - here's
+
 Target journal: Nature Sustainability
@@
 The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking this important metric.
-As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fishery, 6.2% are underfished ,and 34.2% are overfished [@fao2020].
+As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished ,and 34.2% are overfished [@fao2020].
 While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment.
@@
 While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-The SOFIA report bases its estimates of included unassessed stocks mostly on expert elicitation.
+The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion.
 While local experts often are well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable.
@@
 
-Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance such as catch-per-unit-effort (CPUE).
+Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE).
 While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect.
@@
 
-However, @free2020 demonstrated that these catch-only models can often produce both imprecise and biased estimates of B/B~MSY~.
+However, Free et al.
+2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status (B/B~MSY~).
 These issues become apparent when we consider some of the macro-level predictions made by these models.
+The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020].
+A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.
+
+xx spend time on this tomorrow xx
 Examining stock assessment data from RAM from the NE Pacific, the NE Atlantic and the Mediterranean, the NE Pacific is in the best state (mean RAM b/b~MSY~ = 1.8), NE Atlantic (mean RAM b/b~MSY~ =1.3) and the Mediterranean the worst (mean RAM b/b~MSY~ =0.5).
@@
 The Costello et al.
-20162 model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and SE Asia seem to be overoptimistic, and the NE Pacific should be better by comparison.
+2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and SE Asia seem to be overoptimistic, and the NE Pacific should be better by comparison.
 The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific.
@@
 
+xx move some of this to intro and chop downXX
 We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
@@
 
-```{}
-
-```
+XX add stock assessment estimates for both regions? XX Maybe just do matching stocks?
 
@@
 
+ xx dig into this, which stocks are in that region? XX
+
+```{r}
+a  = ram_comp_data %>% 
+  filter(fao_area_code == 61)
+
+unique(a$common_name)
+```
+
 ```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent B/B~MSY~ by FAO statistical area from different data sources.
@@
 
-Global-level assessment are critical for guiding management agendas for the world's oceans.
+<!-- Global-level assessment are critical for guiding management agendas for the world's oceans.
 Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world.
@@
 While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.
+-->
+
+Global-level assessment are critical for guiding management agendas for the world's oceans.
+Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world.
+While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone, at the global level no combination of readily available data was able to substantially and reliably clarify our picture of global fisheries status.
 
@@
 It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
-We have stated that addition of the globally available datasets compiled here did not substantially improve estimates of stock status, but in some instances using these data did provide some improvement over use of catch-only style methods.
+We have stated that addition of the globally available datasets compiled here did not substantially improve estimates of stock status, but in some instances using these data did provide some improvement over use of catch-only style methods; the difficulty comes in attempting to apply data types uniformly across the globe.
 However, none of the data sources used here came close to the benchmark performance provided by our fits to the RAM data, either in overall performance or in capturing the geographic distribution of stock status.
-While it is unreasonable to expect any global-scale dataset to be able to perform as well as data pulled from RAM assessments themselves, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general trends in stock status in time and space.
+While it is unreasonable to expect any global-scale data to be able to perform as well as data pulled from RAM assessments themselves, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general trends in stock status in time and space.
 That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the data themselves.
@@
 In terms of data however, information on fishing mortality rates (F/F~MSY~) in the most recent year was almost as useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot).
-Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance.
+Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance (xx uep, since in some cases the index was flatish but the stock status was low, so a model fit just to the trend in the abundance gets the trend right but the absolute value wrong.
+having F/Fmsy data helps anchor that).
 This is because we varied both the length and quality of the index of abundance, and on average an even noisy estimate of recent F/F~MSY~ may be more valuable than a short and noisy abundance index.
@@
 As a result, a growing number of methods have focused on using length-composition data to provide estimates of stock status, including measures of fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
-These methods have a distinct advantage for many fisheries in that they can in theory function with as little as one year of data.
+These methods have an advantage for many fisheries in that they can in theory function with as little as one year of data.
 As a result, many fisheries around the world are enacting programs to collect and utilize length composition data for fisheries assessment.
+While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries this may provide an overlooked source of fisheries data at a global scale.
 
@@
 All analysis were conducted in the R programming language [@rcoreteam2019].
-Unless otherwise stated all model fitting was conducted using stan [@carpenter2017], implemented through template model builder [@kristensen2016] by the tmbstan package [@monnahan2018].
+Model fitting was conducted using Rcpp [eddelbuettel2011] and stan [@carpenter2017], implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018].
 The `sraplus` package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
 
-## `sraplus`
+## Model Overview
+
+The core of `sraplus` is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018.
+While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
+The purpose of `sraplus` is not to make substantial improvements in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status.
+Detailed examples of how to operate the code of the model can be found [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus).
+Here we describe the overall methods used to produce the results in this paper and highlight key details.
+
+`sraplus` can be used in two general ways: as a mechanism for revealing implicit priors on fishery status through a sample-importance-resampling algorithm, and as a means for fitting indices of abundance informed by priors.
+To illustrate we will walk through a series of ways that `sraplus` could be applied to a case study fishery.
+
+Suppose that the only thing we observe from a fishery is a catch history.
+Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past (i.e.
+that catches have never been greater than biomass).
+We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
+Beyond that though, baring additional information we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0 and less than or equal to carrying capacity *k*.
+
+What hapens when we impose informative priors on the population parameters?
+For example at a bare minimum, we know that carrying capacity must be less than infinity, and we can certainly make inferences based on knowledge from other better studied fisheries of similar species on what a potential upper bound on carrying capacity might be.
+Once we have generated informative priors on our population parameters, we have implicitly provided priors on the biomass and therefore status of the fishery over its entire history: if our priors on carrying capacity only extend a small fraction beyond the minimum value needed to ensure the existence of the population, our implicit prior is that the population is severely depleted.
+If our prior is that carrying capacity is orders of magnitude greater than the observed catches, then our implicit prior is that the population is relatively unexploited.
+
+For our case study fishery, we can use `sraplus` as a filter to say, conditional on our priors, the catch history, and the population model, what is our implied prior on stock status of the fishery over time (see Section XX and Table.\@ref(tab:prior-tab) for descriptions of model parameters).
+While user defined priors on population parameters can be supplied, `sraplus` has a number of built in mechanisms for setting parameters based on the taxonomy of the species in question and the magnitude of the catch history.
+This form of analysis is often extended though to include an explicit prior on stock status in the most recent year [@walters2006].
+
+For our case study, expert opinion tells us that the stock is likely between XX and XX% of carrying capacity in the most recent year.
+While we can use `sraplus` to include this addition information, doing so introduces a nontrivial statistical problem: we have now placed two priors on the same outcome (current stock status), one explicit and one implicit in our growth parameter priors.
+The net effect of this is that once these priors are run through `sraplus`, our post-model-pre-data "estimate" of stock status will appear to have "updated", creating an illusion that we have learned something about stock status.
+Remember though that we have absolutely no data on which to update our prior on stock status beyond the minimum floor needed to ensure that the population exists.
+
+`sraplus` implements an approximate solution to this problem.
+Rather than using combinations of implied and explicit priors, `sraplus` asks, given an explicit prior on recent stock status, what prior distributions of growth parameters produce our explicit prior on recent stock status, given the catch history and model assumptions.
+This helps resolve some of the statistical issues related to placing multiple priors on the same parameter, and makes clear where any information on recent stock status is coming from.
+To return to our case study then, uncorrected, combining our explicit priors on recent stock status with our population parameter priors reveals an implicit prior on recent stock status that spans XX to XX.
+Using our correction, the implicit prior on recent stock status more or less matches our explicit prior (Fig.XX).
+
+While `sraplus` lets users supply their best estimate of recent stock status for a given fishery, the model also allows users to supply data on the swept area ratio (SAR , @amoroso2018) and/or fisheries management index (FMI, @melnychuk2017) scores for a given fishery.
+SAR is a measure of the trawling intensity in a given region, and as such for species susceptible to trawl gears can be an effective index of fishing pressure.
+FMI reflects expert opinion on the research, management, enforcement, and socioeconomic capacity of a given fishery, which together have significant associations with stock status.
+`sraplus` uses relationships between these SAR and FMI scores and observed F/F~MSY~  and B/B~MSY~ values to generate a posterior predictive distribution of F/F~MSY~ or B/B~MSY~ for the fishery in question, which serves as the priors for `sraplus`.
+
+To continue our case study example, suppose that we no longer believe we have good data on current stock status, but do have a reasonable estimate of swept area ratio.
+`sraplus` now uses the posterior predictive F/F~MSY~ generated by the SAR value as the prior for recent F/F~MSY~ for the fishery in question.
+This process allows users to translate something that they can measure, SAR, into something that they perhaps cannot (F/F~MSY~).
+The model then determines the growth parameter priors that conditional on the catch history and model structure produce the SAR generated F/F~MSY~ prior
+
+Up to this point we still have not fit the model to any data; sraplus is simply helping us visualize the implications of our priors in terms of stock status over time and relevant reference points such as MSY.
+In order to actually learn something new about the stock in question, we need to update our priors with data.
+`sraplus` allows for two methods for accomplishing this: the model can be fit to an observed index of abundance (for example a fishery independent survey), or fit to a catch-per-unit effort index calculated from the observed catch and an index of fishing effort.
+In the latter case, the model must make an assumption about the rate of efficiency increase of a unit of fishing effort over time.
+When data are available the model can be provided priors on recent fishery status derived from expert opinion or FMI/SAR values as well.
+
+pull out example from SAR fits
+```{r}
+sar <- sraplus::sar %>% 
+  filter(stockid %in% unique(ram_data$stockid))
+
+ex <- ram_data %>% 
+  filter(stockid == "PLAICIS")
+
+ex %>% 
+  ggplot(aes(year, b_v_bmsy)) + 
+  geom_point()
+
+# fit using catch heuristic
+
+# fit using SAR
+
+# fit using effort index
+
+```
+
+## Population Model
 
@@
 
-```{r}
+```{r prior-tab}
 cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
@@
   row.names = F,
-  caption = "Estimated parameters, abbreviations, and default priors"
+  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
+) %>%
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+```
+
+The population growth equation is
+
+\begin{equation}
+  f(x)=\begin{cases}
+      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
+, & \text{if $B_t>0.25 \times K$}.\\
+     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
+  \end{cases}
+  (\#eq:hockey)
+\end{equation}
+
+Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of Bmsy/K to shift (when *m* is 2 the Bmsy / K = 0.5 and the model acts as a Schaefer model, lower values of m shift the production function left, higher values right), *c* is catch, and *p* is process error.
+The Pella-Tomlinson model does have an issue where growth rates can become unrealistically large when the population reaches low sizes.
+We deal with this problem by following the methods described in @winker2018.
+to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
+Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see @thorson2018).
+
+<!-- When the SIR method is used we use the supplied catch history.
+However, for the maximum likelihood approach, convergence is greatly facilitated by estimating catches ($\hat{c_t}$) rather than using the observed catches directly.
+$\hat{c_t}$ is given as -->
+
+<!-- \begin{equation} -->
+<!--   \hat{c_t} = u_{t}B_{t} -->
+<!--   (\#eq:chat) -->
+<!-- \end{equation} -->
+
+<!-- where $u_t$ is estimated through -->
+
+<!-- \begin{equation} -->
+<!--   u_{t} = \frac{1}{1 + e^{-u^*_t}} -->
+<!--   (\#eq:invf) -->
+<!-- \end{equation} -->
+
+<!-- Equation \@ref(eq:invf) ensures that $u_t$ is between 0 and 1, while the model estimates $u^*_t$ to improve performance.  -->
+
+<!-- $\hat{c_t}$ is then fit to the observed catches through the likelihood -->
+
+<!-- \begin{equation} -->
+<!--   log\left(c_t\right) \sim normal\left(log\left(\hat{c_t}\right),0.05\right) -->
+<!--   (\#eq:catch) -->
+<!-- \end{equation} -->
+
+<!-- Where we assume an observation coefficient of variation of 5\%.
+We estimate catch rather than directly using the observed catch in order to improve model performance.
+As priors on depletion, or indices of abundance, pull the estimated population closer to zero, there is a greater and greater chance that the observed catch history will result in a collapse of the population (catch in time *t* being greater than biomass in time *t*); It is much harder to find parameters that produce a final depletion of 1\% of carrying capacity while not crashing the population prior to then given a catch history, than it is to produce a final depletion of 99\% of carrying capacity.
+We can impose restrictions and penalties on population crashes, but this distorts the gradients of the model, causing the model to often become stuck once a few crashes have occurred.
+Since $u_t$ is constrained between 0 and 1, $\hat{c_t}$ will always be "viable", and by then fitting to the observed catches, the model has an easier time finding parameters that produce a near match to the observed catches while not crashing the population.
+As of now none of the data used in this exercise have missing years of catch data, but the model can handle those cases in the future due to the random walk priors imposed on $u_t$ detailed below.
+-->
+
+<!-- In order to help separate out observation error in the catch and process error in the population model, when catch is being estimated we assume that fishing mortality $u_t$ follows a random walk on average, assigning each $u_t$ a prior of -->
+
+<!-- \begin{equation} -->
+<!--   log(u_t) \sim normal(log(u_{t - 1}),0.1)) -->
+<!--   (\#eq:rwalk) -->
+<!-- \end{equation} -->
+
+We  allow for process error (in the manner of the stochastic stock reduction analysis error suggested by @walters2006).
+This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
+Process error *p* is assumed to be log-normally distributed, such that
+
+\begin{equation}
+  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
+  (\#eq:procerror)
+\end{equation}
+
+All of our estimates are Bayesian in nature.
+We can break the use of our model into two distinct categories: with data and without.
+By "data", we refer to measurements which are confronted with model estimates within a likelihood function.
+In our context, these include fishery-independent survey data, or a CPUE index.
+When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
+Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population, given the constraints of the model.
+
+We use a Bayesian methods as it suits the philosophy of our analysis: incorporating prior knowledge about stock status drawn from expert opinion or other similar stocks, and where possible confronting those priors with data.
+However, this process does introduce a problem, in that it a) certain parts of the prior probability distribution have zero support in the data (e.g.
+carrying capacity's so small that they crash the population) and b) they create priors on both the *input* (e.g.
+growth rates and carrying capacity) and *outputs* (recent stock status), effectively putting two prior distributions for the same outcome, a problem termed Borel's Paradox XX punt citation.
+This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
+However, particularly for the data-less version of our model, Borel's Paradox poses a particular problem.
+
+Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
+Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of r and k together with the model produces a deterministic stock status in each time step.
+Suppose then we wish to also provide a prior on stock status in the most recent year, for example to reflect local expert opinion based on sound reasoning.
+We have now placed two priors on stock status in the most recent year, one implicit and one explicit, violating Borel's Paradox.
+The net effect of this is that once these priors are run through the model, through for example a sample-importance-resampling (SIR) algorithm as generally applied in stock-reduction analysis, our post-model-pre-data "estimate" of stock status will appear to have "updated".
+This creates the illusion that we have learned something about stock status, when in fact we have simply used our model to reveal our true implied prior on stock status.
+This is technically incorrect, since we do not have a likelihood in the model at this point, and as such per Bayes Theorem have not updated our prior.
+Of more practical importance, this "implied prior" will always reflect a more optimistic stock status then our original prior on stock status, due to the constraints of the population model.
+
+To resolve this, @puntXX (either cite that "on assessment of the bering-chukchi-beufort sea stock of bowhead whales paper that I can't find a date for, or see if he has a new paper on thisXX) suggests a "backwards" fitting approach, in which the model estimates growth parameters and a terminal depletion, and given the model backs out a value of *k* that satisfies the priors (conditional on an assumption initial stock status).
+We do not adopt this approach here due to computational issues.
+Adopting the "backwards" approach requires numerically searching for the plausible *k* during each iteration of the model, which is time consuming.
+In addition, the backwards approach is not without its own complications, for example the requirement of a tolerance around the value of terminal depletion produced by a given *k* and the actual value selected by the model in that time step.
+This discrepancy can occur when a given terminal stock status is not technically possible given the model; to give an extreme example under this model no value of *k* can produce a terminal stock status of 0 if catches are 0 for the entire history of the fishery (assuming *k* > 0).
+
+We present a novel approximation to this problem here.
+Our solution amounts to a two-step SIR algorithm.
+We first run the standard SIR algorithm as described above.
+We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
+
+$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
+
+We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
+
+$$p(n_i) = \frac{p(bin_i)}{N_i}$$
+
+And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
+The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
+In effect, this answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
+
+<!-- @raftery1993 -->
+
+## FMI and SAR Based Priors
+
+All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [xx citation].
+These models are intended as a reasonably robust means of translating available data into usable prior for our model.
+Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with fishery status indicators of interest.
+Rather, each regression was tested to ensure that it is unlike that, given the same data, and alternative model would be unlikely to perform much better than those presented here XX this might be too much defenseXX
+
+For each statistical prior, we fit three different models, relating the data of interesting, e.g. swept area ratio (SAR), to each of *B/B~MSY~*, *F/F~MSY~*, and *F*
+
+### Swept Area Ratio Priors
+
+@amoroso2018 provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
+This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
+As illustrated in @amoroso2018, there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
+Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
+if all trawl-able areas are trawled twice a year then the SAR will be 2.
+Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
+
+The final selected model relating SAR to to stock status metrics was
+
+$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
+
+## Fisheries Management Index Priors
+
+The Fisheries Management Index (FMI), as presented in @melnychuk2017, utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
+These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
+The higher the score, the better the expert judges that a given metric is met in that fishery.
+Importantly, FMI surveys can be filled out in the absence of stock assessments.
+This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to @osio2015 and @cope2015).
+
+The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
+
+$$log(value_i) \sim N(s(research_i) + s(management_i) +
+s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
+
+## Catch-Only Priors
+
+Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
+While these catch-only methods have been shown to have series shortcomings [@free2020], as a point of reference we use two forms of catch-only priors: a statistical and CMSY, as presented implemented by @palomares2020.
+
+For the CMSY based methods, we used a ported version of the CMSY model available [here](https://github.com/DanOvando/portedcmsy).
+For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [boettiger2012].
+Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
+
+For the statistical approach, we utilize data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
+To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
+Emergent clusters show for example one build around "one way trip" style catch histories, others with a boom and bust, etc.
+
+We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history.
+This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term).
+For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status
+
+$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma) $$
+
+For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data
+
+$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
+where *fyear* is the year of the fishery, starting from 0.
+
+## Effort Data
+
+While the FMI and SAR data show some potential for estimating U/U~MSY~ and B/B~MSY~ values, they are clearly not an ideal measure of stock status.
+An alternative is to actually develop indices of abundance that we can fit models to.
+@rousseau2019 provides a useful starting place for this exercise, presenting estimates of fishing effort throughout the globe from 1950 to 2015.
+Merging these data with capture statistics can provide a crude CPUE index that can then be used to fit a surplus production model.
+Developing an index of abundance in this manner is clearly a very rough analysis only intended to illustrate broad concepts.
+Any analysis of CPUE data can be severely affected by changes in technology, skill, hyperdepletion, hyperstability, changes in the spatial footprint of the fishery, and selectivity [@maunder2004a; @walters2003].
+In the case of this analysis, a kilowatt day of fishing with 1950s technology clearly represents a different amount of fishing power than a kilowatt day of fishing in 2012.
+To account for this somewhat, @rousseau2019 include a 2.6% annual "technological creep" factor, which we incorporate in our analysis.
+
+@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6\% technology rate).
+This assumes that every unit increase in effort is linearly greater than the unit of effort before it.
+When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
+This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
+To accommodate this, we generate a catch per effective harvest rate index of abundance, as
+
+$$I_t = \frac{catch_t}{(1 - e^{-q_tE_t})}$$
+
+Where $q_t$ can has a technology rate component $\tau$
+
+$$q_t = q_{t-1} \times (1  + \tau)$$
+
+## Value of Information Calculations
+
+Details of value of information calculates (fitting to messed up versions of RAM) XX need to add this in XX
+
+# Supplemental Information
+
+## Population Model
+
+The core of our model is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018.
+While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
+
+```{r si-prior-tab}
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
+  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
+  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
+  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
+  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
+  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))
+
+knitr::kable(
+  cd,
+  "latex",
+  align = "l",
+  booktabs = TRUE,
+  escape = F,
+  row.names = F,
+  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
 ) %>%
@@
 
-## Priors and Data
-
-We briefly describe the construction of priors and nature of available data.
-All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [xx citation].
-These models are intended as a reasonably robust means of translating available data into usable prior for our model.
-Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with fishery status indicators of interest.
-Rather, each regression was tested to ensure that it is unlike that, given the same data, and alternative model would be unlikely to perform much better than those presented here XX this might be too much defenseXX
-
-For each statistical prior, we fit three different models, relating the data of interesting, e.g. swept area ratio (SAR), to each of *B/B~MSY~*, *F/F~MSY~*, and *F*
-
-## Swept Area Ratio Priors
-
-@amoroso2018 provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
-This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
-As illustrated in @amoroso2018, there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
-Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
-if all trawl-able areas are trawled twice a year then the SAR will be 2.
-Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
-
-The final selected model relating SAR to to stock status metrics was
-
-$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
-
-```{r sar-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of swept area ratio (SAR) on F/F~MSY~"}
-fit <- sraplus::sar_models$fit[sraplus::sar_models$metric == "mean_uumsy"][[1]]
-
-fit_data <- fit$data
-
-pp <- posterior_predict(fit, draws = 1000)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
-  scale_x_continuous(name = "Observed Log F/Fmsy") + 
-  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-
-```
-
-## Fisheries Management Index Priors
-
-The Fisheries Management Index (FMI), as presented in @melnychuk2017, utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
-These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
-The higher the score, the better the expert judges that a given metric is met in that fishery.
-Importantly, FMI surveys can be filled out in the absence of stock assessments.
-This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to @osio2015 and @cope2015).
-
-The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
-
-$$log(value_i) \sim N(s(research_i) + s(management_i) +
-s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
-
-```{r fmi-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of fisheries management index (FMI) on F/F~MSY~"}
-
-fit <- sraplus::fmi_models$fit[sraplus::fmi_models$metric == "mean_uumsy"][[1]]
-
-fit_data <- fit$data
-
-pp <- posterior_predict(fit, draws = 1000)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
-  scale_x_continuous(name = "Observed Log F/Fmsy") + 
-  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-```
-
-## Catch-Only Priors
-
-Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
-While these catch-only methods have been shown to have series shortcomings [@free2020], as a point of reference we use two forms of catch-only priors: a statistical and a heuristic approach.
-
-For the heuristic approach, we borrow from the original methods of @martell2013 as presented in @andersno2017, and define a heuristic on initial and terminal depletion where
-
-```r
-        if (first(catch) / max(catch) < 0.2){
-
-        initial_depletion = 0.7
-
-        } else {
-
-        initial_depletion = 0.4
-
-        }
-
-        if (last(catch) / max(catch) > 0.5){
-
-        terminal_depletion <- 0.6
-
-        } else {
-
-        terminal_depletion <- 0.2
-
-        }
-
-```
-
-For the statistical approach, we utilize data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
-To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
-Emergent clusters show for example one build around "one way trip" style catch histories, others with a boom and bust, etc.
-
-We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history.
-This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term).
-For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status
-
-$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma) $$
-
-For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data
-
-$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
-where *fyear* is the year of the fishery, starting from 0.
-
-```{r catch-reg,fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) B/B~MSY~ for regression of catch on B/B~MSY~"}
-
-fit <- sraplus::catch_b_model
-
-fit_data <- fit$data
-
-pp <- posterior_predict(fit, draws = 1000)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_scatter_avg(y = fit_data$log_value, yrep = pp) + 
-  coord_flip() +
-  scale_y_continuous(name = "Observed Log B/Bmsy") + 
-  scale_x_continuous(name = "Posterior Predictive B/Bmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-```
-
-## Effort Data
-
-While the FMI and SAR data show some potential for estimating U/U~MSY~ and B/B~MSY~ values, they are clearly not an ideal measure of stock status.
-An alternative is to actually develop indices of abundance that we can fit models to.
-@rousseau2019 provides a useful starting place for this exercise, presenting estimates of fishing effort throughout the globe from 1950 to 2015.
-Merging these data with capture statistics can provide a crude CPUE index that can then be used to fit a surplus production model.
-Developing an index of abundance in this manner is clearly a very rough analysis only intended to illustrate broad concepts.
-Any analysis of CPUE data can be severely affected by changes in technology, skill, hyperdepletion, hyperstability, changes in the spatial footprint of the fishery, and selectivity [@maunder2004a; @walters2003].
-In the case of this analysis, a kilowatt day of fishing with 1950s technology clearly represents a different amount of fishing power than a kilowatt day of fishing in 2012.
-To account for this somewhat, @rousseau2019 include a 2.6% annual "technological creep" factor, which we incorporate in our analysis.
-
-@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6\% technology rate).
-This assumes that every unit increase in effort is linearly greater than the unit of effort before it.
-When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
-This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
-To accommodate this, we generate a catch per effective harvest rate index of abundance, as
-
-$$I_t = \frac{catch_t}{(1 - e^{-q_tE_t})}$$
-
-Where $q_t$ can has a technology rate component $\tau$
-
-$$q_t = q_{t-1} \times (1  + \tau)$$
-
-## Value of Information Calculations
-
-Details of value of information calculates (fitting to messed up versions of RAM) XX need to add this in XX
-
-# Supplemental Figures
-
 ```{r fao-acc-map, fig.cap = "Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
@@
 
+## SAR
+
+## FMI
+
+```{r fmi-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of fisheries management index (FMI) on F/F~MSY~"}
+
+fit <- sraplus::fmi_models$fit[sraplus::fmi_models$metric == "mean_uumsy"][[1]]
+
+fit_data <- fit$data
+
+pp <- posterior_predict(fit, draws = 1000)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
+  scale_x_continuous(name = "Observed Log F/Fmsy") + 
+  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+```
+
+```{r sar-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of swept area ratio (SAR) on F/F~MSY~"}
+fit <- sraplus::sar_models$fit[sraplus::sar_models$metric == "mean_uumsy"][[1]]
+
+fit_data <- fit$data
+
+pp <- posterior_predict(fit, draws = 1000)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
+  scale_x_continuous(name = "Observed Log F/Fmsy") + 
+  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+
+```
+
+## Catch
+
+```{r catch-reg,fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) B/B~MSY~ for regression of catch on B/B~MSY~"}
+
+fit <- sraplus::catch_b_model
+
+fit_data <- fit$data %>% 
+  group_by(stockid) %>% 
+  filter(year == max(year))
+
+pp <- posterior_predict(fit, draws = 1000, newdata = fit_data)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_scatter_avg(y = fit_data$log_value, yrep = pp) + 
+  coord_flip() +
+  scale_y_continuous(name = "Observed Log B/Bmsy") + 
+  scale_x_continuous(name = "Posterior Predictive B/Bmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+```
+
 # References

### 2020-08-25 functions/fit-fao.R

@@ -18,7 +18,7 @@ fit_fao <-
            draws = 2e6,
            refresh = 0,
            thin_draws = TRUE, 
-           cmsy_cores = 3,
+           cmsy_cores = 1,
            results_path,
            write_results) {
     #

### 2020-08-25 make-assessing-global-fisheries.R

@@ -71,15 +71,15 @@ run_continent_examples <- FALSE
 
 run_ei_example <- FALSE
 
-run_sofia_comparison <- TRUE
+run_sofia_comparison <- FALSE
 
 run_ram_tests <- FALSE
 
-run_ram_comparison <- TRUE
+run_ram_comparison <- FALSE
 
 knit_paper <- TRUE
 
-warning("Running full analysis takes upwards of 12 hours on 4 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 4 cores is not recommended.")
+warning("Running full analysis takes upwards of 24 hours on 2 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 2 cores is not recommended (memory routinley runs out on a 16 core 36GB machine when cores > 2)")
 
 engine <-  "stan"
 
@@ -1611,7 +1611,8 @@ if (run_sofia_comparison == TRUE) {
         default_initial_state_cv = NA,
         min_effort_year = 1975,
         engine = "stan",
-        cores = 2,
+        cores = 1,# run in to lots of memory problems if greater than 1
+        cmsy_cores = 1, # run in to lots of memory problems if greater than 1
         write_results = FALSE, 
         results_path = results_path,
         .progress = TRUE,
@@ -1761,13 +1762,14 @@ fao_areas <- fao_areas %>%
   mutate(geometry = map(data, st_union)) %>%
   select(-data)
 
+
 fao_areas = fao_areas %>%
   unnest(cols = geometry) %>%
   ungroup() %>%
   sf::st_as_sf() %>%
-  sf::st_simplify() %>%
-  mutate(fao_area_code = as.numeric(f_area))
-
+  # sf::st_simplify() %>%
+  mutate(fao_area_code = as.numeric(f_area)) #%>% 
+  # st_transform(crs = "+proj=moll")
 
 fao_area_accuracy <- fao_areas %>%
   left_join(fao_sraplus_acc %>% mutate(f_area = as.character(fao_area_code)), by = "f_area") %>%
@@ -1783,7 +1785,9 @@ world_map <-
     type = 'land',
     category = 'physical',
     returnclass = "sf"
-  )
+  ) #%>% 
+  # st_transform(crs = "+proj=moll")
+  
 
 fao_acc_map_plot <- fao_area_accuracy %>%
[…]

### 2020-08-25 functions/fit-fao.R
Cmsy (#5)
* results updated to reflect new CMSY method

@@ -1,5 +1,6 @@
 fit_fao <-
   function(data,
+           id,
            support_data,
            model = "sraplus_tmb",
            engine = "tmb",
@@ -16,9 +17,17 @@ fit_fao <-
            reg_cv = NA,
            draws = 2e6,
            refresh = 0,
-           thin_draws = TRUE) {
+           thin_draws = TRUE, 
+           cmsy_cores = 1,
+           results_path,
+           write_results) {
     #
     #     data <- test$data[[1]]
+    #     
+    
+    if (write_results & !dir.exists(file.path(results_path, "faofits"))){
+      dir.create((file.path(results_path, "faofits")))
+    }
     
     ogengine <- engine
     
@@ -27,6 +36,44 @@ fit_fao <-
     catches <- data$catch
     
     years <- data$year
+    
+    
+    # fit cmsy
+    # 
+    sfc <- purrr::safely(portedcmsy::funct_cmsy)
+    
+    ported_cmsy <- sfc(
+      catches = catches[years >= 1950],
+      catch_years = years[years >= 1950],
+      stock = scientific_name,
+      common_name = NA,
+      scientific_name = scientific_name,
+      res = unique(data$resilience),
+      start.yr = years[years >= 1950][1],
+      end.yr = dplyr::last(years[years >= 1950]),
+      cores = cmsy_cores
+    )
+    if (is.null(ported_cmsy$error)) {
+      cmsy_results <- ported_cmsy$result %>%
+        mutate(c_div_msy  = ct / msy) %>%
+        select(year, bt, b_bmsy, f_fmsy, c_div_msy) %>%
+        rename(depletion = bt,
+               b_div_bmsy = b_bmsy,
+               u_div_umsy = f_fmsy) %>%
+        pivot_longer(-year, names_to = "variable", values_to = "mean") %>%
+        mutate(
+          sd = NA,
+          lower = NA,
+          upper = NA,
+          data = "cmsy"
+        )
+      
+      cmsy_worked <- TRUE
+    } else {
+      cmsy_worked <- FALSE
+    }
+    
+    # fit to u
     
     u_years <- seq_along(years)[!is.na(data$mean_u_umsy)]
     
@@ -595,18 +642,27 @@ fit_fao <-
         else{
           .
         }
+      }  %>% {
+        if (cmsy_worked) {
+          bind_rows(., cmsy_results)
+        } else {
+          .
+        }
       }
     
[…]

### 2020-08-25 functions/fit-ram.R

@@ -9,10 +9,12 @@ fit_ram <-
            q_slope_prior = 0.025,
            estimate_proc_error = TRUE,
            estimate_qslope = FALSE,
+           estimate_shape = FALSE,
            default_initial_state = NA,
            default_initial_state_cv = 0.2,
            chains = 1,
            cores = 1,
+           cmsy_cores = 1,
            reg_cv = NA,
            draws = 2e6,
            refresh = 0,
@@ -20,6 +22,9 @@ fit_ram <-
     #
     #     data <- test$data[[1]]
     # browser()
+    
+    sfs <- safely(fit_sraplus)
+    
     ogengine <- engine
     
     scientific_name <- unique(data$scientific_name)
@@ -27,6 +32,44 @@ fit_ram <-
     catches <- data$catch
     
     years <- data$year
+    
+    sfc <- purrr::safely(portedcmsy::funct_cmsy)
+    # browser()
+    ported_cmsy <- sfc(
+      catches = catches[years >= 1950],
+      catch_years = years[years >= 1950],
+      stock = scientific_name,
+      common_name = unique(data$common_name),
+      scientific_name = unique(data$scientific_name),
+      res = unique(data$resilience),
+      start.yr = years[years >= 1950][1],
+      end.yr = dplyr::last(years[years >= 1950]),
+      cores = cmsy_cores
+    )
+    write_rds(ported_cmsy, "wtf.rds")
+    
+    if (is.null(ported_cmsy$error)){
+      
+      cmsy_results <- ported_cmsy$result %>% 
+        mutate(c_div_msy  = ct / msy) %>% 
+        select(year, bt, b_bmsy, f_fmsy, c_div_msy) %>% 
+        rename(depletion = bt, b_div_bmsy = b_bmsy, u_div_umsy = f_fmsy) %>% 
+        pivot_longer(-year, names_to = "variable", values_to = "mean") %>% 
+        mutate(sd = NA, lower = NA, upper = NA, data = "cmsy")
+      
+      # tests <- cmsy_results %>% 
+      #   bind_rows(basic_fit_results)
+      # 
+      # tests %>% 
+      #   filter(variable == "c_div_msy") %>% 
+      #   select(year, mean, data) %>% 
+      #   ggplot(aes(year, mean, color = data)) + 
+      #   geom_line()
+      
+      cmsy_worked <- TRUE 
+    } else {
+      cmsy_worked <- FALSE
+    }
[…]

### 2020-08-25 make-assessing-global-fisheries.R

@@ -9,6 +9,7 @@ library(rstan)
 library(mvtnorm)
 library(FishLife)
 library(furrr)
+library(doParallel)
 library(future)
 library(viridis)
 library(sraplus)
@@ -23,13 +24,16 @@ library(tidybayes)
 library(rstan)
 library(rstanarm)
 library(sf)
+library(portedcmsy)
+
 Sys.unsetenv("PKG_CXXFLAGS")
+
+sraplus::get_tmb_model()
 
 options(dplyr.summarise.inform = FALSE)
 
 rstan::rstan_options(auto_write = TRUE)
 
-sraplus::get_tmb_model()
 # options -----------------------------------------------------------------
 
 min_years_catch <- 20
@@ -42,7 +46,7 @@ draws <- 3000
 
 min_draws <- 2000 # minimum number of unique SIR draws
 
-n_cores <- 3
+n_cores <- 2
 # number of cores for parallel processing
 
 # options(mc.cores = 1)
@@ -75,7 +79,7 @@ run_ram_comparison <- FALSE
 
 knit_paper <- TRUE
 
-warning("Running full analysis takes upwards of 12 hours on 4 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 4 cores is not recommended.")
+warning("Running full analysis takes upwards of 24 hours on 2 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 2 cores is not recommended (memory routinley runs out on a 16 core 36GB machine when cores > 2)")
 
 engine <-  "stan"
 
@@ -121,7 +125,7 @@ if (dir.exists(here::here("results", results_name)) == FALSE) {
 }
 
 
-prepare_sofia_data(lookup_fmi_names = FALSE)
+prepare_sofia_data(lookup_fmi_names = TRUE)
 
 
 has_total_biomass = ram_data %>%
@@ -762,14 +766,34 @@ total_stocks %>%
 
 # future::plan(future::multiprocess, workers = 4)
 
+
+comp_stocks <- total_stocks %>%
+  select(scientific_name) %>%
+  unique() %>%
+  mutate(fishbase_vuln = map_dbl(
+    scientific_name,
+    ~ rfishbase::species(.x, fields = "Vulnerability")$Vulnerability
+  )) %>%
+  mutate(
+    resilience = dplyr::case_when(
+      fishbase_vuln < 33 ~ "Low",
[…]

### 2020-08-27 documents/ovando-etal-assessing-global-fisheries.Rmd
testing TMB fix on windows

@@
 
-What hapens when we impose informative priors on the population parameters?
+What happens when we impose informative priors on the population parameters?
 For example at a bare minimum, we know that carrying capacity must be less than infinity, and we can certainly make inferences based on knowledge from other better studied fisheries of similar species on what a potential upper bound on carrying capacity might be.
@@
 While we can use `sraplus` to include this addition information, doing so introduces a nontrivial statistical problem: we have now placed two priors on the same outcome (current stock status), one explicit and one implicit in our growth parameter priors.
-The net effect of this is that once these priors are run through `sraplus`, our post-model-pre-data "estimate" of stock status will appear to have "updated", creating an illusion that we have learned something about stock status.
-Remember though that we have absolutely no data on which to update our prior on stock status beyond the minimum floor needed to ensure that the population exists.
+The net effect of this is that once these priors are run through `sraplus`, our post-model-pre-data "estimate" of stock status will appear to have been updated, creating an illusion that we have learned something about stock status.
+However, we have absolutely no data with which to update our prior on stock status beyond the requirement that the population exists.
 
 `sraplus` implements an approximate solution to this problem.
-Rather than using combinations of implied and explicit priors, `sraplus` asks, given an explicit prior on recent stock status, what prior distributions of growth parameters produce our explicit prior on recent stock status, given the catch history and model assumptions.
-This helps resolve some of the statistical issues related to placing multiple priors on the same parameter, and makes clear where any information on recent stock status is coming from.
+Rather than using combinations of implied and explicit priors, `sraplus` asks, given an explicit prior on recent stock status, what prior distributions of population parameters produce our explicit prior on recent stock status, given the other prior distributions, catch history and model assumptions.
+This helps resolve some of the statistical issues related to placing multiple priors on the same parameter (recent stock status), and makes clear where any information on recent stock status is coming from.
 To return to our case study then, uncorrected, combining our explicit priors on recent stock status with our population parameter priors reveals an implicit prior on recent stock status that spans XX to XX.
 Using our correction, the implicit prior on recent stock status more or less matches our explicit prior (Fig.XX).
+
+```{r}
+
+id = unique(ram_comp_data$stockid)
+
+ex <- ram_data %>% 
+  filter(stockid %in% id[[1]])
+
+ex %>% 
+  ggplot(aes(year, b_v_bmsy)) + 
+  geom_point()
+
+# fit using catch heuristic
+
+co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
+                                 catch = ex$catch, 
+                                 years = ex$year,
+                                 terminal_state = 0.25,
+                                 terminal_state_cv = 0.25,
+                                 b_ref_type = "k")
+
+plot_driors(co_driors)
+
+biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
+                                      estimate_shape = TRUE)
+
+co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
+                                      estimate_shape = TRUE)
+
+plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit)
+
+```
 
@@
 `sraplus` uses relationships between these SAR and FMI scores and observed F/F~MSY~  and B/B~MSY~ values to generate a posterior predictive distribution of F/F~MSY~ or B/B~MSY~ for the fishery in question, which serves as the priors for `sraplus`.
-
-To continue our case study example, suppose that we no longer believe we have good data on current stock status, but do have a reasonable estimate of swept area ratio.
-`sraplus` now uses the posterior predictive F/F~MSY~ generated by the SAR value as the prior for recent F/F~MSY~ for the fishery in question.
+This process lets us convert information on stock status present in other available data into a form that aligns with the model structure of `sraplus`.
+
+To continue our case study, suppose that we no longer believe we have good data on current stock status, but do have a reasonable estimate of swept area ratio.
+`sraplus` can now use the posterior predictive $\hat{F/F~MSY~}$ generated by the SAR value as the prior for recent F/F~MSY~ for the fishery in question.
 This process allows users to translate something that they can measure, SAR, into something that they perhaps cannot (F/F~MSY~).
-The model then determines the growth parameter priors that conditional on the catch history and model structure produce the SAR generated F/F~MSY~ prior
+The model then determines the growth parameter priors that conditional on the catch history and model structure produce the SAR generated $\hat{F/F~MSY~}$ prior distribution.
 
@@
 sar <- sraplus::sar %>% 
-  filter(stockid %in% unique(ram_data$stockid))
+  filter(stockid %in% unique(ram_data$stockid)) %>% 
+  arrange(desc(sar)) %>% 
+  select(stockid, everything()) %>% 
+  slice(10)
 
 ex <- ram_data %>% 
-  filter(stockid == "PLAICIS")
+  filter(stockid %in% unique(sar$stockid))
 
@@
 
+co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
+                                 catch = ex$catch, 
+                                 years = ex$year,
+                                 terminal_state = 0.25,
+                                 terminal_state_cv = 0.25,
+                                 b_ref_type = "k")
+
+plot_driors(co_driors)
+
+biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
+                                      estimate_shape = TRUE)
+
+co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
+                                      estimate_shape = TRUE)
+
+plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit)
+
 # fit using SAR
 
-# fit using effort index
+sar_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
+                                 catch = ex$catch, 
+                                 years = ex$year,
+                                 sar = sar$sar[sar$stockid == unique(ex$stockid)])
+
+plot_driors(sar_driors)
+
+sar_fit <- fit_sraplus(sar_driors, estimate_shape = TRUE)
+
+plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit, `sar fit` = sar_fit)
+
+# fit using biomass index
+
+index_driors <- sar_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
+                                 catch = ex$catch, 
+                                 years = ex$year,
+                                 sar = sar$sar[sar$stockid == unique(ex$stockid)],
+                                 index = ex$b_v_bmsy[!is.na(ex$b_v_bmsy)], 
+                                 index_years = ex$year[!is.na(ex$b_v_bmsy)])
+
+index_fit <- fit_sraplus(index_driors, engine = "stan", estimate_shape = TRUE)
+
+plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit, `sar fit` = sar_fit,
+             `index fit` = index_fit,
+             plotvars = "b_div_bmsy") + 
+  geom_point(data = ex, aes(year, b_v_bmsy))
 

### 2020-08-28 Untitled.Rmd
methods almost fixed

---
title: "test"
author: "Dan"
date: "8/27/2020"
output:
  bookdown::pdf_document2: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## R Markdown

Table.\@ref(tab:param-tab)

```{r param-tab}

cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
  rbind(c("Shape parameter", "m", "Drawn from @horson2012")) %>%
  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.05,0.5)$")) %>%
  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$")) %>% 
  rbind(c("Initial State", "$B0$", "$logn(0.025,0.2)$"))

knitr::kable(
  cd,
  "latex",
  align = "l",
  escape = F,
  row.names = F,
  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
) %>%
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

### 2020-08-28 documents/ovando-etal-assessing-global-fisheries.Rmd
methods almost fixed

@@
 output: 
+  bookdown::pdf_document2: default
   bookdown::word_document2:
     reference_docx: word-template.docx
-  bookdown::pdf_document2: default
 params:
@@
 
-## Model Overview
-
-The core of `sraplus` is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018.
+`sraplus` contains too many options to cover within this methods section.
+We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus.
+Below we describe the structure of the population model underpinning `sraplus`, the estimation model, and a brief example demonstrating the use of the model.
+
+## Population Model
+
+The core of `sraplus` is a Pella-Tomlinson [@pella1969] production model constructed in the manner of @winker2018.
 While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
-The purpose of `sraplus` is not to make substantial improvements in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status.
-Detailed examples of how to operate the code of the model can be found [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus).
-Here we describe the overall methods used to produce the results in this paper and highlight key details.
+The purpose of `sraplus` is not to make substantial improvements in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status
+
+The population growth equation is
+
+\begin{equation}
+  f(x)=\begin{cases}
+      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
+, & \text{if $B_t>0.25 \times K$}.\\
+     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
+  \end{cases}
+  (\#eq:hockey)
+\end{equation}
+
+Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of B~msy~/K to shift.
+When *m* is two Bmsy / K = 0.5 and the model acts as a Schaefer model, lower values of m shift the production function left, higher values right.
+*c* is catch, and *p* is process error.
+Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
+We deal with this problem by following the methods described in @winker2018.
+to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
+
+We  allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by @walters2006).
+This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
+Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see @thorson2018).
+
+Process error *p* is assumed to be log-normally distributed, such that
+
+\begin{equation}
+  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
+  (\#eq:procerror)
+\end{equation}
+
+## Estimation Model
+
+All of our estimates are Bayesian in nature.
+We can break the use of `sraplus` into two distinct categories: with data and without.
+By "data", we refer to measurements which are used to confront model estimates within a likelihood function.
+In our context, these include fishery-independent survey data, or a CPUE index.
+When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
+Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e.
+results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model.
+
+The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).
+*r* and *k* are the only two parameters that are always estimated.
+Estimation of every other parameter can be turned on or off.
+When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user.
+For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *k*, *m*, $\sigma_{obs}$, $\gamma$, and *B0*.
+*q* is also estimated when needed.
+
+```{r param-tab}
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
+  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
+  rbind(c("Shape parameter", "m", "Drawn from @horson2012")) %>%
+  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
+  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.05,0.5)$")) %>%
+  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$")) %>% 
+  rbind(c("Initial State", "$B0$", "$logn(0.025,0.2)$"))
+
+knitr::kable(
+  cd,
+  "latex",
+  align = "l",
+  escape = F,
+  row.names = F,
+  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
+) %>%
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+```
+
+`sraplus` can be run in two forms: either as a stock reduction analysis [SRA,@walters2006], or fit to an index of abundance (fishery dependent or independent).
+Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
+A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
+`sraplus` allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~.
+We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history.
+Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected.
+The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
+It is critical to note that a stock reduction analysis is only a mechanism for revealing implicit priors on fishery status and parameters through a sample-importance-resampling algorithm: the model cannot learn anything new about stock status in the absence of additional data to fit to, beyond the prior on stock status that is fed to it.
+The usefullness of this process then is in filling in the gap of parameters of interest (e.g.
+maximum sustainable yield, MSY), and providing historic values of stock status.
+Any model results based on data sources listed in Table.\@ref(tab:dat-description) that do not contain "cpue" or "RAM data" is estimated through stock reduction analysis.
+
+When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018].
+The model fits to a direct estimate of abundance (e.g.
+a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood is
+
+$$log(\pmb{A_t}) \sim normal(f(r,k,m,B0,\pmb{p},\pmb{C}) \times q, sigma_{obs{}}) $$
+where *f* is the Pella-Tomlinson produciton model (Equation 1.XX).
+When an effort index is available, `sraplus` constructs an index of abundance based on the catch and effort data.
+@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6\% technology rate).
+This assumes that every unit increase in effort is linearly greater than the unit of effort before it.
+When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
+This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
+To accommodate this, we generate a catch per effective harvest rate index of abundance, as
+
+$$cpue_t = \frac{catch_t}{(1 - e^{-q_tE_t})}$$
+
+Where $q_t$ can has a technology rate component $\tau$
+
+$$q_t = q_{t-1} \times (1  + \tau)$$
+
+We then fit to the index of abundance per
+
+$$log(cpue_t) \sim normal(f(r,k,m,B0,\pmb{p},\pmb{C}), sigma_{obs{}})$$
+
+## Priors
+
+Priors for all estimable parameters can be left at default values or set by the user.
+The shape parameter is usually not reliably estimable given available data for surplus production models, however, @thorson2012 provides estimate the ratio of B~MSY~ to K for many fish taxa.
+While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter drawn from @thorson2012 based on the genus of the species in question.
+
+We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus.
+
+### Prior Predictive Tuning
+
+Suppose that the only thing we observe from a fishery is a catch history.
+Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
+We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
+Beyond that though, baring additional information we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0 and less than or equal to carrying capacity *k*.
+
+In the absence of any data to fit to, the SRA algorithm works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief.
+This creates a problem for the Bayesian nature of our analysis though.
+Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
+Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of r and k together with the model produces a deterministic stock status in each time step.
+Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox.
+This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
+However, particularly for the SRA version of `sraplus`, Borel's Paradox poses a particular problem.
+
+The SRA algorithm works in two steps.
+First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep).
+From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status.
+If the bulk of the prior on terminal stock status was concentrated at 50% of *k*, combinations of *r* and *k* that produce terminal stock status near 50% of *k* are sampled proportionally more frequently.
+However, lower values of terminal stock status have fewer candidate values of *r* and *k*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
+Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *k* combinations: since under this model the population cannot be greater than carrying capacity, as for example *k* approaches infinity terminal stock status asymptotes at close to 100% of *k*.
+The net result of this is that even though individual combinations of *r* and *k* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
+As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.
+
+We present a novel approximation to this problem here.
+Our solution amounts to a two-step SIR algorithm.
+We first run the standard SIR algorithm as described above.
+We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
+
+$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
+
+We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
+
+$$p(n_i) = \frac{p(bin_i)}{N_i}$$
+
+And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
+The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
+In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
+This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot)).
+
+```{r bias-plot, fig.cap = "Post-model-pre-data distribution of depletion (biomass relative to carrying capacity) from raw SRA algorithm (untuned, top row), from SRA algorithm with approximate tuning applied (tuned, middle row), compared to the supplised prior on depletion (bottom row).
+Black vertical line indicates median value.
+"}
+
+id = unique(ram_comp_data$stockid)
+
+ex <- ram_data %>% 
+  filter(stockid %in% id[[1]])
+
+# fit using catch heuristic
+
+co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
+                                 catch = ex$catch, 
+                                 years = ex$year,
+                                 terminal_state = 0.25,
+                                 terminal_state_cv = 0.5,
+                                 b_ref_type = "k")
+
+biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
+                                      estimate_shape = TRUE)
+
+biased_depletion <- biased_co_fit$fit %>% 
+  filter(variable == "dep_t") %>%
+  filter(year == max(year)) %>% 
+  mutate(set = "Untuned Post-Model-Pre-Data") %>% 
+  select(year, value, set)
+
+co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
+                                      estimate_shape = TRUE)
+
+tuned_depletion <- co_fit$fit %>% 
+  filter(year == max(year), variable == "dep_t") %>% 
+  mutate(set = "Tuned Post-Model-Pre-Data") %>% 
+  select(year, value, set)
+
+prior <- tuned_depletion %>% 
+  mutate(value = rlnorm(n(), log(0.25), 0.5)) %>% 
+    mutate(set = "Prior")
+
+comps <- biased_depletion %>% 
+  bind_rows(tuned_depletion) %>% 
+  bind_rows(prior) %>% 
+  group_by(set) %>% 
+  mutate(valrank = percent_rank(value)) %>% 
+  filter(between(valrank, .05, .95))
+
+comps %>% 
+  ggplot(aes(x = value, y = set)) + 
+  ggridges::geom_density_ridges(alpha = 0.9,
+                                quantile_lines = TRUE, quantiles = 2) +
+  scale_x_continuous(limits = c(0, 1.2), name = "Depletion (B/K)") + 
+  scale_y_discrete(name = '')
+
+```
+
+### Priors Informed by Outside Data
+
+Along with allowing users to supply their own priors, the `sraplus` package contains three classes of built in methods for converting information on stock status from additional outside data into a form usable as a prior by `sraplus`.
+We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
+We then train a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
+Given values of these variables for a new fishery then, `sraplus` uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by `sraplus`.
+
+All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [xx citation].
+These models are intended as a reasonably robust means of translating available data into usable prior for our model.
+Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with fishery status indicators of interest.
+Rather, each regression was tested to ensure that it is unlikely that, given the same data, and alternative model would be unlikely to perform much better than those presented here.
+
+#### Catch-Only Priors
+
+Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
+While these catch-only methods have been shown to have series shortcomings [@free2020], as a point of reference we use two forms of catch-only priors: a statistical and CMSY, as presented implemented by @palomares2020.
+
+For the CMSY based methods, we used a ported version of the CMSY model available [here](https://github.com/DanOvando/portedcmsy).
+For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [boettiger2012].
+Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
+
+For the statistical approach, we utilize data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
+To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
+Emergent clusters show for example one build around "one way trip" style catch histories, others with a boom and bust, etc.
+
+We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history.
+This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term).
+For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status
+
+$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma) $$
+
+For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data
+
+$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
+where *fyear* is the year of the fishery, starting from 0.
+
+#### Fisheries Management Index Priors
+
+The Fisheries Management Index (FMI), as presented in @melnychuk2017, utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
+These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
+The higher the score, the better the expert judges that a given metric is met in that fishery.
+Importantly, FMI surveys can be filled out in the absence of stock assessments.
+This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to @osio2015 and @cope2015).
+
+The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
+
+$$log(value_i) \sim N(s(research_i) + s(management_i) +
+s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
+
+#### Swept Area Ratio Priors
+
+@amoroso2018 provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
+This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
+As illustrated in @amoroso2018, there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
+Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
+if all trawl-able areas are trawled twice a year then the SAR will be 2.
+Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
+
+The final selected model relating SAR to to stock status metrics was
+
+$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
+
+## Example Use of `sraplus`
 
@@
 To illustrate we will walk through a series of ways that `sraplus` could be applied to a case study fishery.
-
-Suppose that the only thing we observe from a fishery is a catch history.
-Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past (i.e.
-that catches have never been greater than biomass).
-We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0 and less than or equal to carrying capacity *k*.
 
@@
 
-For our case study fishery, we can use `sraplus` as a filter to say, conditional on our priors, the catch history, and the population model, what is our implied prior on stock status of the fishery over time (see Section XX and Table.\@ref(tab:prior-tab) for descriptions of model parameters).
+For our case study fishery, we can use `sraplus` as a filter to say, conditional on our priors, the catch history, and the population model, what is our implied prior on stock status of the fishery over time (see Section XX and for descriptions of model parameters).
 While user defined priors on population parameters can be supplied, `sraplus` has a number of built in mechanisms for setting parameters based on the taxonomy of the species in question and the magnitude of the catch history.
@@
 
-```{r}
-
-id = unique(ram_comp_data$stockid)
-
-ex <- ram_data %>% 
-  filter(stockid %in% id[[1]])
-
-ex %>% 
-  ggplot(aes(year, b_v_bmsy)) + 
-  geom_point()
-
-# fit using catch heuristic
-
-co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
-                                 catch = ex$catch, 
-                                 years = ex$year,
-                                 terminal_state = 0.25,
-                                 terminal_state_cv = 0.25,
-                                 b_ref_type = "k")
-
-plot_driors(co_driors)
-
-biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
-                                      estimate_shape = TRUE)
-
-co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
-                                      estimate_shape = TRUE)
-
-plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit)
-
-```
-
 While `sraplus` lets users supply their best estimate of recent stock status for a given fishery, the model also allows users to supply data on the swept area ratio (SAR , @amoroso2018) and/or fisheries management index (FMI, @melnychuk2017) scores for a given fishery.
@@
 
-pull out example from SAR fits
-```{r}
-sar <- sraplus::sar %>% 
-  filter(stockid %in% unique(ram_data$stockid)) %>% 
-  arrange(desc(sar)) %>% 
-  select(stockid, everything()) %>% 
-  slice(10)
-
-ex <- ram_data %>% 
-  filter(stockid %in% unique(sar$stockid))
-
-ex %>% 
-  ggplot(aes(year, b_v_bmsy)) + 
-  geom_point()
-
-# fit using catch heuristic
-
-co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
-                                 catch = ex$catch, 
-                                 years = ex$year,
-                                 terminal_state = 0.25,
-                                 terminal_state_cv = 0.25,
-                                 b_ref_type = "k")
-
-plot_driors(co_driors)
-
-biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
-                                      estimate_shape = TRUE)
-
-co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
-                                      estimate_shape = TRUE)
-
-plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit)
-
-# fit using SAR
-
-sar_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
-                                 catch = ex$catch, 
-                                 years = ex$year,
-                                 sar = sar$sar[sar$stockid == unique(ex$stockid)])
-
-plot_driors(sar_driors)
-
-sar_fit <- fit_sraplus(sar_driors, estimate_shape = TRUE)
-
-plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit, `sar fit` = sar_fit)
-
-# fit using biomass index
-
-index_driors <- sar_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
-                                 catch = ex$catch, 
-                                 years = ex$year,
-                                 sar = sar$sar[sar$stockid == unique(ex$stockid)],
-                                 index = ex$b_v_bmsy[!is.na(ex$b_v_bmsy)], 
-                                 index_years = ex$year[!is.na(ex$b_v_bmsy)])
-
-index_fit <- fit_sraplus(index_driors, engine = "stan", estimate_shape = TRUE)
-
-plot_sraplus(`biased fit` = biased_co_fit, `unbiased fit` = co_fit, `sar fit` = sar_fit,
-             `index fit` = index_fit,
-             plotvars = "b_div_bmsy") + 
-  geom_point(data = ex, aes(year, b_v_bmsy))
-
-```
+We use a Bayesian methods as it suits the philosophy of our analysis: incorporating prior knowledge about stock status drawn from expert opinion or other similar stocks, and where possible confronting those priors with data.
+However, this process does introduce a problem, in that it a) certain parts of the prior probability distribution have zero support in the data (e.g.
+carrying capacity's so small that they crash the population) and b) they create priors on both the *input* (e.g.
+growth rates and carrying capacity) and *outputs* (recent stock status), effectively putting two prior distributions for the same outcome, a problem termed Borel's Paradox XX punt citation.
+This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
+However, particularly for the data-less version of our model, Borel's Paradox poses a particular problem.
+
+Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
+Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of r and k together with the model produces a deterministic stock status in each time step.
+Suppose then we wish to also provide a prior on stock status in the most recent year, for example to reflect local expert opinion based on sound reasoning.
+We have now placed two priors on stock status in the most recent year, one implicit and one explicit, violating Borel's Paradox.
+The net effect of this is that once these priors are run through the model, through for example a sample-importance-resampling (SIR) algorithm as generally applied in stock-reduction analysis, our post-model-pre-data "estimate" of stock status will appear to have "updated".
+This creates the illusion that we have learned something about stock status, when in fact we have simply used our model to reveal our true implied prior on stock status.
+This is technically incorrect, since we do not have a likelihood in the model at this point, and as such per Bayes Theorem have not updated our prior.
+Of more practical importance, this "implied prior" will always reflect a more optimistic stock status then our original prior on stock status, due to the constraints of the population model.
+
+To resolve this, @puntXX (either cite that "on assessment of the bering-chukchi-beufort sea stock of bowhead whales paper that I can't find a date for, or see if he has a new paper on thisXX) suggests a "backwards" fitting approach, in which the model estimates growth parameters and a terminal depletion, and given the model backs out a value of *k* that satisfies the priors (conditional on an assumption initial stock status).
+We do not adopt this approach here due to computational issues.
+Adopting the "backwards" approach requires numerically searching for the plausible *k* during each iteration of the model, which is time consuming.
+In addition, the backwards approach is not without its own complications, for example the requirement of a tolerance around the value of terminal depletion produced by a given *k* and the actual value selected by the model in that time step.
+This discrepancy can occur when a given terminal stock status is not technically possible given the model; to give an extreme example under this model no value of *k* can produce a terminal stock status of 0 if catches are 0 for the entire history of the fishery (assuming *k* > 0).
+
+We present a novel approximation to this problem here.
+Our solution amounts to a two-step SIR algorithm.
+We first run the standard SIR algorithm as described above.
+We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
+
+$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
+
+We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
+
+$$p(n_i) = \frac{p(bin_i)}{N_i}$$
+
+And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
+The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
+In effect, this answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
+
+<!-- @raftery1993 -->
+
+## Value of Information Calculations
+
+Details of value of information calculates (fitting to messed up versions of RAM) XX need to add this in XX
+
+# Supplemental Information
 
@@
 
-```{r prior-tab}
+```{r si-prior-tab}
 cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
@@
   \end{cases}
-  (\#eq:hockey)
+  (\#eq:sihockey)
 \end{equation}
@@
   p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
-  (\#eq:procerror)
+  (\#eq:siprocerror)
 \end{equation}
@@
 
-To resolve this, @puntXX (either cite that "on assessment of the bering-chukchi-beufort sea stock of bowhead whales paper that I can't find a date for, or see if he has a new paper on thisXX) suggests a "backwards" fitting approach, in which the model estimates growth parameters and a terminal depletion, and given the model backs out a value of *k* that satisfies the priors (conditional on an assumption initial stock status).
-We do not adopt this approach here due to computational issues.
-Adopting the "backwards" approach requires numerically searching for the plausible *k* during each iteration of the model, which is time consuming.
-In addition, the backwards approach is not without its own complications, for example the requirement of a tolerance around the value of terminal depletion produced by a given *k* and the actual value selected by the model in that time step.
-This discrepancy can occur when a given terminal stock status is not technically possible given the model; to give an extreme example under this model no value of *k* can produce a terminal stock status of 0 if catches are 0 for the entire history of the fishery (assuming *k* > 0).
-
-We present a novel approximation to this problem here.
-Our solution amounts to a two-step SIR algorithm.
-We first run the standard SIR algorithm as described above.
-We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
-
-$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
-
-We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
-
-$$p(n_i) = \frac{p(bin_i)}{N_i}$$
-
-And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
-The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
-In effect, this answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
-
-<!-- @raftery1993 -->
-
-## FMI and SAR Based Priors
-
-All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [xx citation].
-These models are intended as a reasonably robust means of translating available data into usable prior for our model.
-Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with fishery status indicators of interest.
-Rather, each regression was tested to ensure that it is unlike that, given the same data, and alternative model would be unlikely to perform much better than those presented here XX this might be too much defenseXX
-
-For each statistical prior, we fit three different models, relating the data of interesting, e.g. swept area ratio (SAR), to each of *B/B~MSY~*, *F/F~MSY~*, and *F*
-
-### Swept Area Ratio Priors
-
-@amoroso2018 provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
-This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
-As illustrated in @amoroso2018, there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
-Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
-if all trawl-able areas are trawled twice a year then the SAR will be 2.
-Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
-
-The final selected model relating SAR to to stock status metrics was
-
-$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
-
-## Fisheries Management Index Priors
-
-The Fisheries Management Index (FMI), as presented in @melnychuk2017, utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
-These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
-The higher the score, the better the expert judges that a given metric is met in that fishery.
-Importantly, FMI surveys can be filled out in the absence of stock assessments.
-This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to @osio2015 and @cope2015).
-
-The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
-
-$$log(value_i) \sim N(s(research_i) + s(management_i) +
-s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
-
-## Catch-Only Priors
-
-Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
-While these catch-only methods have been shown to have series shortcomings [@free2020], as a point of reference we use two forms of catch-only priors: a statistical and CMSY, as presented implemented by @palomares2020.
-
-For the CMSY based methods, we used a ported version of the CMSY model available [here](https://github.com/DanOvando/portedcmsy).
-For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [boettiger2012].
-Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
-
-For the statistical approach, we utilize data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
-To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
-Emergent clusters show for example one build around "one way trip" style catch histories, others with a boom and bust, etc.
-
-We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history.
-This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term).
-For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status
-
-$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma) $$
-
-For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data
-
-$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
-where *fyear* is the year of the fishery, starting from 0.
-
-## Effort Data
-
-While the FMI and SAR data show some potential for estimating U/U~MSY~ and B/B~MSY~ values, they are clearly not an ideal measure of stock status.
-An alternative is to actually develop indices of abundance that we can fit models to.
-@rousseau2019 provides a useful starting place for this exercise, presenting estimates of fishing effort throughout the globe from 1950 to 2015.
-Merging these data with capture statistics can provide a crude CPUE index that can then be used to fit a surplus production model.
-Developing an index of abundance in this manner is clearly a very rough analysis only intended to illustrate broad concepts.
-Any analysis of CPUE data can be severely affected by changes in technology, skill, hyperdepletion, hyperstability, changes in the spatial footprint of the fishery, and selectivity [@maunder2004a; @walters2003].
-In the case of this analysis, a kilowatt day of fishing with 1950s technology clearly represents a different amount of fishing power than a kilowatt day of fishing in 2012.
-To account for this somewhat, @rousseau2019 include a 2.6% annual "technological creep" factor, which we incorporate in our analysis.
-
-@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6\% technology rate).
-This assumes that every unit increase in effort is linearly greater than the unit of effort before it.
-When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
-This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
-To accommodate this, we generate a catch per effective harvest rate index of abundance, as
-
-$$I_t = \frac{catch_t}{(1 - e^{-q_tE_t})}$$
-
-Where $q_t$ can has a technology rate component $\tau$
-
-$$q_t = q_{t-1} \times (1  + \tau)$$
-
-## Value of Information Calculations
-
-Details of value of information calculates (fitting to messed up versions of RAM) XX need to add this in XX
-
-# Supplemental Information
-
-## Population Model
-
-The core of our model is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018.
-While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
-
-```{r si-prior-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
-  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
-  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
-  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
-  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
-  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))
-
-knitr::kable(
-  cd,
-  "latex",
-  align = "l",
-  booktabs = TRUE,
-  escape = F,
-  row.names = F,
-  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
-) %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
-
-The population growth equation is
-
-\begin{equation}
-  f(x)=\begin{cases}
-      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
-, & \text{if $B_t>0.25 \times K$}.\\
-     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
-  \end{cases}
-  (\#eq:hockey)
-\end{equation}
-
-Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of Bmsy/K to shift (when *m* is 2 the Bmsy / K = 0.5 and the model acts as a Schaefer model, lower values of m shift the production function left, higher values right), *c* is catch, and *p* is process error.
-The Pella-Tomlinson model does have an issue where growth rates can become unrealistically large when the population reaches low sizes.
-We deal with this problem by following the methods described in @winker2018.
-to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
-Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see @thorson2018).
-
-<!-- When the SIR method is used we use the supplied catch history.
-However, for the maximum likelihood approach, convergence is greatly facilitated by estimating catches ($\hat{c_t}$) rather than using the observed catches directly.
-$\hat{c_t}$ is given as -->
-
-<!-- \begin{equation} -->
-<!--   \hat{c_t} = u_{t}B_{t} -->
-<!--   (\#eq:chat) -->
-<!-- \end{equation} -->
-
-<!-- where $u_t$ is estimated through -->
-
-<!-- \begin{equation} -->
-<!--   u_{t} = \frac{1}{1 + e^{-u^*_t}} -->
-<!--   (\#eq:invf) -->
-<!-- \end{equation} -->
-
-<!-- Equation \@ref(eq:invf) ensures that $u_t$ is between 0 and 1, while the model estimates $u^*_t$ to improve performance.  -->
-
-<!-- $\hat{c_t}$ is then fit to the observed catches through the likelihood -->
-
-<!-- \begin{equation} -->
-<!--   log\left(c_t\right) \sim normal\left(log\left(\hat{c_t}\right),0.05\right) -->
-<!--   (\#eq:catch) -->
-<!-- \end{equation} -->
-
-<!-- Where we assume an observation coefficient of variation of 5\%.
-We estimate catch rather than directly using the observed catch in order to improve model performance.
-As priors on depletion, or indices of abundance, pull the estimated population closer to zero, there is a greater and greater chance that the observed catch history will result in a collapse of the population (catch in time *t* being greater than biomass in time *t*); It is much harder to find parameters that produce a final depletion of 1\% of carrying capacity while not crashing the population prior to then given a catch history, than it is to produce a final depletion of 99\% of carrying capacity.
-We can impose restrictions and penalties on population crashes, but this distorts the gradients of the model, causing the model to often become stuck once a few crashes have occurred.
-Since $u_t$ is constrained between 0 and 1, $\hat{c_t}$ will always be "viable", and by then fitting to the observed catches, the model has an easier time finding parameters that produce a near match to the observed catches while not crashing the population.
-As of now none of the data used in this exercise have missing years of catch data, but the model can handle those cases in the future due to the random walk priors imposed on $u_t$ detailed below.
--->
-
-<!-- In order to help separate out observation error in the catch and process error in the population model, when catch is being estimated we assume that fishing mortality $u_t$ follows a random walk on average, assigning each $u_t$ a prior of -->
-
-<!-- \begin{equation} -->
-<!--   log(u_t) \sim normal(log(u_{t - 1}),0.1)) -->
-<!--   (\#eq:rwalk) -->
-<!-- \end{equation} -->
-
-We  allow for process error (in the manner of the stochastic stock reduction analysis error suggested by @walters2006).
-This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
-Process error *p* is assumed to be log-normally distributed, such that
-
-\begin{equation}
-  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
-  (\#eq:procerror)
-\end{equation}
-
-All of our estimates are Bayesian in nature.
-We can break the use of our model into two distinct categories: with data and without.
-By "data", we refer to measurements which are confronted with model estimates within a likelihood function.
-In our context, these include fishery-independent survey data, or a CPUE index.
-When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
-Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population, given the constraints of the model.
-
-We use a Bayesian methods as it suits the philosophy of our analysis: incorporating prior knowledge about stock status drawn from expert opinion or other similar stocks, and where possible confronting those priors with data.
-However, this process does introduce a problem, in that it a) certain parts of the prior probability distribution have zero support in the data (e.g.
-carrying capacity's so small that they crash the population) and b) they create priors on both the *input* (e.g.
-growth rates and carrying capacity) and *outputs* (recent stock status), effectively putting two prior distributions for the same outcome, a problem termed Borel's Paradox XX punt citation.
-This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, particularly for the data-less version of our model, Borel's Paradox poses a particular problem.
-
-Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
-Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of r and k together with the model produces a deterministic stock status in each time step.
-Suppose then we wish to also provide a prior on stock status in the most recent year, for example to reflect local expert opinion based on sound reasoning.
-We have now placed two priors on stock status in the most recent year, one implicit and one explicit, violating Borel's Paradox.
-The net effect of this is that once these priors are run through the model, through for example a sample-importance-resampling (SIR) algorithm as generally applied in stock-reduction analysis, our post-model-pre-data "estimate" of stock status will appear to have "updated".
-This creates the illusion that we have learned something about stock status, when in fact we have simply used our model to reveal our true implied prior on stock status.
-This is technically incorrect, since we do not have a likelihood in the model at this point, and as such per Bayes Theorem have not updated our prior.
-Of more practical importance, this "implied prior" will always reflect a more optimistic stock status then our original prior on stock status, due to the constraints of the population model.
-
 The SRA algorithm works in two steps.

### 2020-09-10 documents/ovando-etal-assessing-global-fisheries.Rmd
run on squid

@@
 output: 
-  bookdown::pdf_document2: default
   bookdown::word_document2:
     reference_docx: word-template.docx
+  bookdown::pdf_document2: default
 params:
@@
 While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor.
-Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results.
+Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
 This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources.
@@
 These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection.
-Our results provide further evidence that catch data by themselves cannot provide reliable estimates of fishery health.
+Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random quess.
 Our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
@@
 
-pub_theme <- theme_ipsum(base_size = 10,
-                         axis_text_size = 10,
-                         axis_title_size = 12) +
-    theme(panel.spacing = unit(.5,"lines"))
-
-theme_set(pub_theme)
-
-```
-
-# rishi notes
-
-* more assertive on what the data needs really are
-  * where to apply globally 
-* For places with less coverage, what would we actually recommend there 
-* what do we know about catchabiliy
-- e.g. 51, fleets became industrial 
-  - hammer down on that's the issue with the effort 
-  - how do we translate effort into f bey area
-  - here's
-
-Target journal: Nature Sustainability
-
-https://www.nature.com/natsustain/about/content
-
-3500 words main text
-
-50 references (main text)
-
-3000 words (max) methods
-
-# Introduction
-
-The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
-A substantial part of achieving this goal rests on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems.
-Achieving this goal depends in part on our ability to effectively track the status of global marine fish stocks.
-The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking this important metric.
-As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished ,and 34.2% are overfished [@fao2020].
-While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment.
-Although revised periodically over time, the method has kept revisions to a minimally necessary levels for the purposes of temporal consistency and comparability.
-With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.
-
-49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020].
-The SOFIA process uses these formal assessments for their determinations of stock status wherever possible.
-However, that leaves 51% of global fisheries landings lacking in formal stock assessments.
-While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion.
-While local experts often are well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable.
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
-
-Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE).
-While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect.
-Even when data are available though, running an individualized stock assessment is not a "just press play" process.
-Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment.
-This process takes considerable time and expertise.
-
-Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment.
-Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (@free2020 and references therein).
-These models seek to infer stock status, for example in terms of biomass *b* relative to the biomass at maximum sustainable yield *b~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
-In addition, given that a global assessment by its nature requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time.
-As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks (high variance, low bias).Each of these methods sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.
-
-However, Free et al.
-2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status (B/B~MSY~).
-These issues become apparent when we consider some of the macro-level predictions made by these models.
-The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020].
-A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.
-
-xx spend time on this tomorrow xx
-Examining stock assessment data from RAM from the NE Pacific, the NE Atlantic and the Mediterranean, the NE Pacific is in the best state (mean RAM b/b~MSY~ = 1.8), NE Atlantic (mean RAM b/b~MSY~ =1.3) and the Mediterranean the worst (mean RAM b/b~MSY~ =0.5).
-We do not have formal stock assessment data from S.
-and SE Asia (Western Central Pacific and Eastern Indian Ocean), but other data sources such as surveys and expert opinion suggest this region should have more depleted fisheries than Northeast Pacific.
-The Costello et al.
-2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and SE Asia seem to be overoptimistic, and the NE Pacific should be better by comparison.
-The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific.
-The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).
-
-```{r status-tab}
-
-readr::read_csv(here("data","status-table.csv")) %>% 
-  knitr::kable(caption = "Status by FAO Statistical area from four different assessment methods",
-               booktabs = TRUE) %>% 
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-
-```
-
-The coming decades are a critical period for the future of global oceans.
-Climate change, population growth, and economic expansion will all place increasing pressure on fisheries.
-While our understanding of global fisheries has improved greatly,  mitigating these threats, as called for by SDG 14, will require an improved ability to track the status of fisheries around the world, particularly those stocks that are currently unassessed.
-At the same time, new sets of globally distributed data have recently become available, along with advances in statistical modeling methods.
-We leverage these opportunities to provide a global assessment of our perception of global stock status resulting from different use of different data sources.
-Our results demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection and interpretation effort.
-
-# Results
-
-xx move some of this to intro and chop downXX
-We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
-The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability.
-However, catches are not the only broadly available source of fishery data.
-We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@melnychuk2017], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c].
-`sraplus` provides a flexible framework for integrating these data into valid statistical model.
-`FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model.
-We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~.
-Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.
-
-<!-- ## Case Studies  -->
-
-<!-- Two case studies provide examples of how augmenting catch data alone with local and globally available data can provide dramatically different estimates of local stock status.
-For our first case study, we augment local catch statistics from the Eastern Bay of Bengal (XX details on these data) with catch-per-unit effort data from a local study, as well as a local estimate of swept-area-ratio for the region.
-We can then compare the estimates of stock status that come from using the catch-data alone (with a heuristic algorithm in the manner of @martell2013), with the catch data along with priors on terminal depletion informed by the swept-area-ratio data, and using the swept-area ratio data together with a local abundance index.
-By it's nature the catch-only heuristic produces a bi-model distribution, with most fisheries reported to have a B/B~MSY~ of 1.5 and an F/F~MSY~ less than 1 (Fig.\@ref(fig:east-india)-heuristic).
-Local experts put the swept-area-ratio (SAR) value for the region at an incredibly high 17.
-Using this value to inform a prior on terminal stock status dramatically changes our picture of stock status in the Eastern Bay of Bengal, with now all stocks being overfished and experiencing high levels of overfishing (Fig.\@ref(fig:east-india)-sar).
--->
-
- <!-- Evaluating the validity of data-limited assessments methods can be challenging.
-Simulation testing provides one avenue, but it can be difficult to create a simulation environment that adequately captures the problems of real-world ecosystems and data.
-We can also compare results between data-limited assessment and more data (or resource) rich gold-standard stock assessments.
-However, formally stock assessed fisheries represent a very biased subset of global fisheries, and it is entirely possible that a method that performs well for the kinds of species and fisheries that are commonly stock assessed might struggle in the kinds of fisheries that typically are unassessed.
--->
-
- <!-- Alternatively though, while we may lack formal assessment for many regions of the world, stakeholders familiar with a region often have a general sense of the state of fisheries in a region.
-As outlined in Table.\@ref(tab:status-tab), while we often have no proof that results finding that for example fisheries in south-east Asia have similar stock status to those in Alaska, very few experts would agree that this is a plausible result XX tough statement to make in a paper xx.
-We use two case studies here to present the potential of additional, broadly applicable, data to potentially improve on "catch-only" methods.
--->
-
-<!-- If all the stocks in question were demersal, shared similar life histories, and were all caught in the same multi-species fleet, this result might make sense.
-However, imposing the same prior on terminal stock status for all fisheries in a region is not ideal.
-Supplementing these data with a local index of abundance provides a more nuanced result, with most stocks estimated as having B/B~MSY~ less than 1, but with a much wider range of fishing mortality rates reflecting difference in recent catch histories among individual stocks (Fig.\@ref(fig:east-india)-"cpue + sar").
-While we do not know the true state of stocks in the Eastern Bay of Bengal, inclusion of local data provide a picture more in line with local perception XX.
--->
-
-<!-- ```{r east-india, fig.cap="Estimates of stock status for fisheries in the Eastern Bay of Bengal using the catch heuristic, fishing mortality priors informed by swept area ratios, and a local index of abundance together with swept area ratio.
-Each point is a stock."} -->
-<!-- ei_kobe_plot  -->
-<!-- ``` -->
-
-We used `sraplus` to assess the status of stocks in Europe and Northeast Asia using catch data from the FAO augmented with regional effort indices constructed from [@rousseau2019] (xx technically this is the updated dataset xx).
-Catch histories for Europe show a steady increase reaching a peak in the 1990s, and a decline and stabilization since.
-Current catches in Europe are much higher than those at the start of the catch history.
-Catches in northeast Asia have steadily increased over time, with some stabilization as of late.
-For both regions, recent catches expressed as a fraction of maximum catch, they key metric from the perspective of a catch-only heuristic, are both high.
-However, the two regions exhibit markedly different effort histories according to @rousseau2019, with effort in Europe peaking with catches and subsequently declining, while effort in northeast Asia has increased steadily over time.
-As a result, using a "catch-only" heuristic, both Europe and Northeast Asia are estimated to have very similar stock status.
-Adding in an index of abundance, calculated for each stock as catch divided by the effective harvest rate based on the effort data for the region, produces a very different picture of stock status and introduces some contrast between the two regions.
-Based on the catch and effort data, `sraplus` estimates that more stocks in Northeast Asia are overfished / experiencing overfishing than in Europe.
-However, the shared effort series across these stocks also produces a large number of effectively unfished stocks; a highly unlikely result for both regions (Fig.\@ref(fig:eu-v-na)).
-
-XX add stock assessment estimates for both regions? XX Maybe just do matching stocks?
-
-```{r eu-v-na, fig.cap="Estimates of stock status in Europe and Northeast Asia using either a simple catch heuristic as presented in @martell2013 or an abundance index based on continental-scale effort trends reported in @rousseau2019 Each point is a stock, point size is a function of stock size."}
-continent_kobe_plot
-```
-
-## Conflicting Estimates of Global Status
-
-```{r perf-tab}
+  exs <- read_rds(path = file.path(results_path, "raw-case-study-fits.rds"))
+
 breaks <- c(0.8, 1.2)
@@
 
-knitr::kable(perf, digits = 2,
-             caption = "Global summary statistics of models using different sources of data.
-mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).",
-              format = "pipe")
-
-```
-
-Extending this idea to the global level, we  used a variety of data sources to estimate stock status for fisheries listed in RAM.
-Our goal is to simulate a macro-level global assessment, where data are available at regional levels, but not for specific fisheries.
-As a proof of concept, we first fit models to abundance indices from each stock in RAM.
-We assess performance by three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
-Classification accuracy is calculated as the proportion of times that a given dataset results in a stock falling into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
-There are two main reasons why an assessment model might produce poor results: model misspecification and poor data.
-Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries.
-Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`.
-Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality.
-Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.
-
-<!-- The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias.
-The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time.
-Looking across space, we see clear geographic patterns in model performance.
-The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada).
-The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.
+fao_area_codes <- fao %>% 
+  ungroup() %>% 
+  select(fao_area, fao_area_code) %>% 
+  unique()
+
+pub_theme <- theme_ipsum(base_size = 10,
+                         axis_text_size = 10,
+                         axis_title_size = 12) +
+    theme(panel.spacing = unit(.5,"lines"))
+
+theme_set(pub_theme)
+
+```
+
+<!-- # rishi notes -->
+
+<!-- * more assertive on what the data needs really are -->
+<!--   * where to apply globally  -->
+<!-- * For places with less coverage, what would we actually recommend there  -->
+<!-- * what do we know about catchabiliy -->
+<!-- - e.g. 51, fleets became industrial  -->
+<!--   - hammer down on that's the issue with the effort  -->
+<!--   - how do we translate effort into f bey area -->
+<!--   - here's  -->
+
+# Publication Requirements
+
+Target Journal: [Nature Sustainability](https://www.nature.com/natsustain/about/content)
+
+3500 words main text
+
+50 references (main text)
+
+3000 words (max) methods
+
+# Introduction
+
+The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
+A substantial part of achieving this goal depends on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems.
+Meeting these targets depends in part on our ability to effectively track the status of global marine fish stocks.
+The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking the global state of fisheries.
+As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020].
+While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment.
+Although revised periodically over time, the method has kept revisions to minimally necessary levels for the purposes of temporal consistency and comparability.
+With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.
+
+49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020].
+The SOFIA process uses these formal assessments for their determinations of stock status wherever possible.
+However, that leaves 51% of global fisheries landings lacking in formal stock assessments.
+While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
+The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion.
+While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable (and expert opinion can also be misinformed).
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018;@thorson2012a].
+
+Formal stock assessments require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE).
+While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist [@maureaud2020], and would be prohibitively expensive to collect.
+Even when data are available though, running an individualized stock assessment is not a "just press play" process.
+Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment.
+This process takes considerable time and expertise.
+
+Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment.
+Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (Free et al.
+2020 [@free2020] and references therein).
+These models seek to infer stock status, for example in terms of biomass *B* relative to the biomass at maximum sustainable yield *B~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
+In addition, a global assessment requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time.
+As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks.
+Each of the published catch-based methods reviewed in Free et al.
+2020 [@free2020] sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.
+
+However, Free et al.
+2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~.
+These issues become apparent when we consider some of the macro-level predictions made by these models.
+The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020].
+A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.
+
+Examining stock assessment data from RAM from the  Northeast Pacific,the Northeast Atlantic and the Mediterranean & Black Sea, the Northeast Pacific is in the best state (mean RAM B/B~MSY~ = 1.61), NE Atlantic (median RAM b/b~MSY~ =1.27) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.52) [@hilborn2020].
+The Costello et al.
+2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison.
+The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or NOrtheast Pacific.
+The catch based approach @pauly2007, which compares catch in the last year to maximum catch and equates these to stock status, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).
+While the methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks,
+
+```{r status-tab}
+
+o_by_a <- sofia_status %>% 
+  left_join(fao_area_codes, by = "fao_area_code") %>% 
+  group_by(fao_area) %>% 
+  summarise(`FAO % Overfished` = percent(mean(status == "O")))
+
+ss <-
+  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
+                    sheet = "state_space_results_08_28_2018") %>%
+  janitor::clean_names()
+
+ss_lookup <-
+  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
+                    sheet = "ss_to_fao") %>%
+  janitor::clean_names() %>% 
+  mutate(tmp = map2(region, fao_area_codes, ~data.frame(region = .x, fao_area_code = as.integer(str_split(.y, pattern = ',',simplify = TRUE))))) %>% 
+  select(tmp) %>% 
+  unnest(cols = tmp) %>% 
+    left_join(fao_area_codes, by = "fao_area_code")
+
+state_space_results <- ss_lookup %>% 
+  left_join(ss, by = "region") %>% 
+  filter(variable == "BvB") %>% 
+  group_by(fao_area) %>% 
+  filter(year == max(year)) %>% 
+  summarise(`B/BMSY RAM` = round(mean(dlm_geomean),2)) %>% 
+  ungroup()
+
+# ram_by_region %>% 
+#   ggplot(aes(b_v_bmsy)) +
+#   geom_histogram() + 
+#   facet_wrap(~fao_area)
+
+readr::read_csv(here("data","status-table.csv")) %>% 
+  select(-`B/BMSY RAM`) %>%  
+  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
+  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
+  arrange(`FAO % Overfished`) %>% 
+  select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
+  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM through Hilborn et al. 2020, Costello et al. 2016, 
+               Rosenberg et al. 2018, and Pauly 2007",
+               booktabs = TRUE,
+               format = "pipe") %>% 
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+
+```
+
+In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fisheries?
+We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
+The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability.
+However, catches are not the only broadly available source of fishery data.
+We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management through Fisheries Management Index (FMI) scores [@melnychuk2017], trawl-fishing footprints through swept-area-ratio (SAR) values [@amoroso2018], and life history [@thorson2017c].
+`sraplus` provides a flexible framework for integrating these data into valid statistical model.
+
+The coming decades are a critical period for the future of global oceans.
+Climate change, population growth, and economic expansion will all place increasing pressure on fisheries.
+While our understanding of global fisheries has improved greatly,  mitigating these threats, as called for by SDG 14, will require an improved ability to track the status of fisheries around the world, particularly those stocks that are currently unassessed.
+At the same time, new sets of globally distributed data have recently become available, along with advances in statistical modeling methods.
+We leverage these opportunities to provide a global assessment of our perception of global stock status resulting from different use of different data sources.
+Our results demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection and interpretation effort.
+
+# Results
+
+We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
+From there, we assess the performance of models fit using different kinds of broadly available data (combinations of catches, effort, fisheries management index, and swept area ratio values) relative values reported in the RAM database.
+
+ <!-- `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model.
+We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~.
+Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.
 -->
 
-Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves.
-These becomes even clearer once we consider the performance of `sraplus` models fit to globally available datasets.
-Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)).
-However, this is somewhat an artifact of the data.
-The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4.
-The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1.
-This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise.
-Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
-The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])`.
-
-Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process.
-We also see that across the models performance (or lack thereof) is not consistent in space: different data to best/worst for different FAO regions.
-For example, the catch heuristic and models fit to nominal CPUE data both substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
-We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
-Overall, while some data sources performed slightly better than others by some metrics in some places, there was no consistent spatial pattern to this performance, and none of the datasets were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
-
-```{r dat-description}
+XX discussion with Jim.
+How can we frame this.
+I would actually push back the idea that we have learned about stock status.
+I buy that the catch data can update your prior on growth parameters, but by the rules we're playing with here, it shouldn't update our prior on stock status should it.
+Now, the only thing I can't wrap my head aroudn is the case where catches are all zero, and you have a prior that the stock is collapsed.
+I guess my question is from a strikely Bayes theorem perspective, where in here is your prior on stock status being updated, since there's no likelihood as written, except for the binomial process that says that the stock exists.
+
+For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index and swept area ratio scores.
+We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock.
+We used these data to demonstrate how `sraplus` can add incremental amounts of information to an assessment, and the potential value of doing so.
+We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status.
+We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`.
+It is critical to note that in both of these cases the model has learned nothing new about recent stock status that was not encoded in the supplied priors, catch data, and functional form of the model; `sraplus` simply helps reveal our implicit priors on other parameters (e.g.
+maximum sustainable yield, MSY), and provides timeseries of measures such as B/B~MSY~.
+That, in the absence of data to actually fit to such as an index of abundance, methods such as CMSY or sraplus can learn anything new about a stock's status that is not already implicitly coded in the priors, catch data, or specified functional relationships, is a common misconception (see [Methods](#Methods) a for detailed explanation of why this is).
+Lastly, we used the reconstructed effort data from @rousseau2019 to create an index of abundance for each stock (see [Estimation Model](#Estimation Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.
+
+Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
+CMSY, using no stock-specific information but the catch history and the estimated resilience of the target species, badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
+Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
+Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
+This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values.
+An abundance index based on catch and regional effort cannot capture this trend, since it interprets the extremely low catches (near zero) as low biomass since we do not have sufficiently fine-scale effort data to reflect the reduction in effort that goes hand-in-hand with regulated catch reductions (Fig.\@ref(fig:cs-plot)).
+
+```{r cs-plot, fig.cap="Estimates of stock status for case study fisheries using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index and Swept Area Ratio scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
+Each point is a stock, point size is a function of stock size.
+The shaded hexagons in the background show the distribution of stock status estimates from the RAM Legacy Stock Assessment Database for each of the stocks included in this figure, with darker colors representing increased concentrations of RAM stocks."}
+
+ex_kobe_plot
+```
+
+That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
+In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
+Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
+
+While SAR and FMI data are available for a wide number of stocks, they are not ubiquitous enough to provide stock-specific priors for the number of fisheries needed for a global assessment.
+In addition, predicting stock status based on SAR, FMI, or catch patterns requires that the same patterns that exist within the training data as exist in other stocks to be used for prediction, which is far from clear given the highly managed nature of the stocks for which we have reliable estimates of stock status.
+We can extend the ideas presented in the case study, using combinations of catch, SAR, FMI, and effort data to produce estimates of stock status at the global level.
+
+As our test case we use `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
+B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
+We then pair the catch histories for these stocks with regional-level swept area ratio, fisheries management index, and effort data.
+For any given stock we use the area-weighted swept area ratio (where area refers to the spatial area of the SAR study in question) for the FAO region of the stock.
+For fisheries management index we use the mean component FMI values for the species group and FAO region in question (or FAO region if no matches exist at the species group level).
+This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
+We then constructed an effort index for the stock in question as the total effort from @rousseau2019 estimated in all countries that report catch of the species in question within its FAO region, weighting the effort of each country by its contribution to the total catch of that species in that FAO region.
+
+As a proof of concept, we first estimated B/B~MSY~ of our candidate RAM stocks by using `sraplus` to fit to an abundance index drawn directly from RAM.
+We then fit a range of models utilizing different combinations of the regional-level data made available to them (Table.\@ref(tab:dat-desc)).
+We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
+Classification accuracy is calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
+
+```{r dat-desc}
 dat_description <- tribble(~`Data Name`, ~Description,
@@
 
- xx dig into this, which stocks are in that region? XX
-
-```{r}
-a  = ram_comp_data %>% 
-  filter(fao_area_code == 61)
-
-unique(a$common_name)
-```
+Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries.
+Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`.
+This creates two main reasons why `sraplus` might produce poor results at a global level: model misspecification and poor data quality.
+Model misspecification errors might result from the surplus production model underlying `sraplus` being fundamentally unable to capture the more complicated dynamics resulting from the full age-structured population models common in RAM.
+Errors resulting from poor quality data mean that even if the underlying model is a reasonable approximation of the real world, the model may produce incorrect results due to the data supplied to the model simply not containing accurate information on the phenomenon in question.
+
+Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality.
+Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+This exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
+
+<!-- The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias.
+The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time.
+Looking across space, we see clear geographic patterns in model performance.
+The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada).
+The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.
+-->
+
+Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves.
+These becomes clearer once we consider the performance of `sraplus` models fit to combinations of our broadly available datasets.
+Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)).
+However, this is somewhat an artifact of the data.
+The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4.
+The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1.
+This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise.
+Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
+The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`.
+Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])` (since fisheries are not uniformly distributed across each bin).
+
+Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process.
+Across the models performance (or lack thereof) is not consistent in space: use of different data performed best or worst for different FAO regions.
+For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
+We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
+Overall, while some data sources performed slightly better than others by some metrics in some places, there was no consistent spatial pattern to this performance, and no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
+
+```{r perf-tab}
+
+knitr::kable(perf, digits = 2,
+             caption = "Global performance statistics in the most recent year available of models using different sources of data.
+mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
+Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.",
+              format = "pipe")
+
+```
+
+ <!-- xx dig into this, which stocks are in that region? XX -->
 
@@
 CPUE refers to an index of abundance based on the Rousseau effort data.
-Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
-The FAO Report 569 is our reference status estimate provided by the FAO.
+cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
+For both CPUE series 'nominal' assumes a 0% technology rate, otherwise a 2.6% technology rate is assumed.
 ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
-fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
+fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio.
+CMSY uses the methods from @froese2017."}
 ram_mpe_map_plot
@@
 CPUE refers to an index of abundance based on the Rousseau effort data.
-Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
-The FAO Report 569 is our reference status estimate provided by the FAO.
+cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
+For both CPUE series 'nominal' assumes a 0% technology rate, otherwise a 2.6% technology rate is assumed.
 ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
-fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
+fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio.
+CMSY uses the methods from @froese2017."}
 ram_mape_map_plot
@@
 CPUE refers to an index of abundance based on the Rousseau effort data.
-Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
-The FAO Report 569 is our reference status estimate provided by the FAO.
+cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
+For both CPUE series 'nominal' assumes a 0% technology rate, otherwise a 2.6% technology rate is assumed.
 ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
-fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio."}
+fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio.
+CMSY uses the methods from @froese2017."}
 
@@
 # Discussion
+
+<!-- XX somewhere in here make clear that you're not learning anything from the catch data XX -->
+
+<!-- XX Acknolwedge the need for action the the frustration of being told "get more data" XX -->
 
@@
 Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world.
-While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone, at the global level no combination of readily available data was able to substantially and reliably clarify our picture of global fisheries status.
+While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models using each of the available datasets beside the RAM-derived indices produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).
 
@@
 It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
-We have stated that addition of the globally available datasets compiled here did not substantially improve estimates of stock status, but in some instances using these data did provide some improvement over use of catch-only style methods; the difficulty comes in attempting to apply data types uniformly across the globe.
-However, none of the data sources used here came close to the benchmark performance provided by our fits to the RAM data, either in overall performance or in capturing the geographic distribution of stock status.
+In some instances using the data presented here did provide some improvement over use of catch-only style methods; the difficulty comes in attempting to apply data types uniformly across the globe.
 While it is unreasonable to expect any global-scale data to be able to perform as well as data pulled from RAM assessments themselves, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general trends in stock status in time and space.
@@
 
-This is not to say that the data presented here are not valuable.
+Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions.
 The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e.
-those that sufficiently resemble the data on which the regressions were trained).
-Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with relatively constant fishing technologies and a broadly selective fishing fleet.
-For example, despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis.
+those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)).
+Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
+Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
 But, applying one SAR value to all stocks in a region causes inaccurate estimates of stock status.
-The challenge comes in attempting to indiscriminately apply these data to all areas: doing so results in substantially incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
+Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
 
@@
 In world of finite resources though, it is important to consider what sources of data might provide the greatest value in improving our estimates of global stock status.
-We fit 3,000XX sraplus models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
+We fit 3,000 `sraplus` models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
 We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).
 
-XX will simplify this plot down to key things and clean up the variable names XX
-
-Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 20%.
-In terms of data however, information on fishing mortality rates (F/F~MSY~) in the most recent year was almost as useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot).
-Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance (xx uep, since in some cases the index was flatish but the stock status was low, so a model fit just to the trend in the abundance gets the trend right but the absolute value wrong.
-having F/Fmsy data helps anchor that).
-This is because we varied both the length and quality of the index of abundance, and on average an even noisy estimate of recent F/F~MSY~ may be more valuable than a short and noisy abundance index.
-
-```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types and model options on root mean squared error of B/Bmsy in the most recent 5 years of data available for each model fit.
-Dark blue line shows median, blue shaded area the 66% credible interval, full rea the 95% credible interval."}
+Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 15%.
+Information on fishing mortality rates (F/F~MSY~) in the most recent year was equally useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot).
+Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance.
+While having access to a complete timeseries of an abundance index had a higher probability of reducing RMSE than increasing it, in many cases having only an abundance index resulted in a correct fit to the overall trend in abundance, but an error-prone estimate of B/B~MSY~ itself due to incorrectly estimating either the shape parameter, the initial state of the stock, or both.
+However, using an index of abundance that only covered the most recent quarter of a fishery's history actually increased RMSE by over 10% on average.
+
+```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of B/B~MSY~ in the most recent 5 years of data available for each model fit.
+Distributino is full posterior probability distribution.
+Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th."}
 b_voi_plot + 
@@
 
-Given the demonstrated value of F/Fmsy values in improving estimates of stock status, how might we go about obtaining estimates of fishing mortality rates at a global scale?
-We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/Fmsy for individual fisheries (Fig.SXX).
+As we have stated here using catch data alone we cannot learn anything new about recent stock status beyond what is encoded in our priors.
+So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest).
+Given the demonstrated value of F/F~MSY~ values in improving estimates of stock status, and the increased feasibility in estimating F/F~MSY~ at a local scale relative to B/B~MSY~, how might we go about obtaining estimates of fishing mortality rates at a global scale?
+We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/F~MSY~ for individual fisheries (see Supplemental Information).
 While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
@@
 As a result, many fisheries around the world are enacting programs to collect and utilize length composition data for fisheries assessment.
-While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries this may provide an overlooked source of fisheries data at a global scale.
+While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries this may provide an overlooked source of fisheries data at a global scale, at least as an improvement over relying on catch-alone or regional proxies.
 
@@
 Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which when paired with catch data and where possible indicies of abundance could meaningfully improve our understanding of global fisheries [@thorson2015f].
-Such a database could be constructed in a top-down manner in a manner similar to the catch records compiled by the FAO (xx maybe top down isn't best descriptor here xx,or a bottom-up platform, such as the [MERMAID](https://datamermaid.org/) platform for coral reef data, could be developed that would allow users around the world to self-contribute length composition studies.
-
-This is not to say that we should also invest in global strategies for calculating abundance indicies.
-Where possible, repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//).
-However for many areas of the world effective fishery independent estimates of stock abundance are unlikely to be feasible.
+Such a database could be constructed in a top-down manner in a manner similar to the catch records compiled by the FAO,or a bottom-up platform, such as the [MERMAID](https://datamermaid.org/) platform for coral reef data, could be developed that would allow users around the world to self-contribute length composition studies.
+
+We should of course prioritize collection of survey data worldwide, high-quality synthesis and standardization of these data, as well as their collection in an international database.
+This assessment assumes that we have perfect knowledge of for example F/F~MSY~, when in fact any estimates derived from for example length-based methods are likely to be highly error-prone.
+However, our results demonstrate that we must prioritize aggregation of long-running but less publicly-accessible survey data, or expect to wait for some time until new surveys are capable of yielding meaningful information on their own about stock status.
+Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//).
+Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have previously had stock assessments.
+
+For many areas of the world effective fishery independent estimates of stock abundance are unlikely to be feasible in the near future.
 Effort databases such as @rousseau2019, can be combined with catch histories to to construct fishery-dependent abundance indicies.
 While there are substantial difficulties and biases that can arise from improper use of fishery-dependent abundance data, improved understanding of the evolution of global fishing effort could be very useful in improving global estimates of fishery health.
-While we did include effort reconstructions in our analysis, and did not find it to provide substantial improvements in assessment results, we attribute this largely to two difficulties.
-Use of effort data requires making an assumption about the rate at which the fishing power of a unit of fishing effort has changed over time.
+While we did include effort reconstructions in our analysis, and did not find it to provide substantial improvements in assessment results at the macro scale, we attribute this largely to two difficulties.
+
+First, use of effort data requires making an assumption about the rate at which the fishing power of a unit of fishing effort has changed over time.
 @rousseau2019 used a fixed global rate of 2.6% to describe the increasing trend in fishing power, but as shown in their paper and in our analysis, any estimates of stock status using these effort data will be critically sensitive to the value of this parameter.
-Global studies of technology creep could help inform more realistic values and functional forms for the evolution of fishing skill in particular fisheries and regions.
-Secondly, the effort reconstructions as employed here are still at very coarse scales, requiring us to apply one effort series for all the stocks within a given region.
-Going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
+Global studies of technology creep could help inform more realistic values and functional forms for the evolution of fishing skill in particular fisheries and regions, and also require accounting for regional complications including density-dependent catchability [@thorson2010].
+Second, the effort reconstructions as employed here are still at very coarse scales, requiring us to apply one effort series for all the stocks within a given region.
+Going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
+Along with supporting methods used here, access to fine-scale effort dynamics data can help inform data-limited assessment based on catch or length-composition data XX @thorson2013;@ovando2018XX
 
 Expanded training of fisheries scientists around the globe is another critical need.
-Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and critically evaluate the results of any model based on local expertise.
-This is why stock assessments even in the data rich world are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
+Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise.
+This is why stock assessments even in the data-rich world are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
 Empowering a global network of fisheries scientists through training and peer-support would both improve the health and management of local stocks and provide a means for ensuring that global estimates of stock status are as accurate as they can be.
@@
 The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
-However, this process still leaves a substantial number of fisheries and global catch unassessed xx still need to dial in the numbers here regarding what proportion of global catch is directly covered by SOFIA assessments xx.
+However, this process still leaves a substantial number of fisheries and global catch unassessed.
 Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a clear solution to this problem.
-We demonstrate here that the bottleneck to improving estimates of global stock status depends on investing in an improved global network of fishery data and fisheries scientists.
+Improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists.
 
@@
 We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus.
-Below we describe the structure of the population model underpinning `sraplus`, the estimation model, and a brief example demonstrating the use of the model.
+Below we describe the structure of the population model underpinning `sraplus`, the estimation models used, and the construction of priors used in this paper.
 
@@
   \end{cases}
-  (\#eq:hockey)
+  (\#eq:pt)
 \end{equation}
 
-Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of B~msy~/K to shift.
-When *m* is two Bmsy / K = 0.5 and the model acts as a Schaefer model, lower values of m shift the production function left, higher values right.
-*c* is catch, and *p* is process error.
+Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of B~MSY~/K to shift.
+When *m* is two B~MSY~ / K = 0.5, lower values of *m* shift the production function left, higher values right.
+$\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors.
 Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
@@
 Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e.
-results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model.
+results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches.
+This step updates the prior distribution XX of population parameters XX by eliminating combinations of priors that are impossible for a given catch history.
 
@@
 
+[@thorson2017c]
+
 ```{r param-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
-  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
-  rbind(c("Shape parameter", "m", "Drawn from @horson2012")) %>%
-  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
-  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.05,0.5)$")) %>%
-  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$")) %>% 
-  rbind(c("Initial State", "$B0$", "$logn(0.025,0.2)$"))
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "Prior predictive tuning") %>%
+  rbind(c("Growth rate", "r", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
+  rbind(c("Shape parameter", "m", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
+  rbind(c("Catchability", "q", "$logn(1e^{-3}, 1)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,1)$")) %>%
+  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.5,0.25)$")) %>%
+  rbind(c("Initial State", "$B0$", "Posterior probability dist. of catch-based regressions"))
 
@@
   cd,
-  "latex",
   align = "l",
@@
   row.names = F,
-  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
+  caption = "Name, abbreviations, and priors distribution for parameters potentially estimated by sraplus in this manuscript.",
+  format = "pipe"
 ) %>%
@@
 The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
-It is critical to note that a stock reduction analysis is only a mechanism for revealing implicit priors on fishery status and parameters through a sample-importance-resampling algorithm: the model cannot learn anything new about stock status in the absence of additional data to fit to, beyond the prior on stock status that is fed to it.
-The usefullness of this process then is in filling in the gap of parameters of interest (e.g.
-maximum sustainable yield, MSY), and providing historic values of stock status.
-Any model results based on data sources listed in Table.\@ref(tab:dat-description) that do not contain "cpue" or "RAM data" is estimated through stock reduction analysis.
+
+It is critical to note that a stock reduction analysis is only a mechanism for revealing implicit priors on fishery status and parameters through a sample-importance-resampling algorithm: the model cannot learn anything new about stock status in the absence of additional data to fit to, beyond the prior on stock status that is fed to it and the constraints imposed by structural assumptions.
+The usefulness of this process then is in filling in the gap of parameters of interest (e.g.
+maximum sustainable yield), and providing historic values of stock status.
+Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" is estimated through stock reduction analysis.
+All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
 
 When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018].
+By default the model uses 2000 draws with a 1000 step warm-up and one chain.
+Any detailed fit for a particular fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria.
 The model fits to a direct estimate of abundance (e.g.
@@
 
-$$log(\pmb{A_t}) \sim normal(f(r,k,m,B0,\pmb{p},\pmb{C}) \times q, sigma_{obs{}}) $$
-where *f* is the Pella-Tomlinson produciton model (Equation 1.XX).
+$$log(a_t) \sim normal(f(r,k,m,B0,\pmb{p},\pmb{c}) \times q, sigma_{obs})$$
+
+where *f* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
 When an effort index is available, `sraplus` constructs an index of abundance based on the catch and effort data.
-@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6\% technology rate).
-This assumes that every unit increase in effort is linearly greater than the unit of effort before it.
+@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6\% annualized technology rate).
+This assumes that every unit increase in effort is log-linearly greater than the unit of effort before it.
 When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
@@
 
-$$cpue_t = \frac{catch_t}{(1 - e^{-q_tE_t})}$$
+$$cpue_t = \frac{catch_t}{(1 - e^{-f_t})}$$
+
+$$f_t = q_tE_t$$
 
@@
 
-$$log(cpue_t) \sim normal(f(r,k,m,B0,\pmb{p},\pmb{C}), sigma_{obs{}})$$
+$$log(cpue_t) \sim normal(f(r,k,m,B0,\pmb{p},\pmb{C}), \sigma_{obs})$$
+
+### CMSY
+
+In addition to the results from `sraplus`, we include a set of results produced by the CMSY method [@froese2017].
+For computational efficiency, we used a ported version of the CMSY model available at https://github.com/DanOvando/portedcmsy.
+The only modification made is to convert the underlying population model to C++ for faster computation.
+For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [boettiger2012].
+Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
 
@@
 The shape parameter is usually not reliably estimable given available data for surplus production models, however, @thorson2012 provides estimate the ratio of B~MSY~ to K for many fish taxa.
-While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter drawn from @thorson2012 based on the genus of the species in question.
+While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on @thorson2012 for on the genus of the species in question.
 
@@
 We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0 and less than or equal to carrying capacity *k*.
+Beyond that though, baring additional information we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0.
 
@@
 Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
-Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of r and k together with the model produces a deterministic stock status in each time step.
+Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *k* together with the model and the catch history produces a deterministic stock status in each time step.
 Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox.
@@
 
-We present a novel approximation to this problem here.
+We present an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000].
 Our solution amounts to a two-step SIR algorithm.
@@
 biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
-                                      estimate_shape = TRUE)
+                                      estimate_shape = FALSE)
 
@@
 co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
-                                      estimate_shape = TRUE)
+                                      estimate_shape = FALSE)
 
@@
 
-Along with allowing users to supply their own priors, the `sraplus` package contains three classes of built in methods for converting information on stock status from additional outside data into a form usable as a prior by `sraplus`.
+Along with allowing users to supply their own priors, the `sraplus` package contains three built-in methods used in this manuscript for converting information on stock status from additional outside data into a form usable as a prior by `sraplus`.
 We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
-We then train a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
+We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
 Given values of these variables for a new fishery then, `sraplus` uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by `sraplus`.
 
-All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [xx citation].
-These models are intended as a reasonably robust means of translating available data into usable prior for our model.
-Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with fishery status indicators of interest.
-Rather, each regression was tested to ensure that it is unlikely that, given the same data, and alternative model would be unlikely to perform much better than those presented here.
+All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
+The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by `sraplus`.
+Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with the fishery status indicators of interest.
+Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.
 
@@
 
-Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
-While these catch-only methods have been shown to have series shortcomings [@free2020], as a point of reference we use two forms of catch-only priors: a statistical and CMSY, as presented implemented by @palomares2020.
-
-For the CMSY based methods, we used a ported version of the CMSY model available [here](https://github.com/DanOvando/portedcmsy).
-For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [boettiger2012].
-Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
-
-For the statistical approach, we utilize data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
+Many of the current methods for estimating global stock status of unassessed stocks are based on predicting stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
+While these catch-only methods have been shown to have serious shortcomings [@free2020], we include them as a point of reference given their ubiquity in the global assessment literature.
+
+We used data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
 To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
-Emergent clusters show for example one build around "one way trip" style catch histories, others with a boom and bust, etc.
+Emergent clusters show for example one built around a downward "one way trip" style catch histories, others with a boom and bust pattern, others with stable but fluctuating catches.
 
@@
 
-$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma) $$
+$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma)$$
 
@@
 $$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
+
 where *fyear* is the year of the fishery, starting from 0.
@@
 
-## Example Use of `sraplus`
-
-`sraplus` can be used in two general ways: as a mechanism for revealing implicit priors on fishery status through a sample-importance-resampling algorithm, and as a means for fitting indices of abundance informed by priors.
-To illustrate we will walk through a series of ways that `sraplus` could be applied to a case study fishery.
-
-What happens when we impose informative priors on the population parameters?
-For example at a bare minimum, we know that carrying capacity must be less than infinity, and we can certainly make inferences based on knowledge from other better studied fisheries of similar species on what a potential upper bound on carrying capacity might be.
-Once we have generated informative priors on our population parameters, we have implicitly provided priors on the biomass and therefore status of the fishery over its entire history: if our priors on carrying capacity only extend a small fraction beyond the minimum value needed to ensure the existence of the population, our implicit prior is that the population is severely depleted.
-If our prior is that carrying capacity is orders of magnitude greater than the observed catches, then our implicit prior is that the population is relatively unexploited.
-
-For our case study fishery, we can use `sraplus` as a filter to say, conditional on our priors, the catch history, and the population model, what is our implied prior on stock status of the fishery over time (see Section XX and for descriptions of model parameters).
-While user defined priors on population parameters can be supplied, `sraplus` has a number of built in mechanisms for setting parameters based on the taxonomy of the species in question and the magnitude of the catch history.
-This form of analysis is often extended though to include an explicit prior on stock status in the most recent year [@walters2006].
-
-For our case study, expert opinion tells us that the stock is likely between XX and XX% of carrying capacity in the most recent year.
-While we can use `sraplus` to include this addition information, doing so introduces a nontrivial statistical problem: we have now placed two priors on the same outcome (current stock status), one explicit and one implicit in our growth parameter priors.
-The net effect of this is that once these priors are run through `sraplus`, our post-model-pre-data "estimate" of stock status will appear to have been updated, creating an illusion that we have learned something about stock status.
-However, we have absolutely no data with which to update our prior on stock status beyond the requirement that the population exists.
-
-`sraplus` implements an approximate solution to this problem.
-Rather than using combinations of implied and explicit priors, `sraplus` asks, given an explicit prior on recent stock status, what prior distributions of population parameters produce our explicit prior on recent stock status, given the other prior distributions, catch history and model assumptions.
-This helps resolve some of the statistical issues related to placing multiple priors on the same parameter (recent stock status), and makes clear where any information on recent stock status is coming from.
-To return to our case study then, uncorrected, combining our explicit priors on recent stock status with our population parameter priors reveals an implicit prior on recent stock status that spans XX to XX.
-Using our correction, the implicit prior on recent stock status more or less matches our explicit prior (Fig.XX).
-
-While `sraplus` lets users supply their best estimate of recent stock status for a given fishery, the model also allows users to supply data on the swept area ratio (SAR , @amoroso2018) and/or fisheries management index (FMI, @melnychuk2017) scores for a given fishery.
-SAR is a measure of the trawling intensity in a given region, and as such for species susceptible to trawl gears can be an effective index of fishing pressure.
-FMI reflects expert opinion on the research, management, enforcement, and socioeconomic capacity of a given fishery, which together have significant associations with stock status.
-`sraplus` uses relationships between these SAR and FMI scores and observed F/F~MSY~  and B/B~MSY~ values to generate a posterior predictive distribution of F/F~MSY~ or B/B~MSY~ for the fishery in question, which serves as the priors for `sraplus`.
-This process lets us convert information on stock status present in other available data into a form that aligns with the model structure of `sraplus`.
-
-To continue our case study, suppose that we no longer believe we have good data on current stock status, but do have a reasonable estimate of swept area ratio.
-`sraplus` can now use the posterior predictive $\hat{F/F~MSY~}$ generated by the SAR value as the prior for recent F/F~MSY~ for the fishery in question.
-This process allows users to translate something that they can measure, SAR, into something that they perhaps cannot (F/F~MSY~).
-The model then determines the growth parameter priors that conditional on the catch history and model structure produce the SAR generated $\hat{F/F~MSY~}$ prior distribution.
-
-Up to this point we still have not fit the model to any data; sraplus is simply helping us visualize the implications of our priors in terms of stock status over time and relevant reference points such as MSY.
-In order to actually learn something new about the stock in question, we need to update our priors with data.
-`sraplus` allows for two methods for accomplishing this: the model can be fit to an observed index of abundance (for example a fishery independent survey), or fit to a catch-per-unit effort index calculated from the observed catch and an index of fishing effort.
-In the latter case, the model must make an assumption about the rate of efficiency increase of a unit of fishing effort over time.
-When data are available the model can be provided priors on recent fishery status derived from expert opinion or FMI/SAR values as well.
-
-We use a Bayesian methods as it suits the philosophy of our analysis: incorporating prior knowledge about stock status drawn from expert opinion or other similar stocks, and where possible confronting those priors with data.
-However, this process does introduce a problem, in that it a) certain parts of the prior probability distribution have zero support in the data (e.g.
-carrying capacity's so small that they crash the population) and b) they create priors on both the *input* (e.g.
-growth rates and carrying capacity) and *outputs* (recent stock status), effectively putting two prior distributions for the same outcome, a problem termed Borel's Paradox XX punt citation.
-This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, particularly for the data-less version of our model, Borel's Paradox poses a particular problem.
-
-Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
-Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of r and k together with the model produces a deterministic stock status in each time step.
-Suppose then we wish to also provide a prior on stock status in the most recent year, for example to reflect local expert opinion based on sound reasoning.
-We have now placed two priors on stock status in the most recent year, one implicit and one explicit, violating Borel's Paradox.
-The net effect of this is that once these priors are run through the model, through for example a sample-importance-resampling (SIR) algorithm as generally applied in stock-reduction analysis, our post-model-pre-data "estimate" of stock status will appear to have "updated".
-This creates the illusion that we have learned something about stock status, when in fact we have simply used our model to reveal our true implied prior on stock status.
-This is technically incorrect, since we do not have a likelihood in the model at this point, and as such per Bayes Theorem have not updated our prior.
-Of more practical importance, this "implied prior" will always reflect a more optimistic stock status then our original prior on stock status, due to the constraints of the population model.
-
-To resolve this, @puntXX (either cite that "on assessment of the bering-chukchi-beufort sea stock of bowhead whales paper that I can't find a date for, or see if he has a new paper on thisXX) suggests a "backwards" fitting approach, in which the model estimates growth parameters and a terminal depletion, and given the model backs out a value of *k* that satisfies the priors (conditional on an assumption initial stock status).
-We do not adopt this approach here due to computational issues.
-Adopting the "backwards" approach requires numerically searching for the plausible *k* during each iteration of the model, which is time consuming.
-In addition, the backwards approach is not without its own complications, for example the requirement of a tolerance around the value of terminal depletion produced by a given *k* and the actual value selected by the model in that time step.
-This discrepancy can occur when a given terminal stock status is not technically possible given the model; to give an extreme example under this model no value of *k* can produce a terminal stock status of 0 if catches are 0 for the entire history of the fishery (assuming *k* > 0).
-
-We present a novel approximation to this problem here.
-Our solution amounts to a two-step SIR algorithm.
-We first run the standard SIR algorithm as described above.
-We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
-
-$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
-
-We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
-
-$$p(n_i) = \frac{p(bin_i)}{N_i}$$
-
-And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
-The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
-In effect, this answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
-
-<!-- @raftery1993 -->
-
 ## Value of Information Calculations
 
-Details of value of information calculates (fitting to messed up versions of RAM) XX need to add this in XX
-
-# Supplemental Information
-
-## Population Model
-
-The core of our model is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018.
-While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
-
-```{r si-prior-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
-  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
-  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
-  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
-  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
-  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))
-
-knitr::kable(
-  cd,
-  "latex",
-  align = "l",
-  booktabs = TRUE,
-  escape = F,
-  row.names = F,
-  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
-) %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
-
-The population growth equation is
-
-\begin{equation}
-  f(x)=\begin{cases}
-      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
-, & \text{if $B_t>0.25 \times K$}.\\
-     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
-  \end{cases}
-  (\#eq:sihockey)
-\end{equation}
-
-Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of Bmsy/K to shift (when *m* is 2 the Bmsy / K = 0.5 and the model acts as a Schaefer model, lower values of m shift the production function left, higher values right), *c* is catch, and *p* is process error.
-The Pella-Tomlinson model does have an issue where growth rates can become unrealistically large when the population reaches low sizes.
-We deal with this problem by following the methods described in @winker2018.
-to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
-Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see @thorson2018).
-
-<!-- When the SIR method is used we use the supplied catch history.
-However, for the maximum likelihood approach, convergence is greatly facilitated by estimating catches ($\hat{c_t}$) rather than using the observed catches directly.
-$\hat{c_t}$ is given as -->
-
-<!-- \begin{equation} -->
-<!--   \hat{c_t} = u_{t}B_{t} -->
-<!--   (\#eq:chat) -->
-<!-- \end{equation} -->
-
-<!-- where $u_t$ is estimated through -->
-
-<!-- \begin{equation} -->
-<!--   u_{t} = \frac{1}{1 + e^{-u^*_t}} -->
-<!--   (\#eq:invf) -->
-<!-- \end{equation} -->
-
-<!-- Equation \@ref(eq:invf) ensures that $u_t$ is between 0 and 1, while the model estimates $u^*_t$ to improve performance.  -->
-
-<!-- $\hat{c_t}$ is then fit to the observed catches through the likelihood -->
-
-<!-- \begin{equation} -->
-<!--   log\left(c_t\right) \sim normal\left(log\left(\hat{c_t}\right),0.05\right) -->
-<!--   (\#eq:catch) -->
-<!-- \end{equation} -->
-
-<!-- Where we assume an observation coefficient of variation of 5\%.
-We estimate catch rather than directly using the observed catch in order to improve model performance.
-As priors on depletion, or indices of abundance, pull the estimated population closer to zero, there is a greater and greater chance that the observed catch history will result in a collapse of the population (catch in time *t* being greater than biomass in time *t*); It is much harder to find parameters that produce a final depletion of 1\% of carrying capacity while not crashing the population prior to then given a catch history, than it is to produce a final depletion of 99\% of carrying capacity.
-We can impose restrictions and penalties on population crashes, but this distorts the gradients of the model, causing the model to often become stuck once a few crashes have occurred.
-Since $u_t$ is constrained between 0 and 1, $\hat{c_t}$ will always be "viable", and by then fitting to the observed catches, the model has an easier time finding parameters that produce a near match to the observed catches while not crashing the population.
-As of now none of the data used in this exercise have missing years of catch data, but the model can handle those cases in the future due to the random walk priors imposed on $u_t$ detailed below.
--->
-
-<!-- In order to help separate out observation error in the catch and process error in the population model, when catch is being estimated we assume that fishing mortality $u_t$ follows a random walk on average, assigning each $u_t$ a prior of -->
-
-<!-- \begin{equation} -->
-<!--   log(u_t) \sim normal(log(u_{t - 1}),0.1)) -->
-<!--   (\#eq:rwalk) -->
-<!-- \end{equation} -->
-
-We  allow for process error (in the manner of the stochastic stock reduction analysis error suggested by @walters2006).
-This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
-Process error *p* is assumed to be log-normally distributed, such that
-
-\begin{equation}
-  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
-  (\#eq:siprocerror)
-\end{equation}
-
-All of our estimates are Bayesian in nature.
-We can break the use of our model into two distinct categories: with data and without.
-By "data", we refer to measurements which are confronted with model estimates within a likelihood function.
-In our context, these include fishery-independent survey data, or a CPUE index.
-When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
-Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population, given the constraints of the model.
-
-We use a Bayesian methods as it suits the philosophy of our analysis: incorporating prior knowledge about stock status drawn from expert opinion or other similar stocks, and where possible confronting those priors with data.
-However, this process does introduce a problem, in that it a) certain parts of the prior probability distribution have zero support in the data (e.g.
-carrying capacity's so small that they crash the population) and b) they create priors on both the *input* (e.g.
-growth rates and carrying capacity) and *outputs* (recent stock status), effectively putting two prior distributions for the same outcome, a problem termed Borel's Paradox XX punt citation.
-This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, particularly for the data-less version of our model, Borel's Paradox poses a particular problem.
-
-Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
-Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of r and k together with the model produces a deterministic stock status in each time step.
-Suppose then we wish to also provide a prior on stock status in the most recent year, for example to reflect local expert opinion based on sound reasoning.
-We have now placed two priors on stock status in the most recent year, one implicit and one explicit, violating Borel's Paradox.
-The net effect of this is that once these priors are run through the model, through for example a sample-importance-resampling (SIR) algorithm as generally applied in stock-reduction analysis, our post-model-pre-data "estimate" of stock status will appear to have "updated".
-This creates the illusion that we have learned something about stock status, when in fact we have simply used our model to reveal our true implied prior on stock status.
-This is technically incorrect, since we do not have a likelihood in the model at this point, and as such per Bayes Theorem have not updated our prior.
-Of more practical importance, this "implied prior" will always reflect a more optimistic stock status then our original prior on stock status, due to the constraints of the population model.
-
-The SRA algorithm works in two steps.
-First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep).
-From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status, e.g.
-if the bulk of the prior on terminal stock status was concentrated at 50% of *k*, combinations of *r* and *k* that produce terminal stock status near 50% of *k* are sampled proportionally more frequently.
-However, lower values of terminal stock status have fewer candidate values of *r* and *k*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
-Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *k* combinations: since under this model the population cannot be greater than carrying capacity, as for example *k* approaches infinity terminal stock status asymptotes at close to 100% of *k*.
-The net result of this is that even though individual combinations of *r* and *k* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
-As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.
-
-To resolve this, @puntXX (either cite that "on assessment of the bering-chukchi-beufort sea stock of bowhead whales paper that I can't find a date for, or see if he has a new paper on thisXX) suggests a "backwards" fitting approach, in which the model estimates growth parameters and a terminal depletion, and given the model backs out a value of *k* that satisfies the priors (conditional on an assumption initial stock status).
-We do not adopt this approach here due to computational issues.
-Adopting the "backwards" approach requires numerically searching for the plausible *k* during each iteration of the model, which is time consuming.
-In addition, the backwards approach is not without its own complications, for example the requirement of a tolerance around the value of terminal depletion produced by a given *k* and the actual value selected by the model in that time step.
-This discrepancy can occur when a given terminal stock status is not technically possible given the model; to give an extreme example under this model no value of *k* can produce a terminal stock status of 0 if catches are 0 for the entire history of the fishery (assuming *k* > 0).
-
-We present a novel approximation to this problem here.
-Our solution amounts to a two-step SIR algorithm.
-We first run the standard SIR algorithm as described above.
-We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
-
-$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
-
-We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
-
-$$p(n_i) = \frac{p(bin_i)}{N_i}$$
-
-And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
-The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
-In effect, this answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
-
-XX insert figure if there's space illustrating this.
-
-<!-- @raftery1993 -->
-
-```{r fao-acc-map, fig.cap = "Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
-ram-data refers to catch and abundance index drawn from RAM.
-CPUE refers to an index of abundance based on the Rousseau effort data.
-Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
-The FAO Report 569 is our reference status estimate provided by the FAO.
-ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
-fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio"}
-
-fao_acc_map_plot
-
-```
+The value-of-information (VOI) calculations presented in Fig.\@ref(fig:voi-plot) help illustrate what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
+The VOI analysis was performed by using `sraplus` to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
+We generate fits for 3000 combinations of a RAM stock and available data.
+For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
+For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
+The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
+
+Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
+We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
+
+$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
+
+Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
+
+# References
+<div id="refs"></div>
+
+# Supplemental Material
 
 ```{r}
-
-rand_fits <- fao_sraplus_comp_data %>% 
-  filter(data == "catch_only") %>% 
-  mutate(bin = sample(c("over", "fully", "under"), n(), replace = TRUE),
-         data = "guess") %>% 
-  mutate(correct = bin == fao_bin)
-
-fao_sraplus_acc <- fao_sraplus_comp_data %>% 
-  bind_rows(rand_fits) %>% 
-  group_by(data) %>% 
-  summarise(accuracy = mean(correct, na.rm = TRUE)) %>% 
-  arrange(desc(accuracy))
-
-knitr::kable(fao_sraplus_acc,
-             caption = "Accuracy of different data sources in classifying stocks reported in the FAO SOFIA report to FAO status categories.
-Guess corresponds to randomly assigning a stock to one of the FAO status categories.",
-             format = "pipe", 
-             digits = 2)
-
-```
-
-```{r}
-
-rand_fits <- fao_sraplus_comp_data %>% 
-  filter(data == "catch_only") %>% 
-  mutate(bin = sample(c("over", "fully", "under"), n(), replace = TRUE),
-         data = "guess") %>% 
-  mutate(correct = bin == fao_bin)
-
-fao_sraplus_area_acc <- fao_sraplus_comp_data %>% 
-  bind_rows(rand_fits) %>% 
-  group_by(data, fao_area_code) %>% 
-  summarise(accuracy = mean(correct, na.rm = TRUE)) %>% 
-  arrange(desc(accuracy)) %>% 
-  pivot_wider(names_from = data, values_from = accuracy) %>% 
-  arrange(fao_area_code)
-
-knitr::kable(fao_sraplus_area_acc,
-             caption = "Accuracy of different data sources in classifying stocks reported in the FAO SOFIA report to FAO status categories by FAO major statistical area.
-Guess corresponds to randomly assigning a stock to one of the FAO status categories.",
-             format = "pipe", 
-             digits = 2)
-
-```
-
-## SAR
-
-## FMI
-
-```{r fmi-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of fisheries management index (FMI) on F/F~MSY~"}
-
-fit <- sraplus::fmi_models$fit[sraplus::fmi_models$metric == "mean_uumsy"][[1]]
-
-fit_data <- fit$data
-
-pp <- posterior_predict(fit, draws = 1000)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
-  scale_x_continuous(name = "Observed Log F/Fmsy") + 
-  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-```
-
-```{r sar-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of swept area ratio (SAR) on F/F~MSY~"}
-fit <- sraplus::sar_models$fit[sraplus::sar_models$metric == "mean_uumsy"][[1]]
-
-fit_data <- fit$data
-
-pp <- posterior_predict(fit, draws = 1000)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
-  scale_x_continuous(name = "Observed Log F/Fmsy") + 
-  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-
-```
-
-## Catch
-
-```{r catch-reg,fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) B/B~MSY~ for regression of catch on B/B~MSY~"}
-
-fit <- sraplus::catch_b_model
-
-fit_data <- fit$data %>% 
-  group_by(stockid) %>% 
-  filter(year == max(year))
-
-pp <- posterior_predict(fit, draws = 1000, newdata = fit_data)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_scatter_avg(y = fit_data$log_value, yrep = pp) + 
-  coord_flip() +
-  scale_y_continuous(name = "Observed Log B/Bmsy") + 
-  scale_x_continuous(name = "Posterior Predictive B/Bmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-```
-
-# References
+plot_prior_posterior(co_fit, co_driors)
+```

### 2020-09-10 functions/fit-case-studies.R

@@ -1,41 +1,26 @@
-fit_continent_examples <-
-  function(scientific_name,
+fit_case_studies <-
+  function(
            data,
            model = "sraplus_tmb",
-           engine = "tmb",
-           fmi_f_reg = NA,
-          estimate_shape = TRUE,
-          estimate_proc_error = TRUE,
-          estimate_qslope = FALSE,
-          use_baranov = TRUE,
-          first_effort_year = 1950,
-          chains = 1,
-          cores = 1,
-          cmsy_cores = 4,
-          q_slope_prior = 0.025) {
-    # heuristic example
-    # i = 2
-    # engine = "tmb"
-    # model = "sraplus_tmb"
-    # data <- huh$data[[1]]
-    # scientific_name <- huh$scientific_name[[1]]
-    # catches <- data$total_catch
-
-    # estimate_proc_error = FALSE
-    # estimate_qslope = TRUE
-    # estimate_shape = TRUE
-    # use_baranov = TRUE
-    # first_effort_year = 1950
-
-    
-    catches <- data$total_catch
+           engine = "stan",
+           estimate_shape = TRUE,
+           estimate_proc_error = TRUE,
+           estimate_qslope = FALSE,
+           use_baranov = TRUE,
+           first_effort_year = 1950,
+           chains = 1,
+           cores = 1,
+           cmsy_cores = 4,
+           q_slope_prior = 0.025) {
+ 
+    catches <- data$catch
     
     years <- data$year
     
- 
-  
+    scientific_name <- data$scientificname[[1]]
+    
     # fit cmsy
-  
+    
     sfc <- purrr::safely(portedcmsy::funct_cmsy)
     
     ported_cmsy <- sfc(
@@ -56,16 +41,7 @@ fit_continent_examples <-
         select(year, bt, b_bmsy, f_fmsy, c_div_msy) %>% 
         rename(depletion = bt, b_div_bmsy = b_bmsy, u_div_umsy = f_fmsy) %>% 
         pivot_longer(-year, names_to = "variable", values_to = "mean") %>% 
-        mutate(sd = NA, lower = NA, upper = NA, data = "cmsy")
-      
-      # tests <- cmsy_results %>% 
-      #   bind_rows(basic_fit_results)
-      # 
-      # tests %>% 
-      #   filter(variable == "c_div_msy") %>% 
-      #   select(year, mean, data) %>% 
-      #   ggplot(aes(year, mean, color = data)) + 
[…]

### 2020-09-10 make-assessing-global-fisheries.R

@@ -65,19 +65,15 @@ write_results <- FALSE
 
 process_fits <- FALSE
 
-# other run options things
-
-run_continent_examples <- FALSE
-
-run_ei_example <- FALSE
+run_case_studies <- FALSE
 
 run_sofia_comparison <- FALSE
 
 run_ram_tests <- FALSE
 
-run_ram_comparison <- FALSE
+run_ram_comparison <- TRUE
 
-knit_paper <- TRUE
+knit_paper <- FALSE
 
 warning("Running full analysis takes upwards of 24 hours on 2 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 2 cores is not recommended (memory routinley runs out on a 16 core 36GB machine when cores > 2)")
 
@@ -166,12 +162,12 @@ if (run_voi_models == TRUE) {
     initial_state_cv =  runif(draws, 0.05, 0.2),
     use_terminal_state = sample(c(TRUE, FALSE),  draws, replace = TRUE),
     terminal_state_cv = runif(draws, 0.05, 0.2),
-    error_cv = sample(c(0, 1), draws, replace = TRUE),
+    error_cv = sample(c(0), draws, replace = TRUE),
     stockid = sample(unique(ram_data$stockid), draws, replace = TRUE),
     use_u_priors =  sample(c(TRUE, FALSE),  draws, replace = TRUE),
     u_cv = runif(draws, 0.05, 0.2),
     index_window = sample(c(1, 0.25, .5), draws, replace = TRUE),
-    index_freq = sample(c(1, 2), draws, replace = TRUE),
+    index_freq = sample(c(1), draws, replace = TRUE),
     u_window = sample(c("snapshot", "recent", "complete"), draws, replace = TRUE),
     b_ref_type = sample(c("k", "b"),  draws, replace = TRUE),
     f_ref_type = sample(c("f", "fmsy"),  draws, replace = TRUE),
@@ -250,7 +246,7 @@ if (run_voi_models == TRUE) {
           estimate_initial_state = !initial_state_type == "unfished"
         ),
         safely(run_assessments),
-        assessments = c("cmsy", "sraplus"),
+        assessments = c("sraplus"),
         model = "sraplus_tmb",
         results_name = results_name,
         write_results = write_results,
@@ -431,10 +427,23 @@ if (run_voi_models == TRUE) {
   
   
[…]

### 2020-09-10 documents/ovando-etal-assessing-global-fisheries.Rmd
jims edits

@@
 Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random quess.
-Our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
+Obtaining accurate estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of high-priority data at a global scale, not on the development of new modeling methods alone.
 always_allow_html: true
+linkcolor: blue
 ---
@@
 
-XX discussion with Jim.
-How can we frame this.
-I would actually push back the idea that we have learned about stock status.
-I buy that the catch data can update your prior on growth parameters, but by the rules we're playing with here, it shouldn't update our prior on stock status should it.
-Now, the only thing I can't wrap my head aroudn is the case where catches are all zero, and you have a prior that the stock is collapsed.
-I guess my question is from a strikely Bayes theorem perspective, where in here is your prior on stock status being updated, since there's no likelihood as written, except for the binomial process that says that the stock exists.
-
 For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index and swept area ratio scores.
@@
 We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`.
-It is critical to note that in both of these cases the model has learned nothing new about recent stock status that was not encoded in the supplied priors, catch data, and functional form of the model; `sraplus` simply helps reveal our implicit priors on other parameters (e.g.
-maximum sustainable yield, MSY), and provides timeseries of measures such as B/B~MSY~.
-That, in the absence of data to actually fit to such as an index of abundance, methods such as CMSY or sraplus can learn anything new about a stock's status that is not already implicitly coded in the priors, catch data, or specified functional relationships, is a common misconception (see [Methods](#Methods) a for detailed explanation of why this is).
+It is critical to note that in both of these cases the catch data on its own contains no information about stock status; we can only produce estimates of stock status by combining catch data with explicit priors and specified functional relationships [@thorson2013].
+This is important since it means that we must look to our priors and functional forms to understand where our insight stock status is coming from, not the catch data alone.
 Lastly, we used the reconstructed effort data from @rousseau2019 to create an index of abundance for each stock (see [Estimation Model](#Estimation Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.
@@
 Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world.
-While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models using each of the available datasets beside the RAM-derived indices produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).
+While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).
 
@@
 Going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
-Along with supporting methods used here, access to fine-scale effort dynamics data can help inform data-limited assessment based on catch or length-composition data XX @thorson2013;@ovando2018XX
+Along with supporting methods used here, access to fine-scale effort dynamics data can help inform data-limited assessment based on catch or length-composition data [@ovando2018; @thorson2013].
 
@@
 results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches.
-This step updates the prior distribution XX of population parameters XX by eliminating combinations of priors that are impossible for a given catch history.
+This step updates the prior distribution of population parameters by eliminating combinations of priors that are impossible for a given catch history and a specified functional form.
 
@@
 When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user.
-For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *k*, *m*, $\sigma_{obs}$, $\gamma$, and *B0*.
+For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *k*, $\sigma_{obs}$, $\gamma$, and *B0*.
 *q* is also estimated when needed.
-
-[@thorson2017c]
 
@@
 The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
-
-It is critical to note that a stock reduction analysis is only a mechanism for revealing implicit priors on fishery status and parameters through a sample-importance-resampling algorithm: the model cannot learn anything new about stock status in the absence of additional data to fit to, beyond the prior on stock status that is fed to it and the constraints imposed by structural assumptions.
-The usefulness of this process then is in filling in the gap of parameters of interest (e.g.
-maximum sustainable yield), and providing historic values of stock status.
-Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" is estimated through stock reduction analysis.
+Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis.
 All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
@@
 Priors for all estimable parameters can be left at default values or set by the user.
-The shape parameter is usually not reliably estimable given available data for surplus production models, however, @thorson2012 provides estimate the ratio of B~MSY~ to K for many fish taxa.
-While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on @thorson2012 for on the genus of the species in question.
+The shape parameter is usually not reliably estimable given available data for surplus production models, however, Thorson et al.
+2012 [@thorson2012] provides estimate the ratio of B~MSY~ to K for many fish taxa.
+While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on Thorson et al.
+2012 [@thorson2012] for the genus of the species in question.
 
@@
 We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0.
+Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0.
 
@@
 
-We present an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000].
+We use an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000].
 Our solution amounts to a two-step SIR algorithm.
@@
 In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
-This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot)).
+This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.S1).
 
@@
 biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
-                                      estimate_shape = FALSE)
+                                      estimate_shape = FALSE, estimate_proc_error = TRUE)
 
@@
 co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
-                                      estimate_shape = FALSE)
+                                      estimate_shape = FALSE, 
+                               estimate_proc_error = TRUE)
 
@@
 
-```{r}
-plot_prior_posterior(co_fit, co_driors)
+```{r ppplot, fig.cap="Prior-posterior plots of fits for case study fishery in Fig.6"}
+
+a = get_prior_posterior(co_fit, co_driors)
+
+a$prior_posterior %>% 
+  ggplot(aes(x = variable, y = mean, ymin = lower, ymax = upper, color = source)) + 
+  geom_pointrange(position = position_dodge(width = 0.1)) + 
+  facet_wrap(~variable, scales = "free") + 
+  theme(axis.text.x = element_blank()) + 
+  scale_color_discrete(name = '') + 
+  scale_y_continuous(name = "Value") + 
+  scale_x_discrete(name = '') + 
+  theme_minimal()
+
 ```

### 2020-10-02 documents/ovando-etal-assessing-global-fisheries.Rmd
merging back to master

@@
 These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection.
-Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random quess.
+Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random guess.
 Obtaining accurate estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of high-priority data at a global scale, not on the development of new modeling methods alone.
@@
 <!--   - here's  -->
+
+# check in call notes
+
+Ray: setting the stage: what we want is a way to assess the status of stocks and see how we're doing relative to SDG.
+
+We need to discuss a sampling strategy: how can we get down to a set of priority stock. A step of expert knowledge to refine what stocks you would focus in on.
+
+Regressions based approach, focusing in effort on stocks
+
+Basically a paragraph laying out that we need to do this, and discussing how we might go about doing that. Equal weight to each stock vs proportional to biomass
+
+add in a bit more to the discussion on guidelines
+
+add in error in the index for the VOI analysis
+
+Apples and apples comparisons between model and design developments.
+How much is the value of the effort data relative to actually improving the population model.
+Production model is ill-suited to some of these questions.
+
+fishing effort uncertainty of the effort
+
+how credible is the information
+
+probably don't need to bother with the SAUP
+
+stock assessments
+
+the stock will collapse next year
+
+nansen data
 
@@
 Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have previously had stock assessments.
+xx bring this home xx
 
 For many areas of the world effective fishery independent estimates of stock abundance are unlikely to be feasible in the near future.
-Effort databases such as @rousseau2019, can be combined with catch histories to to construct fishery-dependent abundance indicies.
+XXbe more specific here, and clearXX.
+.Effort databases such as @rousseau2019, can be combined with catch histories to to construct fishery-dependent abundance indicies.
 While there are substantial difficulties and biases that can arise from improper use of fishery-dependent abundance data, improved understanding of the evolution of global fishing effort could be very useful in improving global estimates of fishery health.

### 2020-10-02 make-assessing-global-fisheries.R

@@ -25,6 +25,7 @@ library(rstan)
 library(rstanarm)
 library(sf)
 library(portedcmsy)
+extrafont::loadfonts()
 
 Sys.unsetenv("PKG_CXXFLAGS")
 
@@ -450,21 +451,21 @@ if (run_voi_models == TRUE) {
     filter(fit_name != "cmsy")
   
   
-  voi_data %>%
-    filter(use_index == FALSE,
-           metric == "b_v_bmsy",
-           b_ref_type == "b") %>%
-    ggplot(aes(rmse, fill = use_terminal_state)) +
-    geom_density(position = "dodge")
+  # voi_data %>%
+  #   filter(use_index == FALSE,
+  #          metric == "b_v_bmsy",
+  #          b_ref_type == "b") %>%
+  #   ggplot(aes(rmse, fill = use_terminal_state)) +
+  #   geom_density(position = "dodge")
   
   
   
-  noerror_data %>%
-    filter(use_index == FALSE,
-           metric == "b_v_bmsy",
-           b_ref_type == "b") %>%
-    ggplot(aes(observed, predicted, color = use_terminal_state)) +
-    geom_point()
+  # noerror_data %>%
+  #   filter(use_index == FALSE,
+  #          metric == "b_v_bmsy",
+  #          b_ref_type == "b") %>%
+  #   ggplot(aes(observed, predicted, color = use_terminal_state)) +
+  #   geom_point()
   
   obs_v_pred_plot <- noerror_data %>%
     filter(metric == "b_v_bmsy") %>%
@@ -757,9 +758,9 @@ comp_stocks <- exs %>%
   select(scientificname, resilience)
 
 
-exs %>%
-  ggplot(aes(year, b_v_bmsy, color = stockid)) +
-  geom_line(show.legend = FALSE)
+# exs %>%
+#   ggplot(aes(year, b_v_bmsy, color = stockid)) +
+#   geom_line(show.legend = FALSE)
 
 total_nominal_effort <- rous_data %>%
   left_join(effort_region_to_country, by = "country") %>%
@@ -876,9 +877,9 @@ sraplus_v_truth <- tmp %>%
   left_join(truth, by = c("stockid","year")) %>% 
   filter(data == "cpue") 
 
-sraplus_v_truth %>% 
-  ggplot(aes(catch, b_div_bmsy)) + 
-  geom_point(size = 2)
+# sraplus_v_truth %>% 
[…]

### 2020-10-02 documents/si-ovando-etal-assessing-global-fisheries.Rmd
Pre internal review edits (#6)
* testing TMB fix on windows

* merging back to master

---
title: "Supplemental Information: Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::pdf_document2: default
  bookdown::word_document2:
    reference_docx: word-template.docx
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. Our results provide further evidence that catch data by themselves cannot provide reliable estimates of fishery health. Our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
library(tidyverse)
library(viridis)
library(sf)
library(hrbrthemes)
library(extrafont)
library(scales)
library(patchwork)
library(rstanarm)
library(bayesplot)
library(here)
library(sraplus)
extrafont::loadfonts()

functions <- list.files(here::here("functions"))

purrr::walk(functions, ~ source(here::here("functions", .x)))

results_name <- params$results_name

results_path <- here("results",results_name)

min_years_catch <- params$min_years_catch

prepare_sofia_data(lookup_fmi_names = FALSE)

ram_comp_data <- ram_data %>%
  mutate(has_things = !(is.na(catch) | catch == 0)) %>%
  filter(has_things) %>%
  group_by(stockid) %>%
  mutate(delta_year = as.integer(year - lag(year))) %>%
  mutate(delta_year = case_when(year == min(year) ~ as.integer(1),
                                TRUE ~ delta_year)) %>%
  mutate(missing_gaps = any(delta_year > 1)) %>%
  filter(missing_gaps == FALSE) %>%
  group_by(stockid) %>%
  mutate(n = length(catch)) %>%
  filter(n >= min_years_catch) %>%
  ungroup()

ram_comp_data <- ram_comp_data %>%
  select(
    year,
    stockid,
    country,
    primary_FAOarea,
    scientificname,
    commonname,
    year,
    catch,
    b_v_bmsy,
    u_v_umsy,
    total_biomass
  ) %>%
  rename(scientific_name = scientificname) %>% 
  mutate(fao_area_code = as.numeric(primary_FAOarea)) %>% 
  left_join(fao_species, by = "scientific_name")

assess_ram_fits <- read_rds(path = file.path(results_path,"assess_ram_fits.rds"))

ram_v_sraplus_area <- read_rds(path = file.path(results_path,"ram_v_sraplus_area.rds"))

ram_comp_data <- read_rds(path = file.path(results_path,"ram_comp_data.rds"))

fao_sraplus_comp_data <- read_rds(path = file.path(results_path, "fao_sraplus_comp_data.rds"))

load(file = file.path(results_path, "paper_plots.Rdata"))

load(file = file.path(results_path, "voi_fits.RData"))

pub_theme <- theme_ipsum(base_size = 10,
                         axis_text_size = 10,
                         axis_title_size = 12) +
    theme(panel.spacing = unit(.5,"lines"))

theme_set(pub_theme)

```

# Supplemental Information

## Population Model

The core of our model is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018. While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.

```{r si-prior-tab}
cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))

knitr::kable(
  cd,
  "latex",
  align = "l",
  booktabs = TRUE,
  escape = F,
  row.names = F,
  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
) %>%
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

The population growth equation is

\begin{equation}
  f(x)=\begin{cases}
      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
, & \text{if $B_t>0.25 \times K$}.\\
     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
  \end{cases}
  (\#eq:sihockey)
\end{equation}

Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of Bmsy/K to shift (when *m* is 2 the Bmsy / K = 0.5 and the model acts as a Schaefer model, lower values of m shift the production function left, higher values right), *c* is catch, and *p* is process error. The Pella-Tomlinson model does have an issue where growth rates can become unrealistically large when the population reaches low sizes. We deal with this problem by following the methods described in @winker2018. to reduce the production of the population when it falls below a threshold of 25% of carrying capacity. Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see @thorson2018).

<!-- When the SIR method is used we use the supplied catch history. However, for the maximum likelihood approach, convergence is greatly facilitated by estimating catches ($\hat{c_t}$) rather than using the observed catches directly. $\hat{c_t}$ is given as -->

<!-- \begin{equation} -->
<!--   \hat{c_t} = u_{t}B_{t} -->
<!--   (\#eq:chat) -->
<!-- \end{equation} -->

<!-- where $u_t$ is estimated through -->

<!-- \begin{equation} -->
<!--   u_{t} = \frac{1}{1 + e^{-u^*_t}} -->
<!--   (\#eq:invf) -->
<!-- \end{equation} -->

<!-- Equation \@ref(eq:invf) ensures that $u_t$ is between 0 and 1, while the model estimates $u^*_t$ to improve performance.  -->

<!-- $\hat{c_t}$ is then fit to the observed catches through the likelihood -->

<!-- \begin{equation} -->
<!--   log\left(c_t\right) \sim normal\left(log\left(\hat{c_t}\right),0.05\right) -->
<!--   (\#eq:catch) -->
<!-- \end{equation} -->

<!-- Where we assume an observation coefficient of variation of 5\%. We estimate catch rather than directly using the observed catch in order to improve model performance. As priors on depletion, or indices of abundance, pull the estimated population closer to zero, there is a greater and greater chance that the observed catch history will result in a collapse of the population (catch in time *t* being greater than biomass in time *t*); It is much harder to find parameters that produce a final depletion of 1\% of carrying capacity while not crashing the population prior to then given a catch history, than it is to produce a final depletion of 99\% of carrying capacity. We can impose restrictions and penalties on population crashes, but this distorts the gradients of the model, causing the model to often become stuck once a few crashes have occurred. Since $u_t$ is constrained between 0 and 1, $\hat{c_t}$ will always be "viable", and by then fitting to the observed catches, the model has an easier time finding parameters that produce a near match to the observed catches while not crashing the population. As of now none of the data used in this exercise have missing years of catch data, but the model can handle those cases in the future due to the random walk priors imposed on $u_t$ detailed below.  -->

<!-- In order to help separate out observation error in the catch and process error in the population model, when catch is being estimated we assume that fishing mortality $u_t$ follows a random walk on average, assigning each $u_t$ a prior of -->

<!-- \begin{equation} -->
<!--   log(u_t) \sim normal(log(u_{t - 1}),0.1)) -->
<!--   (\#eq:rwalk) -->
<!-- \end{equation} -->

We  allow for process error (in the manner of the stochastic stock reduction analysis error suggested by @walters2006). This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average. Process error *p* is assumed to be log-normally distributed, such that

\begin{equation}
  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
  (\#eq:siprocerror)
\end{equation}

All of our estimates are Bayesian in nature. We can break the use of our model into two distinct categories: with data and without. By "data", we refer to measurements which are confronted with model estimates within a likelihood function. In our context, these include fishery-independent survey data, or a CPUE index. When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters). Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population, given the constraints of the model.

We use a Bayesian methods as it suits the philosophy of our analysis: incorporating prior knowledge about stock status drawn from expert opinion or other similar stocks, and where possible confronting those priors with data. However, this process does introduce a problem, in that it a) certain parts of the prior probability distribution have zero support in the data (e.g. carrying capacity's so small that they crash the population) and b) they create priors on both the *input* (e.g. growth rates and carrying capacity) and *outputs* (recent stock status), effectively putting two prior distributions for the same outcome, a problem termed Borel's Paradox XX punt citation. This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood. However, particularly for the data-less version of our model, Borel's Paradox poses a particular problem.

Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*. Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of r and k together with the model produces a deterministic stock status in each time step. Suppose then we wish to also provide a prior on stock status in the most recent year, for example to reflect local expert opinion based on sound reasoning. We have now placed two priors on stock status in the most recent year, one implicit and one explicit, violating Borel's Paradox. The net effect of this is that once these priors are run through the model, through for example a sample-importance-resampling (SIR) algorithm as generally applied in stock-reduction analysis, our post-model-pre-data "estimate" of stock status will appear to have "updated". This creates the illusion that we have learned something about stock status, when in fact we have simply used our model to reveal our true implied prior on stock status. This is technically incorrect, since we do not have a likelihood in the model at this point, and as such per Bayes Theorem have not updated our prior. Of more practical importance, this "implied prior" will always reflect a more optimistic stock status then our original prior on stock status, due to the constraints of the population model.

The SRA algorithm works in two steps. First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep). From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status, e.g. if the bulk of the prior on terminal stock status was concentrated at 50% of *k*, combinations of *r* and *k* that produce terminal stock status near 50% of *k* are sampled proportionally more frequently. However, lower values of terminal stock status have fewer candidate values of *r* and *k*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step. Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *k* combinations: since under this model the population cannot be greater than carrying capacity, as for example *k* approaches infinity terminal stock status asymptotes at close to 100% of *k*. The net result of this is that even though individual combinations of *r* and *k* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled. As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.

To resolve this, @puntXX (either cite that "on assessment of the bering-chukchi-beufort sea stock of bowhead whales paper that I can't find a date for, or see if he has a new paper on thisXX) suggests a "backwards" fitting approach, in which the model estimates growth parameters and a terminal depletion, and given the model backs out a value of *k* that satisfies the priors (conditional on an assumption initial stock status). We do not adopt this approach here due to computational issues. Adopting the "backwards" approach requires numerically searching for the plausible *k* during each iteration of the model, which is time consuming. In addition, the backwards approach is not without its own complications, for example the requirement of a tolerance around the value of terminal depletion produced by a given *k* and the actual value selected by the model in that time step. This discrepancy can occur when a given terminal stock status is not technically possible given the model; to give an extreme example under this model no value of *k* can produce a terminal stock status of 0 if catches are 0 for the entire history of the fishery (assuming *k* > 0).

We present a novel approximation to this problem here. Our solution amounts to a two-step SIR algorithm. We first run the standard SIR algorithm as described above. We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.

$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$

We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*

$$p(n_i) = \frac{p(bin_i)}{N_i}$$

And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$. The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status. In effect, this answers the question "given the model, what combinations of parameters produce my prior on recent stock status".

XX insert figure if there's space illustrating this.

<!-- @raftery1993 -->

```{r fao-acc-map, fig.cap = "Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources. ram-data refers to catch and abundance index drawn from RAM. CPUE refers to an index of abundance based on the Rousseau effort data. Cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data. The FAO Report 569 is our reference status estimate provided by the FAO. ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region. fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio"}

fao_acc_map_plot

```

```{r}

rand_fits <- fao_sraplus_comp_data %>% 
  filter(data == "catch_only") %>% 
  mutate(bin = sample(c("over", "fully", "under"), n(), replace = TRUE),
         data = "guess") %>% 
  mutate(correct = bin == fao_bin)

fao_sraplus_acc <- fao_sraplus_comp_data %>% 
  bind_rows(rand_fits) %>% 
  group_by(data) %>% 
  summarise(accuracy = mean(correct, na.rm = TRUE)) %>% 
  arrange(desc(accuracy))

knitr::kable(fao_sraplus_acc,
             caption = "Accuracy of different data sources in classifying stocks reported in the FAO SOFIA report to FAO status categories. Guess corresponds to randomly assigning a stock to one of the FAO status categories.",
             format = "pipe", 
             digits = 2)

```

```{r}

rand_fits <- fao_sraplus_comp_data %>% 
  filter(data == "catch_only") %>% 
  mutate(bin = sample(c("over", "fully", "under"), n(), replace = TRUE),
         data = "guess") %>% 
  mutate(correct = bin == fao_bin)

fao_sraplus_area_acc <- fao_sraplus_comp_data %>% 
  bind_rows(rand_fits) %>% 
  group_by(data, fao_area_code) %>% 
  summarise(accuracy = mean(correct, na.rm = TRUE)) %>% 
  arrange(desc(accuracy)) %>% 
  pivot_wider(names_from = data, values_from = accuracy) %>% 
  arrange(fao_area_code)

knitr::kable(fao_sraplus_area_acc,
             caption = "Accuracy of different data sources in classifying stocks reported in the FAO SOFIA report to FAO status categories by FAO major statistical area. Guess corresponds to randomly assigning a stock to one of the FAO status categories.",
             format = "pipe", 
             digits = 2)

```

## SAR

## FMI

```{r fmi-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of fisheries management index (FMI) on F/F~MSY~"}

fit <- sraplus::fmi_models$fit[sraplus::fmi_models$metric == "mean_uumsy"][[1]]

fit_data <- fit$data

pp <- posterior_predict(fit, draws = 1000)

r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
  ggplot(aes(r2)) + 
  geom_histogram() + 
  labs(subtitle = "B) Bayesian R2")

pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
  scale_x_continuous(name = "Observed Log F/Fmsy") + 
  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
  labs(subtitle = "A) Observed vs. Posterior Predictive") +
 pub_theme

pi_plot + r2_plot
```

```{r sar-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of swept area ratio (SAR) on F/F~MSY~"}
fit <- sraplus::sar_models$fit[sraplus::sar_models$metric == "mean_uumsy"][[1]]

fit_data <- fit$data

pp <- posterior_predict(fit, draws = 1000)

r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
  ggplot(aes(r2)) + 
  geom_histogram() + 
  labs(subtitle = "B) Bayesian R2")

pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
  scale_x_continuous(name = "Observed Log F/Fmsy") + 
  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
  labs(subtitle = "A) Observed vs. Posterior Predictive") +
 pub_theme

pi_plot + r2_plot

```

## Catch

```{r catch-reg,fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) B/B~MSY~ for regression of catch on B/B~MSY~"}

fit <- sraplus::catch_b_model

fit_data <- fit$data %>% 
  group_by(stockid) %>% 
  filter(year == max(year))

pp <- posterior_predict(fit, draws = 1000, newdata = fit_data)

r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
  ggplot(aes(r2)) + 
  geom_histogram() + 
  labs(subtitle = "B) Bayesian R2")

pi_plot <- ppc_scatter_avg(y = fit_data$log_value, yrep = pp) + 
  coord_flip() +
  scale_y_continuous(name = "Observed Log B/Bmsy") + 
  scale_x_continuous(name = "Posterior Predictive B/Bmsy") +
  labs(subtitle = "A) Observed vs. Posterior Predictive") +
 pub_theme

pi_plot + r2_plot
```

### 2020-10-02 functions/prepare-sofia-data.R

@@ -1,6 +1,6 @@
 prepare_sofia_data <- function(min_years_catch = 20,
                                crazy_b = 4,
-                               crazy_u = 5,
+                               crazy_u = 20,
                                lookup_fmi_names = FALSE,
                                q = 1e-6) {
   # min_years_catch = 20
@@ -837,5 +837,59 @@ prepare_sofia_data <- function(min_years_catch = 20,
     mutate(year = as.numeric(str_replace_all(year, "\\D", "")))
  
   assign("ei_capture",ei_capture, envir = .GlobalEnv )
+  
+  
+  # sofia status
+  
+  sofia_status <-
+    read_csv(here("data", "StockSTatus_FAOSofia.csv")) %>%
+    janitor::clean_names() %>%
+    select(area, sp_group, name, species, x2017) %>%
+    rename(
+      status = x2017,
+      bad_fao_code = area,
+      isscaap_number = sp_group,
+      common_name = name,
+      scientific_name = species
+    ) %>%
+    mutate(fao_area_code = as.numeric(bad_fao_code)) %>%
+    filter(!is.na(status), status != "?",!is.na(fao_area_code)) %>%
+    mutate(
+      stock_complex = paste(common_name, isscaap_number, fao_area_code, sep = '_'),
+      full_status = status
+    )
+  
+  stats <-
+    str_trim(str_split(sofia_status$status, pattern = "[[:punct:]]", simplify = TRUE)[, 1])
+  
+  
+  sofia_status$status <- stats
+  
+  # there are many mispellings of scientific names in SOFIA that need correcting
+  
+  correct_sofia <- sofia_status %>% 
+    select(common_name, scientific_name) %>% 
+    rename(bad_scientific_name = scientific_name) %>% 
+    unique() %>% 
+    mutate(sciname = map(common_name, ~ taxize::comm2sci(com = .x, db = "worms")[[1]]))
+  
+  correct_sofia <-  correct_sofia %>% 
+    mutate(sci_match = map_lgl(sciname, ~length(.x) > 0)) %>% 
+    filter(sci_match == TRUE) %>% 
+    mutate(good_scientific_name = map_chr(sciname, ~.x[1])) %>% 
+    select(common_name,good_scientific_name )
+  
+  
+  sofia_status <- sofia_status %>%
+    left_join(correct_sofia, by = "common_name") %>%
+    mutate(scientific_name = ifelse(
+      is.na(good_scientific_name),
[…]

### 2020-10-02 make-assessing-global-fisheries.R

@@ -25,6 +25,7 @@ library(rstan)
 library(rstanarm)
 library(sf)
 library(portedcmsy)
+extrafont::loadfonts()
 
 Sys.unsetenv("PKG_CXXFLAGS")
 
@@ -65,19 +66,15 @@ write_results <- FALSE
 
 process_fits <- FALSE
 
-# other run options things
-
-run_continent_examples <- FALSE
-
-run_ei_example <- FALSE
+run_case_studies <- FALSE
 
 run_sofia_comparison <- FALSE
 
 run_ram_tests <- FALSE
 
-run_ram_comparison <- FALSE
+run_ram_comparison <- TRUE
 
-knit_paper <- TRUE
+knit_paper <- FALSE
 
 warning("Running full analysis takes upwards of 24 hours on 2 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 2 cores is not recommended (memory routinley runs out on a 16 core 36GB machine when cores > 2)")
 
@@ -166,12 +163,12 @@ if (run_voi_models == TRUE) {
     initial_state_cv =  runif(draws, 0.05, 0.2),
     use_terminal_state = sample(c(TRUE, FALSE),  draws, replace = TRUE),
     terminal_state_cv = runif(draws, 0.05, 0.2),
-    error_cv = sample(c(0, 1), draws, replace = TRUE),
+    error_cv = sample(c(0), draws, replace = TRUE),
     stockid = sample(unique(ram_data$stockid), draws, replace = TRUE),
     use_u_priors =  sample(c(TRUE, FALSE),  draws, replace = TRUE),
     u_cv = runif(draws, 0.05, 0.2),
     index_window = sample(c(1, 0.25, .5), draws, replace = TRUE),
-    index_freq = sample(c(1, 2), draws, replace = TRUE),
+    index_freq = sample(c(1), draws, replace = TRUE),
     u_window = sample(c("snapshot", "recent", "complete"), draws, replace = TRUE),
     b_ref_type = sample(c("k", "b"),  draws, replace = TRUE),
     f_ref_type = sample(c("f", "fmsy"),  draws, replace = TRUE),
@@ -250,7 +247,7 @@ if (run_voi_models == TRUE) {
           estimate_initial_state = !initial_state_type == "unfished"
         ),
         safely(run_assessments),
-        assessments = c("cmsy", "sraplus"),
+        assessments = c("sraplus"),
         model = "sraplus_tmb",
[…]

### 2020-10-02 tmp.R

@@ -1,0 +1,145 @@
+run_case_studies <- TRUE
+
+exs <- ram_data %>%
+  filter(
+    stockid %in% unique(sraplus::fmi$stockid) &
+      stockid %in% unique(sraplus::sar$stockid)
+  ) %>%
+  mutate(fao_area_code = as.numeric(primary_FAOarea),
+         country = as.character(primary_country)) %>%
+  mutate(country =
+           countrycode::countrycode(country, "country.name", "un.name.en"))
+
+comp_stocks <- exs %>%
+  select(scientificname) %>%
+  unique() %>%
+  mutate(fishbase_vuln = map_dbl(
+    scientific_name,
+    ~ rfishbase::species(.x, fields = "Vulnerability")$Vulnerability
+  )) %>%
+  mutate(
+    resilience = dplyr::case_when(
+      fishbase_vuln < 33 ~ "Low",
+      fishbase_vuln >= 66 ~ "High",
+      TRUE ~ "Medium"
+    )
+  ) %>%
+  select(scientificname, resilience)
+
+
+exs %>%
+  ggplot(aes(year, b_v_bmsy, color = stockid)) +
+  geom_line(show.legend = FALSE)
+
+total_nominal_effort <- rous_data %>%
+  left_join(effort_region_to_country, by = "country") %>%
+  rename(fao_area_code = area) %>%
+  group_by(fao_area_code, year, country) %>%
+  summarise(total_effort = sum(effort_cell_reported_nom, na.rm = TRUE)) %>%
+  ungroup()
+
+matched_fmi <- sraplus::fmi %>%
+  filter(stockid %in% unique(exs$stockid)) %>%
+  group_by(stockid) %>%
+  nest() %>%
+  ungroup() %>%
+  mutate(tmp = map(data, ~ map_df(.x[,c("research", "management", "enforcement","socioeconomics")], mean))) %>%
+  select(-data) %>%
+  unnest(cols = tmp) %>%
+  select(stockid, research, management, enforcement, socioeconomics)
+
+
+matched_sar <- sraplus::sar %>%
+  filter(stockid %in% unique(exs$stockid)) %>%
+  group_by(stockid) %>%
+  nest() %>%
+  ungroup() %>%
+  mutate(tmp = map(data, ~ purrrlyr::dmap_at(.x, c("sar"), mean))) %>%
+  select(-data) %>%
+  unnest(cols = tmp) %>%
[…]

### 2020-10-15 documents/ovando-etal-assessing-global-fisheries-long.Rmd
waiting to hear back before merging to master (#7)

--- 2020-10-02 documents/ovando-etal-assessing-global-fisheries.Rmd
@@
 <!-- * more assertive on what the data needs really are -->
+
 <!--   * where to apply globally  -->
+
 <!-- * For places with less coverage, what would we actually recommend there  -->
+
 <!-- * what do we know about catchabiliy -->
+
 <!-- - e.g. 51, fleets became industrial  -->
+
 <!--   - hammer down on that's the issue with the effort  -->
+
 <!--   - how do we translate effort into f bey area -->
+
 <!--   - here's  -->
@@
 
-49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020].
+Forty nine percent of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020].
 The SOFIA process uses these formal assessments for their determinations of stock status wherever possible.
@@
 While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable (and expert opinion can also be misinformed).
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018;@thorson2012a].
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
 
-Formal stock assessments require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE).
+Formal stock assessments require at a minimum robust timeseries of captures and a reliable index of abundance and/or length composition data, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE).
 While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist [@maureaud2020], and would be prohibitively expensive to collect.
@@
 
-Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment.
+When applied in bulk to the question of global stock status, data-limited assessment methods typically make compromises on both the data requirements and individual attention needed for formal stock assessment.
 Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (Free et al.
@@
 In addition, a global assessment requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time.
-As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks.
+As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, data-limited methods applied across thousands of stocks at the global level typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks.
 Each of the published catch-based methods reviewed in Free et al.
-2020 [@free2020] sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.
+2020 [@free2020] sought to address specific gaps in our understanding of global fisheries, and in doing so offer unique strengths and weaknesses.
 
@@
 The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020].
-A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.
+While it must noted that the estimates in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the best available estimates of fishery status from RAM to those predicted by potentially less reliable methods intended for use when insufficient data are available for a full stock assessment model.
 
@@
 The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or NOrtheast Pacific.
-The catch based approach @pauly2007, which compares catch in the last year to maximum catch and equates these to stock status, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).
-While the methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks,
+The catch based approach [@pauly2007], which compares catch in the last year to maximum catch and equates these to stock status, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).
+The methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
+However, the lack of consistency across heavily assessed regions such as the NOrtheast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.
 
@@
 At the same time, new sets of globally distributed data have recently become available, along with advances in statistical modeling methods.
-We leverage these opportunities to provide a global assessment of our perception of global stock status resulting from different use of different data sources.
+We leverage these opportunities to show how our perception of global stock status changes depending on the data source used in an assessment.
 Our results demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection and interpretation effort.
@@
 # Results
+
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Data Source                                 | Short Name | Data Use                                                                 | Caveats                                                                                                         |
++=============================================+============+==========================================================================+=================================================================================================================+
+| Catch data [@fao2020]                       | catches    | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Fisheries Management Index [@melnychuk2017] | FMI        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Swept Area Ratio [@amoroso2018]             | SAR        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Reconstructed effort data[@rousseau2019b]   | effort     | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors.
+Assumed rate of technology creep reported in individual sections |
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+
+: Broadly available data used to inform estimates of global stock status
 
@@
 
-For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index and swept area ratio scores.
-We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock.
+For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index (a measure of management capacity) and swept area ratio (a measure of trawl fishing pressure) scores.
+We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al.
+2019 [@rousseau2019] to each stock.
 We used these data to demonstrate how `sraplus` can add incremental amounts of information to an assessment, and the potential value of doing so.
-We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status.
+We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status, due to its widespread use [XX do we really want to cite those papers...XX].
 We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`.
-It is critical to note that in both of these cases the catch data on its own contains no information about stock status; we can only produce estimates of stock status by combining catch data with explicit priors and specified functional relationships [@thorson2013].
-This is important since it means that we must look to our priors and functional forms to understand where our insight stock status is coming from, not the catch data alone.
-Lastly, we used the reconstructed effort data from @rousseau2019 to create an index of abundance for each stock (see [Estimation Model](#Estimation Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.
+Lastly, we used the reconstructed effort data from [@rousseau2019] to create an index of abundance for each stock (see [Estimation Model](#Estimation%20Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.
 
 Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
-CMSY, using no stock-specific information but the catch history and the estimated resilience of the target species, badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
+CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
 Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
 Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
-This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values.
-An abundance index based on catch and regional effort cannot capture this trend, since it interprets the extremely low catches (near zero) as low biomass since we do not have sufficiently fine-scale effort data to reflect the reduction in effort that goes hand-in-hand with regulated catch reductions (Fig.\@ref(fig:cs-plot)).
+This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values (Fig.\@ref(fig:cs-plot)).
 
@@
 
-That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
+<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
 In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
 Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
+-->
 
@@
 
-As our test case we use `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
+Moving beyond our case study, our goal is to determine the ability of FMI, SAR, or effort data to improve estimates of global stock status.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
 B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
-We then pair the catch histories for these stocks with regional-level swept area ratio, fisheries management index, and effort data.
+We then paired the catch histories for these RAM stocks with regional-level swept area ratio, fisheries management index, and effort data.
 For any given stock we use the area-weighted swept area ratio (where area refers to the spatial area of the SAR study in question) for the FAO region of the stock.
@@
 This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
-We then constructed an effort index for the stock in question as the total effort from @rousseau2019 estimated in all countries that report catch of the species in question within its FAO region, weighting the effort of each country by its contribution to the total catch of that species in that FAO region.
+We then constructed an effort index for the stock in question as the total effort from Rousseau et al.
+2019 [@rousseau2019] estimated in all countries that report catch of the species in question within its FAO region, weighting the effort of each country by its contribution to the total catch of that species in that FAO region.
 
@@
         "catch-only", "priors on terminal B/Bmsy informed by catch history",
-        "heuristic", "priors on initial and terminal B/Bmsy informed by Cmsy heuristc")
+        "heuristic", "priors on initial and terminal B/Bmsy informed by CMSY heuristc")
 
@@
 Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries.
-Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`.
-This creates two main reasons why `sraplus` might produce poor results at a global level: model misspecification and poor data quality.
-Model misspecification errors might result from the surplus production model underlying `sraplus` being fundamentally unable to capture the more complicated dynamics resulting from the full age-structured population models common in RAM.
-Errors resulting from poor quality data mean that even if the underlying model is a reasonable approximation of the real world, the model may produce incorrect results due to the data supplied to the model simply not containing accurate information on the phenomenon in question.
+Since no global database of length or age composition data exists, most global assessment methods use surplus production models in the manner used by `sraplus`.
+This creates two main reasons why `sraplus` might produce poor results at a global level:  first, model misspecification errors might result from the surplus production model underlying `sraplus` being fundamentally unable to capture the more complicated dynamics resulting from the full age-structured population models common in RAM.
+Second, errors resulting from poor quality data might mean that even if the underlying model is a reasonable approximation of the real world, the model may produce incorrect results due to the data supplied to the model simply not containing accurate information on the phenomenon in question.
 
-Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality.
-Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+Overall the `sraplus` estimates of B/B~MSY~  resulting from using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
 This exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
-
-<!-- The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias.
-The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time.
-Looking across space, we see clear geographic patterns in model performance.
-The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada).
-The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.
--->
 
@@
 The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])` (since fisheries are not uniformly distributed across each bin).
+Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])`.
 
 Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process.
-Across the models performance (or lack thereof) is not consistent in space: use of different data performed best or worst for different FAO regions.
+Across the models, performance (or lack thereof) was not consistent in space: use of different data performed best or worst for different FAO regions.
 For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
@@
 
-Global-level assessment are critical for guiding management agendas for the world's oceans.
+Global-level assessment are critical for guiding management agendas for the world's oceans, and tracking critical indicators such as the United Nations Sustainable Development Goals.
 Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world.
-While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).
+While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).
 
-What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends of course on the needs of those using the assessment outputs.
+What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality dependson the needs of those using the assessment outputs.
 It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
@@
 those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)).
-Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
+Effort data such as those reconstructed by Rousseau et al.
+2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
 Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
-But, applying one SAR value to all stocks in a region causes inaccurate estimates of stock status.
+
+But, we must simulataneously consider the ideas of data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
 Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
 
-It is easy enough to call for expanded data collection at a global scale.
-In world of finite resources though, it is important to consider what sources of data might provide the greatest value in improving our estimates of global stock status.
+What sources of data might provide the greatest value in improving our estimates of global stock status?
 We fit 3,000 `sraplus` models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
@@
 
-Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 15%.
-Information on fishing mortality rates (F/F~MSY~) in the most recent year was equally useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot).
+Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates by on average 15%.
+Information on fishing mortality rates (F/F~MSY~) in the most recent year was equally useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ (which is conditional on the recent fishing selectivity regime) was actually more useful (Fig.\@ref(fig:voi-plot)).
 Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance.
@@
 
-As we have stated here using catch data alone we cannot learn anything new about recent stock status beyond what is encoded in our priors.
+<!-- Using catch data alone does not reveal about recent stock status beyond what is encoded in our priors and functional forms.
 So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest).
@@
 While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
+-->
 
@@
 
-We should of course prioritize collection of survey data worldwide, high-quality synthesis and standardization of these data, as well as their collection in an international database.
-This assessment assumes that we have perfect knowledge of for example F/F~MSY~, when in fact any estimates derived from for example length-based methods are likely to be highly error-prone.
+We must also prioritize collection of survey data worldwide, high-quality synthesis and standardization of these data, as well as their collection in an international database.
+XX This assessment assumes that we have perfect knowledge of for example F/F~MSY~, when in fact any estimates derived from for example length-based methods are likely to be highly error-prone.
 However, our results demonstrate that we must prioritize aggregation of long-running but less publicly-accessible survey data, or expect to wait for some time until new surveys are capable of yielding meaningful information on their own about stock status.
@@
 XXbe more specific here, and clearXX.
-.Effort databases such as @rousseau2019, can be combined with catch histories to to construct fishery-dependent abundance indicies.
+.
+Effort databases such as Rousseau et al.
+2019 [@rousseau2019], can be combined with catch histories to to construct fishery-dependent abundance indicies.
 While there are substantial difficulties and biases that can arise from improper use of fishery-dependent abundance data, improved understanding of the evolution of global fishing effort could be very useful in improving global estimates of fishery health.
@@
 
+XX Prioritizing and selecting stocks for assessment, with light towards SDG XX
+
 First, use of effort data requires making an assumption about the rate at which the fishing power of a unit of fishing effort has changed over time.
-@rousseau2019 used a fixed global rate of 2.6% to describe the increasing trend in fishing power, but as shown in their paper and in our analysis, any estimates of stock status using these effort data will be critically sensitive to the value of this parameter.
+[@rousseau2019] used a fixed global rate of 2.6% to describe the increasing trend in fishing power, but as shown in their paper and in our analysis, any estimates of stock status using these effort data will be critically sensitive to the value of this parameter.
 Global studies of technology creep could help inform more realistic values and functional forms for the evolution of fishing skill in particular fisheries and regions, and also require accounting for regional complications including density-dependent catchability [@thorson2010].
 Second, the effort reconstructions as employed here are still at very coarse scales, requiring us to apply one effort series for all the stocks within a given region.
-Going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
+Going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of [@rousseau2019] might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
 Along with supporting methods used here, access to fine-scale effort dynamics data can help inform data-limited assessment based on catch or length-composition data [@ovando2018; @thorson2013].
@@
 
-The core of `sraplus` is a Pella-Tomlinson [@pella1969] production model constructed in the manner of @winker2018.
+The core of `sraplus` is a Pella-Tomlinson [@pella1969] production model constructed in the manner of [@winker2018].
 While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
@@
 
-\begin{equation}
-  f(x)=\begin{cases}
-      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
-, & \text{if $B_t>0.25 \times K$}.\\
-     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
-  \end{cases}
-  (\#eq:pt)
-\end{equation}
+\\begin{equation} f(x)=\\begin{cases} B\_{t + 1} = \\left(B\_{t} + B\_{t}\\frac{r}{m - 1}\\left(1 - \\left(\\frac{B\_t}{K}\\right)\^{m- 1}\\right) - \\hat{c\_t}\\right)p\_t , & \\text{if \$B\_t\>0.25 \\times K\$}.\\\\ B\_{t + 1} = \\left(B\_{t} + \\frac{B\_{t}}{0.25 \\times K}\\left(B\_{t}\\frac{r}{m - 1}\\left(1 - \\left(\\frac{B\_t}{K}\\right)\^{m- 1}\\right) - \\hat{c\_t}\\right)\\right)p\_t, & \\text{otherwise}.
+\\end{cases} (\\\#eq:pt) \\end{equation}
 
@@
 Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
-We deal with this problem by following the methods described in @winker2018.
+We deal with this problem by following the methods described in [@winker2018].
 to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
 
-We  allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by @walters2006).
+We allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by [@walters2006]).
 This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
-Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see @thorson2018).
+Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).
 
@@
 
-\begin{equation}
-  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
-  (\#eq:procerror)
-\end{equation}
+\\begin{equation} p\_t \\sim e\^{normal\\left(-\\sigma\_{proc}\^2/2,\\sigma\_{proc}\\right)} (\\\#eq:procerror) \\end{equation}
 
@@
 When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
-Under this mode, the model is essentially a stock-reduction analysis model, in the manner of @walters2006, in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e.
+Under this mode, the model is essentially a stock-reduction analysis model, in the manner of [@walters2006], in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e.
 results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches.
@@
 The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).
-*r* and *k* are the only two parameters that are always estimated.
+*r* and *K* are the only two parameters that are always estimated.
 Estimation of every other parameter can be turned on or off.
 When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user.
-For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *k*, $\sigma_{obs}$, $\gamma$, and *B0*.
+For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*.
 *q* is also estimated when needed.
@@
 
-`sraplus` can be run in two forms: either as a stock reduction analysis [SRA,@walters2006], or fit to an index of abundance (fishery dependent or independent).
+`sraplus` can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent).
 Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
@@
 
-$$log(a_t) \sim normal(f(r,k,m,B0,\pmb{p},\pmb{c}) \times q, sigma_{obs})$$
+$$log(a_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}) \times q, sigma_{obs})$$
 
@@
 When an effort index is available, `sraplus` constructs an index of abundance based on the catch and effort data.
-@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6\% annualized technology rate).
-This assumes that every unit increase in effort is log-linearly greater than the unit of effort before it.
+[@rousseau2019] measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology rate).
+This rate of technology creep assumes that every unit increase in effort is log-linearly greater than the unit of effort before it.
 When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
@@
 
-$$log(cpue_t) \sim normal(f(r,k,m,B0,\pmb{p},\pmb{C}), \sigma_{obs})$$
+$$log(cpue_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{C}), \sigma_{obs})$$
 
@@
 In addition to the results from `sraplus`, we include a set of results produced by the CMSY method [@froese2017].
-For computational efficiency, we used a ported version of the CMSY model available at https://github.com/DanOvando/portedcmsy.
+For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>.
 The only modification made is to convert the underlying population model to C++ for faster computation.
@@
 The shape parameter is usually not reliably estimable given available data for surplus production models, however, Thorson et al.
-2012 [@thorson2012] provides estimate the ratio of B~MSY~ to K for many fish taxa.
+2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa.
 While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on Thorson et al.
@@
 This creates a problem for the Bayesian nature of our analysis though.
-Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
-Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *k* together with the model and the catch history produces a deterministic stock status in each time step.
+Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
+Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
 Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox.
@@
 From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status.
-If the bulk of the prior on terminal stock status was concentrated at 50% of *k*, combinations of *r* and *k* that produce terminal stock status near 50% of *k* are sampled proportionally more frequently.
-However, lower values of terminal stock status have fewer candidate values of *r* and *k*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
-Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *k* combinations: since under this model the population cannot be greater than carrying capacity, as for example *k* approaches infinity terminal stock status asymptotes at close to 100% of *k*.
-The net result of this is that even though individual combinations of *r* and *k* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
+If the bulk of the prior on terminal stock status was concentrated at 50% of *K*, combinations of *r* and *K* that produce terminal stock status near 50% of *K* are sampled proportionally more frequently.
+However, lower values of terminal stock status have fewer candidate values of *r* and *K*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
+Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *K* combinations: since under this model the population cannot be greater than carrying capacity, as for example *K* approaches infinity terminal stock status asymptotes at close to 100% of *K*.
+The net result of this is that even though individual combinations of *r* and *K* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
 As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.
@@
 And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
-The net result of this is a post-model-pre-data distribution of parameters *r* and *k*  that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
+The net result of this is a post-model-pre-data distribution of parameters *r* and *K* that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
 In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
@@
 
-The Fisheries Management Index (FMI), as presented in @melnychuk2017, utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
+The Fisheries Management Index (FMI), as presented in [@melnychuk2017], utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
 These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
@@
 Importantly, FMI surveys can be filled out in the absence of stock assessments.
-This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to @osio2015 and @cope2015).
+This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to [@osio2015] and [@cope2015]).
 
@@
 
-@amoroso2018 provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
+[@amoroso2018] provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
 This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
-As illustrated in @amoroso2018, there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
+As illustrated in [@amoroso2018], there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
 Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
@@
 # References
-<div id="refs"></div>
+
+::: {#refs}
+:::
 

### 2020-10-15 make-assessing-global-fisheries.R

@@ -72,7 +72,7 @@ run_sofia_comparison <- FALSE
 
 run_ram_tests <- FALSE
 
-run_ram_comparison <- TRUE
+run_ram_comparison <- FALSE
 
 knit_paper <- FALSE
 
@@ -872,6 +872,37 @@ ex_kobe_plot <- tmp %>%
   scale_fill_continuous(low = "gainsboro", high = "black") +
   ggsci::scale_color_d3(name = "Region") +
   guides(size = FALSE)
+
+
+tmp2 <-  tmp %>%
+  pivot_longer(contains("div"), names_to = "variable", values_to = "value")
+
+truth2 <- truth %>%
+  select(stockid, year, contains("_v_")) %>%
+  pivot_longer(contains("_v_"), names_to = "variable", values_to = "truth")  %>%
+  mutate(variable = str_replace_all(variable, "_v_", "_div_"))
+
+
+tmp2 <- tmp2 %>%
+  left_join(truth2, by = c("stockid", "year", "variable")) %>%
+  mutate(
+    variable = case_when(
+      variable == "b_div_bmsy" ~ "B/Bmsy",
+      variable == "u_div_umsy" ~ "F/Fmsy",
+      TRUE ~ variable
+    )
+  )
+
+ex_scatter_plot <- tmp2 %>% 
+  ggplot(aes(truth, value, size = lifetime_catch)) + 
+  geom_vline(aes(xintercept = 0)) + 
+  geom_hline(aes(yintercept = 0)) +
+  geom_abline(aes(slope = 1, intercept = 0),linetype = 2) +
+  geom_point(alpha = 0.75) + 
+  facet_grid(variable ~ data, scales = "free_y")  + 
+  scale_size(trans = "sqrt", name = "Lifetime Catch") +
+  scale_x_continuous(name = "RAM Value", expand = expansion(add = c(0, .1))) + 
+  scale_y_continuous("Estimated Value", expand = expansion(add = c(0, .1)))
 
 sraplus_v_truth <- tmp %>% 
   left_join(truth, by = c("stockid","year")) %>% 

### 2020-10-16 documents/ovando-etal-assessing-global-fisheries.Rmd
pre submission draft

--- 2020-09-10 documents/ovando-etal-assessing-global-fisheries.Rmd
@@
 ---
-title: "Global Assessments of Fishery Status need Better Data more than Better Models"
+title: |
+  | Global Assessments of Fishery Status need 
+  | Better Data more than Better Models
 author: 
@@
 abstract: |
-  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals.
-While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor.
-Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
-This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources.
-How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets?
-We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status.
-We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data.
-These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection.
-Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random quess.
-Obtaining accurate estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of high-priority data at a global scale, not on the development of new modeling methods alone.
+  Assessments of the global state of fisheries play an important role in tracking the implementation of the United Nations Sustainable Development Goals.
+While we have reliable estimates of stock status for fisheries accounting for 49% of global catch, our knowledge of the state of the remaining 51%, the worlds 'unassessed' fisheries, is poor.
+Numerous high-profile publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
+Here, we show that despite numerous efforts, our understanding of the status of global fisheries remains poor, even when new sources of broadly available data are added.
+Obtaining accurate estimates of stock status for the world's fisheries depends on prioritizing the collection of high-priority data at around the globe, rather than the development of new modeling methods alone.
 always_allow_html: true
@@
 
-rand_fits <- assess_ram_fits %>% 
-  filter(data == "ram-data") %>% 
-  mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
-         data = "guess")
+# rand_fits <- assess_ram_fits %>%
+#   filter(data == "ram-data") %>%
+#   mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
+#          data = "guess")
 
 perf <- assess_ram_fits %>% 
-  bind_rows(rand_fits) %>% 
+  # bind_rows(rand_fits) %>% 
   mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
@@
   ) %>% 
-  arrange((mape))
+  arrange((mape)) %>% 
+  rename("Data Used" = "data")
 
@@
 <!-- * more assertive on what the data needs really are -->
+
 <!--   * where to apply globally  -->
+
 <!-- * For places with less coverage, what would we actually recommend there  -->
+
 <!-- * what do we know about catchabiliy -->
+
 <!-- - e.g. 51, fleets became industrial  -->
+
 <!--   - hammer down on that's the issue with the effort  -->
+
 <!--   - how do we translate effort into f bey area -->
+
 <!--   - here's  -->
 
-# Publication Requirements
+<!-- https://www.nature.com/natsustain/about/content -->
+<!-- # check in call notes -->
 
-Target Journal: [Nature Sustainability](https://www.nature.com/natsustain/about/content)
+<!-- Ray: setting the stage: what we want is a way to assess the status of stocks and see how we're doing relative to SDG. -->
 
-3500 words main text
+<!-- We need to discuss a sampling strategy: how can we get down to a set of priority stock. A step of expert knowledge to refine what stocks you would focus in on. -->
 
-50 references (main text)
+<!-- Regressions based approach, focusing in effort on stocks -->
 
-3000 words (max) methods
+<!-- Basically a paragraph laying out that we need to do this, and discussing how we might go about doing that. Equal weight to each stock vs proportional to biomass -->
+
+<!-- add in a bit more to the discussion on guidelines -->
+
+<!-- add in error in the index for the VOI analysis -->
+
+<!-- Apples and apples comparisons between model and design developments.
+How much is the value of the effort data relative to actually improving the population model.
+Production model is ill-suited to some of these questions.
+-->
+
+<!-- fishing effort uncertainty of the effort -->
+
+<!-- how credible is the information -->
+
+<!-- probably don't need to bother with the SAUP -->
+
+<!-- stock assessments -->
+
+<!-- the stock will collapse next year -->
+
+<!-- nansen data -->
+
+<!-- # Publication Requirements -->
+
+<!-- Target Journal: [Nature Sustainability](https://www.nature.com/natsustain/about/content) -->
+
+<!-- 3500 words main text -->
+
+<!-- 50 references (main text) -->
+
+<!-- 3000 words (max) methods -->
 
@@
 
-The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
-A substantial part of achieving this goal depends on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems.
+The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
 Meeting these targets depends in part on our ability to effectively track the status of global marine fish stocks.
-The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking the global state of fisheries.
+The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the most widely used source for tracking the global state of fisheries.
 As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020].
-While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment.
-Although revised periodically over time, the method has kept revisions to minimally necessary levels for the purposes of temporal consistency and comparability.
-With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.
+The SOFIA assessment was designed in the 1970s based on the then available data and methods, With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods, fresh approaches are needed to meet the new demand for tracking progress towards the SDG goals and supporting the research targets set out in the UNited Nations Decade on Oceans and Science plan.
 
-49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020].
+Forty nine percent of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database [@ricard2012], a repository of best-available estimates of fishery status [@hilborn2020].
 The SOFIA process uses these formal assessments for their determinations of stock status wherever possible.
@@
 While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
+
 The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion.
-While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable (and expert opinion can also be misinformed).
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018;@thorson2012a].
-
-Formal stock assessments require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE).
-While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist [@maureaud2020], and would be prohibitively expensive to collect.
-Even when data are available though, running an individualized stock assessment is not a "just press play" process.
-Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment.
-This process takes considerable time and expertise.
-
-Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment.
-Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (Free et al.
+While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable, in order to increase reproducibility, transparency, and potential reliability.
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
+Due to data limitations, all of these global assessment efforts use forms of "catch-only" stock assessment models (Free et al.
 2020 [@free2020] and references therein).
 These models seek to infer stock status, for example in terms of biomass *B* relative to the biomass at maximum sustainable yield *B~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
-In addition, a global assessment requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time.
-As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks.
-Each of the published catch-based methods reviewed in Free et al.
-2020 [@free2020] sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.
 
@@
 These issues become apparent when we consider some of the macro-level predictions made by these models.
-The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020].
-A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.
+While the estimates B/B~MSY~ and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable methods intended for use when insufficient data are available for a full stock assessment model.
 
-Examining stock assessment data from RAM from the  Northeast Pacific,the Northeast Atlantic and the Mediterranean & Black Sea, the Northeast Pacific is in the best state (mean RAM B/B~MSY~ = 1.61), NE Atlantic (median RAM b/b~MSY~ =1.27) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.52) [@hilborn2020].
-The Costello et al.
-2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison.
-The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or NOrtheast Pacific.
-The catch based approach @pauly2007, which compares catch in the last year to maximum catch and equates these to stock status, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).
-While the methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks,
+Costello et al.
+2016 [@costello2016a] finds similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison.
+The super-erensemble method [@anderson2017a] used in Rosenberg et al.
+2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific.
+The Pauly 2007 [@pauly2007] catch based approach finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
+The methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
+However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.
 
@@
   select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
-  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM through Hilborn et al. 2020, Costello et al. 2016, 
+  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
                Rosenberg et al. 2018, and Pauly 2007",
@@
 In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fisheries?
-We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
-The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability.
-However, catches are not the only broadly available source of fishery data.
-We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management through Fisheries Management Index (FMI) scores [@melnychuk2017], trawl-fishing footprints through swept-area-ratio (SAR) values [@amoroso2018], and life history [@thorson2017c].
-`sraplus` provides a flexible framework for integrating these data into valid statistical model.
-
-The coming decades are a critical period for the future of global oceans.
-Climate change, population growth, and economic expansion will all place increasing pressure on fisheries.
-While our understanding of global fisheries has improved greatly,  mitigating these threats, as called for by SDG 14, will require an improved ability to track the status of fisheries around the world, particularly those stocks that are currently unassessed.
-At the same time, new sets of globally distributed data have recently become available, along with advances in statistical modeling methods.
-We leverage these opportunities to provide a global assessment of our perception of global stock status resulting from different use of different data sources.
-Our results demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection and interpretation effort.
+We use a flexible surplus-production based stock assessment software package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
+We show that our understanding of global fishery status is poor, and that improvements depend on an redoubled effort at global data collection.
 
@@
 
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Data Source                                 | Short Name | Data Use                                                                 | Caveats                                                                                                         |
++=============================================+============+==========================================================================+=================================================================================================================+
+| Catch data [@fao2020]                       | catches    | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Fisheries Management Index [@melnychuk2017] | FMI        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Swept Area Ratio [@amoroso2018]             | SAR        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Reconstructed effort data[@rousseau2019b]   | effort     | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors.
+Assumed rate of technology creep reported in individual sections |
++---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+
+: (\#tab:data) Broadly available data used to inform estimates of global stock status
+
 We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
-From there, we assess the performance of models fit using different kinds of broadly available data (combinations of catches, effort, fisheries management index, and swept area ratio values) relative values reported in the RAM database.
+From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index (FMI) scores (a measure of management capacity), and swept area ratio (SAR) values (a measure of trawl fishing intensity) (Table.\@ref(tab:data)).
 
@@
 
-For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index and swept area ratio scores.
-We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock.
-We used these data to demonstrate how `sraplus` can add incremental amounts of information to an assessment, and the potential value of doing so.
-We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status.
-We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`.
-It is critical to note that in both of these cases the catch data on its own contains no information about stock status; we can only produce estimates of stock status by combining catch data with explicit priors and specified functional relationships [@thorson2013].
-This is important since it means that we must look to our priors and functional forms to understand where our insight stock status is coming from, not the catch data alone.
-Lastly, we used the reconstructed effort data from @rousseau2019 to create an index of abundance for each stock (see [Estimation Model](#Estimation Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.
+For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
+We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al.
+2019 [@rousseau2019] to each stock.
+We first used the default catch history heuristics internal to CMSY [@froese2017] to estimate stock status.
+We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus.
+Lastly, we used the reconstructed effort data from [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
+While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results (Fig.\@ref(fig:cs-plot)).
 
-Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
-CMSY, using no stock-specific information but the catch history and the estimated resilience of the target species, badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
+<!-- Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
+CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
 Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
 Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
-This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values.
-An abundance index based on catch and regional effort cannot capture this trend, since it interprets the extremely low catches (near zero) as low biomass since we do not have sufficiently fine-scale effort data to reflect the reduction in effort that goes hand-in-hand with regulated catch reductions (Fig.\@ref(fig:cs-plot)).
+This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values (Fig.\@ref(fig:cs-plot)).
+-->
 
-```{r cs-plot, fig.cap="Estimates of stock status for case study fisheries using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index and Swept Area Ratio scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
-Each point is a stock, point size is a function of stock size.
-The shaded hexagons in the background show the distribution of stock status estimates from the RAM Legacy Stock Assessment Database for each of the stocks included in this figure, with darker colors representing increased concentrations of RAM stocks."}
-
-ex_kobe_plot
+```{r cs-plot, fig.cap="RAM values of B/B~MSY~ and F/F~MSY~ (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
+Each point is a stock.
+Black dashed line shows the 1:1 relationship."}
+ex_scatter_plot
 ```
 
-That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
+<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
 In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
 Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
+-->
 
-While SAR and FMI data are available for a wide number of stocks, they are not ubiquitous enough to provide stock-specific priors for the number of fisheries needed for a global assessment.
-In addition, predicting stock status based on SAR, FMI, or catch patterns requires that the same patterns that exist within the training data as exist in other stocks to be used for prediction, which is far from clear given the highly managed nature of the stocks for which we have reliable estimates of stock status.
-We can extend the ideas presented in the case study, using combinations of catch, SAR, FMI, and effort data to produce estimates of stock status at the global level.
+We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
+B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
+We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
+This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
 
-As our test case we use `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
-B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
-We then pair the catch histories for these stocks with regional-level swept area ratio, fisheries management index, and effort data.
-For any given stock we use the area-weighted swept area ratio (where area refers to the spatial area of the SAR study in question) for the FAO region of the stock.
-For fisheries management index we use the mean component FMI values for the species group and FAO region in question (or FAO region if no matches exist at the species group level).
-This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
-We then constructed an effort index for the stock in question as the total effort from @rousseau2019 estimated in all countries that report catch of the species in question within its FAO region, weighting the effort of each country by its contribution to the total catch of that species in that FAO region.
-
-As a proof of concept, we first estimated B/B~MSY~ of our candidate RAM stocks by using `sraplus` to fit to an abundance index drawn directly from RAM.
-We then fit a range of models utilizing different combinations of the regional-level data made available to them (Table.\@ref(tab:dat-desc)).
+As a proof of concept, we also estimated B/B~MSY~ of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
+We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
+2017 [@froese2017] (See Table.5).
 We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
@@
 
-```{r dat-desc}
-dat_description <- tribble(~`Data Name`, ~Description,
-        "RAM-Data", "Fit to abundance index from RAM",
-        "sar", "Prior on terminal F/Fmsy set by regional swept area ratio",
-        "fmi", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
-        "cpue", "Fit to CPUE index created from RAM catch and regional effort index. 2.5% tech. creep",
-        "cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.5% tech. creep",
-        "nominal-cpue", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
-         "nominal-cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
-        "catch-only", "priors on terminal B/Bmsy informed by catch history",
-        "heuristic", "priors on initial and terminal B/Bmsy informed by Cmsy heuristc")
-
-knitr::kable(dat_description,
-             caption = "Data sources used for terminal stock status estimate",
-             format = "pipe")
-# %>%
-#   kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-#   kableExtra::row_spec(0, bold = T)
-```
-
-Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries.
-Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`.
-This creates two main reasons why `sraplus` might produce poor results at a global level: model misspecification and poor data quality.
-Model misspecification errors might result from the surplus production model underlying `sraplus` being fundamentally unable to capture the more complicated dynamics resulting from the full age-structured population models common in RAM.
-Errors resulting from poor quality data mean that even if the underlying model is a reasonable approximation of the real world, the model may produce incorrect results due to the data supplied to the model simply not containing accurate information on the phenomenon in question.
-
-Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality.
-Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-This exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
-
-<!-- The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias.
-The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time.
-Looking across space, we see clear geographic patterns in model performance.
-The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada).
-The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.
--->
+Overall the sraplus estimates of B/B~MSY~ resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
 
 Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves.
-These becomes clearer once we consider the performance of `sraplus` models fit to combinations of our broadly available datasets.
-Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)).
+This becomes clearer once we consider the performance of sraplus models fit to combinations of our broadly available datasets.
+Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)).
 However, this is somewhat an artifact of the data.
-The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4.
-The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1.
-This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise.
+The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near one.
+This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise.
 Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
-The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])` (since fisheries are not uniformly distributed across each bin).
+The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`.
+Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.
 
 Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process.
-Across the models performance (or lack thereof) is not consistent in space: use of different data performed best or worst for different FAO regions.
+Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions.
 For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
 We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
-Overall, while some data sources performed slightly better than others by some metrics in some places, there was no consistent spatial pattern to this performance, and no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
+Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
 
@@
 mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
-Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.",
+Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.
+See Table.5 for details of data types.",
               format = "pipe")
@@
 ram-data refers to catch and abundance index drawn from RAM.
-CPUE refers to an index of abundance based on the Rousseau effort data.
-cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
-For both CPUE series 'nominal' assumes a 0% technology rate, otherwise a 2.6% technology rate is assumed.
+CPUE refers to an index of abundance based on reconstructed effort data.
+cpue-plus uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, otherwise a 2.6% technology creep is assumed.
 ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
 fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio.
-CMSY uses the methods from @froese2017."}
+CMSY uses the methods from Froese et al.
+2017 [@froese2017]."}
 ram_mpe_map_plot
@@
 ram-data refers to catch and abundance index drawn from RAM.
-CPUE refers to an index of abundance based on the Rousseau effort data.
-cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
-For both CPUE series 'nominal' assumes a 0% technology rate, otherwise a 2.6% technology rate is assumed.
+CPUE refers to an index of abundance based on reconstructed effort data.
+cpue-plus uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, otherwise a 2.6% technology creep is assumed.
 ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
 fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio.
-CMSY uses the methods from @froese2017."}
+CMSY uses the methods from Froese et al.
+2017 [@froese2017]."}
 ram_mape_map_plot
@@
 ram-data refers to catch and abundance index drawn from RAM.
-CPUE refers to an index of abundance based on the Rousseau effort data.
-cpue-plus uses CPUE along with fisheries management index (FMI) and/or swept area ratio data.
-For both CPUE series 'nominal' assumes a 0% technology rate, otherwise a 2.6% technology rate is assumed.
+CPUE refers to an index of abundance based on reconstructed effort data.
+cpue-plus uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, otherwise a 2.6% technology creep is assumed.
 ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
 fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio.
-CMSY uses the methods from @froese2017."}
+CMSY uses the methods from Froese et al.
+2017 [@froese2017]."}
 
 ram_acc_map_plot 
+```
+
+The poor global performance of models fit to currently accesible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
+What sources of data might provide the greatest value in improving our estimates of global stock status?
+We used sraplus together with the RAM database to estimate the average reduction in error resulting from having access to different kinds of data (Fig.\@ref(fig:voi-plot)).
+While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average.
+We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].
+
+```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of B/B~MSY~ in the most recent 5 years of data available for each model fit.
+Distribution is full posterior probability distribution.
+Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
+Change is relative to the mean performance of a catch-only heuristic model."}
+b_voi_plot + 
+  geom_vline(aes(xintercept = 0), alpha = 0.75)
 ```
@@
 
-Global-level assessment are critical for guiding management agendas for the world's oceans.
+Global-level assessment are critical for guiding management agendas for the world's oceans, and tracking critical indicators such as the United Nations Sustainable Development Goals.
 Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world.
-While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).
+While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).
 
-What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends of course on the needs of those using the assessment outputs.
+What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends on the needs of those using the assessment outputs.
 It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
 In some instances using the data presented here did provide some improvement over use of catch-only style methods; the difficulty comes in attempting to apply data types uniformly across the globe.
-While it is unreasonable to expect any global-scale data to be able to perform as well as data pulled from RAM assessments themselves, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general trends in stock status in time and space.
-That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the data themselves.
-
-We have stated that improving global estimates of stock status depends on improved data, not improved models *per se*.
-This is not to suggest that improvements in stock assessment methods, even for relatively simple models such as the surplus production types used here, are not of use.
-We certainly do not claim that the modeling choices we have made here represent the best version possible.
-At the tactical level, modeling improvements can make meaningful changes in management outcomes.
-However, at the global strategic level, the evidence from multiple efforts at assessing global fisheries strongly suggests that any improvements to be made from model improvement are marginal compared to the fundamental uncertainties stemming from currently available data.
+While it is unreasonable to expect any global-scale data to be able to perform as well as detailed stock assessments reported in RAM, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
+That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the quality and use of data themselves.
 
@@
 those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)).
-Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
+Effort data such as those reconstructed by Rousseau et al.
+2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
 Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
-But, applying one SAR value to all stocks in a region causes inaccurate estimates of stock status.
+
+But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
 Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
 
-It is easy enough to call for expanded data collection at a global scale.
-In world of finite resources though, it is important to consider what sources of data might provide the greatest value in improving our estimates of global stock status.
-We fit 3,000 `sraplus` models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
-We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).
+Our value-of-information analysis also shows though the high utility of having access to even a recent snapshot of F/F~MSY~ (Fig.\@ref(fig:voi-plot)).
+Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
+Another avenue would be to prioritize the development of a global repository for length and age composition data.
+Given appropriate conditions, these length measurements can be used to estimate local fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
+While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
+Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f;@rudd2017].
 
-Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 15%.
-Information on fishing mortality rates (F/F~MSY~) in the most recent year was equally useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot).
-Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance.
-While having access to a complete timeseries of an abundance index had a higher probability of reducing RMSE than increasing it, in many cases having only an abundance index resulted in a correct fit to the overall trend in abundance, but an error-prone estimate of B/B~MSY~ itself due to incorrectly estimating either the shape parameter, the initial state of the stock, or both.
-However, using an index of abundance that only covered the most recent quarter of a fishery's history actually increased RMSE by over 10% on average.
-
-```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of B/B~MSY~ in the most recent 5 years of data available for each model fit.
-Distributino is full posterior probability distribution.
-Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th."}
-b_voi_plot + 
-  geom_vline(aes(xintercept = 0), alpha = 0.75)
-```
-
-As we have stated here using catch data alone we cannot learn anything new about recent stock status beyond what is encoded in our priors.
+<!-- Using catch data alone does not reveal about recent stock status beyond what is encoded in our priors and functional forms.
 So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest).
@@
 While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
+-->
 
-Another avenue would be to prioritize the development of a global repository for length and age composition data.
-The data-limited assessment world has recently been developing along two almost parallel tracks.
-Catch-based methods have most commonly been employed in efforts to understand global fisheries, and on occasion for management of stocks mandated to manage catch but lacking sufficient data for a formal stock assessment [@berkson2015].
-Along with general performance questions around catch-only models, no matter what they require a complete, or at least near complete timeseries of captures.
-This rules catch-based methods out for many small scale fisheries around the world.
-As a result, a growing number of methods have focused on using length-composition data to provide estimates of stock status, including measures of fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
-These methods have an advantage for many fisheries in that they can in theory function with as little as one year of data.
-As a result, many fisheries around the world are enacting programs to collect and utilize length composition data for fisheries assessment.
-While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries this may provide an overlooked source of fisheries data at a global scale, at least as an improvement over relying on catch-alone or regional proxies.
-
-However, while we have standardized databases for global marine catch [@fao2020] and life history estimates [@froese2000; @thorson2017c], no comparable repository for length composition data exists.
-Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which when paired with catch data and where possible indicies of abundance could meaningfully improve our understanding of global fisheries [@thorson2015f].
-Such a database could be constructed in a top-down manner in a manner similar to the catch records compiled by the FAO,or a bottom-up platform, such as the [MERMAID](https://datamermaid.org/) platform for coral reef data, could be developed that would allow users around the world to self-contribute length composition studies.
-
-We should of course prioritize collection of survey data worldwide, high-quality synthesis and standardization of these data, as well as their collection in an international database.
-This assessment assumes that we have perfect knowledge of for example F/F~MSY~, when in fact any estimates derived from for example length-based methods are likely to be highly error-prone.
-However, our results demonstrate that we must prioritize aggregation of long-running but less publicly-accessible survey data, or expect to wait for some time until new surveys are capable of yielding meaningful information on their own about stock status.
+We must also prioritize collection and curation of fish population survey data worldwide.
 Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//).
 Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have previously had stock assessments.
-
-For many areas of the world effective fishery independent estimates of stock abundance are unlikely to be feasible in the near future.
-Effort databases such as @rousseau2019, can be combined with catch histories to to construct fishery-dependent abundance indicies.
-While there are substantial difficulties and biases that can arise from improper use of fishery-dependent abundance data, improved understanding of the evolution of global fishing effort could be very useful in improving global estimates of fishery health.
-While we did include effort reconstructions in our analysis, and did not find it to provide substantial improvements in assessment results at the macro scale, we attribute this largely to two difficulties.
-
-First, use of effort data requires making an assumption about the rate at which the fishing power of a unit of fishing effort has changed over time.
-@rousseau2019 used a fixed global rate of 2.6% to describe the increasing trend in fishing power, but as shown in their paper and in our analysis, any estimates of stock status using these effort data will be critically sensitive to the value of this parameter.
-Global studies of technology creep could help inform more realistic values and functional forms for the evolution of fishing skill in particular fisheries and regions, and also require accounting for regional complications including density-dependent catchability [@thorson2010].
-Second, the effort reconstructions as employed here are still at very coarse scales, requiring us to apply one effort series for all the stocks within a given region.
-Going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
-Along with supporting methods used here, access to fine-scale effort dynamics data can help inform data-limited assessment based on catch or length-composition data [@ovando2018; @thorson2013].
+Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of [@rousseau2019] might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes
 
@@
 This is why stock assessments even in the data-rich world are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
-Empowering a global network of fisheries scientists through training and peer-support would both improve the health and management of local stocks and provide a means for ensuring that global estimates of stock status are as accurate as they can be.
+Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
 
@@
 All analysis were conducted in the R programming language [@rcoreteam2019].
-Model fitting was conducted using Rcpp [eddelbuettel2011] and stan [@carpenter2017], implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018].
-The `sraplus` package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
+Model fitting was conducted using Rcpp [@eddelbuettel2011] and stan [@carpenter2017], implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018].
+The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
 
-`sraplus` contains too many options to cover within this methods section.
+sraplus contains too many options to cover within this methods section.
 We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus.
-Below we describe the structure of the population model underpinning `sraplus`, the estimation models used, and the construction of priors used in this paper.
+Below we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.
 
@@
 
-The core of `sraplus` is a Pella-Tomlinson [@pella1969] production model constructed in the manner of @winker2018.
+The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of [@winker2018].
 While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
-The purpose of `sraplus` is not to make substantial improvements in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status
+The purpose of sraplus is not to make substantial improvements in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status
 
@@
 All of our estimates are Bayesian in nature.
-We can break the use of `sraplus` into two distinct categories: with data and without.
+We can break the use of sraplus into two distinct categories: with data and without.
 By "data", we refer to measurements which are used to confront model estimates within a likelihood function.
@@
 
-`sraplus` can be run in two forms: either as a stock reduction analysis [SRA,@walters2006], or fit to an index of abundance (fishery dependent or independent).
+sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent).
 Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
 A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
-`sraplus` allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~.
+sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~.
 We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history.
@@
 
-$$log(a_t) \sim normal(f(r,k,m,B0,\pmb{p},\pmb{c}) \times q, sigma_{obs})$$
+$$log(a_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}) \times q, sigma_{obs})$$
 
 where *f* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
-When an effort index is available, `sraplus` constructs an index of abundance based on the catch and effort data.
-@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6\% annualized technology rate).
-This assumes that every unit increase in effort is log-linearly greater than the unit of effort before it.
+When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data.
+[@rousseau2019] measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep).
+This rate of technology creep assumes that every unit increase in effort is log-linearly greater than the unit of effort before it.
 When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
@@
 
-Where $q_t$ can has a technology rate component $\tau$
+Where $q_t$ can has a technology creep component $\tau$
 
@@
 
-In addition to the results from `sraplus`, we include a set of results produced by the CMSY method [@froese2017].
-For computational efficiency, we used a ported version of the CMSY model available at https://github.com/DanOvando/portedcmsy.
+In addition to the results from sraplus, we include a set of results produced by the CMSY method [@froese2017].
+For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>.
 The only modification made is to convert the underlying population model to C++ for faster computation.
-For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [boettiger2012].
+For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012].
 Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
+
+In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation.
+The core different between CMSY and sraplus as utilized here then is that when we refer to CMSY we mean CMSY run using the default heuristics used by the algorithm to construct priors on initial, intermediate, and terminal stock status.
+In contrast, sraplus fit using for example FMI data uses the empirical models described here to generate priors on stock status.
+In addition, all sraplus runs use a prior-predictive tuning procedure to resolve issues related to Borel's paradox, described below.
 
@@
 This creates a problem for the Bayesian nature of our analysis though.
-Consider a production model with two parameters, a growth rate *r* and a carrying capacity *k*.
-Once we specify prior distributions on r and k, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *k* together with the model and the catch history produces a deterministic stock status in each time step.
+Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
+Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
 Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox.
 This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, particularly for the SRA version of `sraplus`, Borel's Paradox poses a particular problem.
+However, particularly for the SRA version of sraplus, Borel's Paradox poses a particular problem.
 
@@
 
-Along with allowing users to supply their own priors, the `sraplus` package contains three built-in methods used in this manuscript for converting information on stock status from additional outside data into a form usable as a prior by `sraplus`.
+Along with allowing users to supply their own priors, the sraplus package contains three built-in methods used in this manuscript for converting information on stock status from additional outside data into a form usable as a prior by sraplus.
 We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
 We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
-Given values of these variables for a new fishery then, `sraplus` uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by `sraplus`.
+Given values of these variables for a new fishery then, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus.
 
 All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
-The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by `sraplus`.
+The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.
 Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with the fishery status indicators of interest.
@@
 
+What sources of data might provide the greatest value in improving our estimates of global stock status?
+We fit 3,000 sraplus models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
+We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).
+
 The value-of-information (VOI) calculations presented in Fig.\@ref(fig:voi-plot) help illustrate what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The VOI analysis was performed by using `sraplus` to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
+The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
 We generate fits for 3000 combinations of a RAM stock and available data.
@@
 # References
-<div id="refs"></div>
+
+::: {#refs}
+:::
 
 # Supplemental Material
+
+```{r dat-desc}
+dat_description <- tribble(~`Data Name`, ~Description,
+        "RAM-Data", "Fit to abundance index from RAM",
+        "sar", "Prior on terminal F/Fmsy set by regional swept area ratio",
+        "fmi", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
+        "cpue", "Fit to CPUE index created from RAM catch and regional effort index. 2.6% technology creep",
+        "cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
+        "nominal-cpue", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
+         "nominal-cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
+        "catch-only", "priors on terminal B/Bmsy informed by catch history",
+        "heuristic", "priors on initial and terminal B/Bmsy informed by CMSY heuristc")
+
+knitr::kable(dat_description,
+             caption = "Data sources used for terminal stock status estimate",
+             format = "pipe")
+# %>%
+#   kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+#   kableExtra::row_spec(0, bold = T)
+```
 

### 2020-10-16 make-assessing-global-fisheries.R

@@ -894,13 +894,13 @@ tmp2 <- tmp2 %>%
   )
 
 ex_scatter_plot <- tmp2 %>% 
-  ggplot(aes(truth, value, size = lifetime_catch)) + 
+  ggplot(aes(truth, value)) + 
   geom_vline(aes(xintercept = 0)) + 
   geom_hline(aes(yintercept = 0)) +
   geom_abline(aes(slope = 1, intercept = 0),linetype = 2) +
   geom_point(alpha = 0.75) + 
   facet_grid(variable ~ data, scales = "free_y")  + 
-  scale_size(trans = "sqrt", name = "Lifetime Catch") +
+  # scale_size(trans = "sqrt", name = "Lifetime Catch") +
   scale_x_continuous(name = "RAM Value", expand = expansion(add = c(0, .1))) + 
   scale_y_continuous("Estimated Value", expand = expansion(add = c(0, .1)))
 
@@ -1938,7 +1938,7 @@ assess_ram_fits <- ram_status_fits %>%
 null_model <- assess_ram_fits %>% 
   filter(data == "cmsy") %>% 
   mutate(mean = sample(c(.4,1,1.6), n(), replace = TRUE)) %>% 
-  mutate(data = "Guess") %>% 
+  mutate(data = "guess") %>% 
   mutate(resid = ram_b_v_bmsy - mean,
          ae = abs(resid))
 

### 2020-10-20 documents/ovando-etal-assessing-global-fisheries-long.Rmd
submitted version

@@
 params:
-  results_name: ["v0.5"]
+  results_name: ["v1.0"]
   min_years_catch: [25]

### 2020-10-20 documents/ovando-etal-assessing-global-fisheries.Rmd
submitted version

@@
   Assessments of the global state of fisheries play an important role in tracking the implementation of the United Nations Sustainable Development Goals.
-While we have reliable estimates of stock status for fisheries accounting for 49% of global catch, our knowledge of the state of the remaining 51%, the worlds 'unassessed' fisheries, is poor.
+While we have reliable estimates of stock status for fisheries accounting for 49% of global catch, our knowledge of the state of the remaining 51%, the world's 'unassessed' fisheries, is poor.
 Numerous high-profile publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
@@
   mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
+  mutate(data = as.factor(data)) %>% 
+  mutate(data = forcats::fct_recode(data, 
+                                    "RAM Index" = "ram-data",
+                                    "SAR" = "sar",
+                                    "FMI" = "fmi",
+                                    "Effective CPUE" = "cpue",
+                                    "Effective CPUE+" = "cpue-plus",
+                                    "Nominal CPUE" = "nominal-cpue",
+                                    "Nominal CPUE+" ="nominal-cpue-plus",
+                                    "CMSY" = "cmsy",
+                                    "Guess" = "guess",
+                                    "RAM U/Umsy" = "u_umsy"
+
+                                    )) %>% 
   group_by(data) %>% 
@@
 2016 [@costello2016a] finds similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison.
-The super-erensemble method [@anderson2017a] used in Rosenberg et al.
+The super-ensemble method [@anderson2017a] used in Rosenberg et al.
 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific.
@@
 We use a flexible surplus-production based stock assessment software package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
-We show that our understanding of global fishery status is poor, and that improvements depend on an redoubled effort at global data collection.
+We show that our understanding of global fishery status is poor, and that improvements depend on a redoubled effort at global data collection and synthesis.
 
@@
 2019 [@rousseau2019] to each stock.
-We first used the default catch history heuristics internal to CMSY [@froese2017] to estimate stock status.
+As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available.
 We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus.
@@
 
-As a proof of concept, we also estimated B/B~MSY~ of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
+We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
 We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
-2017 [@froese2017] (See Table.5).
+2017 [@froese2017] (See Table.\@ref(tab:dat-desc)).
 We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
-Classification accuracy is calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
+Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
 
@@
 ```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent B/B~MSY~ by FAO statistical area from different data sources.
-ram-data refers to catch and abundance index drawn from RAM.
-CPUE refers to an index of abundance based on reconstructed effort data.
-cpue-plus uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, otherwise a 2.6% technology creep is assumed.
-ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
-fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio.
+RAM Index refers to catch and abundance index drawn from RAM.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
 CMSY uses the methods from Froese et al.
-2017 [@froese2017]."}
+2017 [@froese2017].
+Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
 ram_mpe_map_plot
@@
 ```{r mape-map, fig.cap = "Median absolute percent error in most recent B/B~MSY~ by FAO statistical area from different data sources.
-ram-data refers to catch and abundance index drawn from RAM.
-CPUE refers to an index of abundance based on reconstructed effort data.
-cpue-plus uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, otherwise a 2.6% technology creep is assumed.
-ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
-fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio.
+RAM Index refers to catch and abundance index drawn from RAM.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
 CMSY uses the methods from Froese et al.
-2017 [@froese2017]."}
+2017 [@froese2017].
+Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
 ram_mape_map_plot
@@
 ```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
-ram-data refers to catch and abundance index drawn from RAM.
-CPUE refers to an index of abundance based on reconstructed effort data.
-cpue-plus uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, otherwise a 2.6% technology creep is assumed.
-ram_u_umsy assumes all fisheries in the region share a common U/Umsy series with formally assessed fisheries in the region.
-fmi uses fmi scores to develop a prior on recent fishing mortality rates, sar does the same but based on swept area ratio.
+RAM Index refers to catch and abundance index drawn from RAM.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
 CMSY uses the methods from Froese et al.
-2017 [@froese2017]."}
+2017 [@froese2017].
+Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
 
@@
 
-The poor global performance of models fit to currently accesible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
+\newpage
+
+The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
 What sources of data might provide the greatest value in improving our estimates of global stock status?
@@
 That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the quality and use of data themselves.
+
+We chose to test the performance of methods against estimated values reported in RAM.
+A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed stocks in RAM.
+The clear and systemically poor performance of the methods tested here when applied to RAM stocks would require though that these methods have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks if they are to be expected to provide reasonable picture of global stock status.
+Free et al.
+2020's [@free2020] simulation testing of the types of methods included here suggests that this is unlikely to be true.
+z
 
@@
 f(x)=\begin{cases}
-B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
+B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)p_t
 , & \text{if $B_t>0.25 \times K$}.\\
-B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
+B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)\right)p_t, & \text{otherwise}.
 \end{cases}
@@
 ```{r param-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "Prior predictive tuning") %>%
-  rbind(c("Growth rate", "r", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
-  rbind(c("Shape parameter", "m", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
-  rbind(c("Catchability", "q", "$logn(1e^{-3}, 1)$")) %>%
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "$K$", `Default Prior` = "Prior predictive tuning") %>%
+  rbind(c("Growth rate", "$r$", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
+  rbind(c("Shape parameter", "$m$", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
+  rbind(c("Catchability", "$q$", "$logn(1e^{-3}, 1)$")) %>%
   rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,1)$")) %>%
@@
 
-$$cpue_t = \frac{catch_t}{(1 - e^{-f_t})}$$
+$$cpue_t = \frac{c_t}{(1 - e^{-f_t})}$$
 
@@
 
-$$log(cpue_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{C}), \sigma_{obs})$$
+$$log(cpue_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}), \sigma_{obs})$$
 
@@
 We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0.
+Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than zero.
 
@@
 Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
-Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox.
+Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
 This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
@@
 
-# Supplemental Material
+# Supplementary Information
+
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\setcounter{figure}{0}
+
+\renewcommand{\thetable}{S\arabic{table}}
+
+\setcounter{table}{0}
+
+## Data Sources Used
 
@@
 dat_description <- tribble(~`Data Name`, ~Description,
-        "RAM-Data", "Fit to abundance index from RAM",
-        "sar", "Prior on terminal F/Fmsy set by regional swept area ratio",
-        "fmi", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
-        "cpue", "Fit to CPUE index created from RAM catch and regional effort index. 2.6% technology creep",
-        "cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
-        "nominal-cpue", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
-         "nominal-cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
-        "catch-only", "priors on terminal B/Bmsy informed by catch history",
-        "heuristic", "priors on initial and terminal B/Bmsy informed by CMSY heuristc")
+        "RAM Index", "Fit to abundance index from RAM",
+        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
+        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
+        "EFfective CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 2.6% technology creep",
+        "Effective CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
+        "Nominal CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
+         "Nominal CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
+        "Guess", "Priors on terminal B/Bmsy randomly sampled from .4,1,1.6")
 
@@
 
-```{r ppplot, fig.cap="Prior-posterior plots of fits for case study fishery in Fig.6"}
+## Sample Prior Posterior Plots
+
+```{r ppplot, fig.cap="Prior posterior plots of fits for case study fishery in Fig.6"}
 
@@
 ```
+
+## Population Model
+
+The core of our model is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018.
+While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
+
+```{r si-prior-tab}
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
+  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
+  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
+  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
+  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
+  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))
+
+knitr::kable(
+  cd,
+  "latex",
+  align = "l",
+  booktabs = TRUE,
+  escape = F,
+  row.names = F,
+  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
+) %>%
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+```
+
+The population growth equation is
+
+\begin{equation}
+  f(x)=\begin{cases}
+      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
+, & \text{if $B_t>0.25 \times K$}.\\
+     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
+  \end{cases}
+  (\#eq:sihockey)
+\end{equation}
+
+## Prior Generating Regressions
+
+### Swept Area Ratio
+
+```{r sar-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of swept area ratio (SAR) on F/F~MSY~"}
+fit <- sraplus::sar_models$fit[sraplus::sar_models$metric == "mean_uumsy"][[1]]
+
+fit_data <- fit$data
+
+pp <- posterior_predict(fit, draws = 1000)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
+  scale_x_continuous(name = "Observed Log F/Fmsy") + 
+  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+
+```
+
+### Fisheries Management Index
+
+```{r fmi-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of fisheries management index (FMI) on F/F~MSY~"}
+
+fit <- sraplus::fmi_models$fit[sraplus::fmi_models$metric == "mean_uumsy"][[1]]
+
+fit_data <- fit$data
+
+pp <- posterior_predict(fit, draws = 1000)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
+  scale_x_continuous(name = "Observed Log F/Fmsy") + 
+  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+```
+
+## Catch History
+
+```{r catch-reg,fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) B/B~MSY~ for regression of catch on B/B~MSY~"}
+
+fit <- sraplus::catch_b_model
+
+fit_data <- fit$data %>% 
+  group_by(stockid) %>% 
+  filter(year == max(year))
+
+pp <- posterior_predict(fit, draws = 1000, newdata = fit_data)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_scatter_avg(y = fit_data$log_value, yrep = pp) + 
+  coord_flip() +
+  scale_y_continuous(name = "Observed Log B/Bmsy") + 
+  scale_x_continuous(name = "Posterior Predictive B/Bmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+```

### 2020-10-20 make-assessing-global-fisheries.R

@@ -52,7 +52,7 @@ n_cores <- 2
 
 # options(mc.cores = 1)
 
-results_name <- "v0.5"
+results_name <- "v1.0"
 
 
 results_description <-
@@ -899,7 +899,7 @@ ex_scatter_plot <- tmp2 %>%
   geom_hline(aes(yintercept = 0)) +
   geom_abline(aes(slope = 1, intercept = 0),linetype = 2) +
   geom_point(alpha = 0.75) + 
-  facet_grid(variable ~ data, scales = "free_y")  + 
+  facet_grid(variable ~ data, scales = "free_x")  + 
   # scale_size(trans = "sqrt", name = "Lifetime Catch") +
   scale_x_continuous(name = "RAM Value", expand = expansion(add = c(0, .1))) + 
   scale_y_continuous("Estimated Value", expand = expansion(add = c(0, .1)))
@@ -2028,6 +2028,19 @@ fao_area_ram_status %>%
   arrange(mape)
 
 
+ram_labeller <- c(
+  "ram-data" = "RAM Index",
+  "sar" = "SAR",
+  "fmi" = "FMI",
+  "cpue" = "Effective CPUE",
+  "cpue-plus" = "Effective CPUE+",
+  "nominal-cpue" = "Nominal CPUE",
+  "nominal-cpue-plus" = "Nominal CPUE+",
+  "cmsy" = "CMSY",
+  "guess" = "Guess",
+  "u_umsy" = "RAM U/Umsy"
+)
+
 ram_mpe_map_plot <- fao_area_ram_status %>%
   filter(!is.na(data)) %>% 
   mutate(data = fct_reorder(data, abs(mpe), .fun = median)) %>% 
@@ -2039,7 +2052,7 @@ ram_mpe_map_plot <- fao_area_ram_status %>%
     color = "black",
     size = 0.01
   ) +
-  facet_wrap(~ data) +
+  facet_wrap(~ data, labeller = labeller(data = ram_labeller)) +
   scale_fill_gradient2(
     low = "steelblue",
     high = "tomato",
@@ -2072,7 +2085,7 @@ ram_mape_map_plot <- fao_area_ram_status %>%
     color = "black",
     size = 0.01
   ) +
-  facet_wrap(~ data) +
+  facet_wrap(~ data, labeller = labeller(data = ram_labeller)) +
   scale_fill_gradient(
     low = "white",
     # mid = "orange",
[…]

### 2020-10-20 make-assessing-global-fisheries.R
Longo edits (#8)
* submitted version

@@ -52,7 +52,7 @@ n_cores <- 2
 
 # options(mc.cores = 1)
 
-results_name <- "v0.5"
+results_name <- "v1.0"
 
 
 results_description <-
@@ -894,13 +894,13 @@ tmp2 <- tmp2 %>%
   )
 
 ex_scatter_plot <- tmp2 %>% 
-  ggplot(aes(truth, value, size = lifetime_catch)) + 
+  ggplot(aes(truth, value)) + 
   geom_vline(aes(xintercept = 0)) + 
   geom_hline(aes(yintercept = 0)) +
   geom_abline(aes(slope = 1, intercept = 0),linetype = 2) +
   geom_point(alpha = 0.75) + 
-  facet_grid(variable ~ data, scales = "free_y")  + 
-  scale_size(trans = "sqrt", name = "Lifetime Catch") +
+  facet_grid(variable ~ data, scales = "free_x")  + 
+  # scale_size(trans = "sqrt", name = "Lifetime Catch") +
   scale_x_continuous(name = "RAM Value", expand = expansion(add = c(0, .1))) + 
   scale_y_continuous("Estimated Value", expand = expansion(add = c(0, .1)))
 
@@ -1938,7 +1938,7 @@ assess_ram_fits <- ram_status_fits %>%
 null_model <- assess_ram_fits %>% 
   filter(data == "cmsy") %>% 
   mutate(mean = sample(c(.4,1,1.6), n(), replace = TRUE)) %>% 
-  mutate(data = "Guess") %>% 
+  mutate(data = "guess") %>% 
   mutate(resid = ram_b_v_bmsy - mean,
          ae = abs(resid))
 
@@ -2028,6 +2028,19 @@ fao_area_ram_status %>%
   arrange(mape)
 
 
+ram_labeller <- c(
+  "ram-data" = "RAM Index",
+  "sar" = "SAR",
+  "fmi" = "FMI",
+  "cpue" = "Effective CPUE",
+  "cpue-plus" = "Effective CPUE+",
+  "nominal-cpue" = "Nominal CPUE",
+  "nominal-cpue-plus" = "Nominal CPUE+",
+  "cmsy" = "CMSY",
+  "guess" = "Guess",
+  "u_umsy" = "RAM U/Umsy"
+)
+
 ram_mpe_map_plot <- fao_area_ram_status %>%
   filter(!is.na(data)) %>% 
   mutate(data = fct_reorder(data, abs(mpe), .fun = median)) %>% 
[…]

### 2020-12-09 documents/ovando-etal-assessing-global-fisheries.Rmd
nature comms resubmission

@@
 title: |
-  | Global Assessments of Fishery Status need 
-  | Better Data more than Better Models
+  | Status of Global Unassessed Fisheries will Remain Highly Uncertain
+  | without Better Data
 author: 
@@
 bibliography: ["../references.bib"]
-csl: nature.csl
+csl: nature-communications.csl
 output: 
-  bookdown::pdf_document2: default
+  bookdown::pdf_document2:
+    latex_engine: xelatex
   bookdown::word_document2:
-    reference_docx: word-template.docx
+    reference_docx: SA_MS_Template_2020.docx
 params:
@@
   Assessments of the global state of fisheries play an important role in tracking the implementation of the United Nations Sustainable Development Goals.
-While we have reliable estimates of stock status for fisheries accounting for 49% of global catch, our knowledge of the state of the remaining 51%, the world's 'unassessed' fisheries, is poor.
+While we have reliable estimates of stock status for fisheries accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fisheries remains highly uncertain.
 Numerous high-profile publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
-Here, we show that despite numerous efforts, our understanding of the status of global fisheries remains poor, even when new sources of broadly available data are added.
-Obtaining accurate estimates of stock status for the world's fisheries depends on prioritizing the collection of high-priority data at around the globe, rather than the development of new modeling methods alone.
+Here, we show that despite numerous efforts, our understanding of the status of global fisheries remains incomplete, even when new sources of broadly available data are added.
+Obtaining accurate estimates of stock status for the world's fisheries depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
 always_allow_html: true
 linkcolor: blue
+# header-includes:
+#   - \usepackage{draftwatermark}
+#   - \SetWatermarkText{DRAFT}
 ---
@@
 <!-- https://www.nature.com/natsustain/about/content -->
+
 <!-- # check in call notes -->
@@
 Meeting these targets depends in part on our ability to effectively track the status of global marine fish stocks.
-The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the most widely used source for tracking the global state of fisheries.
+The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the dedicated SDG 14 Indicator on fisheries sustainability, and the FAO's report on the State of World Fisheries and Aquaculture (SOFIA) report is the most widely used source for tracking the global state of fisheries.
 As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020].
-The SOFIA assessment was designed in the 1970s based on the then available data and methods, With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods, fresh approaches are needed to meet the new demand for tracking progress towards the SDG goals and supporting the research targets set out in the UNited Nations Decade on Oceans and Science plan.
-
-Forty nine percent of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database [@ricard2012], a repository of best-available estimates of fishery status [@hilborn2020].
-The SOFIA process uses these formal assessments for their determinations of stock status wherever possible.
-However, that leaves 51% of global fisheries landings lacking in formal stock assessments.
-While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-
-The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion.
-While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable, in order to increase reproducibility, transparency, and potential reliability.
+The SOFIA assessment was designed in the 1970s based on the then available data and methods, With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods,the SOFIA global assessment methods need to be further strengthened to meet the demand for estimation and tracking progress towards the SDG goals.
+
+The FAO's SOFIA index captures 70% of the landings of all fisheries in the world, with statements on on individual stock status made on the basis of formal stock assessments where available, and through a combination of other methods where not [@fao2020].
+In this paper we evaluate the ability of alternative data-limited assessment models to improve the methods used by the FAO for these unassessed stocks.
+While these unassessed stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
+
+The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed.
+While these methods combined with local knowledge can be well informed as to the status of their fish stocks, a more quantitatively vetted process would be desirable in order to increase reproducibility, transparency, and potential reliability.
 Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
@@
 2016 [@costello2016a] finds similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison.
-The super-ensemble method [@anderson2017a] used in Rosenberg et al.
+The superensemble method [@anderson2017a] used in Rosenberg et al.
 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific.
-The Pauly 2007 [@pauly2007] catch based approach finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
+The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
 The methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
@@
 We use a flexible surplus-production based stock assessment software package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
-We show that our understanding of global fishery status is poor, and that improvements depend on a redoubled effort at global data collection and synthesis.
+We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis.
 
@@
 
-+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
 | Data Source           |  Short Name | Data Use |  Caveats |
-+=============================================+============+==========================================================================+=================================================================================================================+
+|-----------------------|-------------|------|------|
 | Catch data [@fao2020] | catches     | Priors on stock status, scaling of population size, exploitation history    |    Heuristics or regressions used to translate shape of catch history into priors on stock status  |
-+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Fisheries Management Index [@melnychuk2017] | FMI        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
-+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Swept Area Ratio [@amoroso2018]             | SAR        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
-+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Reconstructed effort data[@rousseau2019b]   | effort     | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors.
-Assumed rate of technology creep reported in individual sections |
-+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-
-: (\#tab:data) Broadly available data used to inform estimates of global stock status
+|    Fisheries Management Index [@melnychuk2017]     | FMI     | Priors on most recent F/F~MSY~ values   |Priors produced by regression trained on data from RAM Legacy Stock Assessment Database|
+| Swept Area Ratio [@amoroso2018]      | SAR     | Priors on most recent F/F~MSY~ values |Priors produced by regression trained on data from RAM Legacy Stock Assessment Database|
+|  Reconstructed effort data [@rousseau2019b] | effort | Combined with catch data to create an index of abundance| Total reconstructed effort across all sectors.
+Assumed rate of technology creep reported in individual sections|
+Table:(\#tab:data) Data sources included across model fits.
 
@@
 Lastly, we used the reconstructed effort data from [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
-While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results (Fig.\@ref(fig:cs-plot)).
+While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both B/B~MSY~ and F/F~MSY~ (Fig.\@ref(fig:cs-plot)).
 
@@
 We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
 B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
@@
 Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
-The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`.
+The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`.
 Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.
@@
 Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or wept area ratio (SAR) data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
 For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
@@
 Global-level assessment are critical for guiding management agendas for the world's oceans, and tracking critical indicators such as the United Nations Sustainable Development Goals.
-Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world.
-While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).
+Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains uncertain in many parts of the world.
+While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status (Table.\@ref(tab:perf-tab)).
 
@@
 2020's [@free2020] simulation testing of the types of methods included here suggests that this is unlikely to be true.
-z
 
@@
 While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
-Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f;@rudd2017].
+Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f; @rudd2017].
 
@@
 
-The core of our model is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018.
+The core of our model is a Pella-Tomlinson [@pella1969] production model in the manner of [@winker2018].
 While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.

### 2020-12-09 documents/test.Rmd
nature comms resubmission

---
title: |
  | Status of Unassessed Fish Stocks will Remain Highly Uncertain
  | until Better Data are Available
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: SA_MS_Template_2020.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fisheries remains highly uncertain Numerous high-profile publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. Here, we show that despite numerous efforts, our understanding of the status of global fisheries remains incpmplete, even when new sources of broadly available data are added. Obtaining accurate estimates of stock status for the world's fisheries depends on prioritizing the collection of high-priority data at around the globe, rather than the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
header-includes:
  - \usepackage{draftwatermark}
  - \SetWatermarkText{DRAFT}
---

@fao2020

### 2021-01-12 documents/ovando-etal-assessing-global-fisheries.Rmd
trimming methods back down

@@
 ---
-title: |
-  | Status of Global Unassessed Fisheries will Remain Highly Uncertain
-  | without Better Data
+title: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
 author: 
@@
 bibliography: ["../references.bib"]
-csl: nature-communications.csl
+csl: proceedings-of-the-royal-society-b.csl
 output: 
+  bookdown::word_document2:
+    reference_docx: template.docx
   bookdown::pdf_document2:
     latex_engine: xelatex
-  bookdown::word_document2:
-    reference_docx: SA_MS_Template_2020.docx
 params:
@@
 abstract: |
-  Assessments of the global state of fisheries play an important role in tracking the implementation of the United Nations Sustainable Development Goals.
-While we have reliable estimates of stock status for fisheries accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fisheries remains highly uncertain.
+  Assessments of the global state of fish populations play an important role in tracking the implementation of the United Nations Sustainable Development Goals.
+While we have reliable estimates of stock status for fisheries accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain.
 Numerous high-profile publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
-Here, we show that despite numerous efforts, our understanding of the status of global fisheries remains incomplete, even when new sources of broadly available data are added.
+Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added.
 Obtaining accurate estimates of stock status for the world's fisheries depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
@@
 linkcolor: blue
-# header-includes:
-#   - \usepackage{draftwatermark}
-#   - \SetWatermarkText{DRAFT}
 ---
@@
 
+data limited are not just messier versions of data rich
+
+80% in trouble
+
 <!-- # rishi notes -->
@@
 
-# Introduction
-
-The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
-Meeting these targets depends in part on our ability to effectively track the status of global marine fish stocks.
-The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the dedicated SDG 14 Indicator on fisheries sustainability, and the FAO's report on the State of World Fisheries and Aquaculture (SOFIA) report is the most widely used source for tracking the global state of fisheries.
-As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020].
-The SOFIA assessment was designed in the 1970s based on the then available data and methods, With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods,the SOFIA global assessment methods need to be further strengthened to meet the demand for estimation and tracking progress towards the SDG goals.
-
-The FAO's SOFIA index captures 70% of the landings of all fisheries in the world, with statements on on individual stock status made on the basis of formal stock assessments where available, and through a combination of other methods where not [@fao2020].
-In this paper we evaluate the ability of alternative data-limited assessment models to improve the methods used by the FAO for these unassessed stocks.
-While these unassessed stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-
-The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed.
-While these methods combined with local knowledge can be well informed as to the status of their fish stocks, a more quantitatively vetted process would be desirable in order to increase reproducibility, transparency, and potential reliability.
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
+# Introduction {.unnumbered}
+
+The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020].
+Meeting the SDG 14 targets depends in part on improving our ability to effectively track the status of global marine fish populations.
+
+The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's report on the State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries.
+The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020].
+Where possible, statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database [@ricard2012].
+The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the biological state of "assessed" fish populations around the globe.
+
+That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed".
+The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020].
+While these methods combined with local knowledge can be well informed as to the status of their fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods.
+With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods,the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
+
+In this paper we evaluate the ability of alternative data-limited assessment models to improve our estimate of the population status of unassessed fish stocks around the world.
+While these unassessed stocks are generally individually smaller than the typically larger and more valuable fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
 Due to data limitations, all of these global assessment efforts use forms of "catch-only" stock assessment models (Free et al.
 2020 [@free2020] and references therein).
-These models seek to infer stock status, for example in terms of biomass *B* relative to the biomass at maximum sustainable yield *B~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
+These models seek to infer stock status, for example biomass B relative to the biomass at maximum sustainable yield B~MSY~, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
 
@@
 The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
-The methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
+The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
 However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.
@@
 
-In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fisheries?
-We use a flexible surplus-production based stock assessment software package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
+In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fish populations?
+We use a flexible surplus-production based stock assessment software package, [sraplus](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
 We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis.
 
-# Results
+# Methods {.unnumbered}
+
+We compared estimates of stock status, defined by B/B~MSY~ obtained from sraplus using various forms of broadly available data to the best-available estimates of B/B~MSY~ available for those fish populations reported in the RAM database.
+All analysis were conducted in the R programming language [@rcoreteam2019].
+Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018].
+The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
+We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus.
+Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.
+
+## Data Sources
+
+We restricted ourselves to databases with wide coverage around the world, representing the types of information that can be leveraged in efforts to assess the status of global fisheries (Table.\@ref(tab:data)).
+While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for assessment of global unassessed fisheries.
 
@@
 
+## Experiment Structure
+
+We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global fish populations.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
+B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
+We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
+This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
+
+We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
+We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
+2017 [@froese2017] (See Table.\@ref(tab:dat-desc)).
+We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
+Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
+
+## Population Model {.unnumbered}
+
+The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of [@winker2018].
+While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
+The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status
+
+The population growth equation is
+
+\begin{equation}
+f(x)=\begin{cases}
+B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)p_t
+, & \text{if $B_t>0.25 \times K$}.\\
+B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)\right)p_t, & \text{otherwise}.
+\end{cases}
+(\#eq:pt)
+\end{equation}
+
+Where $B_t$ is biomass at time *t*, *K* is carrying capacity ,*r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of *B~MSY~/K* to shift.
+When *m* is two *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right.
+$\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors.
+Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
+We deal with this problem by following the methods described in [@winker2018].
+to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
+
+We allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by [@walters2006]).
+This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
+Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).
+
+Process error *p* is assumed to be log-normally distributed, such that
+
+\begin{equation}
+  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
+(\#eq:procerror)
+\end{equation}
+
+## Estimation Model {.unnumbered}
+
+All of our estimates are Bayesian in nature.
+sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent).
+Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
+A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
+sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~.
+We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history.
+Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected.
+The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
+Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis.
+All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
+
+<!-- .
+We can break the use of sraplus into two distinct categories: with data to fit to and without.
+By "data", we refer to measurements which are used to confront model estimates within a likelihood function.
+In our context, these include fishery-independent survey data, or a CPUE index.
+When there are no data to fit to, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
+Under this mode, the model is essentially a stock-reduction analysis model, in the manner of [@walters2006], in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e.
+results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches.
+This step updates the prior distribution of population parameters by eliminating combinations of priors that are impossible for a given catch history and a specified functional form.
+-->
+
+The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).
+*r* and *K* are the only two parameters that are always estimated.
+Estimation of every other parameter can be turned on or off.
+When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user.
+For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*.
+*q* is also estimated when needed.
+
+```{r param-tab}
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "$K$", `Default Prior` = "Prior predictive tuning") %>%
+  rbind(c("Growth rate", "$r$", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
+  rbind(c("Shape parameter", "$m$", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
+  rbind(c("Catchability", "$q$", "$logn(1e^{-3}, 1)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,1)$")) %>%
+  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.5,0.25)$")) %>%
+  rbind(c("Initial State", "$B0$", "Posterior probability dist. of catch-based regressions"))
+
+knitr::kable(
+  cd,
+  align = "l",
+  escape = F,
+  row.names = F,
+  caption = "Name, abbreviations, and priors distribution for parameters potentially estimated by sraplus in this manuscript.",
+  format = "pipe"
+) %>%
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+```
+
+<!-- sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent).
+Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
+A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
+sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~.
+We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history.
+Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected.
+The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
+Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis.
+All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
+-->
+
+When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018].
+By default the model uses 2000 draws with a 1000 step warm-up and one chain.
+Any detailed fit for a particular fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria.
+The model fits to a direct estimate of abundance (e.g.
+a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood is
+
+$$log(a_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}) \times q, sigma_{obs})$$
+
+where *f* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
+When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data.
+[@rousseau2019] measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep).
+This rate of technology creep assumes that every unit increase in effort is log-linearly greater than the unit of effort before it.
+When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
+This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
+To accommodate this, we generate a catch per effective harvest rate index of abundance, as
+
+$$cpue_t = \frac{c_t}{(1 - e^{-f_t})}$$
+
+$$f_t = q_tE_t$$
+
+Where $q_t$ can has a technology creep component $\tau$
+
+$$q_t = q_{t-1} \times (1  + \tau)$$
+
+We then fit to the index of abundance per
+
+$$log(cpue_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}), \sigma_{obs})$$
+
+### CMSY {.unnumbered}
+
+In addition to the results from sraplus, we include a set of results produced by the CMSY method [@froese2017].
+For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>.
+The only modification made is to convert the underlying population model to C++ for faster computation.
+For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012].
+Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
+
+In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation.
+The core different between CMSY and sraplus as utilized here then is that when we refer to CMSY we mean CMSY run using the default heuristics used by the algorithm to construct priors on initial, intermediate, and terminal stock status.
+In contrast, sraplus fit using for example FMI data uses the empirical models described here to generate priors on stock status.
+In addition, all sraplus runs use a prior-predictive tuning procedure to resolve issues related to Borel's paradox, described below.
+
+## Priors {.unnumbered}
+
+The shape parameter is usually not reliably estimable given available data for surplus production models, however, Thorson et al.
+2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa.
+While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on Thorson et al.
+2012 [@thorson2012] for the genus of the species in question.
+
+We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus.
+
+### Prior Predictive Tuning {.unnumbered}
+
+Suppose that the only thing we observe from a fishery is a catch history.
+Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
+We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
+Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.
+
+In the absence of any data to fit to, the SRA algorithm works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief.
+This creates a problem for the Bayesian nature of our analysis though.
+Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
+Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
+Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
+This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
+However, particularly for the SRA version of sraplus, Borel's Paradox poses a particular problem.
+
+The SRA algorithm works in two steps.
+First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep).
+From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status.
+If the bulk of the prior on terminal stock status was concentrated at 50% of *K*, combinations of *r* and *K* that produce terminal stock status near 50% of *K* are sampled proportionally more frequently.
+However, lower values of terminal stock status have fewer candidate values of *r* and *K*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
+Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *K* combinations: since under this model the population cannot be greater than carrying capacity, as for example *K* approaches infinity terminal stock status asymptotes at close to 100% of *K*.
+The net result of this is that even though individual combinations of *r* and *K* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
+As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.
+
+We use an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000].
+Our solution amounts to a two-step SIR algorithm.
+We first run the standard SIR algorithm as described above.
+We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
+
+$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
+
+We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
+
+$$p(n_i) = \frac{p(bin_i)}{N_i}$$
+
+And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
+The net result of this is a post-model-pre-data distribution of parameters *r* and *K* that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
+In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
+This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.S1).
+
+```{r bias-plot, fig.cap = "Post-model-pre-data distribution of depletion (biomass relative to carrying capacity) from raw SRA algorithm (untuned, top row), from SRA algorithm with approximate tuning applied (tuned, middle row), compared to the supplised prior on depletion (bottom row).
+Black vertical line indicates median value.
+"}
+
+id = unique(ram_comp_data$stockid)
+
+ex <- ram_data %>% 
+  filter(stockid %in% id[[1]])
+
+# fit using catch heuristic
+
+co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
+                                 catch = ex$catch, 
+                                 years = ex$year,
+                                 terminal_state = 0.25,
+                                 terminal_state_cv = 0.5,
+                                 b_ref_type = "k")
+
+biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
+                                      estimate_shape = FALSE, estimate_proc_error = TRUE)
+
+biased_depletion <- biased_co_fit$fit %>% 
+  filter(variable == "dep_t") %>%
+  filter(year == max(year)) %>% 
+  mutate(set = "Untuned Post-Model-Pre-Data") %>% 
+  select(year, value, set)
+
+co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
+                                      estimate_shape = FALSE, 
+                               estimate_proc_error = TRUE)
+
+tuned_depletion <- co_fit$fit %>% 
+  filter(year == max(year), variable == "dep_t") %>% 
+  mutate(set = "Tuned Post-Model-Pre-Data") %>% 
+  select(year, value, set)
+
+prior <- tuned_depletion %>% 
+  mutate(value = rlnorm(n(), log(0.25), 0.5)) %>% 
+    mutate(set = "Prior")
+
+comps <- biased_depletion %>% 
+  bind_rows(tuned_depletion) %>% 
+  bind_rows(prior) %>% 
+  group_by(set) %>% 
+  mutate(valrank = percent_rank(value)) %>% 
+  filter(between(valrank, .05, .95))
+
+comps %>% 
+  ggplot(aes(x = value, y = set)) + 
+  ggridges::geom_density_ridges(alpha = 0.9,
+                                quantile_lines = TRUE, quantiles = 2) +
+  scale_x_continuous(limits = c(0, 1.2), name = "Depletion (B/K)") + 
+  scale_y_discrete(name = '')
+
+```
+
+### Priors Informed by Outside Data {.unnumbered}
+
+Along with allowing users to supply their own priors, the sraplus package contains three built-in methods used in this manuscript for converting information on stock status from additional outside data into a form usable as a prior by sraplus.
+We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
+We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
+Given values of these variables for a new fishery then, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus.
+
+All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
+The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.
+Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with the fishery status indicators of interest.
+Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.
+
+#### Catch-Only Priors {.unnumbered}
+
+Many of the current methods for estimating global stock status of unassessed stocks are based on predicting stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
+While these catch-only methods have been shown to have serious shortcomings [@free2020], we include them as a point of reference given their ubiquity in the global assessment literature.
+
+We used data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
+To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
+Emergent clusters show for example one built around a downward "one way trip" style catch histories, others with a boom and bust pattern, others with stable but fluctuating catches.
+
+We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history.
+This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term).
+For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status
+
+$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma)$$
+
+For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data
+
+$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
+
+where *fyear* is the year of the fishery, starting from 0.
+
+#### Fisheries Management Index Priors {.unnumbered}
+
+The Fisheries Management Index (FMI), as presented in [@melnychuk2017], utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
+These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
+The higher the score, the better the expert judges that a given metric is met in that fishery.
+Importantly, FMI surveys can be filled out in the absence of stock assessments.
+This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to [@osio2015] and [@cope2015]).
+
+The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
+
+$$log(value_i) \sim N(s(research_i) + s(management_i) +
+s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
+
+#### Swept Area Ratio Priors {.unnumbered}
+
+[@amoroso2018] provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
+This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
+As illustrated in [@amoroso2018], there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
+Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
+if all trawl-able areas are trawled twice a year then the SAR will be 2.
+Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
+
+The final selected model relating SAR to to stock status metrics was
+
+$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
+
+## Value of Information Calculations {.unnumbered}
+
+<!-- To determine what sources of data might provide the greatest value in improving our estimates of global stock status, we fit 3,000 sraplus models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
+We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).
+-->
+
+We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
+The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
+We generate fits for 3000 combinations of a RAM stock and available data.
+For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
+For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
+The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
+
+Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
+We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
+
+$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
+
+Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
+
+# Results {.unnumbered}
+
+## Case Study  {.unnumbered}
+
 We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
@@
 
-We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
+## Global Performance  {.unnumbered}
+
+<!-- We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
 We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
@@
 This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
-
-We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
+-->
+
+<!-- We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
 We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
@@
 Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
+-->
 
@@
 
-Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves.
-This becomes clearer once we consider the performance of sraplus models fit to combinations of our broadly available datasets.
+Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves.
+This becomes clearer once we consider the performance of sraplus models fit to combinations of broadly available datasets.
 Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)).
@@
 
-# Discussion
+# Discussion {.unnumbered}
 
@@
 
-Global-level assessment are critical for guiding management agendas for the world's oceans, and tracking critical indicators such as the United Nations Sustainable Development Goals.
-Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains uncertain in many parts of the world.
-While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status (Table.\@ref(tab:perf-tab)).
+Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking vital indicators such as the United Nations Sustainable Development Goals.
+Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fish stocks remains uncertain in many parts of the world.
+While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
 
@@
 It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
-In some instances using the data presented here did provide some improvement over use of catch-only style methods; the difficulty comes in attempting to apply data types uniformly across the globe.
-While it is unreasonable to expect any global-scale data to be able to perform as well as detailed stock assessments reported in RAM, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
-That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the quality and use of data themselves.
+In some instances using the data presented here did provide some improvement over use of catch-only style assessment methods; the difficulty comes in attempting to apply data types uniformly across the globe.
+While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RAM, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
+That use of none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the quality and use of data themselves.
 
@@
 A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed stocks in RAM.
-The clear and systemically poor performance of the methods tested here when applied to RAM stocks would require though that these methods have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks if they are to be expected to provide reasonable picture of global stock status.
+The clear and systemically poor performance of the methods tested here when applied to RAM stocks would require though that these methods have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks if they are to be expected to provide reasonable picture of global fish populations.
 Free et al.
@@
 Effort data such as those reconstructed by Rousseau et al.
-2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
+2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
 Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
@@
 But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
-Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
+Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas of the globe results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
 
 Our value-of-information analysis also shows though the high utility of having access to even a recent snapshot of F/F~MSY~ (Fig.\@ref(fig:voi-plot)).
-Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
+Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct fishery-specific priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
 Another avenue would be to prioritize the development of a global repository for length and age composition data.
@@
 
+# Conclusions {.unnumbered}
+
 The coming decades are a critical time for the future of fisheries and ocean health.
@@
 The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
-However, this process still leaves a substantial number of fisheries and global catch unassessed.
+However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed.
 Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a clear solution to this problem.
@@
 
-# Methods
-
-All analysis were conducted in the R programming language [@rcoreteam2019].
-Model fitting was conducted using Rcpp [@eddelbuettel2011] and stan [@carpenter2017], implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018].
-The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
-
-sraplus contains too many options to cover within this methods section.
-We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus.
-Below we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.
-
-## Population Model
-
-The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of [@winker2018].
-While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
-The purpose of sraplus is not to make substantial improvements in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status
-
-The population growth equation is
-
-\begin{equation}
-f(x)=\begin{cases}
-B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)p_t
-, & \text{if $B_t>0.25 \times K$}.\\
-B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)\right)p_t, & \text{otherwise}.
-\end{cases}
-(\#eq:pt)
-\end{equation}
-
-Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of B~MSY~/K to shift.
-When *m* is two B~MSY~ / K = 0.5, lower values of *m* shift the production function left, higher values right.
-$\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors.
-Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
-We deal with this problem by following the methods described in [@winker2018].
-to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
-
-We allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by [@walters2006]).
-This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
-Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).
-
-Process error *p* is assumed to be log-normally distributed, such that
-
-\begin{equation}
-  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
-(\#eq:procerror)
-\end{equation}
-
-## Estimation Model
-
-All of our estimates are Bayesian in nature.
-We can break the use of sraplus into two distinct categories: with data and without.
-By "data", we refer to measurements which are used to confront model estimates within a likelihood function.
-In our context, these include fishery-independent survey data, or a CPUE index.
-When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
-Under this mode, the model is essentially a stock-reduction analysis model, in the manner of [@walters2006], in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e.
-results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches.
-This step updates the prior distribution of population parameters by eliminating combinations of priors that are impossible for a given catch history and a specified functional form.
-
-The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).
-*r* and *K* are the only two parameters that are always estimated.
-Estimation of every other parameter can be turned on or off.
-When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user.
-For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*.
-*q* is also estimated when needed.
-
-```{r param-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "$K$", `Default Prior` = "Prior predictive tuning") %>%
-  rbind(c("Growth rate", "$r$", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
-  rbind(c("Shape parameter", "$m$", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
-  rbind(c("Catchability", "$q$", "$logn(1e^{-3}, 1)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,1)$")) %>%
-  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.5,0.25)$")) %>%
-  rbind(c("Initial State", "$B0$", "Posterior probability dist. of catch-based regressions"))
-
-knitr::kable(
-  cd,
-  align = "l",
-  escape = F,
-  row.names = F,
-  caption = "Name, abbreviations, and priors distribution for parameters potentially estimated by sraplus in this manuscript.",
-  format = "pipe"
-) %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
-
-sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent).
-Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
-A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
-sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~.
-We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history.
-Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected.
-The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
-Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis.
-All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
-
-When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018].
-By default the model uses 2000 draws with a 1000 step warm-up and one chain.
-Any detailed fit for a particular fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria.
-The model fits to a direct estimate of abundance (e.g.
-a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood is
-
-$$log(a_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}) \times q, sigma_{obs})$$
-
-where *f* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
-When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data.
-[@rousseau2019] measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep).
-This rate of technology creep assumes that every unit increase in effort is log-linearly greater than the unit of effort before it.
-When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
-This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
-To accommodate this, we generate a catch per effective harvest rate index of abundance, as
-
-$$cpue_t = \frac{c_t}{(1 - e^{-f_t})}$$
-
-$$f_t = q_tE_t$$
-
-Where $q_t$ can has a technology creep component $\tau$
-
-$$q_t = q_{t-1} \times (1  + \tau)$$
-
-We then fit to the index of abundance per
-
-$$log(cpue_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}), \sigma_{obs})$$
-
-### CMSY
-
-In addition to the results from sraplus, we include a set of results produced by the CMSY method [@froese2017].
-For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>.
-The only modification made is to convert the underlying population model to C++ for faster computation.
-For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012].
-Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
-
-In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation.
-The core different between CMSY and sraplus as utilized here then is that when we refer to CMSY we mean CMSY run using the default heuristics used by the algorithm to construct priors on initial, intermediate, and terminal stock status.
-In contrast, sraplus fit using for example FMI data uses the empirical models described here to generate priors on stock status.
-In addition, all sraplus runs use a prior-predictive tuning procedure to resolve issues related to Borel's paradox, described below.
-
-## Priors
-
-Priors for all estimable parameters can be left at default values or set by the user.
-The shape parameter is usually not reliably estimable given available data for surplus production models, however, Thorson et al.
-2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa.
-While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on Thorson et al.
-2012 [@thorson2012] for the genus of the species in question.
-
-We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus.
-
-### Prior Predictive Tuning
-
-Suppose that the only thing we observe from a fishery is a catch history.
-Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
-We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than zero.
-
-In the absence of any data to fit to, the SRA algorithm works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief.
-This creates a problem for the Bayesian nature of our analysis though.
-Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
-Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
-Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
-This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, particularly for the SRA version of sraplus, Borel's Paradox poses a particular problem.
-
-The SRA algorithm works in two steps.
-First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep).
-From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status.
-If the bulk of the prior on terminal stock status was concentrated at 50% of *K*, combinations of *r* and *K* that produce terminal stock status near 50% of *K* are sampled proportionally more frequently.
-However, lower values of terminal stock status have fewer candidate values of *r* and *K*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
-Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *K* combinations: since under this model the population cannot be greater than carrying capacity, as for example *K* approaches infinity terminal stock status asymptotes at close to 100% of *K*.
-The net result of this is that even though individual combinations of *r* and *K* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
-As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.
-
-We use an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000].
-Our solution amounts to a two-step SIR algorithm.
-We first run the standard SIR algorithm as described above.
-We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
-
-$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
-
-We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
-
-$$p(n_i) = \frac{p(bin_i)}{N_i}$$
-
-And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
-The net result of this is a post-model-pre-data distribution of parameters *r* and *K* that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
-In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
-This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.S1).
-
-```{r bias-plot, fig.cap = "Post-model-pre-data distribution of depletion (biomass relative to carrying capacity) from raw SRA algorithm (untuned, top row), from SRA algorithm with approximate tuning applied (tuned, middle row), compared to the supplised prior on depletion (bottom row).
-Black vertical line indicates median value.
-"}
-
-id = unique(ram_comp_data$stockid)
-
-ex <- ram_data %>% 
-  filter(stockid %in% id[[1]])
-
-# fit using catch heuristic
-
-co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
-                                 catch = ex$catch, 
-                                 years = ex$year,
-                                 terminal_state = 0.25,
-                                 terminal_state_cv = 0.5,
-                                 b_ref_type = "k")
-
-biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
-                                      estimate_shape = FALSE, estimate_proc_error = TRUE)
-
-biased_depletion <- biased_co_fit$fit %>% 
-  filter(variable == "dep_t") %>%
-  filter(year == max(year)) %>% 
-  mutate(set = "Untuned Post-Model-Pre-Data") %>% 
-  select(year, value, set)
-
-co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
-                                      estimate_shape = FALSE, 
-                               estimate_proc_error = TRUE)
-
-tuned_depletion <- co_fit$fit %>% 
-  filter(year == max(year), variable == "dep_t") %>% 
-  mutate(set = "Tuned Post-Model-Pre-Data") %>% 
-  select(year, value, set)
-
-prior <- tuned_depletion %>% 
-  mutate(value = rlnorm(n(), log(0.25), 0.5)) %>% 
-    mutate(set = "Prior")
-
-comps <- biased_depletion %>% 
-  bind_rows(tuned_depletion) %>% 
-  bind_rows(prior) %>% 
-  group_by(set) %>% 
-  mutate(valrank = percent_rank(value)) %>% 
-  filter(between(valrank, .05, .95))
-
-comps %>% 
-  ggplot(aes(x = value, y = set)) + 
-  ggridges::geom_density_ridges(alpha = 0.9,
-                                quantile_lines = TRUE, quantiles = 2) +
-  scale_x_continuous(limits = c(0, 1.2), name = "Depletion (B/K)") + 
-  scale_y_discrete(name = '')
-
-```
-
-### Priors Informed by Outside Data
-
-Along with allowing users to supply their own priors, the sraplus package contains three built-in methods used in this manuscript for converting information on stock status from additional outside data into a form usable as a prior by sraplus.
-We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
-We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
-Given values of these variables for a new fishery then, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus.
-
-All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
-The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.
-Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with the fishery status indicators of interest.
-Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.
-
-#### Catch-Only Priors
-
-Many of the current methods for estimating global stock status of unassessed stocks are based on predicting stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
-While these catch-only methods have been shown to have serious shortcomings [@free2020], we include them as a point of reference given their ubiquity in the global assessment literature.
-
-We used data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
-To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
-Emergent clusters show for example one built around a downward "one way trip" style catch histories, others with a boom and bust pattern, others with stable but fluctuating catches.
-
-We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history.
-This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term).
-For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status
-
-$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma)$$
-
-For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data
-
-$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
-
-where *fyear* is the year of the fishery, starting from 0.
-
-#### Fisheries Management Index Priors
-
-The Fisheries Management Index (FMI), as presented in [@melnychuk2017], utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
-These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
-The higher the score, the better the expert judges that a given metric is met in that fishery.
-Importantly, FMI surveys can be filled out in the absence of stock assessments.
-This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to [@osio2015] and [@cope2015]).
-
-The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
-
-$$log(value_i) \sim N(s(research_i) + s(management_i) +
-s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
-
-#### Swept Area Ratio Priors
-
-[@amoroso2018] provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
-This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
-As illustrated in [@amoroso2018], there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
-Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
-if all trawl-able areas are trawled twice a year then the SAR will be 2.
-Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
-
-The final selected model relating SAR to to stock status metrics was
-
-$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
-
-## Value of Information Calculations
-
-What sources of data might provide the greatest value in improving our estimates of global stock status?
-We fit 3,000 sraplus models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
-We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).
-
-The value-of-information (VOI) calculations presented in Fig.\@ref(fig:voi-plot) help illustrate what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
-We generate fits for 3000 combinations of a RAM stock and available data.
-For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
-For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
-The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
-
-Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
-We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
-
-$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
-
-Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
-
-# References
+# References {.unnumbered}
 
@@
 
-# Supplementary Information
+# Supplementary Information {.unnumbered}
 
@@
 
-## Data Sources Used
+## Data Sources Used {.unnumbered}
 
@@
 
-## Sample Prior Posterior Plots
+## Sample Prior Posterior Plots {.unnumbered}
 
@@
 
-## Population Model
+## Population Model {.unnumbered}
 
@@
 
-## Prior Generating Regressions
-
-### Swept Area Ratio
+## Prior Generating Regressions {.unnumbered}
+
+### Swept Area Ratio {.unnumbered}
 
@@
 
-### Fisheries Management Index
+### Fisheries Management Index {.unnumbered}
 
@@
 
-## Catch History
+## Catch History {.unnumbered}
 

### 2021-01-16 documents/ovando-etal-assessing-global-fisheries-short.Rmd
getting to the point

--- 2021-01-12 documents/ovando-etal-assessing-global-fisheries.Rmd
@@
 
-The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020].
-Meeting the SDG 14 targets depends in part on improving our ability to effectively track the status of global marine fish populations.
-
-The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's report on the State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries.
+The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
+Marine fisheries represent one of the largest anthropenic impacts on the oceans, and when managed sustainably are a critical source of economic prosperity and food security around the globe.
+As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries.
+While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020].
+This is a major impediment to ensuring the sustainable development of the world's oceans.
+In this paper, we assess why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.
+
+The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries.
 The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020].
-Where possible, statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database [@ricard2012].
+Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012].
 The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the biological state of "assessed" fish populations around the globe.
@@
 The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020].
-While these methods combined with local knowledge can be well informed as to the status of their fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods.
-With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods,the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
-
-In this paper we evaluate the ability of alternative data-limited assessment models to improve our estimate of the population status of unassessed fish stocks around the world.
-While these unassessed stocks are generally individually smaller than the typically larger and more valuable fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
-Due to data limitations, all of these global assessment efforts use forms of "catch-only" stock assessment models (Free et al.
+While these methods combined with local knowledge can provide good insight as to the status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods.
+With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
+
+We evaluate the ability of alternative assessment models to improve our estimate of the population status of unassessed fish stocks around the world.
+While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks lacking the data needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
+Due to data limitations, all of these global assessment efforts use forms of "catch-only" data-limited models (Free et al.
 2020 [@free2020] and references therein).
-These models seek to infer stock status, for example biomass B relative to the biomass at maximum sustainable yield B~MSY~, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
+These models seek to infer stock status, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries XX.
 
 However, Free et al.
-2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~.
+2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*.
 These issues become apparent when we consider some of the macro-level predictions made by these models.
-While the estimates B/B~MSY~ and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable methods intended for use when insufficient data are available for a full stock assessment model.
+While the estimates *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable catch-only models intended for use when insufficient data are available for a full stock assessment model.
 
 Costello et al.
-2016 [@costello2016a] finds similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison.
+2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be overly-optimistic, and the Northeast Pacific should be better by comparison.
 The superensemble method [@anderson2017a] used in Rosenberg et al.
@@
 
-readr::read_csv(here("data","status-table.csv")) %>% 
+tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
   select(-`B/BMSY RAM`) %>%  
@@
   arrange(`FAO % Overfished`) %>% 
-  select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
+  select(`FAO Area`,`FAO % Overfished`, everything())
+
+ tmp %>% 
   knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
@@
 
-In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fish populations?
-We use a flexible surplus-production based stock assessment software package, [sraplus](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
+While each of these prior efforts at assessing global fisheries made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating stock status of the worlds unassessed fisheries.
+In this paper we ask, can augmenting the types of catch-only models employed in global assessments to date with new broadly available data improve our understanding of the state of global fish populations?
+We use a flexible stock assessment software package, [sraplus](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
 We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis.
@@
 
-We compared estimates of stock status, defined by B/B~MSY~ obtained from sraplus using various forms of broadly available data to the best-available estimates of B/B~MSY~ available for those fish populations reported in the RAM database.
+We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database.
 All analysis were conducted in the R programming language [@rcoreteam2019].
@@
 
-## Data Sources
+## Data Sources {.unnumbered}
 
 We restricted ourselves to databases with wide coverage around the world, representing the types of information that can be leveraged in efforts to assess the status of global fisheries (Table.\@ref(tab:data)).
-While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for assessment of global unassessed fisheries.
+While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries.
 
@@
 
-## Experiment Structure
-
-We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global fish populations.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
-B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
+## Experiment Structure {.unnumbered}
+
+We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global unassessed fish populations.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history.
+*B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
 We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
@@
 
-We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
+We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
 We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
@@
 
-The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of [@winker2018].
-While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
+The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of Winker *et al.* 2018 [@winker2018].
+While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the data types evaluated here will support.
 The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status
@@
 
-We allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by [@walters2006]).
+We allow for process error *p* (in the manner of the stochastic stock reduction analysis suggested by Walters *et al.* 2006 [@walters2006]).
 This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
@@
 
-In addition to the results from sraplus, we include a set of results produced by the CMSY method [@froese2017].
+In addition to the results from sraplus, we include a set of results produced by the default settings of the CMSY method [@froese2017].
 For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>.
 The only modification made is to convert the underlying population model to C++ for faster computation.
-For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012].
+For each stock we used all the default options and priors provided and generated by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012].
 Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
 
-In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation.
+<!-- In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation.
 The core different between CMSY and sraplus as utilized here then is that when we refer to CMSY we mean CMSY run using the default heuristics used by the algorithm to construct priors on initial, intermediate, and terminal stock status.
@@
 In addition, all sraplus runs use a prior-predictive tuning procedure to resolve issues related to Borel's paradox, described below.
+-->
 
@@
 
+<!-- We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
+Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.
+-->
+
 Suppose that the only thing we observe from a fishery is a catch history.
-Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
-We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.
-
-In the absence of any data to fit to, the SRA algorithm works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief.
-This creates a problem for the Bayesian nature of our analysis though.
+Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
+In the absence of any data to fit to, the a stock reduction algorithm  (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure.
+This creates a problem for the Bayesian nature of our analysis.
 Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
 Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
-Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
-This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, particularly for the SRA version of sraplus, Borel's Paradox poses a particular problem.
-
-The SRA algorithm works in two steps.
-First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep).
-From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status.
-If the bulk of the prior on terminal stock status was concentrated at 50% of *K*, combinations of *r* and *K* that produce terminal stock status near 50% of *K* are sampled proportionally more frequently.
-However, lower values of terminal stock status have fewer candidate values of *r* and *K*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
-Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *K* combinations: since under this model the population cannot be greater than carrying capacity, as for example *K* approaches infinity terminal stock status asymptotes at close to 100% of *K*.
-The net result of this is that even though individual combinations of *r* and *K* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
-As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.
-
-We use an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000].
-Our solution amounts to a two-step SIR algorithm.
+Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
+
+<!-- This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
+-->
+
+Borel's Paradox poses a particular problem  for the SRA version of sraplus.
+Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020).
+This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm.
+
+We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000].
+Our solution amounts to a two-step sample-importance-resampling (SIR) algorithm.
 We first run the standard SIR algorithm as described above.
 We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
-
-$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
-
-We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
-
-$$p(n_i) = \frac{p(bin_i)}{N_i}$$
-
-And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
-The net result of this is a post-model-pre-data distribution of parameters *r* and *K* that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
-In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
-This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.S1).
-
-```{r bias-plot, fig.cap = "Post-model-pre-data distribution of depletion (biomass relative to carrying capacity) from raw SRA algorithm (untuned, top row), from SRA algorithm with approximate tuning applied (tuned, middle row), compared to the supplised prior on depletion (bottom row).
-Black vertical line indicates median value.
-"}
-
-id = unique(ram_comp_data$stockid)
-
-ex <- ram_data %>% 
-  filter(stockid %in% id[[1]])
-
-# fit using catch heuristic
-
-co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
-                                 catch = ex$catch, 
-                                 years = ex$year,
-                                 terminal_state = 0.25,
-                                 terminal_state_cv = 0.5,
-                                 b_ref_type = "k")
-
-biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
-                                      estimate_shape = FALSE, estimate_proc_error = TRUE)
-
-biased_depletion <- biased_co_fit$fit %>% 
-  filter(variable == "dep_t") %>%
-  filter(year == max(year)) %>% 
-  mutate(set = "Untuned Post-Model-Pre-Data") %>% 
-  select(year, value, set)
-
-co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
-                                      estimate_shape = FALSE, 
-                               estimate_proc_error = TRUE)
-
-tuned_depletion <- co_fit$fit %>% 
-  filter(year == max(year), variable == "dep_t") %>% 
-  mutate(set = "Tuned Post-Model-Pre-Data") %>% 
-  select(year, value, set)
-
-prior <- tuned_depletion %>% 
-  mutate(value = rlnorm(n(), log(0.25), 0.5)) %>% 
-    mutate(set = "Prior")
-
-comps <- biased_depletion %>% 
-  bind_rows(tuned_depletion) %>% 
-  bind_rows(prior) %>% 
-  group_by(set) %>% 
-  mutate(valrank = percent_rank(value)) %>% 
-  filter(between(valrank, .05, .95))
-
-comps %>% 
-  ggplot(aes(x = value, y = set)) + 
-  ggridges::geom_density_ridges(alpha = 0.9,
-                                quantile_lines = TRUE, quantiles = 2) +
-  scale_x_continuous(limits = c(0, 1.2), name = "Depletion (B/K)") + 
-  scale_y_discrete(name = '')
-
-```
+The net result of this is that it allows users to place explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior, rather than creating a complicated and biased prior on stock status based on a mixture of explicit and implicit priors.
+See XX for a detailed explanation of this problem and our solution.
 
@@
 
-Along with allowing users to supply their own priors, the sraplus package contains three built-in methods used in this manuscript for converting information on stock status from additional outside data into a form usable as a prior by sraplus.
+Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus.
 We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
 We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
-Given values of these variables for a new fishery then, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus.
+Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries.
+For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on B/B~MSY~ or F/F~MSY~ in the most recent year of the fishery usable by sraplus.
+See XX for a detailed explanation of this process.
 
@@
 The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.
-Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with the fishery status indicators of interest.
+Given the scope of this analysis, we do not claim that the regressions used here are the best possible model relating these data with the fishery status indicators of interest.
 Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.
-
-#### Catch-Only Priors {.unnumbered}
-
-Many of the current methods for estimating global stock status of unassessed stocks are based on predicting stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
-While these catch-only methods have been shown to have serious shortcomings [@free2020], we include them as a point of reference given their ubiquity in the global assessment literature.
-
-We used data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
-To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
-Emergent clusters show for example one built around a downward "one way trip" style catch histories, others with a boom and bust pattern, others with stable but fluctuating catches.
-
-We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history.
-This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term).
-For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status
-
-$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma)$$
-
-For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data
-
-$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
-
-where *fyear* is the year of the fishery, starting from 0.
-
-#### Fisheries Management Index Priors {.unnumbered}
-
-The Fisheries Management Index (FMI), as presented in [@melnychuk2017], utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
-These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
-The higher the score, the better the expert judges that a given metric is met in that fishery.
-Importantly, FMI surveys can be filled out in the absence of stock assessments.
-This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to [@osio2015] and [@cope2015]).
-
-The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
-
-$$log(value_i) \sim N(s(research_i) + s(management_i) +
-s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
-
-#### Swept Area Ratio Priors {.unnumbered}
-
-[@amoroso2018] provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
-This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
-As illustrated in [@amoroso2018], there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
-Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
-if all trawl-able areas are trawled twice a year then the SAR will be 2.
-Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
-
-The final selected model relating SAR to to stock status metrics was
-
-$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
 
@@
 We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
+The VOI analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RAM, and comparing the estimated values to the values reported in RAM.
 We generate fits for 3000 combinations of a RAM stock and available data.
@@
 
-Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
-We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
-
-$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
-
-Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
-
 # Results {.unnumbered}
@@
 We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
-From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index (FMI) scores (a measure of management capacity), and swept area ratio (SAR) values (a measure of trawl fishing intensity) (Table.\@ref(tab:data)).
+From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index scores, and swept area ratio (SAR) values (Table.\@ref(tab:data)).
 
@@
 
-For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
+As a case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
 We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al.
@@
 We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus.
-Lastly, we used the reconstructed effort data from [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
+Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
 While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both B/B~MSY~ and F/F~MSY~ (Fig.\@ref(fig:cs-plot)).
@@
 
+We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
 Overall the sraplus estimates of B/B~MSY~ resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
@@
 The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.
+Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.
 
@@
 What sources of data might provide the greatest value in improving our estimates of global stock status?
-We used sraplus together with the RAM database to estimate the average reduction in error resulting from having access to different kinds of data (Fig.\@ref(fig:voi-plot)).
+We used sraplus together with the RAM database to estimate the value of information, measured as the average reduction in root mean squared error, resulting from use of different kinds of data (Fig.\@ref(fig:voi-plot)).
+Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available.
+Interestingly though, having access to only an accurate estimate of F/F~MSY~ in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself.
 While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average.
@@
 
-Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking vital indicators such as the United Nations Sustainable Development Goals.
+Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals.
 Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fish stocks remains uncertain in many parts of the world.
-While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
+While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
 
@@
 We chose to test the performance of methods against estimated values reported in RAM.
-A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed stocks in RAM.
+A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RAM.
 The clear and systemically poor performance of the methods tested here when applied to RAM stocks would require though that these methods have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks if they are to be expected to provide reasonable picture of global fish populations.
@@
 2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
-Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
+Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
 
@@
 Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//).
-Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have previously had stock assessments.
+Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments.
 Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of [@rousseau2019] might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes
@@
 Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise.
-This is why stock assessments even in the data-rich world are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
+This is why stock assessments even in data-rich fisheries are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
 Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
@@
 Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a clear solution to this problem.
-Improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists.
+
+<!-- Our results show that improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists. -->
+
+The demonstrated lack of strong information on stock status within catch histories alone means that differences in catch-only model structure and core assumptions between different assessment efforts can produce starkly contrasting conclusions on global stock status, leading to debates that are inconclusive as they are inherently driven by assumptions.
+The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries.
+The FAO is leading efforts to increase technical capacity and monitoring and evaluation infrastructure to better inform fisheries management in places with limited data.
+Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets.
+Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally .
+Catch-only methods may be a good starting point when applied properly, but achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.
 
@@
 #   kableExtra::row_spec(0, bold = T)
-```
-
-## Sample Prior Posterior Plots {.unnumbered}
-
-```{r ppplot, fig.cap="Prior posterior plots of fits for case study fishery in Fig.6"}
-
-a = get_prior_posterior(co_fit, co_driors)
-
-a$prior_posterior %>% 
-  ggplot(aes(x = variable, y = mean, ymin = lower, ymax = upper, color = source)) + 
-  geom_pointrange(position = position_dodge(width = 0.1)) + 
-  facet_wrap(~variable, scales = "free") + 
-  theme(axis.text.x = element_blank()) + 
-  scale_color_discrete(name = '') + 
-  scale_y_continuous(name = "Value") + 
-  scale_x_discrete(name = '') + 
-  theme_minimal()
-
 ```
@@
 
+## Prior Predictive Tuning {.unnumbered}
+
+The SRA algorithm works in two steps.
+First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep).
+From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status.
+If the bulk of the prior on terminal stock status was concentrated at 50% of *K*, combinations of *r* and *K* that produce terminal stock status near 50% of *K* are sampled proportionally more frequently.
+However, lower values of terminal stock status have fewer candidate values of *r* and *K*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
+Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *K* combinations: since under this model the population cannot be greater than carrying capacity, as for example *K* approaches infinity terminal stock status asymptotes at close to 100% of *K*.
+The net result of this is that even though individual combinations of *r* and *K* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
+As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.
+
+We use an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000].
+Our solution amounts to a two-step SIR algorithm.
+We first run the standard SIR algorithm as described above.
+We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
+
+$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
+
+We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
+
+$$p(n_i) = \frac{p(bin_i)}{N_i}$$
+
+And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
+The net result of this is a post-model-pre-data distribution of parameters *r* and *K* that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
+In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
+This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.S1).
+
+```{r bias-plot, fig.cap = "Post-model-pre-data distribution of depletion (biomass relative to carrying capacity) from raw SRA algorithm (untuned, top row), from SRA algorithm with approximate tuning applied (tuned, middle row), compared to the supplised prior on depletion (bottom row).
+Black vertical line indicates median value.
+"}
+
+id = unique(ram_comp_data$stockid)
+
+ex <- ram_data %>% 
+  filter(stockid %in% id[[1]])
+
+# fit using catch heuristic
+
+co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
+                                 catch = ex$catch, 
+                                 years = ex$year,
+                                 terminal_state = 0.25,
+                                 terminal_state_cv = 0.5,
+                                 b_ref_type = "k")
+
+biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
+                                      estimate_shape = FALSE, estimate_proc_error = TRUE)
+
+biased_depletion <- biased_co_fit$fit %>% 
+  filter(variable == "dep_t") %>%
+  filter(year == max(year)) %>% 
+  mutate(set = "Untuned Post-Model-Pre-Data") %>% 
+  select(year, value, set)
+
+co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
+                                      estimate_shape = FALSE, 
+                               estimate_proc_error = TRUE)
+
+tuned_depletion <- co_fit$fit %>% 
+  filter(year == max(year), variable == "dep_t") %>% 
+  mutate(set = "Tuned Post-Model-Pre-Data") %>% 
+  select(year, value, set)
+
+prior <- tuned_depletion %>% 
+  mutate(value = rlnorm(n(), log(0.25), 0.5)) %>% 
+    mutate(set = "Prior")
+
+comps <- biased_depletion %>% 
+  bind_rows(tuned_depletion) %>% 
+  bind_rows(prior) %>% 
+  group_by(set) %>% 
+  mutate(valrank = percent_rank(value)) %>% 
+  filter(between(valrank, .05, .95))
+
+comps %>% 
+  ggplot(aes(x = value, y = set)) + 
+  ggridges::geom_density_ridges(alpha = 0.9,
+                                quantile_lines = TRUE, quantiles = 2) +
+  scale_x_continuous(limits = c(0, 1.2), name = "Depletion (B/K)") + 
+  scale_y_discrete(name = '')
+
+```
+
+## Sample Prior Posterior Plots {.unnumbered}
+
+```{r ppplot, fig.cap="Prior posterior plots of fits for case study fishery in Fig.6"}
+
+a = get_prior_posterior(co_fit, co_driors)
+
+a$prior_posterior %>% 
+  ggplot(aes(x = variable, y = mean, ymin = lower, ymax = upper, color = source)) + 
+  geom_pointrange(position = position_dodge(width = 0.1)) + 
+  facet_wrap(~variable, scales = "free") + 
+  theme(axis.text.x = element_blank()) + 
+  scale_color_discrete(name = '') + 
+  scale_y_continuous(name = "Value") + 
+  scale_x_discrete(name = '') + 
+  theme_minimal()
+
+```
+
 ## Prior Generating Regressions {.unnumbered}
+
+#### Catch-Only Priors {.unnumbered}
+
+Many of the current methods for estimating global stock status of unassessed stocks are based on predicting stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].
+While these catch-only methods have been shown to have serious shortcomings [@free2020], we include them as a point of reference given their ubiquity in the global assessment literature.
+
+We used data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics.
+To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data.
+Emergent clusters show for example one built around a downward "one way trip" style catch histories, others with a boom and bust pattern, others with stable but fluctuating catches.
+
+We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history.
+This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term).
+For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status
+
+$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma)$$
+
+For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data
+
+$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$
+
+where *fyear* is the year of the fishery, starting from 0.
+
+#### Fisheries Management Index Priors {.unnumbered}
+
+The Fisheries Management Index (FMI), as presented in [@melnychuk2017], utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
+These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
+The higher the score, the better the expert judges that a given metric is met in that fishery.
+Importantly, FMI surveys can be filled out in the absence of stock assessments.
+This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to [@osio2015] and [@cope2015]).
+
+The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
+
+$$log(value_i) \sim N(s(research_i) + s(management_i) +
+s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
+
+#### Swept Area Ratio Priors {.unnumbered}
+
+[@amoroso2018] provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
+This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
+As illustrated in [@amoroso2018], there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
+Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
+if all trawl-able areas are trawled twice a year then the SAR will be 2.
+Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
+
+The final selected model relating SAR to to stock status metrics was
+
+$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
 
@@
 ```
+
+## Value of Information Calculation
+
+We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
+The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
+We generate fits for 3000 combinations of a RAM stock and available data.
+For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
+For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
+The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
+
+Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
+We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
+
+$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
+
+Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
 

### 2021-01-16 documents/ovando-etal-assessing-global-fisheries.Rmd
submitted to proc b

@@
   Assessments of the global state of fish populations play an important role in tracking the implementation of the United Nations Sustainable Development Goals.
-While we have reliable estimates of stock status for fisheries accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain.
-Numerous high-profile publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
+While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain.
+Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
 Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added.
-Obtaining accurate estimates of stock status for the world's fisheries depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
+Estimates of fish populations based primarily on catch histories alone frequently performed roughly as well as a random guess.
+Obtaining accurate estimates of stock status for the world's exploited fish populations depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
 always_allow_html: true
@@
 
-data limited are not just messier versions of data rich
+<!-- data limited are not just messier versions of data rich  -->
 
-80% in trouble
+<!-- 80% in trouble  -->
 
@@
 
-The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020].
-Meeting the SDG 14 targets depends in part on improving our ability to effectively track the status of global marine fish populations.
+The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
+Marine fisheries represent one of the largest anthropogenic impacts on the oceans, and when managed sustainably are a critical source of economic prosperity and food security around the globe.
+As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries.
+While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020].
+This is a major impediment to ensuring the sustainable development of the world's oceans.
+In this paper, we assess why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.
 
-The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's report on the State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries.
+The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries.
 The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020].
-Where possible, statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database [@ricard2012].
-The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the biological state of "assessed" fish populations around the globe.
+Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012].
+The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the biological state of assessed fish populations around the globe.
 
@@
 The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020].
-While these methods combined with local knowledge can be well informed as to the status of their fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods.
-With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods,the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
+While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods.
+With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
 
-In this paper we evaluate the ability of alternative data-limited assessment models to improve our estimate of the population status of unassessed fish stocks around the world.
-While these unassessed stocks are generally individually smaller than the typically larger and more valuable fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
-Due to data limitations, all of these global assessment efforts use forms of "catch-only" stock assessment models (Free et al.
+We evaluate the ability of alternative assessment models to improve our estimate of the population status of unassessed fish stocks around the world.
+While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks lacking the data needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
+Due to data limitations, all of these global assessment efforts use forms of "catch-only" data-limited models (Free et al.
 2020 [@free2020] and references therein).
-These models seek to infer stock status, for example biomass B relative to the biomass at maximum sustainable yield B~MSY~, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
+These models seek to infer stock status, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries (e.g.
+@cabral2020,@gaines2018).
 
 However, Free et al.
-2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~.
+2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*.
 These issues become apparent when we consider some of the macro-level predictions made by these models.
-While the estimates B/B~MSY~ and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable methods intended for use when insufficient data are available for a full stock assessment model.
+While the estimates *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable catch-only models intended for use when insufficient data are available for a full stock assessment model.
 
 Costello et al.
-2016 [@costello2016a] finds similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison.
+2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be overly-optimistic, and the Northeast Pacific should be better by comparison.
 The superensemble method [@anderson2017a] used in Rosenberg et al.
@@
   arrange(`FAO % Overfished`) %>% 
-  select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
+    rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
+  select(`FAO Area`,`SOFIA % Overfished`, everything())
+
+ tmp %>% 
   knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
@@
 
-## Experiment Structure
+## Experiment Structure {.unnumbered}
 
-We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global fish populations.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
-B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
+We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global unassessed fish populations.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history.
+*B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
 We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
@@
 
-We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
+We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
 We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
-2017 [@froese2017] (See Table.\@ref(tab:dat-desc)).
+2017 [@froese2017] (See Table.S1).
 We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
@@
 
-We allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by [@walters2006]).
+We allow for process error *p* (in the manner of the stochastic stock reduction analysis (SRA) suggested by Walters *et al.* 2006 [@walters2006]).
 This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
@@
 The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
-Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis.
-All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
+Any model results that are not based on an index of abundance are estimated through stock reduction analysis.
+All stock reduction analysis style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
 
@@
 
+<!-- We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
+Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.
+-->
+
 Suppose that the only thing we observe from a fishery is a catch history.
-Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
-We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.
-
-In the absence of any data to fit to, the SRA algorithm works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief.
-This creates a problem for the Bayesian nature of our analysis though.
+Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
+In the absence of any data to fit to, the a stock reduction algorithm  (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure.
+This creates a problem for the Bayesian nature of our analysis.
 Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
 Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
-Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
-This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, particularly for the SRA version of sraplus, Borel's Paradox poses a particular problem.
+Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
 
-The SRA algorithm works in two steps.
-First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep).
-From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status.
-If the bulk of the prior on terminal stock status was concentrated at 50% of *K*, combinations of *r* and *K* that produce terminal stock status near 50% of *K* are sampled proportionally more frequently.
-However, lower values of terminal stock status have fewer candidate values of *r* and *K*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step.
-Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *K* combinations: since under this model the population cannot be greater than carrying capacity, as for example *K* approaches infinity terminal stock status asymptotes at close to 100% of *K*.
-The net result of this is that even though individual combinations of *r* and *K* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled.
-As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.
+<!-- This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
+-->
 
-We use an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000].
-Our solution amounts to a two-step SIR algorithm.
+Borel's Paradox poses a particular problem  for the SRA version of sraplus.
+Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020).
+This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm.
+
+We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000].
+Our solution amounts to a two-step sample-importance-resampling (SIR) algorithm.
 We first run the standard SIR algorithm as described above.
 We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
-
-$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
-
-We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
-
-$$p(n_i) = \frac{p(bin_i)}{N_i}$$
-
-And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
-The net result of this is a post-model-pre-data distribution of parameters *r* and *K* that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
-In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
-This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.S1).
-
-```{r bias-plot, fig.cap = "Post-model-pre-data distribution of depletion (biomass relative to carrying capacity) from raw SRA algorithm (untuned, top row), from SRA algorithm with approximate tuning applied (tuned, middle row), compared to the supplised prior on depletion (bottom row).
-Black vertical line indicates median value.
-"}
-
-id = unique(ram_comp_data$stockid)
-
-ex <- ram_data %>% 
-  filter(stockid %in% id[[1]])
-
-# fit using catch heuristic
-
-co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
-                                 catch = ex$catch, 
-                                 years = ex$year,
-                                 terminal_state = 0.25,
-                                 terminal_state_cv = 0.5,
-                                 b_ref_type = "k")
-
-biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
-                                      estimate_shape = FALSE, estimate_proc_error = TRUE)
-
-biased_depletion <- biased_co_fit$fit %>% 
-  filter(variable == "dep_t") %>%
-  filter(year == max(year)) %>% 
-  mutate(set = "Untuned Post-Model-Pre-Data") %>% 
-  select(year, value, set)
-
-co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
-                                      estimate_shape = FALSE, 
-                               estimate_proc_error = TRUE)
-
-tuned_depletion <- co_fit$fit %>% 
-  filter(year == max(year), variable == "dep_t") %>% 
-  mutate(set = "Tuned Post-Model-Pre-Data") %>% 
-  select(year, value, set)
-
-prior <- tuned_depletion %>% 
-  mutate(value = rlnorm(n(), log(0.25), 0.5)) %>% 
-    mutate(set = "Prior")
-
-comps <- biased_depletion %>% 
-  bind_rows(tuned_depletion) %>% 
-  bind_rows(prior) %>% 
-  group_by(set) %>% 
-  mutate(valrank = percent_rank(value)) %>% 
-  filter(between(valrank, .05, .95))
-
-comps %>% 
-  ggplot(aes(x = value, y = set)) + 
-  ggridges::geom_density_ridges(alpha = 0.9,
-                                quantile_lines = TRUE, quantiles = 2) +
-  scale_x_continuous(limits = c(0, 1.2), name = "Depletion (B/K)") + 
-  scale_y_discrete(name = '')
-
-```
+The net result of this is that it allows users to place explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior, rather than creating a complicated and biased prior on stock status based on a mixture of explicit and implicit priors.
+See Supplementary Information for a detailed explanation of this problem and our solution.
 
@@
 
-Along with allowing users to supply their own priors, the sraplus package contains three built-in methods used in this manuscript for converting information on stock status from additional outside data into a form usable as a prior by sraplus.
+Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus.
 We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
 We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
-Given values of these variables for a new fishery then, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus.
+Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries.
+For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on B/B~MSY~ or F/F~MSY~ in the most recent year of the fishery usable by sraplus (See Supplementary Information).
 
@@
 The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
-
-Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
-We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
-
-$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
-
-Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
 
@@
 What sources of data might provide the greatest value in improving our estimates of global stock status?
-We used sraplus together with the RAM database to estimate the average reduction in error resulting from having access to different kinds of data (Fig.\@ref(fig:voi-plot)).
-While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average.
+We used sraplus together with the RAM database to estimate the value of information, measured as the average reduction in root mean squared error, resulting from use of different kinds of data (Fig.\@ref(fig:voi-plot)).
+Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available.
+Interestingly though, having access to only an accurate estimate of F/F~MSY~ in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself.
+While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average, due to the lack of historical context for recent trends in abundance.
 We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].
@@
 
-Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking vital indicators such as the United Nations Sustainable Development Goals.
+Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals.
 Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fish stocks remains uncertain in many parts of the world.
-While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
+The hope of efforts using catch-only models to estimate the status of global unassessed fisheries is fundamentally that characteristics of a fishery's catch history and life history are informative as to the status its population.
+Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RAM.
+While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
 
@@
 However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed.
-Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a clear solution to this problem.
-Improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists.
+Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a solution to this problem to has proven to be unbiased and sufficiently precise at a global or regional level.
+
+<!-- Our results show that improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists. -->
+
+The lack of strong information on stock status within catch histories demonstrated here means that differences in models and assumptions between catch-based assessment efforts can produce starkly contrasting conclusions on global stock status, leading to debates that are inconclusive as they are inherently driven by assumptions.
+The FAO is leading efforts to increase technical capacity and monitoring and evaluation infrastructure to improve fisheries management in places with limited data.
+Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets.
+Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally [@costello2020].
+Catch-only methods may be a useful starting point when applied properly, but achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.
+
+<!-- The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries. -->
+
+# Acknowledgements{.unnumbered}
+
+Funding for this work was provided by the Food and Agriculture Organization of the United Nations (FAO).
+We thank participants of the FAO "Methods for Global Assessment" workshop held in Rome, Italy February 2019 for helpful feedback on this project.
+
+# Disclaimer {.unnumbered}
+
+The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of NOAA or the Department of Commerce.
 
@@
 :::
-
-# Supplementary Information {.unnumbered}
-
-\renewcommand{\thefigure}{S\arabic{figure}}
-
-\setcounter{figure}{0}
-
-\renewcommand{\thetable}{S\arabic{table}}
-
-\setcounter{table}{0}
-
-## Data Sources Used {.unnumbered}
-
-```{r dat-desc}
-dat_description <- tribble(~`Data Name`, ~Description,
-        "RAM Index", "Fit to abundance index from RAM",
-        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
-        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
-        "EFfective CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 2.6% technology creep",
-        "Effective CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
-        "Nominal CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
-         "Nominal CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
-        "Guess", "Priors on terminal B/Bmsy randomly sampled from .4,1,1.6")
-
-knitr::kable(dat_description,
-             caption = "Data sources used for terminal stock status estimate",
-             format = "pipe")
-# %>%
-#   kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-#   kableExtra::row_spec(0, bold = T)
-```
-
-## Sample Prior Posterior Plots {.unnumbered}
-
-```{r ppplot, fig.cap="Prior posterior plots of fits for case study fishery in Fig.6"}
-
-a = get_prior_posterior(co_fit, co_driors)
-
-a$prior_posterior %>% 
-  ggplot(aes(x = variable, y = mean, ymin = lower, ymax = upper, color = source)) + 
-  geom_pointrange(position = position_dodge(width = 0.1)) + 
-  facet_wrap(~variable, scales = "free") + 
-  theme(axis.text.x = element_blank()) + 
-  scale_color_discrete(name = '') + 
-  scale_y_continuous(name = "Value") + 
-  scale_x_discrete(name = '') + 
-  theme_minimal()
-
-```
-
-## Population Model {.unnumbered}
-
-The core of our model is a Pella-Tomlinson [@pella1969] production model in the manner of [@winker2018].
-While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
-
-```{r si-prior-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
-  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
-  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
-  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
-  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
-  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))
-
-knitr::kable(
-  cd,
-  "latex",
-  align = "l",
-  booktabs = TRUE,
-  escape = F,
-  row.names = F,
-  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
-) %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
-
-The population growth equation is
-
-\begin{equation}
-  f(x)=\begin{cases}
-      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
-, & \text{if $B_t>0.25 \times K$}.\\
-     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
-  \end{cases}
-  (\#eq:sihockey)
-\end{equation}
-
-## Prior Generating Regressions {.unnumbered}
-
-### Swept Area Ratio {.unnumbered}
-
-```{r sar-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of swept area ratio (SAR) on F/F~MSY~"}
-fit <- sraplus::sar_models$fit[sraplus::sar_models$metric == "mean_uumsy"][[1]]
-
-fit_data <- fit$data
-
-pp <- posterior_predict(fit, draws = 1000)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
-  scale_x_continuous(name = "Observed Log F/Fmsy") + 
-  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-
-```
-
-### Fisheries Management Index {.unnumbered}
-
-```{r fmi-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of fisheries management index (FMI) on F/F~MSY~"}
-
-fit <- sraplus::fmi_models$fit[sraplus::fmi_models$metric == "mean_uumsy"][[1]]
-
-fit_data <- fit$data
-
-pp <- posterior_predict(fit, draws = 1000)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
-  scale_x_continuous(name = "Observed Log F/Fmsy") + 
-  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-```
-
-## Catch History {.unnumbered}
-
-```{r catch-reg,fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) B/B~MSY~ for regression of catch on B/B~MSY~"}
-
-fit <- sraplus::catch_b_model
-
-fit_data <- fit$data %>% 
-  group_by(stockid) %>% 
-  filter(year == max(year))
-
-pp <- posterior_predict(fit, draws = 1000, newdata = fit_data)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_scatter_avg(y = fit_data$log_value, yrep = pp) + 
-  coord_flip() +
-  scale_y_continuous(name = "Observed Log B/Bmsy") + 
-  scale_x_continuous(name = "Posterior Predictive B/Bmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-```

### 2021-01-16 documents/ovando-etal-assessing-global-fisheries-si.Rmd
Proc b (#9)
* trimming methods back down

* getting to the point

* submitted to proc b

--- 2020-12-09 documents/ovando-etal-assessing-global-fisheries.Rmd
@@
 ---
-title: |
-  | Status of Global Unassessed Fisheries will Remain Highly Uncertain
-  | without Better Data
+title: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
+subtitle: Supplementary Information
 author: 
@@
 bibliography: ["../references.bib"]
-csl: nature-communications.csl
+csl: proceedings-of-the-royal-society-b.csl
 output: 
@@
   bookdown::word_document2:
-    reference_docx: SA_MS_Template_2020.docx
+    reference_docx: template.docx
 params:
@@
   min_years_catch: [25]
-abstract: |
-  Assessments of the global state of fisheries play an important role in tracking the implementation of the United Nations Sustainable Development Goals.
-While we have reliable estimates of stock status for fisheries accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fisheries remains highly uncertain.
-Numerous high-profile publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
-Here, we show that despite numerous efforts, our understanding of the status of global fisheries remains incomplete, even when new sources of broadly available data are added.
-Obtaining accurate estimates of stock status for the world's fisheries depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
 always_allow_html: true
 linkcolor: blue
-# header-includes:
-#   - \usepackage{draftwatermark}
-#   - \SetWatermarkText{DRAFT}
 ---
@@
 
-<!-- # rishi notes -->
+# Supplementary Information {.unnumbered}
 
-<!-- * more assertive on what the data needs really are -->
+\renewcommand{\thefigure}{S\arabic{figure}}
 
-<!--   * where to apply globally  -->
+\setcounter{figure}{0}
 
-<!-- * For places with less coverage, what would we actually recommend there  -->
+\renewcommand{\thetable}{S\arabic{table}}
 
-<!-- * what do we know about catchabiliy -->
+\setcounter{table}{0}
 
-<!-- - e.g. 51, fleets became industrial  -->
+## Data Sources Used {.unnumbered}
 
-<!--   - hammer down on that's the issue with the effort  -->
+```{r dat-desc}
+dat_description <- tribble(~`Data Name`, ~Description,
+        "RAM Index", "Fit to abundance index from RAM",
+        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
+        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
+        "EFfective CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 2.6% technology creep",
+        "Effective CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
+        "Nominal CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
+         "Nominal CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
+        "Guess", "Priors on terminal B/Bmsy randomly sampled from .4,1,1.6")
 
-<!--   - how do we translate effort into f bey area -->
+knitr::kable(dat_description,
+             caption = "Data sources used for terminal stock status estimate",
+             format = "pipe")
+# %>%
+#   kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+#   kableExtra::row_spec(0, bold = T)
+```
 
-<!--   - here's  -->
+## Population Model {.unnumbered}
 
-<!-- https://www.nature.com/natsustain/about/content -->
+The core of our model is a Pella-Tomlinson [@pella1969] production model in the manner of [@winker2018].
+While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
 
-<!-- # check in call notes -->
+```{r si-prior-tab}
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
+  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
+  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
+  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
+  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
+  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))
 
-<!-- Ray: setting the stage: what we want is a way to assess the status of stocks and see how we're doing relative to SDG. -->
-
-<!-- We need to discuss a sampling strategy: how can we get down to a set of priority stock. A step of expert knowledge to refine what stocks you would focus in on. -->
-
-<!-- Regressions based approach, focusing in effort on stocks -->
-
-<!-- Basically a paragraph laying out that we need to do this, and discussing how we might go about doing that. Equal weight to each stock vs proportional to biomass -->
-
-<!-- add in a bit more to the discussion on guidelines -->
-
-<!-- add in error in the index for the VOI analysis -->
-
-<!-- Apples and apples comparisons between model and design developments.
-How much is the value of the effort data relative to actually improving the population model.
-Production model is ill-suited to some of these questions.
--->
-
-<!-- fishing effort uncertainty of the effort -->
-
-<!-- how credible is the information -->
-
-<!-- probably don't need to bother with the SAUP -->
-
-<!-- stock assessments -->
-
-<!-- the stock will collapse next year -->
-
-<!-- nansen data -->
-
-<!-- # Publication Requirements -->
-
-<!-- Target Journal: [Nature Sustainability](https://www.nature.com/natsustain/about/content) -->
-
-<!-- 3500 words main text -->
-
-<!-- 50 references (main text) -->
-
-<!-- 3000 words (max) methods -->
-
-# Introduction
-
-The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
-Meeting these targets depends in part on our ability to effectively track the status of global marine fish stocks.
-The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the dedicated SDG 14 Indicator on fisheries sustainability, and the FAO's report on the State of World Fisheries and Aquaculture (SOFIA) report is the most widely used source for tracking the global state of fisheries.
-As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020].
-The SOFIA assessment was designed in the 1970s based on the then available data and methods, With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods,the SOFIA global assessment methods need to be further strengthened to meet the demand for estimation and tracking progress towards the SDG goals.
-
-The FAO's SOFIA index captures 70% of the landings of all fisheries in the world, with statements on on individual stock status made on the basis of formal stock assessments where available, and through a combination of other methods where not [@fao2020].
-In this paper we evaluate the ability of alternative data-limited assessment models to improve the methods used by the FAO for these unassessed stocks.
-While these unassessed stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-
-The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed.
-While these methods combined with local knowledge can be well informed as to the status of their fish stocks, a more quantitatively vetted process would be desirable in order to increase reproducibility, transparency, and potential reliability.
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
-Due to data limitations, all of these global assessment efforts use forms of "catch-only" stock assessment models (Free et al.
-2020 [@free2020] and references therein).
-These models seek to infer stock status, for example in terms of biomass *B* relative to the biomass at maximum sustainable yield *B~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].
-
-However, Free et al.
-2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~.
-These issues become apparent when we consider some of the macro-level predictions made by these models.
-While the estimates B/B~MSY~ and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable methods intended for use when insufficient data are available for a full stock assessment model.
-
-Costello et al.
-2016 [@costello2016a] finds similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison.
-The superensemble method [@anderson2017a] used in Rosenberg et al.
-2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific.
-The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
-The methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
-However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.
-
-```{r status-tab}
-
-o_by_a <- sofia_status %>% 
-  left_join(fao_area_codes, by = "fao_area_code") %>% 
-  group_by(fao_area) %>% 
-  summarise(`FAO % Overfished` = percent(mean(status == "O")))
-
-ss <-
-  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
-                    sheet = "state_space_results_08_28_2018") %>%
-  janitor::clean_names()
-
-ss_lookup <-
-  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
-                    sheet = "ss_to_fao") %>%
-  janitor::clean_names() %>% 
-  mutate(tmp = map2(region, fao_area_codes, ~data.frame(region = .x, fao_area_code = as.integer(str_split(.y, pattern = ',',simplify = TRUE))))) %>% 
-  select(tmp) %>% 
-  unnest(cols = tmp) %>% 
-    left_join(fao_area_codes, by = "fao_area_code")
-
-state_space_results <- ss_lookup %>% 
-  left_join(ss, by = "region") %>% 
-  filter(variable == "BvB") %>% 
-  group_by(fao_area) %>% 
-  filter(year == max(year)) %>% 
-  summarise(`B/BMSY RAM` = round(mean(dlm_geomean),2)) %>% 
-  ungroup()
-
-# ram_by_region %>% 
-#   ggplot(aes(b_v_bmsy)) +
-#   geom_histogram() + 
-#   facet_wrap(~fao_area)
-
-readr::read_csv(here("data","status-table.csv")) %>% 
-  select(-`B/BMSY RAM`) %>%  
-  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
-  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
-  arrange(`FAO % Overfished`) %>% 
-  select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
-  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
-               Rosenberg et al. 2018, and Pauly 2007",
+knitr::kable(
+  cd,
+  "latex",
+  align = "l",
   booktabs = TRUE,
-               format = "pipe") %>% 
+  escape = F,
+  row.names = F,
+  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
+) %>%
   kableExtra::kable_styling(full_width = T, font_size = 9) %>%
   kableExtra::row_spec(0, bold = T)
-
 ```
-
-In this paper we ask, can combining the FAO's catch statistics with other broadly available data improve our understanding of the state of global fisheries?
-We use a flexible surplus-production based stock assessment software package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
-We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis.
-
-# Results
-
-| Data Source           |  Short Name | Data Use |  Caveats |
-|-----------------------|-------------|------|------|
-| Catch data [@fao2020] | catches     | Priors on stock status, scaling of population size, exploitation history    |    Heuristics or regressions used to translate shape of catch history into priors on stock status  |
-|    Fisheries Management Index [@melnychuk2017]     | FMI     | Priors on most recent F/F~MSY~ values   |Priors produced by regression trained on data from RAM Legacy Stock Assessment Database|
-| Swept Area Ratio [@amoroso2018]      | SAR     | Priors on most recent F/F~MSY~ values |Priors produced by regression trained on data from RAM Legacy Stock Assessment Database|
-|  Reconstructed effort data [@rousseau2019b] | effort | Combined with catch data to create an index of abundance| Total reconstructed effort across all sectors.
-Assumed rate of technology creep reported in individual sections|
-Table:(\#tab:data) Data sources included across model fits.
-
-We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
-From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index (FMI) scores (a measure of management capacity), and swept area ratio (SAR) values (a measure of trawl fishing intensity) (Table.\@ref(tab:data)).
-
-<!-- `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model.
-We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~.
-Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.
--->
-
-For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
-We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al.
-2019 [@rousseau2019] to each stock.
-As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available.
-We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus.
-Lastly, we used the reconstructed effort data from [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
-While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both B/B~MSY~ and F/F~MSY~ (Fig.\@ref(fig:cs-plot)).
-
-<!-- Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
-CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
-Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
-Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
-This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values (Fig.\@ref(fig:cs-plot)).
--->
-
-```{r cs-plot, fig.cap="RAM values of B/B~MSY~ and F/F~MSY~ (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
-Each point is a stock.
-Black dashed line shows the 1:1 relationship."}
-ex_scatter_plot
-```
-
-<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
-In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
-Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
--->
-
-We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
-B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
-We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
-This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
-
-We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
-We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
-2017 [@froese2017] (See Table.\@ref(tab:dat-desc)).
-We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
-Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
-
-Overall the sraplus estimates of B/B~MSY~ resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
-
-Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves.
-This becomes clearer once we consider the performance of sraplus models fit to combinations of our broadly available datasets.
-Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)).
-However, this is somewhat an artifact of the data.
-The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near one.
-This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise.
-Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
-The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.
-
-Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process.
-Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions.
-For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
-We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
-Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
-
-```{r perf-tab}
-
-knitr::kable(perf, digits = 2,
-             caption = "Global performance statistics in the most recent year available of models using different sources of data.
-mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
-Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.
-See Table.5 for details of data types.",
-              format = "pipe")
-
-```
-
-<!-- xx dig into this, which stocks are in that region? XX -->
-
-```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent B/B~MSY~ by FAO statistical area from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
-ram_mpe_map_plot
-```
-
-```{r mape-map, fig.cap = "Median absolute percent error in most recent B/B~MSY~ by FAO statistical area from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
-ram_mape_map_plot
-```
-
-```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
-
-ram_acc_map_plot 
-```
-
-\newpage
-
-The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
-What sources of data might provide the greatest value in improving our estimates of global stock status?
-We used sraplus together with the RAM database to estimate the average reduction in error resulting from having access to different kinds of data (Fig.\@ref(fig:voi-plot)).
-While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average.
-We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].
-
-```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of B/B~MSY~ in the most recent 5 years of data available for each model fit.
-Distribution is full posterior probability distribution.
-Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
-Change is relative to the mean performance of a catch-only heuristic model."}
-b_voi_plot + 
-  geom_vline(aes(xintercept = 0), alpha = 0.75)
-```
-
-# Discussion
-
-<!-- XX somewhere in here make clear that you're not learning anything from the catch data XX -->
-
-<!-- XX Acknolwedge the need for action the the frustration of being told "get more data" XX -->
-
-<!-- Global-level assessment are critical for guiding management agendas for the world's oceans.
-Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world.
-Numerous efforts have sought to resolve this problem through development of new modeling methods.
-These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment.
-However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes.
-While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.
--->
-
-Global-level assessment are critical for guiding management agendas for the world's oceans, and tracking critical indicators such as the United Nations Sustainable Development Goals.
-Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains uncertain in many parts of the world.
-While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status (Table.\@ref(tab:perf-tab)).
-
-What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends on the needs of those using the assessment outputs.
-It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
-In some instances using the data presented here did provide some improvement over use of catch-only style methods; the difficulty comes in attempting to apply data types uniformly across the globe.
-While it is unreasonable to expect any global-scale data to be able to perform as well as detailed stock assessments reported in RAM, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
-That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the quality and use of data themselves.
-
-We chose to test the performance of methods against estimated values reported in RAM.
-A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed stocks in RAM.
-The clear and systemically poor performance of the methods tested here when applied to RAM stocks would require though that these methods have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks if they are to be expected to provide reasonable picture of global stock status.
-Free et al.
-2020's [@free2020] simulation testing of the types of methods included here suggests that this is unlikely to be true.
-
-<!-- Our estimates of global stock status vary wildly depending on which source of data we use.
-Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information?
-Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)).
-Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used.
-These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability.
-The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations.
-Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks.
-None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RAM (XX RAM matches?
-need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).
--->
-
-Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions.
-The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e.
-those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)).
-Effort data such as those reconstructed by Rousseau et al.
-2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
-Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
-
-But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
-Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
-
-Our value-of-information analysis also shows though the high utility of having access to even a recent snapshot of F/F~MSY~ (Fig.\@ref(fig:voi-plot)).
-Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
-Another avenue would be to prioritize the development of a global repository for length and age composition data.
-Given appropriate conditions, these length measurements can be used to estimate local fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
-While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
-Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f; @rudd2017].
-
-<!-- Using catch data alone does not reveal about recent stock status beyond what is encoded in our priors and functional forms.
-So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest).
-Given the demonstrated value of F/F~MSY~ values in improving estimates of stock status, and the increased feasibility in estimating F/F~MSY~ at a local scale relative to B/B~MSY~, how might we go about obtaining estimates of fishing mortality rates at a global scale?
-We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/F~MSY~ for individual fisheries (see Supplemental Information).
-While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
--->
-
-We must also prioritize collection and curation of fish population survey data worldwide.
-Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//).
-Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have previously had stock assessments.
-Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of [@rousseau2019] might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes
-
-Expanded training of fisheries scientists around the globe is another critical need.
-Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise.
-This is why stock assessments even in the data-rich world are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
-Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
-
-The coming decades are a critical time for the future of fisheries and ocean health.
-Achieving the United Nations Sustainable Development Goal 14 for the conservation and sustainable use of the world's oceans depends on our ability to effectively assess the status of fish stocks around the world.
-The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
-However, this process still leaves a substantial number of fisheries and global catch unassessed.
-Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a clear solution to this problem.
-Improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists.
-
-# Methods
-
-All analysis were conducted in the R programming language [@rcoreteam2019].
-Model fitting was conducted using Rcpp [@eddelbuettel2011] and stan [@carpenter2017], implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018].
-The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
-
-sraplus contains too many options to cover within this methods section.
-We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus.
-Below we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.
-
-## Population Model
-
-The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of [@winker2018].
-While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
-The purpose of sraplus is not to make substantial improvements in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status
 
@@
   f(x)=\begin{cases}
-B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)p_t
+      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
 , & \text{if $B_t>0.25 \times K$}.\\
-B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)\right)p_t, & \text{otherwise}.
+     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
   \end{cases}
-(\#eq:pt)
+  (\#eq:sihockey)
 \end{equation}
 
-Where $B_t$ is biomass at time *t*, *r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of B~MSY~/K to shift.
-When *m* is two B~MSY~ / K = 0.5, lower values of *m* shift the production function left, higher values right.
-$\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors.
-Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
-We deal with this problem by following the methods described in [@winker2018].
-to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
-
-We allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by [@walters2006]).
-This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
-Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).
-
-Process error *p* is assumed to be log-normally distributed, such that
-
-\begin{equation}
-  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
-(\#eq:procerror)
-\end{equation}
-
-## Estimation Model
-
-All of our estimates are Bayesian in nature.
-We can break the use of sraplus into two distinct categories: with data and without.
-By "data", we refer to measurements which are used to confront model estimates within a likelihood function.
-In our context, these include fishery-independent survey data, or a CPUE index.
-When there are no data, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
-Under this mode, the model is essentially a stock-reduction analysis model, in the manner of [@walters2006], in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e.
-results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches.
-This step updates the prior distribution of population parameters by eliminating combinations of priors that are impossible for a given catch history and a specified functional form.
-
-The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).
-*r* and *K* are the only two parameters that are always estimated.
-Estimation of every other parameter can be turned on or off.
-When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user.
-For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*.
-*q* is also estimated when needed.
-
-```{r param-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "$K$", `Default Prior` = "Prior predictive tuning") %>%
-  rbind(c("Growth rate", "$r$", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
-  rbind(c("Shape parameter", "$m$", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
-  rbind(c("Catchability", "$q$", "$logn(1e^{-3}, 1)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,1)$")) %>%
-  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.5,0.25)$")) %>%
-  rbind(c("Initial State", "$B0$", "Posterior probability dist. of catch-based regressions"))
-
-knitr::kable(
-  cd,
-  align = "l",
-  escape = F,
-  row.names = F,
-  caption = "Name, abbreviations, and priors distribution for parameters potentially estimated by sraplus in this manuscript.",
-  format = "pipe"
-) %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
-
-sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent).
-Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
-A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
-sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~.
-We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history.
-Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected.
-The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
-Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis.
-All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
-
-When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018].
-By default the model uses 2000 draws with a 1000 step warm-up and one chain.
-Any detailed fit for a particular fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria.
-The model fits to a direct estimate of abundance (e.g.
-a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood is
-
-$$log(a_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}) \times q, sigma_{obs})$$
-
-where *f* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
-When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data.
-[@rousseau2019] measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep).
-This rate of technology creep assumes that every unit increase in effort is log-linearly greater than the unit of effort before it.
-When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
-This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
-To accommodate this, we generate a catch per effective harvest rate index of abundance, as
-
-$$cpue_t = \frac{c_t}{(1 - e^{-f_t})}$$
-
-$$f_t = q_tE_t$$
-
-Where $q_t$ can has a technology creep component $\tau$
-
-$$q_t = q_{t-1} \times (1  + \tau)$$
-
-We then fit to the index of abundance per
-
-$$log(cpue_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}), \sigma_{obs})$$
-
-### CMSY
-
-In addition to the results from sraplus, we include a set of results produced by the CMSY method [@froese2017].
-For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>.
-The only modification made is to convert the underlying population model to C++ for faster computation.
-For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012].
-Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
-
-In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation.
-The core different between CMSY and sraplus as utilized here then is that when we refer to CMSY we mean CMSY run using the default heuristics used by the algorithm to construct priors on initial, intermediate, and terminal stock status.
-In contrast, sraplus fit using for example FMI data uses the empirical models described here to generate priors on stock status.
-In addition, all sraplus runs use a prior-predictive tuning procedure to resolve issues related to Borel's paradox, described below.
-
-## Priors
-
-Priors for all estimable parameters can be left at default values or set by the user.
-The shape parameter is usually not reliably estimable given available data for surplus production models, however, Thorson et al.
-2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa.
-While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on Thorson et al.
-2012 [@thorson2012] for the genus of the species in question.
-
-We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus.
-
-### Prior Predictive Tuning
-
-Suppose that the only thing we observe from a fishery is a catch history.
-Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
-We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than zero.
-
-In the absence of any data to fit to, the SRA algorithm works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief.
-This creates a problem for the Bayesian nature of our analysis though.
-Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
-Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
-Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
-This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, particularly for the SRA version of sraplus, Borel's Paradox poses a particular problem.
+## Prior Predictive Tuning {.unnumbered}
 
@@
 
-### Priors Informed by Outside Data
+## Sample Prior Posterior Plots {.unnumbered}
 
-Along with allowing users to supply their own priors, the sraplus package contains three built-in methods used in this manuscript for converting information on stock status from additional outside data into a form usable as a prior by sraplus.
-We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
-We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
-Given values of these variables for a new fishery then, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus.
+```{r ppplot, fig.cap="Prior posterior plots of fits for case study fishery in Fig.6"}
 
-All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
-The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.
-Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with the fishery status indicators of interest.
-Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.
+a = get_prior_posterior(co_fit, co_driors)
 
-#### Catch-Only Priors
+a$prior_posterior %>% 
+  ggplot(aes(x = variable, y = mean, ymin = lower, ymax = upper, color = source)) + 
+  geom_pointrange(position = position_dodge(width = 0.1)) + 
+  facet_wrap(~variable, scales = "free") + 
+  theme(axis.text.x = element_blank()) + 
+  scale_color_discrete(name = '') + 
+  scale_y_continuous(name = "Value") + 
+  scale_x_discrete(name = '') + 
+  theme_minimal()
+
+```
+
+## Prior Generating Regressions {.unnumbered}
+
+#### Catch-Only Priors {.unnumbered}
 
@@
 
-#### Fisheries Management Index Priors
+#### Fisheries Management Index Priors {.unnumbered}
 
@@
 
-#### Swept Area Ratio Priors
+#### Swept Area Ratio Priors {.unnumbered}
 
@@
 
-## Value of Information Calculations
-
-What sources of data might provide the greatest value in improving our estimates of global stock status?
-We fit 3,000 sraplus models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
-We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).
-
-The value-of-information (VOI) calculations presented in Fig.\@ref(fig:voi-plot) help illustrate what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
-We generate fits for 3000 combinations of a RAM stock and available data.
-For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
-For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
-The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
-
-Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
-We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
-
-$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
-
-Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
-
-# References
-
-::: {#refs}
-:::
-
-# Supplementary Information
-
-\renewcommand{\thefigure}{S\arabic{figure}}
-
-\setcounter{figure}{0}
-
-\renewcommand{\thetable}{S\arabic{table}}
-
-\setcounter{table}{0}
-
-## Data Sources Used
-
-```{r dat-desc}
-dat_description <- tribble(~`Data Name`, ~Description,
-        "RAM Index", "Fit to abundance index from RAM",
-        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
-        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
-        "EFfective CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 2.6% technology creep",
-        "Effective CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
-        "Nominal CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
-         "Nominal CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
-        "Guess", "Priors on terminal B/Bmsy randomly sampled from .4,1,1.6")
-
-knitr::kable(dat_description,
-             caption = "Data sources used for terminal stock status estimate",
-             format = "pipe")
-# %>%
-#   kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-#   kableExtra::row_spec(0, bold = T)
-```
-
-## Sample Prior Posterior Plots
-
-```{r ppplot, fig.cap="Prior posterior plots of fits for case study fishery in Fig.6"}
-
-a = get_prior_posterior(co_fit, co_driors)
-
-a$prior_posterior %>% 
-  ggplot(aes(x = variable, y = mean, ymin = lower, ymax = upper, color = source)) + 
-  geom_pointrange(position = position_dodge(width = 0.1)) + 
-  facet_wrap(~variable, scales = "free") + 
-  theme(axis.text.x = element_blank()) + 
-  scale_color_discrete(name = '') + 
-  scale_y_continuous(name = "Value") + 
-  scale_x_discrete(name = '') + 
-  theme_minimal()
-
-```
-
-## Population Model
-
-The core of our model is a Pella-Tomlinson [@pella1969] production model in the manner of [@winker2018].
-While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support.
-
-```{r si-prior-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
-  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
-  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
-  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
-  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
-  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))
-
-knitr::kable(
-  cd,
-  "latex",
-  align = "l",
-  booktabs = TRUE,
-  escape = F,
-  row.names = F,
-  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
-) %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
-
-The population growth equation is
-
-\begin{equation}
-  f(x)=\begin{cases}
-      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
-, & \text{if $B_t>0.25 \times K$}.\\
-     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
-  \end{cases}
-  (\#eq:sihockey)
-\end{equation}
-
-## Prior Generating Regressions
-
-### Swept Area Ratio
+### Swept Area Ratio {.unnumbered}
 
@@
 
-## Catch History
+## Value of Information Calculation {.unnumbered}
+
+We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
+The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
+We generate fits for 3000 combinations of a RAM stock and available data.
+For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
+For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
+The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
+
+Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
+We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
+
+$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
+
+Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
+
+## Catch History {.unnumbered}
 

### 2021-01-16 make-assessing-global-fisheries.R

@@ -2188,8 +2188,15 @@ save(file = here("results", results_name, "paper_plots.Rdata"),
 if (knit_paper == TRUE){
   
   rmarkdown::render(
-    here::here("documents", "ovando-etal-assessing-global-fisheries.Rmd"),
+    here::here("documents", "ovando-etal-assessing-global-fisheries-short.Rmd"),
     params = list(results_name = results_name,
                   min_years_catch = 25)
   )
+  
+  rmarkdown::render(
+    here::here("documents", "ovando-etal-assessing-global-fisheries-si.Rmd"),
+    params = list(results_name = results_name,
+                  min_years_catch = 25)
+  )
+  
 }

### 2021-01-16 make-assessing-global-fisheries.R
updating readme

@@ -58,7 +58,7 @@ results_name <- "v1.0"
 results_description <-
   "publication version of results"
 
-run_voi_models <- TRUE
+run_voi_models <- FALSE
 # sub options for run_voi_models
 fit_models <- FALSE
 
@@ -74,7 +74,7 @@ run_ram_tests <- FALSE
 
 run_ram_comparison <- FALSE
 
-knit_paper <- FALSE
+knit_paper <- TRUE
 
 warning("Running full analysis takes upwards of 24 hours on 2 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 2 cores is not recommended (memory routinley runs out on a 16 core 36GB machine when cores > 2)")
 
@@ -2188,7 +2188,7 @@ save(file = here("results", results_name, "paper_plots.Rdata"),
 if (knit_paper == TRUE){
   
   rmarkdown::render(
-    here::here("documents", "ovando-etal-assessing-global-fisheries-short.Rmd"),
+    here::here("documents", "ovando-etal-assessing-global-fisheries.Rmd"),
     params = list(results_name = results_name,
                   min_years_catch = 25)
   )

### 2021-01-19 documents/ovando-etal-assessing-global-fisheries.Rmd
seems like that did it

@@
 
-Costello et al.
+<!-- Costello et al.
 2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be overly-optimistic, and the Northeast Pacific should be better by comparison.
@@
 However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.
+-->
 
@@
 When *m* is two *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right.
+The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, Thorson et al.
+2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa.
+We fix the shape parameter based on the distributions reported in Thorson et al.
+2012 [@thorson2012] for the genus of the species in question.
+
  $\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors.
@@
 We allow for process error *p* (in the manner of the stochastic stock reduction analysis (SRA) suggested by Walters *et al.* 2006 [@walters2006]).
-This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
+Process error *p* is assumed to be log-normally distributed, such that
+
+<!-- This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
 Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).
-
-Process error *p* is assumed to be log-normally distributed, such that
+-->
 
@@
 All of our estimates are Bayesian in nature.
-sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent).
+sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo [].
 Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
@@
 All stock reduction analysis style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
+For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*.
+*q* is also estimated when needed.
 
@@
 
-The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).
-*r* and *K* are the only two parameters that are always estimated.
+<!-- The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).  -->
+
+<!-- *r* and *K* are the only two parameters that are always estimated.
 Estimation of every other parameter can be turned on or off.
 When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user.
-For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*.
-*q* is also estimated when needed.
-
-```{r param-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "$K$", `Default Prior` = "Prior predictive tuning") %>%
-  rbind(c("Growth rate", "$r$", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
-  rbind(c("Shape parameter", "$m$", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
-  rbind(c("Catchability", "$q$", "$logn(1e^{-3}, 1)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,1)$")) %>%
-  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.5,0.25)$")) %>%
-  rbind(c("Initial State", "$B0$", "Posterior probability dist. of catch-based regressions"))
-
-knitr::kable(
-  cd,
-  align = "l",
-  escape = F,
-  row.names = F,
-  caption = "Name, abbreviations, and priors distribution for parameters potentially estimated by sraplus in this manuscript.",
-  format = "pipe"
-) %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
+-->
 
@@
 
-The shape parameter is usually not reliably estimable given available data for surplus production models, however, Thorson et al.
-2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa.
-While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on Thorson et al.
-2012 [@thorson2012] for the genus of the species in question.
-
-We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus.
+<!-- We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus. -->
 
@@
 
-Suppose that the only thing we observe from a fishery is a catch history.
+<!-- Suppose that the only thing we observe from a fishery is a catch history.
 Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
+-->
+
 In the absence of any data to fit to, the a stock reduction algorithm  (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure.
@@
 
-Borel's Paradox poses a particular problem  for the SRA version of sraplus.
+<!-- Borel's Paradox poses a particular problem  for the SRA version of sraplus.
 Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020).
 This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm.
+-->
 
@@
 For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on B/B~MSY~ or F/F~MSY~ in the most recent year of the fishery usable by sraplus (See Supplementary Information).
-
 All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
 The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.
-Given the scope of this analysis, we do not claim that the regressions used here are the best possible model relating these data with the fishery status indicators of interest.
+
+<!-- Given the scope of this analysis, we do not claim that the regressions used here are the best possible model relating these data with the fishery status indicators of interest.
 Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.
+-->
 
@@
 
-We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
+<!-- We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
 From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index scores, and swept area ratio (SAR) values (Table.\@ref(tab:data)).
+-->
 
@@
 
-As a case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
+We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
 We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al.

### 2021-01-19 documents/ovando-etal-assessing-global-fisheries.Rmd
back to master

@@
  tmp %>% 
-  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
+  knitr::kable(caption = "Estimates of *B/B~MSY~* by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
                Rosenberg et al. 2018, and Pauly 2007",
@@
 The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
-We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus.
 Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.
@@
 All of our estimates are Bayesian in nature.
-sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo [].
+sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo with the No-U-Turn sampler [@hoffman2011].
 Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
 A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
-sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~.
+sraplus allows users to specify the most recent status in units of depletion, *B/B~MSY~*, *F*, or *F/F~MSY~*.
 We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history.
@@
 Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries.
-For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on B/B~MSY~ or F/F~MSY~ in the most recent year of the fishery usable by sraplus (See Supplementary Information).
+For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on *B/B~MSY~* or *F/F~MSY~* in the most recent year of the fishery usable by sraplus (See Supplementary Information).
 All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
@@
 Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
-While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both B/B~MSY~ and F/F~MSY~ (Fig.\@ref(fig:cs-plot)).
+While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both *B/B~MSY~* and *F/F~MSY~* (Fig.\@ref(fig:cs-plot)).
 
@@
 
-```{r cs-plot, fig.cap="RAM values of B/B~MSY~ and F/F~MSY~ (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
+```{r cs-plot, fig.cap="RAM values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
 Each point is a stock.
@@
 We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
-Overall the sraplus estimates of B/B~MSY~ resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+Overall the sraplus estimates of *B/B~MSY~* resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
 This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
@@
 However, this is somewhat an artifact of the data.
-The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near one.
+The status of most stocks in RAM is also relatively good, with recent *B/B~MSY~* values generally near one.
 This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise.
@@
 
-```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent B/B~MSY~ by FAO statistical area from different data sources.
+```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources.
 RAM Index refers to catch and abundance index drawn from RAM.
@@
 2017 [@froese2017].
-Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
+Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
 ram_mpe_map_plot
@@
 
-```{r mape-map, fig.cap = "Median absolute percent error in most recent B/B~MSY~ by FAO statistical area from different data sources.
+```{r mape-map, fig.cap = "Median absolute percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources.
 RAM Index refers to catch and abundance index drawn from RAM.
@@
 2017 [@froese2017].
-Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
+Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
 ram_mape_map_plot
@@
 For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
+RAM *U/U~MSY~* assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
 FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
@@
 2017 [@froese2017].
-Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
+Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
 
@@
 
-```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of B/B~MSY~ in the most recent 5 years of data available for each model fit.
+```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
 Distribution is full posterior probability distribution.

### 2021-01-19 documents/ovando-etal-assessing-global-fisheries-si.Rmd
Chop (#10)
* under 10 printed pages now

@@
 
+```{r param-tab}
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "$K$", `Default Prior` = "Prior predictive tuning") %>%
+  rbind(c("Growth rate", "$r$", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
+  rbind(c("Shape parameter", "$m$", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
+  rbind(c("Catchability", "$q$", "$logn(1e^{-3}, 1)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,1)$")) %>%
+  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.5,0.25)$")) %>%
+  rbind(c("Initial State", "$B0$", "Posterior probability dist. of catch-based regressions"))
+
+knitr::kable(
+  cd,
+  align = "l",
+  escape = F,
+  row.names = F,
+  caption = "Name, abbreviations, and priors distribution for parameters potentially estimated by sraplus in this manuscript.",
+  format = "pipe"
+) %>%
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+```
+
 ```{r si-prior-tab}
@@
 ## Prior Predictive Tuning {.unnumbered}
+
+<!-- We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
+Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.
+-->
+
+<!-- Suppose that the only thing we observe from a fishery is a catch history.
+Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
+-->
+
+In the absence of any data to fit to, the a stock reduction algorithm  (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure.
+This creates a problem for the Bayesian nature of our analysis.
+Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
+Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
+Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
+
+This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
+
+Borel's Paradox poses a particular problem  for the SRA version of sraplus.
+Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020).
+This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm.
+
+We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000].
+Our solution amounts to a two-step sample-importance-resampling (SIR) algorithm.
+We first run the standard SIR algorithm as described above.
+We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
+The net result of this is that it allows users to place explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior, rather than creating a complicated and biased prior on stock status based on a mixture of explicit and implicit priors.
+See Supplementary Information for a detailed explanation of this problem and our solution.
 

### 2021-02-03 documents/ovando-etal-assessing-global-fisheries.Rmd
merging back to master

@@
 bibliography: ["../references.bib"]
-csl: proceedings-of-the-royal-society-b.csl
+csl: pnas.csl
 output: 
+  bookdown::pdf_document2:
+    latex_engine: xelatex
   bookdown::word_document2:
     reference_docx: template.docx
-  bookdown::pdf_document2:
-    latex_engine: xelatex
 params:
@@
 abstract: |
-  Assessments of the global state of fish populations play an important role in tracking the implementation of the United Nations Sustainable Development Goals.
+  Assessments of the global state of fish populations play a critical role in tracking the implementation of the United Nations Sustainable Development Goals.
 While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain.
@@
 linkcolor: blue
+header-includes:
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}\linenumbers
 ---
@@
 
+# Significance Statement  {.unnumbered}
+
+<!-- Authors must submit a 120-word-maximum statement about the significance of their research paper written at a level understandable to an undergraduate-educated scientist outside their field of specialty.
+The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
+-->
+
+The health of ocean ecosystems and the well-being of millions of people depend on our ability to accurately estimate the state of marine fish populations.
+However, these assessments typically require extensive amounts of high-quality data that are not available for large numbers of fisheries that are critical sources of economic and nutritional benefits.
+Many studies have tried to estimate the state of these unassessed fisheries based solely on the history of catches reported for a fishery, but we show here that these catch-based estimates produce unreliable results.
+We demonstrate that improving our understanding of global fisheries will depend on redoubled efforts at data collection and capacity building.
+
+\newpage
 # Introduction {.unnumbered}
@@
 The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
-Marine fisheries represent one of the largest anthropogenic impacts on the oceans, and when managed sustainably are a critical source of economic prosperity and food security around the globe.
+Marine fisheries are one of the largest anthropogenic impacts in the oceans, that when managed sustainably are a critical source of economic prosperity and food security around the globe.
 As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries.
-While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020].
+While our understanding of the state of world fisheries has improved over the last decade, the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their population size relative to reference points [@hilborn2020; @fao2020].
 This is a major impediment to ensuring the sustainable development of the world's oceans.
@@
 
+What do we currently know about the state of fished marine species around the world?
 The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries.
@@
 Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012].
-The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the biological state of assessed fish populations around the globe.
+The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the state of assessed fish populations around the globe.
 
+XX precise numerical outputs XX 
 That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed".
@@
 
-We evaluate the ability of alternative assessment models to improve our estimate of the population status of unassessed fish stocks around the world.
+<!-- We evaluate the ability of alternative assessment models to improve our estimate of the population status of unassessed fish stocks around the world.  -->
+
 While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks lacking the data needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
-Due to data limitations, all of these global assessment efforts use forms of "catch-only" data-limited models (Free et al.
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
+Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (Free et al.
 2020 [@free2020] and references therein).
-These models seek to infer stock status, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries (e.g.
-@cabral2020,@gaines2018).
+These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being our measure of stock status), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries (e.g.
+[@cabral2020],[@gaines2018]).
 
 However, Free et al.
-2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*.
+2020 [@free2020] demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*.
 These issues become apparent when we consider some of the macro-level predictions made by these models.
-While the estimates *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable catch-only models intended for use when insufficient data are available for a full stock assessment model.
+While the estimates *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model.
 
-<!-- Costello et al.
-2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be overly-optimistic, and the Northeast Pacific should be better by comparison.
+Costello et al.
+2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison.
 The superensemble method [@anderson2017a] used in Rosenberg et al.
@@
 The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
-However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.
--->
+However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions resulting from these catch-only efforts at assessing global fisheries is concerning.
 
@@
 
-While each of these prior efforts at assessing global fisheries made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating stock status of the worlds unassessed fisheries.
-In this paper we ask, can augmenting the types of catch-only models employed in global assessments to date with new broadly available data improve our understanding of the state of global fish populations?
+While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating stock status of the individual unassessed fisheries around the world.
+The types of global assessment efforts detailed in Table.\@ref(tab:status-tab) share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure.
+Under this assumption, we can obtain better estimates of the state of unassesssed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries.
+However, critical evaluation of the predictions made by these models (Table.\@ref(tab:status-tab)), as well as works such as @pauly2013 and @branch2011 call into question this very idea.
+In this paper, rather than seeking to improve our knowledge of unassessed fisheries by modifying catch-only models themselves, we ask how much can estimates from catch-only assessments be improved by moving beyond catch-only and augmenting catch histories with additional sources of broadly available fisheries data?
+We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis, not on incremental improvements to assessment methods based solely on catch histories.
+
+To accomplish this we developed a stock assessment software package, sraplus, whose key attribute is not substantial improvements in model fitting methods per say, but the ability to easily incorporate multiple kinds of fishery data in a statistically rigorous manner.
+At its most "data-limited", sraplus can work as a catch-only model in the manner of [@froese2017].
+These catch histories can then be augmented with prior information on stock status derived from increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases.
+The FMI database provides self-reported scores reflecting the strength of fisheries management in many fisheries around the world, which sraplus uses to construct informative priors on stock status, e.g.
+placing a higher probability that fisheries with high FMI scores have better stock status than those with low FMI scores.
+SAR is a measure of the intensity of trawl fishing within a particular area, which sraplus uses to construct an informative prior on the magnitude of fishing mortality.
+While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries.
+As an alternative though, Rousseau et al.
+2019 [@rousseau2019b] present a global reconstruction of fishing effort for countries around the world.
+We pair these effort reconstructions with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status.
+In this way then sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).
+
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Data Source                                 | Short Name       | Data Use                                                                 | Caveats                                                                                                         |
++=============================================+==================+==========================================================================+=================================================================================================================+
+| Catch data [@fao2020]                       | catches          | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Fisheries Management Index [@melnychuk2017] | FMI              | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Swept Area Ratio [@amoroso2018]             | SAR              | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Reconstructed effort data [@rousseau2019b]  | effort           | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors.
+Assumed rate of technology creep reported in individual sections |
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+
+: (\#tab:data) Data sources included across model fits.
+
+<!-- In this paper we ask, can augmenting the types of catch-only models employed in global assessments to date with new broadly available data improve our understanding of the state of global fish populations?
 We use a flexible stock assessment software package, [sraplus](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
-We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis.
+-->
 
-# Methods {.unnumbered}
+<!-- ## Experiment Structure {.unnumbered} -->
 
-We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database.
+We then used sraplus to evaluate the performance of estimates of global fisheries derived by using different combinations of these broadly available data, and as a benchmark compared these results with estimates from the increasingly utilized default settings of the catch-only version of the CMSY described in Froese et al.
+2017 [@froese2017] (See Table.S1).
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history.
+*B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
+We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
+This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
+We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM itself, as measure of the ability of models like sraplus if given perfect information.
+We assessed model performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
+Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
+This experiment structure allows us to assess the ability of catch-only models to estimate the *B/B~MSY~* values reported in RAM, and evaluate how much these estimates can be improved by augmenting the catch histories with additional sources of information.
+
+# Results and Discussion {.unnumbered}
+
+## Case Study {.unnumbered}
+
+We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
+From there, we assess the performance of models fit using different kinds of broadly available data at the global scale.
+We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
+We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al.
+2019 [@rousseau2019] to each stock.
+As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available.
+We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus.
+Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
+
+While CMSY systemically overestimated fishing mortality rates and underestimated stock status for these case study RAM stocks, use of the fishery-specific SAR, FMI, and reginoal effort data, produced substantially more accurate results for both *B/B~MSY~* and *F/F~MSY~* (Fig.\@ref(fig:cs-plot)).
+This is not in and of itself a surprising result, but demonstrates how 1) estimates of stock status based on catch histories alone can be both imprecise and biased and how 2) augmenting catch histories with additional data from the fisheries in question can produce meaningful improvements.
+
+<!-- `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model.
+We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~.
+Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.
+-->
+
+<!-- Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
+CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
+Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
+Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
+This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values (Fig.\@ref(fig:cs-plot)).
+-->
+
+```{r cs-plot, fig.cap="RAM values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
+Each point is a stock.
+Black dashed line shows the 1:1 relationship."}
+ex_scatter_plot
+```
+
+<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
+In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
+Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
+-->
+
+## Global Value of Information {.unnumbered}
+
+<!-- We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
+B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
+We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
+This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
+-->
+
+<!-- We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
+We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
+2017 [@froese2017] (See Table.\@ref(tab:dat-desc)).
+We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
+Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
+-->
+
+We first consider the performance of sraplus when the model is given a perfect index of abundance from RAM for each stock in RAM it is tasked with assessing.
+Assessment models in RAM are typically much more structurally complex than the simple biomass dynamics used in sraplus, and so this "perfect information" test tells us how much of a penalty in model performance we are likely to pay due to model misspecification alone.
+The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RAM were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
+
+Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves.
+We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status beyond those derived from catch-only methods.
+Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)).
+However, this is somewhat an artifact of the data.
+The status of most stocks in RAM is relatively good, with recent *B/B~MSY~* values generally near one.
+This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise.
+Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than the RAM abundance indices are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
+The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`.
+Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.
+
+Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process.
+Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions.
+For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
+We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
+Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
+
+```{r perf-tab}
+
+knitr::kable(perf, digits = 2,
+             caption = "Global performance statistics in the most recent year available of models using different sources of data.
+mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
+Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.
+See Table.5 for details of data types.",
+              format = "pipe")
+
+```
+
+<!-- xx dig into this, which stocks are in that region? XX -->
+
+```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+RAM Index refers to catch and abundance index drawn from RAM.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
+ram_mpe_map_plot
+```
+
+```{r mape-map, fig.cap = "Median absolute percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+RAM Index refers to catch and abundance index drawn from RAM.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
+ram_mape_map_plot
+```
+
+```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
+RAM Index refers to catch and abundance index drawn from RAM.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+RAM *U/U~MSY~* assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
+
+ram_acc_map_plot 
+```
+
+\newpage
+
+The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
+What sources of data might provide the greatest value in improving our estimates of global stock status?
+We used sraplus together with the RAM database to estimate the value of information, measured as the average reduction in root mean squared error, resulting from use of different kinds of data (Fig.\@ref(fig:voi-plot)).
+Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available.
+Interestingly though, having access to only an accurate estimate of F/F~MSY~ in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself.
+While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average, due to the lack of historical context for recent trends in abundance.
+We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].
+
+```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
+Distribution is full posterior probability distribution.
+Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
+Change is relative to the mean performance of a catch-only heuristic model."}
+b_voi_plot + 
+  geom_vline(aes(xintercept = 0), alpha = 0.75)
+```
+
+##  Discussion {.unnumbered}
+
+<!-- XX somewhere in here make clear that you're not learning anything from the catch data XX -->
+
+<!-- XX Acknolwedge the need for action the the frustration of being told "get more data" XX -->
+
+<!-- Global-level assessment are critical for guiding management agendas for the world's oceans.
+Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world.
+Numerous efforts have sought to resolve this problem through development of new modeling methods.
+These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment.
+However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes.
+While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.
+-->
+
+Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals.
+Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fish stocks remains uncertain in many parts of the world.
+The hope of efforts using catch-only models to estimate the status of global unassessed fisheries is fundamentally that we can learn something meaningful about the state of a fished population simply by knowing something about its catch history and life history.
+Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RAM.
+While in some cases addition of globally available data such a FMI scores, SAR values, or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
+
+What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends on the needs of those using the assessment outputs.
+It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
+In some instances using the data presented here did provide some improvement over use of catch-only style assessment methods; the difficulty comes in attempting to apply data types uniformly across the globe.
+While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RAM, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
+That use of none of the datasets collected here can achieve that, and that our test on the RAM data suggests that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the quality and use of data themselves.
+
+We chose to test the performance of methods against estimated values reported in RAM.
+A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RAM.
+The clear and systemically poor performance of the methods tested here when applied to RAM stocks would require though that these methods have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks if they are to be expected to provide reasonable picture of global fish populations.
+Free et al.
+2020's [@free2020] simulation testing of the types of methods included here suggests that this is unlikely to be true.
+
+<!-- Our estimates of global stock status vary wildly depending on which source of data we use.
+Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information?
+Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)).
+Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used.
+These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability.
+The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations.
+Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks.
+None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RAM (XX RAM matches?
+need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).
+-->
+
+Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions.
+The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e.
+those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)).
+Effort data such as those reconstructed by Rousseau et al.
+2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
+Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
+
+But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
+Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas of the globe results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
+
+Our value-of-information analysis also shows though the high utility of having access to even a recent snapshot of F/F~MSY~ (Fig.\@ref(fig:voi-plot)).
+Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct fishery-specific priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
+Another avenue would be to prioritize the development of a global repository for length and age composition data.
+Given appropriate conditions, these length measurements can be used to estimate local fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
+While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
+Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f; @rudd2017].
+
+<!-- Using catch data alone does not reveal about recent stock status beyond what is encoded in our priors and functional forms.
+So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest).
+Given the demonstrated value of F/F~MSY~ values in improving estimates of stock status, and the increased feasibility in estimating F/F~MSY~ at a local scale relative to B/B~MSY~, how might we go about obtaining estimates of fishing mortality rates at a global scale?
+We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/F~MSY~ for individual fisheries (see Supplemental Information).
+While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
+-->
+
+We must also prioritize collection and curation of fish population survey data worldwide.
+Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//).
+Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments.
+Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of [@rousseau2019] might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes
+
+Expanded training of fisheries scientists around the globe is another critical need.
+Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise.
+This is why stock assessments even in data-rich fisheries are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
+Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
+
+## Conclusions {.unnumbered}
+
+The coming decades are a critical time for the future of fisheries and ocean health.
+Achieving the United Nations Sustainable Development Goal 14 for the conservation and sustainable use of the world's oceans depends on our ability to effectively assess the status of fish stocks around the world.
+The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
+However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed.
+Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a solution to this problem to has proven to be unbiased and sufficiently precise at a global or regional level.
+
+<!-- Our results show that improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists. -->
+
+The lack of strong information on stock status within catch histories alone demonstrated here means that differences in models and assumptions between catch-based assessment efforts can produce starkly contrasting conclusions on global stock status, leading to debates that are inconclusive as they are inherently driven by assumptions.
+The FAO is leading efforts to increase technical capacity and monitoring and evaluation infrastructure to improve fisheries management in places with limited data.
+Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets.
+Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally [@costello2020].
+Catch-only methods may be a useful starting point when applied properly, but achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.
+
+<!-- The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries. -->
+
+# Materials and Methods {.unnumbered}
+
+<!-- We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database.
+-->
 All analysis were conducted in the R programming language [@rcoreteam2019].
@@
 
-## Data Sources {.unnumbered}
+<!-- ## Data Sources {.unnumbered} -->
 
-We restricted ourselves to databases with wide coverage around the world, representing the types of information that can be leveraged in efforts to assess the status of global fisheries (Table.\@ref(tab:data)).
-While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries.
-
-| Data Source           |  Short Name | Data Use |  Caveats |
-|-----------------------|-------------|------|------|
-| Catch data [@fao2020] | catches     | Priors on stock status, scaling of population size, exploitation history    |    Heuristics or regressions used to translate shape of catch history into priors on stock status  |
-|    Fisheries Management Index [@melnychuk2017]     | FMI     | Priors on most recent F/F~MSY~ values   |Priors produced by regression trained on data from RAM Legacy Stock Assessment Database|
-| Swept Area Ratio [@amoroso2018]      | SAR     | Priors on most recent F/F~MSY~ values |Priors produced by regression trained on data from RAM Legacy Stock Assessment Database|
-|  Reconstructed effort data [@rousseau2019b] | effort | Combined with catch data to create an index of abundance| Total reconstructed effort across all sectors.
-Assumed rate of technology creep reported in individual sections|
-Table:(\#tab:data) Data sources included across model fits.
-
-## Experiment Structure {.unnumbered}
-
-We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global unassessed fish populations.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history.
-*B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
-We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
-This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
-
-We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
-We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
-2017 [@froese2017] (See Table.S1).
-We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
-Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
+<!-- We restricted ourselves to databases with wide coverage around the world, representing the types of information that can be leveraged in efforts to assess the status of global fisheries (Table.\@ref(tab:data)).
+-->
 
@@
 
-# Results {.unnumbered}
-
-## Case Study  {.unnumbered}
-
-<!-- We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
-From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index scores, and swept area ratio (SAR) values (Table.\@ref(tab:data)).
--->
-
-<!-- `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model.
-We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~.
-Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.
--->
-
-We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
-We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al.
-2019 [@rousseau2019] to each stock.
-As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available.
-We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus.
-Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
-While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both *B/B~MSY~* and *F/F~MSY~* (Fig.\@ref(fig:cs-plot)).
-
-<!-- Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
-CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
-Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
-Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
-This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values (Fig.\@ref(fig:cs-plot)).
--->
-
-```{r cs-plot, fig.cap="RAM values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
-Each point is a stock.
-Black dashed line shows the 1:1 relationship."}
-ex_scatter_plot
-```
-
-<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
-In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
-Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
--->
-
-## Global Performance  {.unnumbered}
-
-<!-- We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
-B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
-We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
-This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
--->
-
-<!-- We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
-We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
-2017 [@froese2017] (See Table.\@ref(tab:dat-desc)).
-We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
-Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
--->
-
-We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
-Overall the sraplus estimates of *B/B~MSY~* resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
-
-Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves.
-This becomes clearer once we consider the performance of sraplus models fit to combinations of broadly available datasets.
-Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)).
-However, this is somewhat an artifact of the data.
-The status of most stocks in RAM is also relatively good, with recent *B/B~MSY~* values generally near one.
-This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise.
-Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
-The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.
-
-Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process.
-Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions.
-For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
-We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
-Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
-
-```{r perf-tab}
-
-knitr::kable(perf, digits = 2,
-             caption = "Global performance statistics in the most recent year available of models using different sources of data.
-mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
-Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.
-See Table.5 for details of data types.",
-              format = "pipe")
-
-```
-
-<!-- xx dig into this, which stocks are in that region? XX -->
-
-```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
-ram_mpe_map_plot
-```
-
-```{r mape-map, fig.cap = "Median absolute percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
-ram_mape_map_plot
-```
-
-```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-RAM *U/U~MSY~* assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
-
-ram_acc_map_plot 
-```
-
-\newpage
-
-The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
-What sources of data might provide the greatest value in improving our estimates of global stock status?
-We used sraplus together with the RAM database to estimate the value of information, measured as the average reduction in root mean squared error, resulting from use of different kinds of data (Fig.\@ref(fig:voi-plot)).
-Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available.
-Interestingly though, having access to only an accurate estimate of F/F~MSY~ in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself.
-While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average, due to the lack of historical context for recent trends in abundance.
-We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].
-
-```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
-Distribution is full posterior probability distribution.
-Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
-Change is relative to the mean performance of a catch-only heuristic model."}
-b_voi_plot + 
-  geom_vline(aes(xintercept = 0), alpha = 0.75)
-```
-
-# Discussion {.unnumbered}
-
-<!-- XX somewhere in here make clear that you're not learning anything from the catch data XX -->
-
-<!-- XX Acknolwedge the need for action the the frustration of being told "get more data" XX -->
-
-<!-- Global-level assessment are critical for guiding management agendas for the world's oceans.
-Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world.
-Numerous efforts have sought to resolve this problem through development of new modeling methods.
-These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment.
-However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes.
-While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.
--->
-
-Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals.
-Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fish stocks remains uncertain in many parts of the world.
-The hope of efforts using catch-only models to estimate the status of global unassessed fisheries is fundamentally that characteristics of a fishery's catch history and life history are informative as to the status its population.
-Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RAM.
-While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
-
-What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends on the needs of those using the assessment outputs.
-It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
-In some instances using the data presented here did provide some improvement over use of catch-only style assessment methods; the difficulty comes in attempting to apply data types uniformly across the globe.
-While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RAM, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
-That use of none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the quality and use of data themselves.
-
-We chose to test the performance of methods against estimated values reported in RAM.
-A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RAM.
-The clear and systemically poor performance of the methods tested here when applied to RAM stocks would require though that these methods have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks if they are to be expected to provide reasonable picture of global fish populations.
-Free et al.
-2020's [@free2020] simulation testing of the types of methods included here suggests that this is unlikely to be true.
-
-<!-- Our estimates of global stock status vary wildly depending on which source of data we use.
-Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information?
-Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)).
-Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used.
-These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability.
-The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations.
-Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks.
-None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RAM (XX RAM matches?
-need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).
--->
-
-Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions.
-The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e.
-those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)).
-Effort data such as those reconstructed by Rousseau et al.
-2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
-Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
-
-But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
-Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas of the globe results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
-
-Our value-of-information analysis also shows though the high utility of having access to even a recent snapshot of F/F~MSY~ (Fig.\@ref(fig:voi-plot)).
-Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct fishery-specific priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
-Another avenue would be to prioritize the development of a global repository for length and age composition data.
-Given appropriate conditions, these length measurements can be used to estimate local fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
-While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
-Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f; @rudd2017].
-
-<!-- Using catch data alone does not reveal about recent stock status beyond what is encoded in our priors and functional forms.
-So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest).
-Given the demonstrated value of F/F~MSY~ values in improving estimates of stock status, and the increased feasibility in estimating F/F~MSY~ at a local scale relative to B/B~MSY~, how might we go about obtaining estimates of fishing mortality rates at a global scale?
-We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/F~MSY~ for individual fisheries (see Supplemental Information).
-While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
--->
-
-We must also prioritize collection and curation of fish population survey data worldwide.
-Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//).
-Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments.
-Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of [@rousseau2019] might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes
-
-Expanded training of fisheries scientists around the globe is another critical need.
-Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise.
-This is why stock assessments even in data-rich fisheries are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
-Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
-
-# Conclusions {.unnumbered}
-
-The coming decades are a critical time for the future of fisheries and ocean health.
-Achieving the United Nations Sustainable Development Goal 14 for the conservation and sustainable use of the world's oceans depends on our ability to effectively assess the status of fish stocks around the world.
-The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
-However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed.
-Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a solution to this problem to has proven to be unbiased and sufficiently precise at a global or regional level.
-
-<!-- Our results show that improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists. -->
-
-The lack of strong information on stock status within catch histories demonstrated here means that differences in models and assumptions between catch-based assessment efforts can produce starkly contrasting conclusions on global stock status, leading to debates that are inconclusive as they are inherently driven by assumptions.
-The FAO is leading efforts to increase technical capacity and monitoring and evaluation infrastructure to improve fisheries management in places with limited data.
-Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets.
-Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally [@costello2020].
-Catch-only methods may be a useful starting point when applied properly, but achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.
-
-<!-- The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries. -->
-
-# Acknowledgements{.unnumbered}
+# Acknowledgements {.unnumbered}
 
 Funding for this work was provided by the Food and Agriculture Organization of the United Nations (FAO).
-We thank participants of the FAO "Methods for Global Assessment" workshop held in Rome, Italy February 2019 for helpful feedback on this project.
-
-# Disclaimer {.unnumbered}
-
-The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of NOAA or the Department of Commerce.
-
-# References {.unnumbered}
+We thank participants of the FAO "Methods for Global Assessment" workshop held in Rome, Italy February 2019 for helpful feedback on this project.The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of NOAA or the Department of Commerce.
 
@@
 :::
+
+# Supporting Information {.unnumbered}
+
+```{r si, child = 'ovando-etal-assessing-global-fisheries-si.Rmd'}
+```

### 2021-02-03 documents/ovando-etal-assessing-global-fisheries-si.Rmd
reordering SI a bit
merging back to master (#11)

@@
----
-title: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
-subtitle: Supplementary Information
-author: 
-  - Daniel Ovando
-  - Ray Hilborn
-  - Cole Monnahan
-  - Merrill Rudd
-  - Rishi Sharma
-  - James Thorson
-  - Yannick Rousseau
-  - Yimin Ye
-date: "`r Sys.Date()`"
-bibliography: ["../references.bib"]
-csl: proceedings-of-the-royal-society-b.csl
-output: 
-  bookdown::pdf_document2:
-    latex_engine: xelatex
-  bookdown::word_document2:
-    reference_docx: template.docx
-params:
-  results_name: ["v1.0"]
-  min_years_catch: [25]
-always_allow_html: true
-linkcolor: blue
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
-                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
-library(tidyverse)
-library(viridis)
-library(sf)
-library(hrbrthemes)
-library(extrafont)
-library(scales)
-library(patchwork)
-library(rstanarm)
-library(bayesplot)
-library(here)
-library(sraplus)
-extrafont::loadfonts()
-
-functions <- list.files(here::here("functions"))
-
-purrr::walk(functions, ~ source(here::here("functions", .x)))
-
-results_name <- params$results_name
-
-results_path <- here("results",results_name)
-
-min_years_catch <- params$min_years_catch
-
-prepare_sofia_data(lookup_fmi_names = FALSE)
-
-ram_comp_data <- ram_data %>%
-  mutate(has_things = !(is.na(catch) | catch == 0)) %>%
-  filter(has_things) %>%
-  group_by(stockid) %>%
-  mutate(delta_year = as.integer(year - lag(year))) %>%
-  mutate(delta_year = case_when(year == min(year) ~ as.integer(1),
-                                TRUE ~ delta_year)) %>%
-  mutate(missing_gaps = any(delta_year > 1)) %>%
-  filter(missing_gaps == FALSE) %>%
-  group_by(stockid) %>%
-  mutate(n = length(catch)) %>%
-  filter(n >= min_years_catch) %>%
-  ungroup()
-
-ram_comp_data <- ram_comp_data %>%
-  select(
-    year,
-    stockid,
-    country,
-    primary_FAOarea,
-    scientificname,
-    commonname,
-    year,
-    catch,
-    b_v_bmsy,
-    u_v_umsy,
-    total_biomass
-  ) %>%
-  rename(scientific_name = scientificname) %>% 
-  mutate(fao_area_code = as.numeric(primary_FAOarea)) %>% 
-  left_join(fao_species, by = "scientific_name")
-
-assess_ram_fits <- read_rds(path = file.path(results_path,"assess_ram_fits.rds"))
-
-ram_v_sraplus_area <- read_rds(path = file.path(results_path,"ram_v_sraplus_area.rds"))
-
-ram_comp_data <- read_rds(path = file.path(results_path,"ram_comp_data.rds"))
-
-fao_sraplus_comp_data <- read_rds(path = file.path(results_path, "fao_sraplus_comp_data.rds"))
-
-load(file = file.path(results_path, "paper_plots.Rdata"))
-
-load(file = file.path(results_path, "voi_fits.RData"))
-
-  exs <- read_rds(path = file.path(results_path, "raw-case-study-fits.rds"))
-
-breaks <- c(0.8, 1.2)
-
-breaks <- c(0, breaks, Inf)
-
-labels <- c("over", "fully", "under")
-
-# rand_fits <- assess_ram_fits %>%
-#   filter(data == "ram-data") %>%
-#   mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
-#          data = "guess")
-
-perf <- assess_ram_fits %>% 
-  # bind_rows(rand_fits) %>% 
-  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
-  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
-  mutate(data = as.factor(data)) %>% 
-  mutate(data = forcats::fct_recode(data, 
-                                    "RAM Index" = "ram-data",
-                                    "SAR" = "sar",
-                                    "FMI" = "fmi",
-                                    "Effective CPUE" = "cpue",
-                                    "Effective CPUE+" = "cpue-plus",
-                                    "Nominal CPUE" = "nominal-cpue",
-                                    "Nominal CPUE+" ="nominal-cpue-plus",
-                                    "CMSY" = "cmsy",
-                                    "Guess" = "guess",
-                                    "RAM U/Umsy" = "u_umsy"
-
-                                    )) %>% 
-  group_by(data) %>% 
-  summarise(
-    mpe = median((mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
-    mape = median(abs(mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
-    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
-  ) %>% 
-  arrange((mape)) %>% 
-  rename("Data Used" = "data")
-
-fao_area_codes <- fao %>% 
-  ungroup() %>% 
-  select(fao_area, fao_area_code) %>% 
-  unique()
-
-pub_theme <- theme_ipsum(base_size = 10,
-                         axis_text_size = 10,
-                         axis_title_size = 12) +
-    theme(panel.spacing = unit(.5,"lines"))
-
-theme_set(pub_theme)
-
-```
-
-# Supplementary Information {.unnumbered}
-
 \renewcommand{\thefigure}{S\arabic{figure}}
@@
 
-## Value of Information Calculation {.unnumbered}
-
-We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
-We generate fits for 3000 combinations of a RAM stock and available data.
-For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
-For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
-The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
-
-Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
-We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
-
-$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
-
-Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
-
-## Catch History {.unnumbered}
+### Catch History {.unnumbered}
 
@@
 ```
+
+## Value of Information Calculation {.unnumbered}
+
+We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
+The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
+We generate fits for 3000 combinations of a RAM stock and available data.
+For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
+For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
+The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
+
+Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
+We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form
+
+$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$
+
+Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$

### 2021-02-04 documents/ovando-etal-assessing-global-fisheries.Rmd
submitted

@@
 output: 
+  bookdown::word_document2:
+    reference_docx: template.docx
   bookdown::pdf_document2:
     latex_engine: xelatex
-  bookdown::word_document2:
-    reference_docx: template.docx
 params:
@@
   mutate(data = as.factor(data)) %>% 
+  filter(data != "u_umsy") %>% 
   mutate(data = forcats::fct_recode(data, 
@@
                                     "Guess" = "guess",
-                                    "RAM U/Umsy" = "u_umsy"
+                                    "RAM F/Fmsy" = "u_umsy"
 
@@
 
-XX precise numerical outputs XX 
 That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed".
+While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
 The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020].
@@
 With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
+In addition, the SOFIA process assigns included unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to to the biomass that would on average result in maximized yield over the long-term.
 
@@
 
-While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
+Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide numerical estimates of the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
 Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (Free et al.
@@
  tmp %>% 
-  knitr::kable(caption = "Estimates of *B/B~MSY~* by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
+  knitr::kable(caption = "Estimates of fish stock status by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
                Rosenberg et al. 2018, and Pauly 2007",
@@
 The types of global assessment efforts detailed in Table.\@ref(tab:status-tab) share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure.
-Under this assumption, we can obtain better estimates of the state of unassesssed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries.
+Under this assumption, we can obtain better estimates of the state of unassessed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries.
 However, critical evaluation of the predictions made by these models (Table.\@ref(tab:status-tab)), as well as works such as @pauly2013 and @branch2011 call into question this very idea.
@@
 2019 [@rousseau2019b] present a global reconstruction of fishing effort for countries around the world.
-We pair these effort reconstructions with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status.
+We pair these effort reconstructions (both nominal effort and effective effort assuming an annual rate of efficiency increase of 2.6%) with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status.
 In this way then sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).
@@
 
-# Results and Discussion {.unnumbered}
+# Results {.unnumbered}
 
@@
 
-While CMSY systemically overestimated fishing mortality rates and underestimated stock status for these case study RAM stocks, use of the fishery-specific SAR, FMI, and reginoal effort data, produced substantially more accurate results for both *B/B~MSY~* and *F/F~MSY~* (Fig.\@ref(fig:cs-plot)).
+While CMSY systemically overestimated fishing mortality rates and underestimated stock status for these case study RAM stocks, use of the fishery-specific SAR, FMI, and regional effort data, produced substantially more accurate results for both *B/B~MSY~* and *F/F~MSY~* (Fig.\@ref(fig:cs-plot)).
 This is not in and of itself a surprising result, but demonstrates how 1) estimates of stock status based on catch histories alone can be both imprecise and biased and how 2) augmenting catch histories with additional data from the fisheries in question can produce meaningful improvements.
@@
 Assessment models in RAM are typically much more structurally complex than the simple biomass dynamics used in sraplus, and so this "perfect information" test tells us how much of a penalty in model performance we are likely to pay due to model misspecification alone.
-The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RAM were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RAM were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "RAM Index"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "RAM Index"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "RAM Index"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
 This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
@@
 Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than the RAM abundance indices are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
-The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.
+The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RAM Index","Guess")]))`.
+Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "Guess"])`.
 
@@
 For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
 FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
@@
 For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-RAM *U/U~MSY~* assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region.
 FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
@@
 
-##  Discussion {.unnumbered}
+#  Discussion {.unnumbered}
 
@@
 Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
-
-## Conclusions {.unnumbered}
 
@@
 
-# Acknowledgements {.unnumbered}
+# Acknowledgements and Funding {.unnumbered}
 

### 2021-02-04 make-assessing-global-fisheries.R

@@ -58,7 +58,7 @@ results_name <- "v1.0"
 results_description <-
   "publication version of results"
 
-run_voi_models <- FALSE
+run_voi_models <- TRUE
 # sub options for run_voi_models
 fit_models <- FALSE
 
@@ -74,9 +74,9 @@ run_ram_tests <- FALSE
 
 run_ram_comparison <- FALSE
 
-knit_paper <- TRUE
+knit_paper <- FALSE
 
-warning("Running full analysis takes upwards of 24 hours on 2 cores. Recommend starting on a Friday night having a nice weekend. Given memory constraints of models, more than 2 cores is not recommended (memory routinley runs out on a 16 core 36GB machine when cores > 2)")
+warning("Running full analysis takes upwards of 24 hours on 2 cores. Recommend starting on a Friday night and then having a nice weekend. Given memory constraints of models, more than 2 cores is not recommended (memory routinely runs out on a 16 core 36GB machine when cores > 2)")
 
 engine <-  "stan"
 
@@ -2010,7 +2010,8 @@ fao_area_ram_status <- fao_areas %>%
   left_join(ram_v_sraplus_area,
             by = c("f_area" = "primary_fao_area")) %>%
   mutate(pe = (median_sraplus - median_ram) / median_ram) %>% 
-  mutate(data = fct_relevel(data, "ram-data"))
+  mutate(data = fct_relevel(data, "ram-data")) %>% 
+  filter(data != "u_umsy")
 
 
 ram_status <- fao_area_ram_status %>% 
@@ -2069,7 +2070,7 @@ ram_mpe_map_plot <- fao_area_ram_status %>%
       frame.colour = "black"
     )
   ) + 
-  theme(legend.position = c(.75,.1),
+  theme(legend.position = "top",
         legend.direction = "horizontal",
         panel.background = element_rect(fill = "white"),
         legend.text = element_text(size = 8))
@@ -2101,7 +2102,7 @@ ram_mape_map_plot <- fao_area_ram_status %>%
       frame.colour = "black"
     )
   ) + 
-  theme(legend.position = c(.75,.1),
+  theme(legend.position = "top",
         legend.direction = "horizontal",
         panel.background = element_rect(fill = "white"))
 
@@ -2131,7 +2132,7 @@ ram_acc_map_plot <- fao_area_ram_status %>%
       ticks.colour = "white",
       frame.colour = "black"
     )) +
-    theme(legend.position = c(.75,.1),
+    theme(legend.position = "top",
[…]

### 2021-04-18 documents/ovando-etal-assessing-global-fisheries-si.Rmd
cole edits in prep for f and f submission

@@
 \setcounter{table}{0}
-
-## Data Sources Used {.unnumbered}
-
-```{r dat-desc}
-dat_description <- tribble(~`Data Name`, ~Description,
-        "RAM Index", "Fit to abundance index from RAM",
-        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
-        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
-        "EFfective CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 2.6% technology creep",
-        "Effective CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
-        "Nominal CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
-         "Nominal CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
-        "Guess", "Priors on terminal B/Bmsy randomly sampled from .4,1,1.6")
-
-knitr::kable(dat_description,
-             caption = "Data sources used for terminal stock status estimate",
-             format = "pipe")
-# %>%
-#   kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-#   kableExtra::row_spec(0, bold = T)
-```
 
@@
 
-The population growth equation is
-
-\begin{equation}
-  f(x)=\begin{cases}
-      B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)p_t
-, & \text{if $B_t>0.25 \times K$}.\\
-     B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - \hat{c_t}\right)\right)p_t, & \text{otherwise}.
-  \end{cases}
-  (\#eq:sihockey)
-\end{equation}
-
 ## Prior Predictive Tuning {.unnumbered}
 
-<!-- We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.
--->
-
-<!-- Suppose that the only thing we observe from a fishery is a catch history.
-Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
--->
-
-In the absence of any data to fit to, the a stock reduction algorithm  (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure.
-This creates a problem for the Bayesian nature of our analysis.
-Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
-Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
-Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
-
-This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-
-Borel's Paradox poses a particular problem  for the SRA version of sraplus.
-Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020).
-This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm.
-
-We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000].
+Our prior predictive tuning regime is similar in spirit to Bayesian melding [@poole2000].
 Our solution amounts to a two-step sample-importance-resampling (SIR) algorithm.
@@
 The net result of this is that it allows users to place explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior, rather than creating a complicated and biased prior on stock status based on a mixture of explicit and implicit priors.
-See Supplementary Information for a detailed explanation of this problem and our solution.
 
@@
 In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
-This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.S1).
+This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.\@ref(fig:ppplot)).
 
@@
 
-```{r ppplot, fig.cap="Prior posterior plots of fits for case study fishery in Fig.6"}
+```{r ppplot, fig.cap="Prior posterior plots of fits for case study fishery"}
 
@@
 
-#### Catch-Only Priors {.unnumbered}
+### Catch-Only Priors {.unnumbered}
 
@@
 
-#### Fisheries Management Index Priors {.unnumbered}
-
-The Fisheries Management Index (FMI), as presented in [@melnychuk2017], utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
-These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
-The higher the score, the better the expert judges that a given metric is met in that fishery.
-Importantly, FMI surveys can be filled out in the absence of stock assessments.
-This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to [@osio2015] and [@cope2015]).
-
-The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
-
-$$log(value_i) \sim N(s(research_i) + s(management_i) +
-s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
-
-#### Swept Area Ratio Priors {.unnumbered}
-
-[@amoroso2018] provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
-This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
-As illustrated in [@amoroso2018], there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
-Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
-if all trawl-able areas are trawled twice a year then the SAR will be 2.
-Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
-
-The final selected model relating SAR to to stock status metrics was
-
-$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
-
-### Swept Area Ratio {.unnumbered}
-
-```{r sar-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of swept area ratio (SAR) on F/F~MSY~"}
-fit <- sraplus::sar_models$fit[sraplus::sar_models$metric == "mean_uumsy"][[1]]
-
-fit_data <- fit$data
-
-pp <- posterior_predict(fit, draws = 1000)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
-  scale_x_continuous(name = "Observed Log F/Fmsy") + 
-  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-
-```
-
-### Fisheries Management Index {.unnumbered}
-
-```{r fmi-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of fisheries management index (FMI) on F/F~MSY~"}
-
-fit <- sraplus::fmi_models$fit[sraplus::fmi_models$metric == "mean_uumsy"][[1]]
-
-fit_data <- fit$data
-
-pp <- posterior_predict(fit, draws = 1000)
-
-r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
-  ggplot(aes(r2)) + 
-  geom_histogram() + 
-  labs(subtitle = "B) Bayesian R2")
-
-pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
-  scale_x_continuous(name = "Observed Log F/Fmsy") + 
-  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
-  labs(subtitle = "A) Observed vs. Posterior Predictive") +
- pub_theme
-
-pi_plot + r2_plot
-```
-
-### Catch History {.unnumbered}
+Fits for the catch-only prior-generating regression are visible in Fig.\@ref(fig:catch-reg).
 
@@
 
+### Fisheries Management Index Priors {.unnumbered}
+
+The Fisheries Management Index (FMI), as presented in [@melnychuk2017], utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place.
+These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic.
+The higher the score, the better the expert judges that a given metric is met in that fishery.
+Importantly, FMI surveys can be filled out in the absence of stock assessments.
+This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to [@osio2015] and [@cope2015]).
+
+The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form
+
+$$log(value_i) \sim N(s(research_i) + s(management_i) +
+s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$
+
+Fits for the FMI prior-generating regression are visible in Fig.\@ref(fig:fmi-fit).
+
+```{r fmi-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of fisheries management index (FMI) on F/F~MSY~"}
+
+fit <- sraplus::fmi_models$fit[sraplus::fmi_models$metric == "mean_uumsy"][[1]]
+
+fit_data <- fit$data
+
+pp <- posterior_predict(fit, draws = 1000)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
+  scale_x_continuous(name = "Observed Log F/Fmsy") + 
+  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+```
+
+### Swept Area Ratio Priors {.unnumbered}
+
+[@amoroso2018] provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas.
+This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts.
+As illustrated in [@amoroso2018], there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~.
+Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g.
+if all trawl-able areas are trawled twice a year then the SAR will be 2.
+Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.
+
+The final selected model relating SAR to to stock status metrics was
+
+$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$
+Fits for the SAR prior-generating regression are visible in Fig.\@ref(fig:sar-fit).
+
+```{r sar-fit, fig.cap = "Observed (x-axis) vs posterior predictive (y-axis) F/F~MSY~ for regression of swept area ratio (SAR) on F/F~MSY~"}
+fit <- sraplus::sar_models$fit[sraplus::sar_models$metric == "mean_uumsy"][[1]]
+
+fit_data <- fit$data
+
+pp <- posterior_predict(fit, draws = 1000)
+
+r2_plot <- tibble(r2 = bayes_R2(fit)) %>% 
+  ggplot(aes(r2)) + 
+  geom_histogram() + 
+  labs(subtitle = "B) Bayesian R2")
+
+pi_plot <- ppc_intervals(y = fit_data$log_value, yrep = pp, x = fit_data$log_value) + 
+  scale_x_continuous(name = "Observed Log F/Fmsy") + 
+  scale_y_continuous(name = "Posterior Predictive F/Fmsy") +
+  labs(subtitle = "A) Observed vs. Posterior Predictive") +
+ pub_theme
+
+pi_plot + r2_plot
+
+```
+
 ## Value of Information Calculation {.unnumbered}
 
-We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM.
-We generate fits for 3000 combinations of a RAM stock and available data.
-For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
-For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
-The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
-
-Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
+We performed a value-of-information (VOI) exercise we assessed performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year.
 We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form

### 2021-04-18 documents/ovando-etal-assessing-global-fisheries.Rmd
cole edits in prep for f and f submission

@@
 bibliography: ["../references.bib"]
-csl: pnas.csl
+csl: fish-and-fisheries.csl
 output: 
+  bookdown::pdf_document2:
+    latex_engine: xelatex
   bookdown::word_document2:
     reference_docx: template.docx
-  bookdown::pdf_document2:
-    latex_engine: xelatex
 params:
@@
 Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added.
-Estimates of fish populations based primarily on catch histories alone frequently performed roughly as well as a random guess.
+Estimates of fish populations based primarily on catch histories alone on average performed 29% better than a random guess.
+But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time.
 Obtaining accurate estimates of stock status for the world's exploited fish populations depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
@@
 
-# Significance Statement  {.unnumbered}
+<!-- # Significance Statement  {.unnumbered} -->
 
-<!-- Authors must submit a 120-word-maximum statement about the significance of their research paper written at a level understandable to an undergraduate-educated scientist outside their field of specialty.
+<!-- <!-- Authors must submit a 120-word-maximum statement about the significance of their research paper written at a level understandable to an undergraduate-educated scientist outside their field of specialty.
 The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
@@
 
-The health of ocean ecosystems and the well-being of millions of people depend on our ability to accurately estimate the state of marine fish populations.
+<!-- The health of ocean ecosystems and the well-being of millions of people depend on our ability to accurately estimate the state of marine fish populations.
 However, these assessments typically require extensive amounts of high-quality data that are not available for large numbers of fisheries that are critical sources of economic and nutritional benefits.
@@
 We demonstrate that improving our understanding of global fisheries will depend on redoubled efforts at data collection and capacity building.
+-->
 
@@
 The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
-Marine fisheries are one of the largest anthropogenic impacts in the oceans, that when managed sustainably are a critical source of economic prosperity and food security around the globe.
+Marine fisheries are one of the largest anthropogenic impacts in the oceans aside from the effects of climate change.
+When managed sustainably fisheries are a critical source of economic prosperity and food security around the globe.
 As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries.
@@
 This is a major impediment to ensuring the sustainable development of the world's oceans.
-In this paper, we assess why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.
+In this paper, we consider why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.
 
@@
 The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries.
-The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020].
+The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished (at or near targets for sustainable food production), 6.2% are underfished, and 34.2% are overfished [@fao2020].
 Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012].
@@
 With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
-In addition, the SOFIA process assigns included unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to to the biomass that would on average result in maximized yield over the long-term.
+In addition, the SOFIA process assigns unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to to the biomass that would on average result in maximized yield over the long-term.
 
@@
 Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide numerical estimates of the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
-Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (Free et al.
-2020 [@free2020] and references therein).
+Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (@free2020 and references therein).
 These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being our measure of stock status), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries (e.g.
-[@cabral2020],[@gaines2018]).
+@cabral2020 and @gaines2018).
 
-However, Free et al.
-2020 [@free2020] demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*.
-These issues become apparent when we consider some of the macro-level predictions made by these models.
-While the estimates *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model.
+<!-- These issues become apparent when we consider some of the macro-level predictions made by these models.
+While the estimates of *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model.
+-->
 
-Costello et al.
-2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison.
-The superensemble method [@anderson2017a] used in Rosenberg et al.
-2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific.
-The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
+<!-- @costello2016a found similar rankings of regions in terms of stock status as RAM, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison.
+The superensemble method [@anderson2017a] used in @rosenberg2018 demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific.
+Simply comparing catch divided by maximum catch, in the manner of @pauly2007, finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
 The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
-However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions resulting from these catch-only efforts at assessing global fisheries is concerning.
+However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly exploited regions as measured by RAM and the FAO resulting from these catch-only efforts at assessing global fisheries is concerning.
+-->
 
-```{r status-tab}
+```{r status-tab, include=FALSE}
 
@@
     rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
-  select(`FAO Area`,`SOFIA % Overfished`, everything())
+  select(`FAO Area`,`SOFIA % Overfished`,`B/BMSY RAM`, everything())
 
@@
   knitr::kable(caption = "Estimates of fish stock status by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
-               Rosenberg et al. 2018, and Pauly 2007",
+               Rosenberg et al. 2018, Catch / max(Catch)",
                booktabs = TRUE,
@@
 
+However, @free2020 demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*.
 While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating stock status of the individual unassessed fisheries around the world.
-The types of global assessment efforts detailed in Table.\@ref(tab:status-tab) share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure.
+The types of global assessment efforts share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure.
 Under this assumption, we can obtain better estimates of the state of unassessed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries.
-However, critical evaluation of the predictions made by these models (Table.\@ref(tab:status-tab)), as well as works such as @pauly2013 and @branch2011 call into question this very idea.
-In this paper, rather than seeking to improve our knowledge of unassessed fisheries by modifying catch-only models themselves, we ask how much can estimates from catch-only assessments be improved by moving beyond catch-only and augmenting catch histories with additional sources of broadly available fisheries data?
+However, further evaluation of the predictions made by these models [@free2020;@bouch2021;@pons2020] as well as works such as @pauly2013 and @branch2011 call into question this very idea.
+
+In this paper we ask, how much can estimates from catch-only assessments be improved by moving beyond catch-only and augmenting catch histories with additional sources of broadly available fisheries data?
+We answer this question by aggregating a set of broadly distributed datasets that might be of use to global fishery assessment.
+We then used our new assessment method, sraplus, to evaluate the performance status estimates derived from different combinations of these broadly available data.
+Wecompared these results with estimates from the increasingly utilized (e.g.
+@palomares2020) default settings of the catch-only version of CMSY described in @froese2017.
 We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis, not on incremental improvements to assessment methods based solely on catch histories.
 
-To accomplish this we developed a stock assessment software package, sraplus, whose key attribute is not substantial improvements in model fitting methods per say, but the ability to easily incorporate multiple kinds of fishery data in a statistically rigorous manner.
-At its most "data-limited", sraplus can work as a catch-only model in the manner of [@froese2017].
+# Methods {.unnumbered}
+
+The key attribute of sraplus is not substantial improvements in model fitting methods per say, but the ability to easily incorporate multiple kinds of fishery data potentially used in SRA style analyses in a statistically rigorous manner.
+We paired these data with fisheries in RAM, and then used sraplus to generate predictions of B/B~MSY~ in the most recent year of each fishery across different combinations of the data types.
+We then compared the predictions generated by sraplus using different types of data, along with predictions for the same fisheries generated by CMSY, to the reported values in RAM.
+
+All analysis were conducted in the R programming language [@rcoreteam2019].
+Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018].
+The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
+Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.
+
+## Data Sources {.unnumbered}
+
+At its most "data-limited", sraplus can work as a catch-only model in the manner of @froese2017.
 These catch histories can then be augmented with prior information on stock status derived from increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases.
@@
 While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries.
-As an alternative though, Rousseau et al.
-2019 [@rousseau2019b] present a global reconstruction of fishing effort for countries around the world.
+As an alternative, @rousseau2019b present a global reconstruction of fishing effort for countries around the world.
 We pair these effort reconstructions (both nominal effort and effective effort assuming an annual rate of efficiency increase of 2.6%) with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status.
-In this way then sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).
+In this way sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).
 
@@
 
-We then used sraplus to evaluate the performance of estimates of global fisheries derived by using different combinations of these broadly available data, and as a benchmark compared these results with estimates from the increasingly utilized default settings of the catch-only version of the CMSY described in Froese et al.
-2017 [@froese2017] (See Table.S1).
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history.
-*B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
-We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
-This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
-We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM itself, as measure of the ability of models like sraplus if given perfect information.
-We assessed model performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
-Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
-This experiment structure allows us to assess the ability of catch-only models to estimate the *B/B~MSY~* values reported in RAM, and evaluate how much these estimates can be improved by augmenting the catch histories with additional sources of information.
-
-# Results {.unnumbered}
-
-## Case Study {.unnumbered}
-
-We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status.
-From there, we assess the performance of models fit using different kinds of broadly available data at the global scale.
-We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
-We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al.
-2019 [@rousseau2019] to each stock.
-As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available.
-We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus.
-Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
-
-While CMSY systemically overestimated fishing mortality rates and underestimated stock status for these case study RAM stocks, use of the fishery-specific SAR, FMI, and regional effort data, produced substantially more accurate results for both *B/B~MSY~* and *F/F~MSY~* (Fig.\@ref(fig:cs-plot)).
-This is not in and of itself a surprising result, but demonstrates how 1) estimates of stock status based on catch histories alone can be both imprecise and biased and how 2) augmenting catch histories with additional data from the fisheries in question can produce meaningful improvements.
-
-<!-- `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model.
-We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~.
-Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.
--->
-
-<!-- Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
-CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
-Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
-Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
-This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values (Fig.\@ref(fig:cs-plot)).
--->
-
-```{r cs-plot, fig.cap="RAM values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
-Each point is a stock.
-Black dashed line shows the 1:1 relationship."}
-ex_scatter_plot
-```
-
-<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
-In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
-Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
--->
-
-## Global Value of Information {.unnumbered}
-
-<!-- We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
-B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
-We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
-This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
--->
-
-<!-- We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM.
-We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al.
-2017 [@froese2017] (See Table.\@ref(tab:dat-desc)).
-We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
-Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
--->
-
-We first consider the performance of sraplus when the model is given a perfect index of abundance from RAM for each stock in RAM it is tasked with assessing.
-Assessment models in RAM are typically much more structurally complex than the simple biomass dynamics used in sraplus, and so this "perfect information" test tells us how much of a penalty in model performance we are likely to pay due to model misspecification alone.
-The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RAM were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "RAM Index"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "RAM Index"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "RAM Index"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
-
-Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves.
-We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status beyond those derived from catch-only methods.
-Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)).
-However, this is somewhat an artifact of the data.
-The status of most stocks in RAM is relatively good, with recent *B/B~MSY~* values generally near one.
-This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise.
-Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than the RAM abundance indices are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
-The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RAM Index","Guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "Guess"])`.
-
-Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process.
-Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions.
-For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
-We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
-Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.
-
-```{r perf-tab}
-
-knitr::kable(perf, digits = 2,
-             caption = "Global performance statistics in the most recent year available of models using different sources of data.
-mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
-Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.
-See Table.5 for details of data types.",
-              format = "pipe")
-
-```
-
-<!-- xx dig into this, which stocks are in that region? XX -->
-
-```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
-ram_mpe_map_plot
-```
-
-```{r mape-map, fig.cap = "Median absolute percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
-ram_mape_map_plot
-```
-
-```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
-
-ram_acc_map_plot 
-```
-
-\newpage
-
-The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
-What sources of data might provide the greatest value in improving our estimates of global stock status?
-We used sraplus together with the RAM database to estimate the value of information, measured as the average reduction in root mean squared error, resulting from use of different kinds of data (Fig.\@ref(fig:voi-plot)).
-Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available.
-Interestingly though, having access to only an accurate estimate of F/F~MSY~ in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself.
-While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average, due to the lack of historical context for recent trends in abundance.
-We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].
-
-```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
-Distribution is full posterior probability distribution.
-Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
-Change is relative to the mean performance of a catch-only heuristic model."}
-b_voi_plot + 
-  geom_vline(aes(xintercept = 0), alpha = 0.75)
-```
-
-#  Discussion {.unnumbered}
-
-<!-- XX somewhere in here make clear that you're not learning anything from the catch data XX -->
-
-<!-- XX Acknolwedge the need for action the the frustration of being told "get more data" XX -->
-
-<!-- Global-level assessment are critical for guiding management agendas for the world's oceans.
-Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world.
-Numerous efforts have sought to resolve this problem through development of new modeling methods.
-These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment.
-However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes.
-While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.
--->
-
-Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals.
-Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fish stocks remains uncertain in many parts of the world.
-The hope of efforts using catch-only models to estimate the status of global unassessed fisheries is fundamentally that we can learn something meaningful about the state of a fished population simply by knowing something about its catch history and life history.
-Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RAM.
-While in some cases addition of globally available data such a FMI scores, SAR values, or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
-
-What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends on the needs of those using the assessment outputs.
-It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
-In some instances using the data presented here did provide some improvement over use of catch-only style assessment methods; the difficulty comes in attempting to apply data types uniformly across the globe.
-While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RAM, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
-That use of none of the datasets collected here can achieve that, and that our test on the RAM data suggests that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the quality and use of data themselves.
-
-We chose to test the performance of methods against estimated values reported in RAM.
-A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RAM.
-The clear and systemically poor performance of the methods tested here when applied to RAM stocks would require though that these methods have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks if they are to be expected to provide reasonable picture of global fish populations.
-Free et al.
-2020's [@free2020] simulation testing of the types of methods included here suggests that this is unlikely to be true.
-
-<!-- Our estimates of global stock status vary wildly depending on which source of data we use.
-Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information?
-Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)).
-Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used.
-These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability.
-The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations.
-Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks.
-None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RAM (XX RAM matches?
-need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).
--->
-
-Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions.
-The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e.
-those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)).
-Effort data such as those reconstructed by Rousseau et al.
-2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
-Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
-
-But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
-Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas of the globe results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
-
-Our value-of-information analysis also shows though the high utility of having access to even a recent snapshot of F/F~MSY~ (Fig.\@ref(fig:voi-plot)).
-Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct fishery-specific priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
-Another avenue would be to prioritize the development of a global repository for length and age composition data.
-Given appropriate conditions, these length measurements can be used to estimate local fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
-While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
-Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f; @rudd2017].
-
-<!-- Using catch data alone does not reveal about recent stock status beyond what is encoded in our priors and functional forms.
-So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest).
-Given the demonstrated value of F/F~MSY~ values in improving estimates of stock status, and the increased feasibility in estimating F/F~MSY~ at a local scale relative to B/B~MSY~, how might we go about obtaining estimates of fishing mortality rates at a global scale?
-We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/F~MSY~ for individual fisheries (see Supplemental Information).
-While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
--->
-
-We must also prioritize collection and curation of fish population survey data worldwide.
-Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//).
-Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments.
-Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) in combination with the reconstruction approaches of [@rousseau2019] might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes
-
-Expanded training of fisheries scientists around the globe is another critical need.
-Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise.
-This is why stock assessments even in data-rich fisheries are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
-Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
-
-The coming decades are a critical time for the future of fisheries and ocean health.
-Achieving the United Nations Sustainable Development Goal 14 for the conservation and sustainable use of the world's oceans depends on our ability to effectively assess the status of fish stocks around the world.
-The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
-However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed.
-Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a solution to this problem to has proven to be unbiased and sufficiently precise at a global or regional level.
-
-<!-- Our results show that improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists. -->
-
-The lack of strong information on stock status within catch histories alone demonstrated here means that differences in models and assumptions between catch-based assessment efforts can produce starkly contrasting conclusions on global stock status, leading to debates that are inconclusive as they are inherently driven by assumptions.
-The FAO is leading efforts to increase technical capacity and monitoring and evaluation infrastructure to improve fisheries management in places with limited data.
-Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets.
-Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally [@costello2020].
-Catch-only methods may be a useful starting point when applied properly, but achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.
-
-<!-- The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries. -->
-
-# Materials and Methods {.unnumbered}
-
 <!-- We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database.
 -->
-All analysis were conducted in the R programming language [@rcoreteam2019].
-Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018].
-The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
-Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.
 
@@
 Where $B_t$ is biomass at time *t*, *K* is carrying capacity ,*r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of *B~MSY~/K* to shift.
-When *m* is two *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right.
+When *m* = 2 *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right.
 The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, Thorson et al.
@@
 
-We allow for process error *p* (in the manner of the stochastic stock reduction analysis (SRA) suggested by Walters *et al.* 2006 [@walters2006]).
-Process error *p* is assumed to be log-normally distributed, such that
+We allow for process error *p_t* (in the manner of the stochastic stock reduction analysis (SRA) suggested by Walters *et al.* 2006 [@walters2006]).
+Process error *p_t* is assumed to be log-normally distributed, such that
 
@@
 sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo with the No-U-Turn sampler [@hoffman2011].
-Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
+Unless there is an abundance index to fit to, the model runs as a stock reduction analysis.
 A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
@@
 The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
-Any model results that are not based on an index of abundance are estimated through stock reduction analysis.
-All stock reduction analysis style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
+All stock reduction analysis runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
 For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*.
@@
 
-When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018].
+When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in Stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018].
 By default the model uses 2000 draws with a 1000 step warm-up and one chain.
-Any detailed fit for a particular fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria.
+Any detailed fit for an individual fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria.
 The model fits to a direct estimate of abundance (e.g.
@@
 
-$$log(a_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}) \times q, sigma_{obs})$$
+$$log(a_t) \sim normal(F(r,K,m,B0,\pmb{p},\pmb{c}) \times q, \sigma_{obs})$$
 
-where *f* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
+where *a_t* is the observed abundance index and *F* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
 When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data.
 [@rousseau2019] measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep).
-This rate of technology creep assumes that every unit increase in effort is log-linearly greater than the unit of effort before it.
+Treating this raw effort data as the denominator in the CPUE calculation assumes that every increase in fishing effort translates to a commensurate increase in fishing mortality.
 When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
-This is due to the fact that the marginal fishing mortality produced by increasing unit of efforts increases decreases as effort approaches infinity (since fishing mortality is bounded between 0 and 1).
+This is due to the fact that the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the fishing mortality rate is bounded between 0% and 100%).
 To accommodate this, we generate a catch per effective harvest rate index of abundance, as
@@
 The only modification made is to convert the underlying population model to C++ for faster computation.
-For each stock we used all the default options and priors provided and generated by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012].
+For each stock we used all the default options and priors provided and generated by CMSY, in the same manner as @palomares2020, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012].
 Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
@@
 
-In the absence of any data to fit to, the a stock reduction algorithm (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure.
+In the absence of any data to fit to, the stock reduction algorithm (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure.
 This creates a problem for the Bayesian nature of our analysis.
@@
 Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
-Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).
+Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See @poole2000 and references therein for a discussion of Borel's Paradox in a fisheries context).
 
-<!-- This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
--->
-
-<!-- Borel's Paradox poses a particular problem  for the SRA version of sraplus.
-Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020).
-This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm.
--->
+This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
+Borel's Paradox poses a particular problem  for the SRA version of sraplus.
+Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020).
+This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the posterior distributions of stock status resulting from the SRA algorithm.
 
@@
 We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
-We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
+We then trained a regression of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
 Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries.
 For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on *B/B~MSY~* or *F/F~MSY~* in the most recent year of the fishery usable by sraplus (See Supplementary Information).
-All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
+All prior regression models were tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
 The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.
@@
 
+## Assessing Performance {.unnumbered}
+
+Simulation testing can be preferable in many ways to comparison against model outputs.
+However it is not possible to simulation test the value of information derived from empirical relationships between variables such as fisheries management strength and fishery outcomes, which our study depends on.
+As such we assess model performance through comparison to best available estimates of stock status available in RAM.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history.
+*B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
+We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
+Our methods approximated a regional-level assessment exercise, where data beyond catch histories are available at regional levels, but not for specific fisheries.
+
+We assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
+B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
+We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
+This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
+We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM itself "RAM Index", as a measure of the ability of models like sraplus if given perfect information
+
+We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with CMSY, and a set based on the RAM Index (Table.\@ref(tab:dat-desc)).
+We assessed model performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
+Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
+This experiment structure allows us to assess the ability of catch-only models to estimate the *B/B~MSY~* values reported in RAM, and evaluate how much these estimates can be improved by augmenting the catch histories with additional sources of information.
+Our benchmark model if a simple "Guess".
+Guess assigns each stock a random *B/B~MSY~* of 1.6, 1, or 0.4, corresponding to the middle value of the FAO status bins of underfished  ($ B/B~MSY~ \geq 1.2$), maximally sustainably fished ($0.8 \geq B/B~MSY~ < 1.2$), overfished  ($0 \geq B/B~MSY~ < 0.8$).
+
+```{r dat-desc}
+dat_description <- tribble(~`Data Name`, ~Description,
+        "RAM Index", "Fit to abundance index from RAM",
+        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
+        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
+        "Efective CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 2.6% technology creep",
+        "Effective CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
+        "Nominal CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
+         "Nominal CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
+        "Guess", "Priors on terminal B/Bmsy randomly sampled from .4,1,1.6")
+
+knitr::kable(dat_description,
+             caption = "Data sources used for terminal stock status estimate",
+             format = "pipe") %>%
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+```
+
 ## Value of Information Calculations {.unnumbered}
+
+We performed a value-of-information assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
+The value-of-information analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RAM, and comparing the estimated values to the values reported in RAM.
+There are too many combinations for us to run the full expansion of possible parameter states.
+To resolve this we generated fits for 3000 combinations of a RAM stock and available data.
+For any one draw, we randomly sampled a RAM stock and a list of available data and data quality.
+For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
+The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
+We then measured the value of information as the average reduction in root mean squared error (RMSE) in B/B~MSY~, resulting from use of different kinds of data.
+
+## Case Study {.unnumbered}
+
+We also ran a case study demonstrating how different kinds of data led to different conclusions about stock status.
+We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
+We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock.
+As a benchmark, we first estimated stock status for these case study fisheries using the default settings of the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available.
+We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus.
+Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
+This case study represents a more "localized" assessment, where external data sources (FMI and SAR) are available at the stock level, rather than at the regional level.
 
@@
 
-We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The VOI analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RAM, and comparing the estimated values to the values reported in RAM.
-We generate fits for 3000 combinations of a RAM stock and available data.
-For any one draw, we randomly sample a RAM stock and a list of available data and data quality.
-For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
-The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
+# Results {.unnumbered}
 
-# Acknowledgements and Funding {.unnumbered}
+## Case Study {.unnumbered}
+
+Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
+CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing ($R^2$ for B/Bmsy = 0.1, $R^2$ for F/F~MSY~ = 0).
+Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces some improvements, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
+Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
+This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values.
+However none of the models fit to the three case-study data combinations were able to explain much of the variation in B/B~MSY~, though the models performed better in some cases for predicting F/F~MSY~ (though this is to be expected as the prior-generating models were trained on data from RAM).
+Models informed by the SAR and FMI values performed the best in terms of B/B~MSY~ ($R^2$ = 0.1).
+The ranking of performance by data sources used was the same for F/F~MSY~, though the range of values was much higher, with the SAR and FMI based result producing an $R^2$ of 0.47 while the default CMSY settings produced an $R^2$ of effectively 0 (Fig.\@ref(fig:cs-plot)).
+
+<!-- This is not in and of itself a surprising result, but demonstrates how 1) estimates of stock status based on catch histories alone can be both imprecise and biased and how 2) augmenting catch histories with additional data from the fisheries in question can produce meaningful improvements.
+-->
+
+<!-- `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model.
+We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~.
+Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.
+-->
+
+```{r cs-plot, fig.cap="RAM values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
+Each point is a stock.
+Black dashed line shows the 1:1 relationship."}
+ex_scatter_plot
+```
+
+<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
+In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
+Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
+-->
+
+## Performance of Regional Fishery Assessments {.unnumbered}
+
+We first considered the performance of sraplus when the model was given a perfect index of abundance from RAM (RAM Index) for each stock in RAM it was tasked with assessing.
+Assessment models in RAM are typically much more structurally complex than the simple biomass dynamics used in sraplus, and so this "perfect information" test tells us how much of a penalty in model performance we are likely to pay due to model misspecification alone.
+The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RAM were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "RAM Index"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "RAM Index"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "RAM Index"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
+
+<!-- Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves.  -->
+
+ <!-- However, this is somewhat an artifact of the data.
+The status of most stocks in RAM is relatively good, with recent *B/B~MSY~* values generally near one.
+This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise.
+-->
+
+We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status beyond those derived from catch-only methods.
+Many of the datasets used produced similar levels of bias as the RAM data (e.g.
+FMI, SAR, Nominal CPUE fits), though notable the "Guess" method actually performed near the top in terms of bias.
+However, this is somewhat an artifact of the data.
+The status of most stocks in RAM is relatively good, with recent *B/B~MSY~* values generally near one.
+As the mean value of the "Guess" method is one, on average the "Guess" model is an unbiased but imprecise measure of stock status in RAM (Table.\@ref(tab:perf-tab)).
+
+Focusing on MAPE (our measure of error rather than bias) and classification accuracy, the error of the models jumps dramatically as soon as data other than the RAM abundance indices are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
+The mean accuracy of the sraplus models across all non-RAM data fits was `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RAM Index","Guess")]))`.
+Note that there are only three bins in the FAO stock status classifications, and a so our "Guess" model has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "Guess"])`.
+This means that the accuracy of our models designed as a proxy for a global assessment process were across all non-RAM index data fits `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RAM Index","Guess")]) / perf$accuracy[perf$"Data Used" == "Guess"] - 1)` more accurate than a random guess, certainly an improvement, but on average assigned fisheries to the wrong FAO status bin `r percent(1 - mean(perf$accuracy[!perf$"Data Used" %in% c("RAM Index","Guess")]))`of the time.
+
+Looking geographically we found a similar pattern of a rapid decrease in performance for models besides those fit to the RAM Index.
+Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions.
+Models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
+We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
+Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistent manner.
+
+```{r perf-tab}
+
+knitr::kable(perf, digits = 2,
+             caption = "Global performance statistics in the most recent year available of models using different sources of data.
+mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
+Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.",
+              format = "pipe")
+
+```
+
+<!-- xx dig into this, which stocks are in that region? XX -->
+
+```{r mpe-map, fig.width=8, fig.height=6, fig.cap="Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+RAM Index refers to catch and abundance index drawn from RAM.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6.
+Panels ordered in asscending (starting from top left) mean MPE at the FAO region level.
+"}
+ram_mpe_map_plot
+```
+
+```{r mape-map, fig.cap = "Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+RAM Index refers to catch and abundance index drawn from RAM.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6.
+Panels ordered in descending (starting from top left) mean MAPE at the FAO region level",fig.width=8, fig.height=6}
+ram_mape_map_plot
+```
+
+```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
+RAM Index refers to catch and abundance index drawn from RAM.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6.
+Panels ordered in descending (starting from top left) mean accuracy at the FAO region level",fig.width=8, fig.height=6}
+
+ram_acc_map_plot 
+```
+
+\newpage
+
+<!-- The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
+-->
+
+## Value of Information {.unnumbered}
+
+What sources of data might provide the greatest value in improving our estimates of global stock status?
+Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available.
+Interestingly though, having access to only an accurate estimate of F/F~MSY~ in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself  (Fig.\@ref(fig:voi-plot)).
+While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average, due to the lack of historical context for recent trends in abundance.
+
+<!-- We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].
+-->
+
+```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
+Distribution is full posterior probability distribution.
+Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
+Change is relative to the mean performance of a catch-only heuristic model."}
+b_voi_plot + 
+  geom_vline(aes(xintercept = 0), alpha = 0.75)
+```
+
+#  Discussion {.unnumbered}
+
+<!-- XX somewhere in here make clear that you're not learning anything from the catch data XX -->
+
+<!-- XX Acknolwedge the need for action the the frustration of being told "get more data" XX -->
+
+<!-- Global-level assessment are critical for guiding management agendas for the world's oceans.
+Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world.
+Numerous efforts have sought to resolve this problem through development of new modeling methods.
+These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment.
+However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes.
+While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.
+-->
+
+<!-- Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RAM -->
+
+Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals.
+The hope of efforts using catch-only models to estimate the status of global unassessed fisheries is fundamentally that we can learn something meaningful about the state of a fished population simply by knowing something about its catch history and life history.
+While in some cases the addition of globally available data such a FMI scores, SAR values, or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
+
+What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends on the needs of those using the assessment outputs.
+It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
+In some instances using the data presented here did provide some improvement over use of catch-only style assessment methods; the difficulty comes in attempting to apply data types uniformly across the globe.
+While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RAM, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
+The overall poor performance of the models tested here shows that improvements in estimates of global stock status depend on improvements in the quality and use of data themselves
+
+<!-- That use of none of the datasets collected here can achieve that, and that our tests fitting directly to abundance indices from RAM itself suggests that model misspecification is not the primary cause of poor performance, tells us that improvements in estimates of global stock status will likely depend on improvements in the quality and use of data themselves.
+-->
+
+Our claim is that achieving substantial progress in assessing the state of global fisheries will require improved data and capacity building.
+This claim cannot be definitively proven; It is possible that some yet-to-be discovered model will provide a reliable means of dramatically improving the accuracy of estimates produced by models using current broadly-available datasets, such as catch-only style methods.
+Should that occur we will be happy to have been proven wrong.
+However, efforts from numerous groups made up of highly skilled scientists tasked with solving this problem have all come up short, particularly when external efforts tried validate their performance [@free2020;@bouch2021;@pons2020].
+This is not intended as an indictment of those past efforts, but a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone.
+Methods tested in those papers have been demonstrated to perform relatively well when given data or priors informative to specific fisheries.
+We revisited this task in this paper, and found that even when augmented with other broadly available datasets, we were unable to produce consistently accurate estimates of stock status around the globe.
+For these reasons we believe that improvements to our understanding of the state of global unassessed fisheries will come from targeted use of existing data, and expanding the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
+
+We chose to test the performance of methods against estimated values reported in RAM.
+A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RAM, in addition to the errors in the RAM estimates themselves.
+For this reality to change our results though would require that the methods tested here have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks.
+@free2020's simulation based based results of these same types of models suggests that this is unlikely to be true.
+
+<!-- Our estimates of global stock status vary wildly depending on which source of data we use.
+Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information?
+Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)).
+Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used.
+These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability.
+The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations.
+Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks.
+None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RAM (XX RAM matches?
+need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).
+-->
+
+Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions.
+The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e.
+those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)).
+Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
+Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
+Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas of the globe results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
+
+Our value-of-information analysis also shows the high utility of having access to even a recent snapshot of F/F~MSY~ (Fig.\@ref(fig:voi-plot)).
+Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct fishery-specific priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
+Another avenue would be to prioritize the development of a global repository for length and age composition data.
+Given appropriate conditions, these length measurements can be used to estimate local fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
+While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
+Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f; @rudd2017].
+
+<!-- Using catch data alone does not reveal about recent stock status beyond what is encoded in our priors and functional forms.
+So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest).
+Given the demonstrated value of F/F~MSY~ values in improving estimates of stock status, and the increased feasibility in estimating F/F~MSY~ at a local scale relative to B/B~MSY~, how might we go about obtaining estimates of fishing mortality rates at a global scale?
+We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/F~MSY~ for individual fisheries (see Supplemental Information).
+While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
+-->
+
+We must also prioritize collection and curation of fish population survey data worldwide.
+Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://www.FishStats.org)( www.FishStats.org).
+Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments.
+Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/)((https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
+However, out value of information exercise indicates that we may have to wait many years for new surveys to provide substantial improvements in status estimates (Fig.\@ref(fig:voi-plot)).
+
+Expanded training of fisheries scientists around the globe is another critical need.
+Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise.
+This is why stock assessments even in data-rich fisheries are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
+Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
+
+The coming decades are a critical time for the future of fisheries and ocean health.
+Achieving the United Nations Sustainable Development Goal 14 for the conservation and sustainable use of the world's oceans depends on our ability to effectively assess the status of fish stocks around the world.
+The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
+However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed.
+Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a solution to this problem which has proven to be unbiased and sufficiently precise at a global or regional level.
+
+<!-- Our results show that improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists. -->
+
+The lack of strong information on stock status within catch histories alone means that differences in models and assumptions between catch-based assessment efforts can produce starkly contrasting conclusions on global stock status, leading to debates that are inconclusive as they are inherently driven by assumptions.
+The FAO is leading efforts to increase technical capacity and monitoring and evaluation infrastructure to improve fisheries management in places with limited data.
+Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets.
+Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally [@costello2020].
+Catch-only methods may be a useful starting point when applied properly, but the source of their information must be clearly communicated, and achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.
+
+<!-- The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries. -->
+
+# Acknowledgements {.unnumbered}
 
 Funding for this work was provided by the Food and Agriculture Organization of the United Nations (FAO).
-We thank participants of the FAO "Methods for Global Assessment" workshop held in Rome, Italy February 2019 for helpful feedback on this project.The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of NOAA or the Department of Commerce.
+We thank participants of the FAO "Methods for Global Assessment" workshop held in Rome, Italy February 2019 for helpful feedback on this project.
+The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of NOAA or the Department of Commerce.
+
+# References {.unnumbered}
 

### 2021-04-18 make-assessing-global-fisheries.R

@@ -4,6 +4,7 @@
 
 library(tidyverse)
 library(ggridges)
+library(ggtext)
 library(gganimate)
 library(rstan)
 library(mvtnorm)
@@ -893,16 +894,23 @@ tmp2 <- tmp2 %>%
     )
   )
 
+ex_performance <- tmp2 %>% 
+  group_by(variable, data) %>% 
+  summarise(r2 = round(yardstick::rsq_vec(truth, value),2)) 
+
+
 ex_scatter_plot <- tmp2 %>% 
   ggplot(aes(truth, value)) + 
   geom_vline(aes(xintercept = 0)) + 
   geom_hline(aes(yintercept = 0)) +
   geom_abline(aes(slope = 1, intercept = 0),linetype = 2) +
-  geom_point(alpha = 0.75) + 
+  geom_point(alpha = 0.5, size = 2) + 
+  ggtext::geom_richtext(data = ex_performance, aes(x = 1, y = 4, label = paste0("R<sup>2</sup> = ",r2))) +
   facet_grid(variable ~ data, scales = "free_x")  + 
   # scale_size(trans = "sqrt", name = "Lifetime Catch") +
   scale_x_continuous(name = "RAM Value", expand = expansion(add = c(0, .1))) + 
   scale_y_continuous("Estimated Value", expand = expansion(add = c(0, .1)))
+
 
 sraplus_v_truth <- tmp %>% 
   left_join(truth, by = c("stockid","year")) %>% 
@@ -2042,11 +2050,16 @@ ram_labeller <- c(
   "u_umsy" = "RAM U/Umsy"
 )
 
+
+collabs <-
+  c(paste0(seq(-100, 75, by = 25),"%"),
+    expression("" >= "100%"))
+
 ram_mpe_map_plot <- fao_area_ram_status %>%
   filter(!is.na(data)) %>% 
   mutate(data = fct_reorder(data, abs(mpe), .fun = median)) %>% 
   ggplot() +
-  geom_sf(aes(fill = mpe), size = .01) +
+  geom_sf(aes(fill = pmin(1,mpe)), size = .01) +
   geom_sf(
     data = world_map,
     fill = "darkgrey",
@@ -2055,14 +2068,13 @@ ram_mpe_map_plot <- fao_area_ram_status %>%
   ) +
   facet_wrap(~ data, labeller = labeller(data = ram_labeller)) +
   scale_fill_gradient2(
-    low = "steelblue",
+    low = "darkblue",
[…]

### 2021-04-24 documents/ovando-etal-assessing-global-fisheries-si.Rmd
submited version

@@
 
-We use an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000].
-Our solution amounts to a two-step SIR algorithm.
-We first run the standard SIR algorithm as described above.
-We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
-
-$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
-
-We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*
-
-$$p(n_i) = \frac{p(bin_i)}{N_i}$$
-
-And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
-The net result of this is a post-model-pre-data distribution of parameters *r* and *K* that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
+The net result of our correction is a post-model-pre-data distribution of life history parameters that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status.
 In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status".
@@
 ```
-
-## Sample Prior Posterior Plots {.unnumbered}
 

### 2021-04-24 documents/ovando-etal-assessing-global-fisheries.Rmd
submited version

@@
 ---
-title: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
-author: 
-  - Daniel Ovando
-  - Ray Hilborn
-  - Cole Monnahan
-  - Merrill Rudd
-  - Rishi Sharma
-  - James Thorson
-  - Yannick Rousseau
-  - Yimin Ye
+title: Improving Estimates of the State of Global Fisheries <br> Depends on Better Data
 date: "`r Sys.Date()`"
@@
   min_years_catch: [25]
-abstract: |
-  Assessments of the global state of fish populations play a critical role in tracking the implementation of the United Nations Sustainable Development Goals.
-While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain.
-Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
-Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added.
-Estimates of fish populations based primarily on catch histories alone on average performed 29% better than a random guess.
-But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time.
-Obtaining accurate estimates of stock status for the world's exploited fish populations depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
 always_allow_html: true
@@
   - \usepackage{lineno}\linenumbers
+toc: false
 ---
@@
   mutate(data = forcats::fct_recode(data, 
-                                    "RAM Index" = "ram-data",
+                                    "RLSADB Index" = "ram-data",
                                     "SAR" = "sar",
@@
                                     "Guess" = "guess",
-                                    "RAM F/Fmsy" = "u_umsy"
+                                    "RLSADB F/Fmsy" = "u_umsy"
 
@@
 
+# Title Page {.unnumbered}
+
+**Title 1**: Improving Estimates of the State of Global Fisheries Depends on Better Data
+
+**Title 2**: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
+
+**Running Title**: Unassessed Fisheries
+
+**Authors**:
+Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^
+
+**Affiliations**:
+
+^1^University of Washington, School of Aquatic and Fishery Sciences
+1122 NE Boat St, Box 355020 Seattle, WA 98195-5020
+Seattle, WA, USA 98103
+
+^2^NOAA Fisheries Alaska Fisheries Science Center, Seattle, WA, USA
+
+^3^Scaleability LLC
+
+^4^Food and Agriculture Organization of the United Nations, Fisheries and Aquaculture,Rome, IT
+
+^5^University of Tasmania, Institute for Marine and Antarctic Studies,Hobart, TAS, AUS
+
+**Authorship**: DO, RH, CM,MR, RS, JT designed model structure and ran analyses. RS, YR, and YE supplied data. All authors contributed to writing of the manuscript
+
+**Conflict of Interest**: RH receives research funding from many groups that have interests in fisheries outcomes including environmental NGOs, foundations, governments and fishing industry groups.
+
 \newpage
+
+# Abstract {.unnumbered}
+
+Implementation of the United Nations Sustainable Development Goals requires assessments of the global state of fish populations.
+While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain.
+Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
+Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added.
+Estimates of fish populations based primarily on catch histories alone on average performed 29% better than a random guess.
+But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time.
+Effective improvement in estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.
+
 # Introduction {.unnumbered}
@@
 The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished (at or near targets for sustainable food production), 6.2% are underfished, and 34.2% are overfished [@fao2020].
-Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012].
-The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the state of assessed fish populations around the globe.
+Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments summarized in the RAM Legacy Stock Assessment Database (RLSADB) [@ricard2012].
+The "assessed" fisheries in RLSADB represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the state of assessed fish populations around the globe.
 
@@
 While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods.
-With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
+With the surge in data availability, such as global assessments of management strength [@melnychuk2017], trawl footprints [@amoroso2018], and fishing effort [@rousseau2019b], the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
 In addition, the SOFIA process assigns unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to to the biomass that would on average result in maximized yield over the long-term.
@@
 Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (@free2020 and references therein).
-These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being our measure of stock status), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries (e.g.
-@cabral2020 and @gaines2018).
+These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being a common measure of stock status), from characteristics of a fishery's catch history.
 
 <!-- These issues become apparent when we consider some of the macro-level predictions made by these models.
-While the estimates of *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model.
+While the estimates of *B/B~MSY~* and other reference points reported in RLSADB are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RLSADB to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model.
 -->
 
-<!-- @costello2016a found similar rankings of regions in terms of stock status as RAM, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison.
+<!-- @costello2016a found similar rankings of regions in terms of stock status as RLSADB, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison.
 The superensemble method [@anderson2017a] used in @rosenberg2018 demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific.
 Simply comparing catch divided by maximum catch, in the manner of @pauly2007, finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
-The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
-However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly exploited regions as measured by RAM and the FAO resulting from these catch-only efforts at assessing global fisheries is concerning.
+The non-RLSADB results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
+However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly exploited regions as measured by RLSADB and the FAO resulting from these catch-only efforts at assessing global fisheries is concerning.
 -->
 
-```{r status-tab, include=FALSE}
+```{r status-tab, include=FALSE, eval=FALSE}
 
@@
   filter(year == max(year)) %>% 
-  summarise(`B/BMSY RAM` = round(mean(dlm_geomean),2)) %>% 
+  summarise(`B/BMSY RLSADB` = round(mean(dlm_geomean),2)) %>% 
   ungroup()
@@
 tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
-  select(-`B/BMSY RAM`) %>%  
+  select(-`B/BMSY RLSADB`) %>%  
   left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
@@
     rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
-  select(`FAO Area`,`SOFIA % Overfished`,`B/BMSY RAM`, everything())
+  select(`FAO Area`,`SOFIA % Overfished`,`B/BMSY RLSADB`, everything())
 
@@
 However, @free2020 demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*.
-While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating stock status of the individual unassessed fisheries around the world.
+While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating the state of the individual unassessed fisheries around the world.
 The types of global assessment efforts share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure.
@@
 We then used our new assessment method, sraplus, to evaluate the performance status estimates derived from different combinations of these broadly available data.
-Wecompared these results with estimates from the increasingly utilized (e.g.
+We compared these results with estimates from the increasingly utilized (e.g.
 @palomares2020) default settings of the catch-only version of CMSY described in @froese2017.
@@
 
-The key attribute of sraplus is not substantial improvements in model fitting methods per say, but the ability to easily incorporate multiple kinds of fishery data potentially used in SRA style analyses in a statistically rigorous manner.
-We paired these data with fisheries in RAM, and then used sraplus to generate predictions of B/B~MSY~ in the most recent year of each fishery across different combinations of the data types.
-We then compared the predictions generated by sraplus using different types of data, along with predictions for the same fisheries generated by CMSY, to the reported values in RAM.
+The basis of this analysis is a new stock assessment software package we call [sraplus](https://github.com/DanOvando/sraplus) (https://github.com/DanOvando/sraplus).
+sraplus is an extension of stochastic stock reduction analysis (SRA) [@kimura1984;@walters2006], which allows users to combine a biomass dynamics model with a variety of data sources (e.g.
+priors on recent stock status, or an index of abundance) in order to produce estimates of the state of a fishery over time.
+The key goal of sraplus is not substantial improvements in model fitting methods per say, but providing the ability to easily incorporate multiple kinds of fishery data potentially used in SRA style analyses in a statistically rigorous manner.
+We paired various types of data with fisheries in RLSADB, and then used sraplus to generate predictions of B/B~MSY~ in the most recent year of each fishery across different combinations of data types.
+We then compared the predictions generated by sraplus using different types of data, along with predictions for the same fisheries generated by CMSY, to the reported values in RLSADB.
 
@@
 At its most "data-limited", sraplus can work as a catch-only model in the manner of @froese2017.
-These catch histories can then be augmented with prior information on stock status derived from increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases.
+These catch histories can then be augmented with prior information on stock status, derived from expert opinion, or using built-in prior generating models based on increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases.
 The FMI database provides self-reported scores reflecting the strength of fisheries management in many fisheries around the world, which sraplus uses to construct informative priors on stock status, e.g.
@@
 In this way sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).
-
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Data Source                                 | Short Name       | Data Use                                                                 | Caveats                                                                                                         |
-+=============================================+==================+==========================================================================+=================================================================================================================+
-| Catch data [@fao2020]                       | catches          | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Fisheries Management Index [@melnychuk2017] | FMI              | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Swept Area Ratio [@amoroso2018]             | SAR              | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Reconstructed effort data [@rousseau2019b]  | effort           | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors.
-Assumed rate of technology creep reported in individual sections |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-
-: (\#tab:data) Data sources included across model fits.
 
@@
 While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the data types evaluated here will support.
-The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status
+
+<!-- The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status -->
 
@@
 When *m* = 2 *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right.
-The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, Thorson et al.
-2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa.
-We fix the shape parameter based on the distributions reported in Thorson et al.
-2012 [@thorson2012] for the genus of the species in question.
+The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, @thorson2012 provides estimates of the ratio of *B~MSY~* to *K* for many fish taxa.
+We fix the shape parameter based on the distributions reported in @thorson2012 for the genus of the species in question.
 
@@
 Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
-We deal with this problem by following the methods described in [@winker2018].
-to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
+We deal with this problem by following the methods described in [@winker2018] to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
 
-We allow for process error *p_t* (in the manner of the stochastic stock reduction analysis (SRA) suggested by Walters *et al.* 2006 [@walters2006]).
-Process error *p_t* is assumed to be log-normally distributed, such that
+We allow for process error $p_t$ (in the manner of the stochastic stock reduction analysis (SRA) suggested by @walters2006). Process error $p_t$ is assumed to be log-normally distributed, such that
 
@@
 sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo with the No-U-Turn sampler [@hoffman2011].
-Unless there is an abundance index to fit to, the model runs as a stock reduction analysis.
+Unless there is an abundance index to fit to, the model runs as an SRA.
 A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
@@
 
-where *a_t* is the observed abundance index and *F* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
+where $a_t$ is the observed abundance index and *F* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
 When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data.
@@
 When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
-This is due to the fact that the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the fishing mortality rate is bounded between 0% and 100%).
+This is due to the fact that in reality the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the realized fishing mortality rate must be between between 0% and 100%).
 To accommodate this, we generate a catch per effective harvest rate index of abundance, as
@@
 
-In the absence of any data to fit to, the stock reduction algorithm (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure.
+In the absence of any data to fit to, sraplus works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history, life history priors, and model structure.
 This creates a problem for the Bayesian nature of our analysis.
@@
 This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-Borel's Paradox poses a particular problem  for the SRA version of sraplus.
-Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020).
+Borel's Paradox poses a particular problem when there are no data to fit to though (i.e.
+when the model is simply filtering through prior distributions in the manner of a traditional SRA).
+Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse.
+In this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020).
 This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the posterior distributions of stock status resulting from the SRA algorithm.
@@
 We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000].
-Our solution amounts to a two-step sample-importance-resampling (SIR) algorithm.
+Our solution amounts to a two-step sampling-importance-resampling (SIR) algorithm.
 We first run the standard SIR algorithm as described above.
-We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.
-The net result of this is that it allows users to place explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior, rather than creating a complicated and biased prior on stock status based on a mixture of explicit and implicit priors.
+We then break the resulting draws into bins based on terminal stock status, and calculate the mean probability density *p* (defined by the prior distributions of estimated parameters) of each bin.
+
+$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
+
+We then divide the mean probability density of bin *i* evenly among each of the draws within that bin *n*
+
+$$p(n_i) = \frac{p(bin_i)}{N_i}$$
+
+And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
+
+The net result of this is that it allows users to place an explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior.
+While the range of possible life history values supplied still influences stock status under this approach, this prior predictive tuning process makes the resulting priors more consistent with explicit priors on recent stock status supplied by the user.
+Users can turn this functionality off and instead base priors on stock status primarily on life history.
 See Supplementary Information for a detailed explanation of this problem and our solution.
@@
 Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus.
-We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database.
+We paired data on catch histories, swept area ratio, and Fisheries Management Index with estimates of stock status from the RLSADB.
 We then trained a regression of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
@@
 The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.
+For all results presented in this paper we used these data to provide priors on *F/F~MSY~*, as we found clearer predictive relationships and subsequent model performance between catch, FMI, and SAR values and *F/F~MSY~* than we did for *B/B~MSY~*.
 
@@
 However it is not possible to simulation test the value of information derived from empirical relationships between variables such as fisheries management strength and fishery outcomes, which our study depends on.
-As such we assess model performance through comparison to best available estimates of stock status available in RAM.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history.
-*B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
-We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
+As such we assess model performance through comparison to best available estimates of stock status available in RLSADB.
+We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RLSADB, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history.
+*B/B~MSY~* values from RLSADB are themselves estimates, not data, but they are the best available information on global stock status.
+We then paired the catch histories for these RLSADB stocks with regional-level SAR, FMI, and effort data.
 Our methods approximated a regional-level assessment exercise, where data beyond catch histories are available at regional levels, but not for specific fisheries.
+We also estimated *B/B~MSY~* values of our candidate RLSADB stocks by using sraplus to fit to an abundance index drawn directly from RLSADB itself "RLSADB Index", as a measure of the ability of models like sraplus if given perfect information
 
-We assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history.
-B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status.
-We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data.
-This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.
-We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM itself "RAM Index", as a measure of the ability of models like sraplus if given perfect information
-
-We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with CMSY, and a set based on the RAM Index (Table.\@ref(tab:dat-desc)).
+We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with CMSY, and a set of runs fit to the RLSADB Index (Table.\@ref(tab:dat-desc)).
 We assessed model performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
 Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
-This experiment structure allows us to assess the ability of catch-only models to estimate the *B/B~MSY~* values reported in RAM, and evaluate how much these estimates can be improved by augmenting the catch histories with additional sources of information.
-Our benchmark model if a simple "Guess".
-Guess assigns each stock a random *B/B~MSY~* of 1.6, 1, or 0.4, corresponding to the middle value of the FAO status bins of underfished  ($ B/B~MSY~ \geq 1.2$), maximally sustainably fished ($0.8 \geq B/B~MSY~ < 1.2$), overfished  ($0 \geq B/B~MSY~ < 0.8$).
-
-```{r dat-desc}
-dat_description <- tribble(~`Data Name`, ~Description,
-        "RAM Index", "Fit to abundance index from RAM",
-        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
-        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
-        "Efective CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 2.6% technology creep",
-        "Effective CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
-        "Nominal CPUE", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
-         "Nominal CPUE+", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
-        "Guess", "Priors on terminal B/Bmsy randomly sampled from .4,1,1.6")
-
-knitr::kable(dat_description,
-             caption = "Data sources used for terminal stock status estimate",
-             format = "pipe") %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
+This experiment structure allows us to assess the ability of catch-only models to estimate the *B/B~MSY~* values reported in RLSADB, and evaluate how much these estimates can be improved by augmenting the catch histories with additional sources of information.
+Our benchmark model is a simple "Guess".
+Guess assigns each stock a random *B/B~MSY~* of 1.6, 1, or 0.4, corresponding to the middle value of the FAO status bins of underfished  ($B/B~MSY~ \geq 1.2$), maximally sustainably fished ($0.8 \geq B/B~MSY~ < 1.2$), overfished  ($0 \geq B/B~MSY~ < 0.8$).
 
@@
 We performed a value-of-information assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The value-of-information analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RAM, and comparing the estimated values to the values reported in RAM.
+The value-of-information analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RLSADB, and comparing the estimated values to the values reported in RLSADB.
 There are too many combinations for us to run the full expansion of possible parameter states.
-To resolve this we generated fits for 3000 combinations of a RAM stock and available data.
-For any one draw, we randomly sampled a RAM stock and a list of available data and data quality.
+To resolve this we generated fits for 3000 combinations of a RLSADB stock and available data.
+For any one draw, we randomly sampled a RLSADB stock and a list of available data and data quality.
 For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
@@
 
-<!-- To determine what sources of data might provide the greatest value in improving our estimates of global stock status, we fit 3,000 sraplus models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
+<!-- To determine what sources of data might provide the greatest value in improving our estimates of global stock status, we fit 3,000 sraplus models to randomly sampled fisheries from RLSADB, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
 We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).
@@
 Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
-CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing ($R^2$ for B/Bmsy = 0.1, $R^2$ for F/F~MSY~ = 0).
-Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces some improvements, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
+CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B~MSY~) and the estimated resilience of the target species (based on Fishlife [@thorson2020]), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing ($R^2$ for B/Bmsy = 0.1, $R^2$ for F/F~MSY~ = 0).
+Adding priors informed by stock-specific swept area ratio and Fisheries Management Index data produces some improvements, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
 Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
 This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values.
-However none of the models fit to the three case-study data combinations were able to explain much of the variation in B/B~MSY~, though the models performed better in some cases for predicting F/F~MSY~ (though this is to be expected as the prior-generating models were trained on data from RAM).
+However none of the models fit to the three case-study data combinations were able to explain much of the variation in B/B~MSY~, though the models performed better in some cases for predicting F/F~MSY~ (though this is to be expected as the prior-generating models were trained on data from RLSADB).
 Models informed by the SAR and FMI values performed the best in terms of B/B~MSY~ ($R^2$ = 0.1).
@@
 
-```{r cs-plot, fig.cap="RAM values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
+```{r cs-plot, fig.cap="RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
 Each point is a stock.
@@
 ex_scatter_plot
+
+ggsave(here("documents","figs","cs-plot.pdf"),ex_scatter_plot)
+
 ```
@@
 In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
-Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
+Since the SAR and FMI regressions were trained on the RLSADB stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
 -->
@@
 
-We first considered the performance of sraplus when the model was given a perfect index of abundance from RAM (RAM Index) for each stock in RAM it was tasked with assessing.
-Assessment models in RAM are typically much more structurally complex than the simple biomass dynamics used in sraplus, and so this "perfect information" test tells us how much of a penalty in model performance we are likely to pay due to model misspecification alone.
-The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RAM were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "RAM Index"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "RAM Index"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "RAM Index"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+We first considered the performance of sraplus when the model was given a perfect index of abundance from RLSADB (RLSADB Index) for each stock in RLSADB it was tasked with assessing.
+Assessment models in RLSADB are typically much more structurally complex than the simple biomass dynamics used in sraplus, and so this "perfect information" test tells us how much of a penalty in model performance we are likely to pay due to model misspecification alone.
+The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RLSADB were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "RLSADB Index"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "RLSADB Index"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "RLSADB Index"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
 This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
@@
  <!-- However, this is somewhat an artifact of the data.
-The status of most stocks in RAM is relatively good, with recent *B/B~MSY~* values generally near one.
+The status of most stocks in RLSADB is relatively good, with recent *B/B~MSY~* values generally near one.
 This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise.
@@
 We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status beyond those derived from catch-only methods.
-Many of the datasets used produced similar levels of bias as the RAM data (e.g.
+Many of the datasets used produced similar levels of bias as the RLSADB data (e.g.
 FMI, SAR, Nominal CPUE fits), though notable the "Guess" method actually performed near the top in terms of bias.
 However, this is somewhat an artifact of the data.
-The status of most stocks in RAM is relatively good, with recent *B/B~MSY~* values generally near one.
-As the mean value of the "Guess" method is one, on average the "Guess" model is an unbiased but imprecise measure of stock status in RAM (Table.\@ref(tab:perf-tab)).
+The status of most stocks in RLSADB is relatively good, with recent *B/B~MSY~* values generally near one.
+As the mean value of the "Guess" method is one, on average the "Guess" model is an unbiased but imprecise measure of stock status in RLSADB (Table.\@ref(tab:perf-tab)).
 
-Focusing on MAPE (our measure of error rather than bias) and classification accuracy, the error of the models jumps dramatically as soon as data other than the RAM abundance indices are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
-The mean accuracy of the sraplus models across all non-RAM data fits was `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RAM Index","Guess")]))`.
+Focusing on MAPE (our measure of error rather than bias) and classification accuracy, the error of the models jumps dramatically as soon as data other than the RLSADB abundance indices are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
+The mean accuracy of the sraplus models across all non-RLSADB data fits was `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]))`.
 Note that there are only three bins in the FAO stock status classifications, and a so our "Guess" model has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "Guess"])`.
-This means that the accuracy of our models designed as a proxy for a global assessment process were across all non-RAM index data fits `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RAM Index","Guess")]) / perf$accuracy[perf$"Data Used" == "Guess"] - 1)` more accurate than a random guess, certainly an improvement, but on average assigned fisheries to the wrong FAO status bin `r percent(1 - mean(perf$accuracy[!perf$"Data Used" %in% c("RAM Index","Guess")]))`of the time.
+This means that the accuracy of our models designed as a proxy for a global assessment process were across all non-RLSADB index data fits `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]) / perf$accuracy[perf$"Data Used" == "Guess"] - 1)` more accurate than a random guess, certainly an improvement, but on average assigned fisheries to the wrong FAO status bin `r percent(1 - mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]))` of the time.
 
-Looking geographically we found a similar pattern of a rapid decrease in performance for models besides those fit to the RAM Index.
+Looking geographically we found a similar pattern of a rapid decrease in performance for models besides those fit to the RLSADB Index.
 Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions.
-Models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map).
+Models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) (Fig.\@ref(fig:mape-map)).
 We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
-Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistent manner.
-
-```{r perf-tab}
-
-knitr::kable(perf, digits = 2,
-             caption = "Global performance statistics in the most recent year available of models using different sources of data.
-mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
-Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.",
-              format = "pipe")
-
-```
+Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RLSADB index data were able to capture the overall state or geographic distribution of stock status represented in RLSADB in a consistent manner.
 
@@
 ```{r mpe-map, fig.width=8, fig.height=6, fig.cap="Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
 Effective CPUE refers to an index of abundance based on reconstructed effort data.
 Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
 FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
@@
 2017 [@froese2017].
-Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-Panels ordered in asscending (starting from top left) mean MPE at the FAO region level.
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
+Panels ordered in ascending (starting from top left) mean MPE at the FAO region level.
 "}
 ram_mpe_map_plot
+
+ggsave(here("documents","figs","mpe-map.pdf"), ram_mpe_map_plot, width = 8, height = 6)
 ```
@@
 ```{r mape-map, fig.cap = "Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
 Effective CPUE refers to an index of abundance based on reconstructed effort data.
 Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
 FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
@@
 2017 [@froese2017].
-Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6.
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
 Panels ordered in descending (starting from top left) mean MAPE at the FAO region level",fig.width=8, fig.height=6}
 ram_mape_map_plot
+
+ggsave(here("documents","figs","mape-map.pdf"),ram_mape_map_plot, width = 8, height = 6)
+
 ```
@@
 ```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
-RAM Index refers to catch and abundance index drawn from RAM.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
 Effective CPUE refers to an index of abundance based on reconstructed effort data.
 Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed.
+For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
 FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
@@
 2017 [@froese2017].
-Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6.
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
 Panels ordered in descending (starting from top left) mean accuracy at the FAO region level",fig.width=8, fig.height=6}
@@
 ram_acc_map_plot
+
+ggsave(here("documents","figs","acc-map.pdf"),ram_acc_map_plot, width = 8, height = 6)
+
 ```
@@
 
-<!-- The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
+<!-- The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RLSADB, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
 -->
@@
 
-What sources of data might provide the greatest value in improving our estimates of global stock status?
 Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available.
@@
 
-```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
+```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error (RMSE) of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
 Distribution is full posterior probability distribution.
@@
 Change is relative to the mean performance of a catch-only heuristic model."}
-b_voi_plot + 
+
+voi_plot <- b_voi_plot + 
   geom_vline(aes(xintercept = 0), alpha = 0.75)
+
+voi_plot
+
+ggsave(here("documents","figs","voi-plot.pdf"),voi_plot, width = 8, height = 6)
+
 ```
@@
 
-<!-- Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RAM -->
+<!-- Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RLSADB -->
 
@@
 The hope of efforts using catch-only models to estimate the status of global unassessed fisheries is fundamentally that we can learn something meaningful about the state of a fished population simply by knowing something about its catch history and life history.
-While in some cases the addition of globally available data such a FMI scores, SAR values, or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
+While in some cases the addition of globally available data such a FMI scores, SAR values, or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RLSADB-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
 
 What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends on the needs of those using the assessment outputs.
-It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient.
+It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficiently accurate.
 In some instances using the data presented here did provide some improvement over use of catch-only style assessment methods; the difficulty comes in attempting to apply data types uniformly across the globe.
-While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RAM, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
+While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RLSADB, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
 The overall poor performance of the models tested here shows that improvements in estimates of global stock status depend on improvements in the quality and use of data themselves
 
-<!-- That use of none of the datasets collected here can achieve that, and that our tests fitting directly to abundance indices from RAM itself suggests that model misspecification is not the primary cause of poor performance, tells us that improvements in estimates of global stock status will likely depend on improvements in the quality and use of data themselves.
+<!-- That use of none of the datasets collected here can achieve that, and that our tests fitting directly to abundance indices from RLSADB itself suggests that model misspecification is not the primary cause of poor performance, tells us that improvements in estimates of global stock status will likely depend on improvements in the quality and use of data themselves.
 -->
@@
 
-We chose to test the performance of methods against estimated values reported in RAM.
-A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RAM, in addition to the errors in the RAM estimates themselves.
-For this reality to change our results though would require that the methods tested here have massively lower bias and higher accuracy for unassessed fisheries than RAM stocks.
+We chose to test the performance of methods against estimated values reported in RLSADB.
+A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RLSADB, in addition to the errors in the RLSADB estimates themselves.
+For this reality to change our results though would require that the methods tested here have massively lower bias and higher accuracy for unassessed fisheries than RLSADB stocks.
 @free2020's simulation based based results of these same types of models suggests that this is unlikely to be true.
+However, while many assessed fisheries have mechanisms to actively manage catches, catch histories in unassessed and unmanaged fisheries may be more likely to reflect underlying stock status than those in heavily regulated fisheries.
 
@@
 Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks.
-None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RAM (XX RAM matches?
-need to add an analysis: match stocks from SOFIA to RAM stocks where possible, plot comparison).
+None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RLSADB (XX RLSADB matches?
+need to add an analysis: match stocks from SOFIA to RLSADB stocks where possible, plot comparison).
 -->
@@
 Given appropriate conditions, these length measurements can be used to estimate local fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
-While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
+While length-based assessments come with a host of assumptions and potential pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
 Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f; @rudd2017].
@@
 We must also prioritize collection and curation of fish population survey data worldwide.
-Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://www.FishStats.org)( www.FishStats.org).
+Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://www.FishStats.org) ( www.FishStats.org).
 Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments.
-Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/)((https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
+Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/)(https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
 However, out value of information exercise indicates that we may have to wait many years for new surveys to provide substantial improvements in status estimates (Fig.\@ref(fig:voi-plot)).
@@
 
+# Data Availability Statement {.unnumbered}
+
+All data and materials needed to reproduce our results are publicly available or queried by code available at https://github.com/DanOvando/assessing-global-fisheries.
+
 # Acknowledgements {.unnumbered}
@@
 We thank participants of the FAO "Methods for Global Assessment" workshop held in Rome, Italy February 2019 for helpful feedback on this project.
-The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of NOAA or the Department of Commerce.
+The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of the FAO, NOAA or the Department of Commerce.
 
@@
 
+\newpage
+# Tables {.unnumbered}
+
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Data Source                                 | Short Name       | Data Use                                                                 | Caveats                                                                                                         |
++=============================================+==================+==========================================================================+=================================================================================================================+
+| Catch data [@fao2020]                       | catches          | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Fisheries Management Index [@melnychuk2017] | FMI              | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Swept Area Ratio [@amoroso2018]             | SAR              | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+| Reconstructed effort data [@rousseau2019b]  | effort           | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors.
+Assumed rate of technology creep reported in individual sections |
++---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+
+: (\#tab:data) Data sources included across model fits.
+
+```{r dat-desc}
+dat_description <- tribble(~`Data Name`, ~Description,
+        "RLSADB Index", "Fit to abundance index from RLSADB",
+        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
+        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
+        "Efective CPUE", "Fit to CPUE index created from RLSADB catch and regional effort index. 2.6% technology creep",
+        "Effective CPUE+", "Fit to CPUE index created from RLSADB catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
+        "Nominal CPUE", "Fit to CPUE index created from RLSADB catch and regional effort index. 0% tech. creep",
+         "Nominal CPUE+", "Fit to CPUE index created from RLSADB catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
+        "Guess", "Priors on terminal B/Bmsy randomly sampled from 0.4,1.0,1.6")
+
+knitr::kable(dat_description,
+             caption = "Data sources used for terminal stock status estimate",
+             format = "pipe") %>%
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+```
+
+```{r perf-tab}
+
+knitr::kable(perf %>% rename(MPE = mpe, MAPE = mape, Accuracy = accuracy), digits = 2,
+             caption = "Global performance statistics in the most recent year available of models using different sources of data.
+MPE = median percent error (bias), MAPE = median absolute percent error (error), Accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
+Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.",
+              format = "pipe")
+
+```
+
+\newpage
+# Figure Legends  {.unnumbered}
+
+Figure 1: RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
+Each point is a stock.
+Black dashed line shows the 1:1 relationship.
+
+Figure 2: Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
+Panels ordered in ascending (starting from top left) mean MPE at the FAO region level.
+
+Figure 3: Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
+Panels ordered in descending (starting from top left) mean MAPE at the FAO region level
+
+Figure 4: Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
+Panels ordered in descending (starting from top left) mean accuracy at the FAO region level
+
+Figure 5: Posterior probability distributions of estimated effect of different data types on root mean squared error (RMSE) of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
+Distribution is full posterior probability distribution.
+Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
+Change is relative to the mean performance of a catch-only heuristic model.
+
+\newpage
+
 # Supporting Information {.unnumbered}

### 2021-04-24 functions/fit-case-studies.R

@@ -11,7 +11,8 @@ fit_case_studies <-
            chains = 1,
            cores = 1,
            cmsy_cores = 4,
-           q_slope_prior = 0.025) {
+           q_slope_prior = 0.025,
+           use_b_reg = FALSE) {
  
     catches <- data$catch
     
@@ -63,7 +64,8 @@ fit_case_studies <-
         terminal_state = NA,
         terminal_state_cv =  0.05,
         use_heuristics = FALSE,
-        b_ref_type = "k"
+        b_ref_type = "k",
+        use_b_reg = use_b_reg
       )
     
     
@@ -101,7 +103,8 @@ fit_case_studies <-
         effort_years = effort_years,
         q_slope_prior = q_slope_prior,
         q_slope_prior_cv = 0.2,
-        b_ref_type = "k"
+        b_ref_type = "k",
+        use_b_reg = use_b_reg
       )
     
     sfs <- safely(fit_sraplus)

### 2021-04-24 make-assessing-global-fisheries.R

@@ -798,7 +798,6 @@ exs <- exs %>%
   left_join(matched_sar, by = "stockid") %>%
   left_join(comp_stocks, by = "scientificname")
 
-
 if (run_case_studies){
   
   exs <- exs %>%
@@ -908,7 +907,7 @@ ex_scatter_plot <- tmp2 %>%
   ggtext::geom_richtext(data = ex_performance, aes(x = 1, y = 4, label = paste0("R<sup>2</sup> = ",r2))) +
   facet_grid(variable ~ data, scales = "free_x")  + 
   # scale_size(trans = "sqrt", name = "Lifetime Catch") +
-  scale_x_continuous(name = "RAM Value", expand = expansion(add = c(0, .1))) + 
+  scale_x_continuous(name = "RLSADB Value", expand = expansion(add = c(0, .1))) + 
   scale_y_continuous("Estimated Value", expand = expansion(add = c(0, .1)))
 
 
@@ -2027,7 +2026,7 @@ ram_status <- fao_area_ram_status %>%
   group_by(f_area) %>% 
   slice(1) %>% 
   select(-data, -median_sraplus) %>% 
-  mutate(data = "RAM Assessment", median_sraplus = median_ram) 
+  mutate(data = "RLSADB Assessment", median_sraplus = median_ram) 
   
   
 
@@ -2038,7 +2037,7 @@ fao_area_ram_status %>%
 
 
 ram_labeller <- c(
-  "ram-data" = "RAM Index",
+  "ram-data" = "RLSADB Index",
   "sar" = "SAR",
   "fmi" = "FMI",
   "cpue" = "Effective CPUE",
@@ -2047,7 +2046,7 @@ ram_labeller <- c(
   "nominal-cpue-plus" = "Nominal CPUE+",
   "cmsy" = "CMSY",
   "guess" = "Guess",
-  "u_umsy" = "RAM U/Umsy"
+  "u_umsy" = "RLSADB U/Umsy"
 )
 
 
@@ -2105,8 +2104,8 @@ ram_mape_map_plot <- fao_area_ram_status %>%
   facet_wrap(~ data, labeller = labeller(data = ram_labeller)) +
   scale_fill_gradient2(
     low = "blue",
-    mid = "yellow",
-    high = "tomato",
+    mid = "tomato",
+    high = "yellow",
     name = "% Error (MAPE)",
     labels = collabs,
     breaks = seq(0,1, by = .25),
[…]

### 2021-04-24 make-assessing-global-fisheries.R
F and F (#12)
* initial edits and formatting for fish and fisheries

@@ -4,6 +4,7 @@
 
 library(tidyverse)
 library(ggridges)
+library(ggtext)
 library(gganimate)
 library(rstan)
 library(mvtnorm)
@@ -797,7 +798,6 @@ exs <- exs %>%
   left_join(matched_sar, by = "stockid") %>%
   left_join(comp_stocks, by = "scientificname")
 
-
 if (run_case_studies){
   
   exs <- exs %>%
@@ -893,16 +893,23 @@ tmp2 <- tmp2 %>%
     )
   )
 
+ex_performance <- tmp2 %>% 
+  group_by(variable, data) %>% 
+  summarise(r2 = round(yardstick::rsq_vec(truth, value),2)) 
+
+
 ex_scatter_plot <- tmp2 %>% 
   ggplot(aes(truth, value)) + 
   geom_vline(aes(xintercept = 0)) + 
   geom_hline(aes(yintercept = 0)) +
   geom_abline(aes(slope = 1, intercept = 0),linetype = 2) +
-  geom_point(alpha = 0.75) + 
+  geom_point(alpha = 0.5, size = 2) + 
+  ggtext::geom_richtext(data = ex_performance, aes(x = 1, y = 4, label = paste0("R<sup>2</sup> = ",r2))) +
   facet_grid(variable ~ data, scales = "free_x")  + 
   # scale_size(trans = "sqrt", name = "Lifetime Catch") +
-  scale_x_continuous(name = "RAM Value", expand = expansion(add = c(0, .1))) + 
+  scale_x_continuous(name = "RLSADB Value", expand = expansion(add = c(0, .1))) + 
   scale_y_continuous("Estimated Value", expand = expansion(add = c(0, .1)))
+
 
 sraplus_v_truth <- tmp %>% 
   left_join(truth, by = c("stockid","year")) %>% 
@@ -2019,7 +2026,7 @@ ram_status <- fao_area_ram_status %>%
   group_by(f_area) %>% 
   slice(1) %>% 
   select(-data, -median_sraplus) %>% 
-  mutate(data = "RAM Assessment", median_sraplus = median_ram) 
+  mutate(data = "RLSADB Assessment", median_sraplus = median_ram) 
   
   
 
@@ -2030,7 +2037,7 @@ fao_area_ram_status %>%
 
 
 ram_labeller <- c(
-  "ram-data" = "RAM Index",
+  "ram-data" = "RLSADB Index",
   "sar" = "SAR",
   "fmi" = "FMI",
[…]

### 2021-06-17 documents/ovando-etal-assessing-global-fisheries.Rmd
fixed new FAO areas mapping problem

@@
 
-assess_ram_fits <- read_rds(path = file.path(results_path,"assess_ram_fits.rds"))
+assess_ram_fits <- read_rds(file = file.path(results_path,"assess_ram_fits.rds"))
 
-ram_v_sraplus_area <- read_rds(path = file.path(results_path,"ram_v_sraplus_area.rds"))
+ram_v_sraplus_area <- read_rds(file = file.path(results_path,"ram_v_sraplus_area.rds"))
 
-ram_comp_data <- read_rds(path = file.path(results_path,"ram_comp_data.rds"))
+ram_comp_data <- read_rds(file = file.path(results_path,"ram_comp_data.rds"))
 
-fao_sraplus_comp_data <- read_rds(path = file.path(results_path, "fao_sraplus_comp_data.rds"))
+fao_sraplus_comp_data <- read_rds(file = file.path(results_path, "fao_sraplus_comp_data.rds"))
 
@@
 
-  exs <- read_rds(path = file.path(results_path, "raw-case-study-fits.rds"))
+  exs <- read_rds(file = file.path(results_path, "raw-case-study-fits.rds"))
 

### 2021-06-17 make-assessing-global-fisheries.R

@@ -25,6 +25,7 @@ library(tidybayes)
 library(rstan)
 library(rstanarm)
 library(sf)
+library(rgdal)
 library(portedcmsy)
 extrafont::loadfonts()
 
@@ -807,11 +808,11 @@ if (run_case_studies){
     mutate(fit = map(data,
                      safely(fit_case_studies)))
   
-  write_rds(exs, path = file.path(results_path, "raw-case-study-fits.rds"))
+  write_rds(exs, file = file.path(results_path, "raw-case-study-fits.rds"))
   
 } else {
   
-  exs <- read_rds(path = file.path(results_path, "raw-case-study-fits.rds"))
+  exs <- read_rds(file = file.path(results_path, "raw-case-study-fits.rds"))
   
 }
 
@@ -1381,7 +1382,7 @@ if (run_sofia_comparison == TRUE) {
   
 } else {
   fao_status_fits <-
-    read_rds(path = file.path(results_path, "fao_status_fits.rds"))
+    read_rds(file = file.path(results_path, "fao_status_fits.rds"))
   
   
   
@@ -1489,8 +1490,10 @@ fao_sraplus_comp_data <-  faofits2 %>%
 
 
 if (!dir.exists(here("data", "FAO_AREAS_NOCOASTLINE"))) {
-  download.file(url = "http://www.fao.org/figis/geoserver/area/ows?service=WFS&request=GetFeature&version=1.0.0&typeName=area:FAO_AREAS_NOCOASTLINE&outputFormat=SHAPE-ZIP",
+
+  download.file(url = "http://www.fao.org/fishery/geoserver/fifao/ows?service=WFS&request=GetFeature&version=1.0.0&typeName=fifao:FAO_AREAS_CWP_NOCOASTLINE&outputFormat=SHAPE-ZIP",
                 destfile = here("data", "FAO_AREAS_NOCOASTLINE.zip"))
+ 
   
   unzip(
     here("data", "FAO_AREAS_NOCOASTLINE.zip"),
@@ -1505,23 +1508,14 @@ fao_sraplus_acc <- fao_sraplus_comp_data %>%
   group_by(data, fao_area_code) %>% 
   summarise(accuracy = mean(correct, na.rm = TRUE))
 
-fao_areas <- sf::st_read(here('data', "FAO_AREAS_NOCOASTLINE")) %>%
-  janitor::clean_names()
-
-fao_areas <- fao_areas %>%
-  group_by(f_area) %>%
-  nest() %>%
[…]

### 2021-06-18 reviewer-replies.Rmd
R1 done

---
title: "Reviewer Replies"
author: "Dan Ovando"
date: "6/16/2021"
output: html_document
---

# Reviewer 1

## Performance metrics

*The stock status of a fish stock is commonly evaluated in terms of biomass and fishing mortality. The authors present all performance metrics for the regional fishery assessments (MPE, MAPE, and accuracy) only in terms of B/BMSY . For the case study, the authors present the stock status in terms of both B/BMSY and F/FMSY . Is there a reason why the authors considered only B/BMSY for the regional fishery assessments? I would be interested in all performance metrics in terms of F/FMSY and recommend to include this quantity for the regional fishery assessments at least for the supporting information.*

Good question. We focused on B/B~MSY~ as this project was originally designed with an eye towards improving the SOFIA status estimates, which consider B/B~MSY~, not F/F~MSY~. However, F/F~MSY~ is clearly of interest as well, particularly as catch-only model F/Fmsy outputs are often used for assessment or projection, in the manner of @costello2016.

We added F/F~MSY~ performance metrics summarized in the same type of table and figures used in the main document to the Supporting Information (we do not include in the main text due to limitations on numbers of available figures and tables). We also briefly mention these results in the results and discussion section. Broadly, we find similar patterns of generally poor and variable performance, with the notable exception that while in terms of B/B~MSY~ the default settings of the version of CMSY used in this paper performed in the middle of the pack in most cases, in terms of F/F~MSY~ CMSY was last in terms of performance across every metric, due to severe positive biases in F/F~MSY~ .

## Discussion of Results

*Why did the authors not apply sraplus without any additional data (SAR, FMI) for the case studies? This would constitute a fairer comparison of the comparison of CMSY and sraplus.*

We have included the requested comparison in the case study section. Under these settings, which we call "Catch Heuristic", sraplus uses the same catch heuristics as CMSY to generate priors on terminal depletion. The only real difference between the two algorithms then is the presence of a prior on midpoint depletion in CMSY, and the use the prior-predictive tuning regime in sraplus. The prior-predictive tuning regime ensures that in the absence of new data to update the priors, the "posterior" more or less match the explicit priors on terminal depletion. This is why it produces two bands of stock status corresponding to the binary values proscribed by the catch-heuristic. CMSY "updates" those priors to produce the join prior-predictive distribution of the life history priors and explicit depletion priors conditional on the model and catch history.  We did not originally include this run in the case study set as our contention is not that the sraplus algorithm is an inherently better catch-only mousetrap than CMSY (though we feel the addition of the prior-predictive tuning procedure is an important option).But, we agree that inclusion of this run in the case studies highlights an important philosophical choice of the default settings of sraplus (that we should be explicit in where our prior in terminal depletion is coming from).

*The results reveal highly variable results and no consistent patterns among regions. Did the authors find any patterns in the assessment performance regarding the life-history parameters of the stocks? This aspect is currently lacking from the discussion.*

Good question. We have not explored this idea empirically, but we now mention it explicitly in the discussion as an area for future research.

*Simiarly for the case studies, I wonder if the authors found any common characteristics for the stocks with larger biases?*

One reason that we identified and more clearly articulate in the results that the effort data wasn't helpful was that construction of a CPUE index based on fishery catches and large-scale estimated effort resulted in very low CPUE values during quota or market constrained periods in a fishery's history. During these periods, according to the CPUE index abundance was very low, but RLSADB reports relatively high biomass levels, causing the model to perform badly. We address this general idea of "why" of performance in the discussion now.

*Lastly, the conclusion that catch-only methods may be useful contradicts the findings of the case study and regional fishery assessments, which indicate a poor performance (high bias, low precision and low classification accuracy) of catch-only methods if no fishery-independent data or informative and unbiased depletion priors are available. If this is one of the main conclusions, the authors should elaborate in more detail what they imply with “when applied properly*

Both reviewers pointed this discrepancy out, and we agree. We have removed that statement, as well feel it contains too many nuances to be adequately covered in the concluding paragraph of the paper. The general point that there may be some uses for these methods is made in the discussion paragraph starting with "What quality of assessment..."

## Specific Comments

Thank you for the edits. All specific comments were accepted unless specifically noted below.

*l. 148pp. Why do you x the shape parameter rather than using the priors recommended by Thorson et al. (2012)?*

We believe the question here is why did we fix the shape parameter the mean values reported by @thorson2012, rather than taking thoes values as informative priors and estimating the shape parameter? We tried this, but it slowed the models down substantially, and often resulted in very poor fits / convergence given the lack of information in the data on this parameter. It also had a tendency to shoot the shape parameter into odd spaces, given the relatively large CVs aroudn the values reported in @thorson2012. We have clarified this in the methods.

*l. 153 Please specify whether you follow Winker et al. (2018) dening BMSY/K = 0.25 (in your case) when BMSY/K < 0:25. If another assumptions is made, this needs to be specied.*

Good catch! In the results in the paper Bmsy / K is never below 0.25, though it comes close for Clupeiformes. However, you are correct that there was not a check for this possibility in the broader sraplus pacakge, so we are adding that in in case users try and specicy a shape parameter such that Bmsy/K < 0.25.

We have included the table below in the SI as well for reference.

```{r shape-tab}

knitr::kable(
  sraplus::ssb0msy_to_ssb0,
  "latex",
  align = "l",
  booktabs = TRUE,
  escape = F,
  row.names = F,
  caption = "Ratio of SSBmsy to SSB0 by taxinomic group from Thorson et al. (2012) used as default priors in sraplus."
) %>%
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

*l. 169 Do you mean proc instead of obs? obs not defined at this point of the manuscript. Where is defined? It is mentioned in Table S1 as the ratio of process and observation error, but not mentioned in the main text. Consider lifting Table S1 to main text and add more explanation to on estimated parameters. Be specic whether proc is estimated or fixed. Are the priors of Table S2 used? Further, consider using Bo instead of B0 and define B0 as it can also be used as virgin biomass.*

Thank you for pointing these out. We have clarified the notation here and included expanded definitions in the text at the section in question. Unfortunately we do not have space to bring in another table to the main results, but we have stated explicitly how the default priors are used and pointed the readers to that table.

*Fig. 5 I am not sure that I understood the differences between the scenarios initial state known" and initial state unfished". Thus, I do not understand, why RMSE is higher for the scenario with the initial state unfished. Please provide more information about these scenarios.*

We have added the missing explanation of these states fo the Value of Information section of the methods.

*Fig. 2-4 Keep same order of models/scenarios for all figures.*

After consultation with the group of authors, we respectfully disagree with this suggestion and as such have left the figures ordered as they are. The order of the panels reflects the performance of each data source by the metric in question, with best performing in the top left and worst in the bottom right. We feel this is helpful to the reader to parse apart the performance across the different data types. However, to clarify this we have made the caption clearer. However if you feel that this still introduces needless confusion we can make the figures in the manner suggested.

*Table S2 Where does the default prior of 1.01 for m come from? Thorson et al. (2012) report an average prior of 1.478 across all taxa.*

Good catch, that was an relic of an older set of results, we have updated the table to state that priors were drawn from @thorson2012, and provided a separate table with the specific values from @thorson2012 used.

# Reviewer 2

Review of “Improving Estimates of the State of Global Fisheries Depends on Better Data” by Ovando et al, submitted to Fish and Fisheries
This is a well written manuscript that describes an interesting analysis and presents valuable results. I recommend publication of this manuscript in Fish and Fisheries after a minor revision.
I have a few comments that I hope assist the authors in improving the manuscript.
General comments
Terms like biomass and notation B can often be used to refer to different things, ie., total biomass, vulnerable biomass, or spawning biomass. A non-structured model such as the SRA used here doesn’t distinguish between these three, and given the model is scaled with catches, B really represents vulnerable biomass; i.e., K = unfished vulnerable biomass, and the estimate of B/BMSY reported by the SRA corresponds with VB/VBMSY in a structured model.
As the authors point out, the assessments in the RLSADB are typically more structurally complex – presumably many of them age-structured – and I assume report B/BMSY as SB/SBMSY given spawning population is generally the focus for conservation outcomes. The relationship between SB/SBMSY and VB/VBMSY would depend on the selectivity pattern relative to maturity and level of fishing mortality (eg VB/VBMSY could go to 0 while SB/SBMSY > 0). From a quick analysis with some operating models I have on hand the ratio of VB/VBMSY and SB/SBMSY ranged from 0.7 to 1.6.
If it is true that the SRA is reporting VB/VBMSY and comparing it to SB/SBMSY from the RLSADB there will be an additional level of bias and variability. This issue applies more widely any time outputs from models with fundamental structure are compared, but I rarely see it acknowledged or discussed.
I don’t expect this would make a significant difference to the results of this study, and I don’t think this is a flaw in the study, but I’m curious if the authors have any thoughts on this.
It seems strange to me that having information on the most recent B/BMSY doesn’t result in lower RMSE for the predicted B/BMSY from the most recent 5 years of data (Figure 5). Presumably B/BMSY wouldn’t change a great deal between years, so it seems counter-intuitive that F/FMSY complete results in a better prediction of B/BMSY than when most recent B/BMSY is known. Perhaps the authors could add a bit of text to explain this.
The paragraph beginning on line 379 is useful, but from my perspective is a little overly generous to catch-only methods. The claim on this paper could be defended more strongly than is implied here. It’s logical that catch-only methods *need* some other data (or assumptions which are essentially assuming data look a certain way) to provide reliable estimate of stock status. E.g., even in unmanaged fisheries, the relationship between catch history and B/BMSY would have to at least depend on the time the stock has been fished, and the relative fishing effort on the stock (e.g., highly targeted, by-catch, etc). This paper makes the important point that providing estimates of stock status for unassessed fisheries requires better data (and capacity building to analyze the data); that is, essentially a stock assessment. Short-cut methods are appealing, but have the danger of
over-simplifying a complex reality and could be detrimental to both sustainability and livelihoods if they don’t reliably categorize fisheries.
Minor comments
25 – typo – “world’s”
39 – should be a comma after ‘sustainably’
Actually – I’d suggest a rephrase – fisheries *are* a critical source of economic prosperity and food security around the globe – that’s why sustainable management is so important
56 – is it possible to say what % of the world’s fisheries this 50% of catch represents? 107 – ‘SRA-style’
121 – move comma out of quotes
152 – remove brackets from front of Winker to date – same with 179
169 – what are sigma obs and gamma? Described later on I think.
188 – should F be upper case?
213 – suggest deleting ‘though’ at end of sentence and replacing with ‘However’ at the beginning
215 – ‘Due to the fact ...’ incomplete sentence
224 – has the SIR algorithm been described above? Hasn’t been mentioned yet – do you mean SRA?
305 – methods said Thorson 2012
306 – R2 values don’t match Figure 1? Fig 1 says R2 = 0.03 for CMSY
302-317 – helpful in this section to refer to figure 1 and panels with figure
313 – only SAR & FMI fit better to F/FMSY right? The others had higher R2 (though too low to consider) for B/BMSY (Figure 1)
325 – ‘high quality data’ – specifically a reliable index of abundance right?
332 – did you consider sub-setting some of the RLSADB assessments to only use the earlier years where management may not have been as effective and B/BMSY in the terminal year less optimistic?
378 – missing full-stop at end of sentence
380 – ‘It’ to ‘it’
396 – ‘heavily managed’ – perhaps ‘relatively well managed ...’
400 – 402 – it would have to depend on the catch history though – ie length of time the stock has been fished (and unmanaged). See General Comments about.
411 – missing space between 0.43 and ‘But’. Also, Fig 1 has R2 of 0.47.

436 - missing space
441 – 449 – It may be worth pointing out that developing management methods that do not explicitly rely on an estimate of stock status could be useful as well, eg spatial closures, size regulations, empirical index-based MPs

### 2021-06-18 documents/ovando-etal-assessing-global-fisheries-si.Rmd
R1 done

@@
   rbind(c("Shape parameter", "$m$", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
-  rbind(c("Catchability", "$q$", "$logn(1e^{-3}, 1)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,1)$")) %>%
-  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.5,0.25)$")) %>%
-  rbind(c("Initial State", "$B0$", "Posterior probability dist. of catch-based regressions"))
+  rbind(c("Catchability", "$q$", "$LN(1e^{-3}, 1)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$LN(.05,1)$")) %>%
+  rbind(c("Ratio of process to observation error", "$\\gamma$", "$LN(.5,0.25)$")) %>%
+  rbind(c("Initial State", "$B/K_0$", "Posterior probability dist. of catch-based regressions"))
 
@@
 ```{r si-prior-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
-  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
-  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
-  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
-  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
-  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))
+cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$LN(10\\times{max(catch)},4)$") %>%
+  rbind(c("Growth rate", "r", "$LN(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
+  rbind(c("Shape parameter", "m", "Thorson et al. (2012)")) %>%
+  rbind(c("Catchability", "q", "$LN(1e^{-3}, 0.3)$")) %>%
+  rbind(c("Observation Error", "$\\sigma_{obs}$", "$LN(.05,2)$")) %>%
+  rbind(c("Process Error", "$\\sigma_{proc}$", "$LN(.05,0.5)$")) %>% 
+  rbind(c("Tech Creep", "$creep$", "$LN(0.025,0.2)$"))
 
@@
   kableExtra::row_spec(0, bold = T)
+```
+
+```{r shape-tab}
+
+knitr::kable(
+  sraplus::ssb0msy_to_ssb0,
+  "latex",
+  align = "l",
+  booktabs = TRUE,
+  escape = F,
+  row.names = F,
+  caption = "Ratio of SSBmsy to SSB0 by taxinomic group from Thorson et al. (2012) used as default priors in sraplus."
+) %>%
+  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
+  kableExtra::row_spec(0, bold = T)
+
 ```
@@
 Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$
+
+## F/F~MSY~ Performance {.unnumbered}
+
+Our results focused on the performance of candidate models in estimating B/B~MSY~, as this reflects the broad mission of the FAO's SOFIA reports to assess the current biomass status of global fisheries.
+However, fishing mortality rates, specifically F/F~MSY~ are also of importance to managers and commonly considered as an output of catch-only models.
+
+As such, we repeated our performance calculations summarized in Figures \@ref(fig: mpe-map)-\@ref(fig:acc-map) but now focused on F/F~MSY~.
+Performance was comparably poor to the B/B~MSY~ based results, with the exception that the default settings of CMSY produced a consistent positive bias in F/F~MSY~.
+
+```{r uperf-tab}
+
+knitr::kable(uperf %>% rename(MPE = mpe, MAPE = mape, Accuracy = accuracy), digits = 2,
+             caption = "Global performance statistics in the most recent year available of models using different sources of data.
+MPE = median percent error (bias), MAPE = median absolute percent error (error), Accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
+Performance is judged relative to *F/F~MSY~* values reported values in RAM Legacy Stock Assessment Database.",
+              format = "pipe")
+
+```
+
+```{r umpe-map, fig.width=8, fig.height=6, fig.cap="Median percent error (MPE, predicted relative to observed) in most recent *F/F~MSY~* by FAO statistical area from different data sources.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assigns a random recent *F/F~MSY~* of 0.4,1, or 1.6.
+Panels ordered in ascending (starting from top left) mean MPE at the FAO region level.
+"}
+ramu_mpe_map_plot
+
+ggsave(here("documents","figs","u-mpe-map.pdf"), ram_mpe_map_plot, width = 8, height = 6)
+```
+
+```{r umape-map, fig.cap = "Median absolute percent error (MAPE) in most recent *F/F~MSY~* by FAO statistical area from different data sources.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assigns a random recent *F/F~MSY~* of 0.4,1, or 1.6.
+Panels ordered in descending (starting from top left) mean MAPE at the FAO region level",fig.width=8, fig.height=6}
+ramu_mape_map_plot
+
+ggsave(here("documents","figs","u-mape-map.pdf"),ram_mape_map_plot, width = 8, height = 6)
+
+```
+
+```{r uacc-map, fig.cap="Mean classification accuracy (assignment to general bin of overfishing, fishing near F~MSY~, and underfishing) by FAO statistical area arising from different data sources.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
+Effective CPUE refers to an index of abundance based on reconstructed effort data.
+Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
+For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
+FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
+CMSY uses the methods from Froese et al.
+2017 [@froese2017].
+Guess assigns a random recent *F/F~MSY~* of 0.4,1, or 1.6.
+Panels ordered in descending (starting from top left) mean accuracy at the FAO region level",fig.width=8, fig.height=6}
+
+ramu_acc_map_plot
+
+ggsave(here("documents","figs","u-acc-map.pdf"),ram_acc_map_plot, width = 8, height = 6)
+
+```

### 2021-06-18 documents/ovando-etal-assessing-global-fisheries.Rmd
R1 done

@@
 
+assess_ramu_fits <- read_rds(file = file.path(results_path,"assess_ramu_fits.rds"))
+
 ram_v_sraplus_area <- read_rds(file = file.path(results_path,"ram_v_sraplus_area.rds"))
@@
 
+uperf <- assess_ramu_fits %>% 
+  # bind_rows(rand_fits) %>% 
+  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
+  mutate(ram_bin = cut(ram_u_v_umsy, breaks = breaks, labels = labels)) %>%
+  mutate(data = as.factor(data)) %>% 
+  filter(data != "u_umsy") %>% 
+  mutate(data = forcats::fct_recode(data, 
+                                    "RLSADB Index" = "ram-data",
+                                    "SAR" = "sar",
+                                    "FMI" = "fmi",
+                                    "Effective CPUE" = "cpue",
+                                    "Effective CPUE+" = "cpue-plus",
+                                    "Nominal CPUE" = "nominal-cpue",
+                                    "Nominal CPUE+" ="nominal-cpue-plus",
+                                    "CMSY" = "cmsy",
+                                    "Guess" = "guess",
+                                    "RLSADB F/Fmsy" = "u_umsy"
+
+                                    )) %>% 
+  group_by(data) %>% 
+  summarise(
+    mpe = median((mean - ram_u_v_umsy ) / ram_u_v_umsy),
+    mape = median(abs(mean - ram_u_v_umsy ) / ram_u_v_umsy),
+    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
+  ) %>% 
+  arrange((mape)) %>% 
+  rename("Data Used" = "data")
+
 fao_area_codes <- fao %>% 
@@
 Implementation of the United Nations Sustainable Development Goals requires assessments of the global state of fish populations.
-While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain.
+While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the world's 'unassessed' fish stocks remains highly uncertain.
 Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results.
@@
 While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods.
-With the surge in data availability, such as global assessments of management strength [@melnychuk2017], trawl footprints [@amoroso2018], and fishing effort [@rousseau2019b], the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
-In addition, the SOFIA process assigns unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to to the biomass that would on average result in maximized yield over the long-term.
+With the surge in data availability, such as global assessments of management strength [@melnychuk2017], trawl footprints [@amoroso2018], and fishing effort [@rousseau2019], the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
+In addition, the SOFIA process assigns unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to the biomass that would on average result in maximized yield over the long-term.
 
@@
 The key goal of sraplus is not substantial improvements in model fitting methods per say, but providing the ability to easily incorporate multiple kinds of fishery data potentially used in SRA style analyses in a statistically rigorous manner.
-We paired various types of data with fisheries in RLSADB, and then used sraplus to generate predictions of B/B~MSY~ in the most recent year of each fishery across different combinations of data types.
+We paired various types of data with fisheries in RLSADB, and then used sraplus to generate predictions of *B/B~MSY~* in the most recent year of each fishery across different combinations of data types.
 We then compared the predictions generated by sraplus using different types of data, along with predictions for the same fisheries generated by CMSY, to the reported values in RLSADB.
@@
 While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries.
-As an alternative, @rousseau2019b present a global reconstruction of fishing effort for countries around the world.
+As an alternative, @rousseau2019 present a global reconstruction of fishing effort for countries around the world.
 We pair these effort reconstructions (both nominal effort and effective effort assuming an annual rate of efficiency increase of 2.6%) with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status.
@@
 
-The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of Winker *et al.* 2018 [@winker2018].
+The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of @winker2018.
 While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the data types evaluated here will support.
@@
 Where $B_t$ is biomass at time *t*, *K* is carrying capacity ,*r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of *B~MSY~/K* to shift.
-When *m* = 2 *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right.
+When *m* = 2, *B~MSY~ / K* = 0.5.
+Lower values of *m* shift the production function left, higher values right.
 The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, @thorson2012 provides estimates of the ratio of *B~MSY~* to *K* for many fish taxa.
-We fix the shape parameter based on the distributions reported in @thorson2012 for the genus of the species in question.
-
-$\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors.
+For each stock we fix the shape parameter based on the distributions reported in @thorson2012 for the genus of the species in question.
+We chose to fix the shape parameter at the mean stock-appropriate values from @thorson2012 rather than estimating the shape parameter with an informative prior since there is so little information regarding the shape in the data considered.
+Attempts to estimate the shape parameter with priors from @thorson2012 frequently resulted in poor model performance.
+$\pmb{c}$ is a vector of catches, and $\pmb{p}$ is vector of process errors.
 Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
@@
 \begin{equation}
-  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
+  p_t \sim e^{N\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
 (\#eq:procerror)
 \end{equation}
+
+where *N* is the normal distribution.
 
@@
 Unless there is an abundance index to fit to, the model runs as an SRA.
-A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
+A stock reduction analysis works by specifying prior distributions on population parameters and, critically, the recent state of the fishery.
 sraplus allows users to specify the most recent status in units of depletion, *B/B~MSY~*, *F*, or *F/F~MSY~*.
@@
 The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
-All stock reduction analysis runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
-For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*.
+All stock reduction analysis runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e^6^ candidate draws.
+For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{proc}$, and *B/K~0~* (initial depletion).
 *q* is also estimated when needed.
+$\sigma_{proc}$ is estimated indirectly through the parameter $\gamma$, the ratio of process (*proc*) to observation (*obs*) error ($\sigma_{obs}$).
+See Table SX for prior distributions for each of these parameters.
 
@@
 
-$$log(a_t) \sim normal(F(r,K,m,B0,\pmb{p},\pmb{c}) \times q, \sigma_{obs})$$
-
-where $a_t$ is the observed abundance index and *F* is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
+$$log(a_t) \sim N(f_{pt}(r,K,m,B0,\pmb{p},\pmb{c}) \times q, \sigma_{obs})$$
+
+where $a_t$ is the observed abundance index and $f_{pt}$ is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
 When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data.
-[@rousseau2019] measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep).
+@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep).
 Treating this raw effort data as the denominator in the CPUE calculation assumes that every increase in fishing effort translates to a commensurate increase in fishing mortality.
@@
 
-$$cpue_t = \frac{c_t}{(1 - e^{-f_t})}$$
-
-$$f_t = q_tE_t$$
+$$cpue_t = \frac{c_t}{(1 - e^{-F_t})}$$
+
+$$F_t = q_tE_t$$
 
@@
 
-$$log(cpue_t) \sim normal(f(r,K,m,B0,\pmb{p},\pmb{c}), \sigma_{obs})$$
+$$log(cpue_t) \sim N(f_{pt}(r,K,m,B0,\pmb{p},\pmb{c}), \sigma_{obs})$$
 
@@
 Our solution amounts to a two-step sampling-importance-resampling (SIR) algorithm.
-We first run the standard SIR algorithm as described above.
+We first run the standard SIR algorithm as described in the Estimation Model section of the methods.
 We then break the resulting draws into bins based on terminal stock status, and calculate the mean probability density *p* (defined by the prior distributions of estimated parameters) of each bin.
@@
 We paired data on catch histories, swept area ratio, and Fisheries Management Index with estimates of stock status from the RLSADB.
-We then trained a regression of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types.
+We then trained a regression of the general form $log(status) \sim N(variable,\sigma)$ for each of these three data types.
 Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries.
@@
 Guess assigns each stock a random *B/B~MSY~* of 1.6, 1, or 0.4, corresponding to the middle value of the FAO status bins of underfished  ($B/B~MSY~ \geq 1.2$), maximally sustainably fished ($0.8 \geq B/B~MSY~ < 1.2$), overfished  ($0 \geq B/B~MSY~ < 0.8$).
+We performed a matching analysis measuring performance in estiamting *F/F~MSY~* as well, with results presented in the SI.
 
@@
 
-We performed a value-of-information assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The value-of-information analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RLSADB, and comparing the estimated values to the values reported in RLSADB.
+We performed a value of information assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
+The value of information analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RLSADB, and comparing the estimated values to the values reported in RLSADB.
 There are too many combinations for us to run the full expansion of possible parameter states.
@@
 The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
-We then measured the value of information as the average reduction in root mean squared error (RMSE) in B/B~MSY~, resulting from use of different kinds of data.
+We then measured the value of information as the average reduction in root mean squared error (RMSE) in *B/B~MSY~* over the most recent five years of the fishery (in order to evaluate the ability of the model to capture recent trends as well as the most recent value), resulting from use of different kinds of data.
+We considered the value of information of having data on: the most recent *B/B~MSY~*, treating initial state of the population as one of unfished ($B/K_{0}=1$), known ($B/K_{0}$ taken from stock assessment), or estimated based on initial shape of the catch history, *F/F~MSY~* values in the most recent year, over the last five years, and over the complete time series, and an abundance index spanning the complete, most recent half, or most recent quarter of the time series.
+We considered the value of information of a longer time series of *F/F~MSY~* than *B/B~MSY~* to consider the potential of augmenting catch-only models with additional data that can inform *F/F~MSY~*, such as length-composition data, that might be available over the history of the fishery.
 
@@
 As a benchmark, we first estimated stock status for these case study fisheries using the default settings of the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available.
-We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus.
+We then used sraplus to generate estimate of stock status based solely on catch heuristics, taking into account the prior-predictive tuning implemented by default in sraplus.
+We also used stock-specific data on SAR and FMI to generate priors on *F/F~MSY~* for each of the stocks, which were then passed to sraplus.
 Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
 This case study represents a more "localized" assessment, where external data sources (FMI and SAR) are available at the stock level, rather than at the regional level.
+We summarized the performance of each model fit in the case study based on root mean squared error (RMSE).
 
@@
 
-Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one.
-CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B~MSY~) and the estimated resilience of the target species (based on Fishlife [@thorson2020]), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing ($R^2$ for B/Bmsy = 0.1, $R^2$ for F/F~MSY~ = 0).
-Adding priors informed by stock-specific swept area ratio and Fisheries Management Index data produces some improvements, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
-Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit.
-This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values.
-However none of the models fit to the three case-study data combinations were able to explain much of the variation in B/B~MSY~, though the models performed better in some cases for predicting F/F~MSY~ (though this is to be expected as the prior-generating models were trained on data from RLSADB).
-Models informed by the SAR and FMI values performed the best in terms of B/B~MSY~ ($R^2$ = 0.1).
-The ranking of performance by data sources used was the same for F/F~MSY~, though the range of values was much higher, with the SAR and FMI based result producing an $R^2$ of 0.47 while the default CMSY settings produced an $R^2$ of effectively 0 (Fig.\@ref(fig:cs-plot)).
+Nearly all of the fisheries used in this case study have *F/F~MSY~* values less than one, and most have *B/B~MSY~* values greater than one.
+Both the catch heuristic implemented in sraplus and CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of *B/B~MSY~*) and the estimated resilience of the target species (based on Fishlife [@thorson2020]), badly miss this trend, predicting instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
+Adding priors informed by stock-specific swept area ratio and Fisheries Management Index data produces some improvements (RMSE below 1 relative to above 1 for all other case studies), capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
+Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit in *B/B~MSY~* and *F/F~MSY~*.
+This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) but high biomass values, creating a mismatch between the CPUE trends based on the effort and catch data and actual biomass.
+However none of the models fit to the four case-study data combinations were able to explain much of the variation in *B/B~MSY~*, though the models performed better in some cases for predicting *F/F~MSY~*.
+Models informed by the SAR and FMI values performed the best in terms of *B/B~MSY~* (RMSE = 0.89).
+The ranking of performance by data sources used was the same for *F/F~MSY~*, though the range of values was much higher, with the SAR and FMI based result producing an RMSE of 0.39 while the catch-only case studies produced *F/F~MSY~* RMSE values greater than 1 (Fig.\@ref(fig:cs-plot)).
+However, this improved performance of the SAR and FMI data is to be expected as the prior-generating models were trained on data from RLSADB, and therefore this is likely an optimistic assessment of the performance of these prior-generating models when applied to new fisheries.
 
@@
 
-```{r cs-plot, fig.cap="RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
-Each point is a stock.
+```{r cs-plot, fig.cap="RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], catch heuristics, priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort (Effort) trends assuming a rate of technological increase of 2.6%.
+Each point is a stock in the RLSADB.
 Black dashed line shows the 1:1 relationship."}
-ex_scatter_plot
+ex_scatter_plot + 
+  theme(strip.text = element_text(size = 8))
 
@@
 The mean accuracy of the sraplus models across all non-RLSADB data fits was `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and a so our "Guess" model has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "Guess"])`.
+Note that there are only three bins in the FAO stock status classifications, and as such our "Guess" model has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "Guess"])`.
 This means that the accuracy of our models designed as a proxy for a global assessment process were across all non-RLSADB index data fits `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]) / perf$accuracy[perf$"Data Used" == "Guess"] - 1)` more accurate than a random guess, certainly an improvement, but on average assigned fisheries to the wrong FAO status bin `r percent(1 - mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]))` of the time.
@@
 Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RLSADB index data were able to capture the overall state or geographic distribution of stock status represented in RLSADB in a consistent manner.
+Performance in estimating *F/F~MSY~ * was similarly variable and poor, with the exception that the default settings of CMSY performed much more consistently poorly in terms of *F/F~MSY~* than *B/B~MSY~* (due to systemic overestimating of *F/F~MSY~*, see SI).
 
@@
 ```{r mpe-map, fig.width=8, fig.height=6, fig.cap="Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+Data source panels are ordered in ascending (starting from top left) mean MPE at the FAO region level.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-Panels ordered in ascending (starting from top left) mean MPE at the FAO region level.
 "}
@@
 ```{r mape-map, fig.cap = "Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+Data source panels are ordered in descending (starting from top left) mean MAPE at the FAO region level.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-Panels ordered in descending (starting from top left) mean MAPE at the FAO region level",fig.width=8, fig.height=6}
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6}
 ram_mape_map_plot
@@
 ```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
-RLSADB Index refers to catch and abundance index drawn from RLSADB.
+Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region level RLSADB Index refers to catch and abundance index drawn from RLSADB.
 Effective CPUE refers to an index of abundance based on reconstructed effort data.
@@
 2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-Panels ordered in descending (starting from top left) mean accuracy at the FAO region level",fig.width=8, fig.height=6}
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6}
 
@@
 Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available.
-Interestingly though, having access to only an accurate estimate of F/F~MSY~ in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself  (Fig.\@ref(fig:voi-plot)).
+Interestingly though, having access to only an accurate estimate of *F/F~MSY~* in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself  (Fig.\@ref(fig:voi-plot)).
 While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average, due to the lack of historical context for recent trends in abundance.
@@
 It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficiently accurate.
+Catch-only based estimates of measures such as MSY are likely to be much more robust than reference points such as *B/B~MSY~* [@martell2013].
+Where catch-only methods are used for estimates of *B/B~MSY~* or *F/F~MSY~*, it is critical that users are aware of the broadly demonstrated poor performance of these methods in the absence of highly reliable fishery-specific information.
 In some instances using the data presented here did provide some improvement over use of catch-only style assessment methods; the difficulty comes in attempting to apply data types uniformly across the globe.
 While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RLSADB, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
-The overall poor performance of the models tested here shows that improvements in estimates of global stock status depend on improvements in the quality and use of data themselves
+The overall poor performance of the models tested here, in terms of estimating both *B/B~MSY~* and *F/F~MSY~*, shows that improvements in estimates of global stock status depend on improvements in the quality and use of data themselves.
 
@@
 However, efforts from numerous groups made up of highly skilled scientists tasked with solving this problem have all come up short, particularly when external efforts tried validate their performance [@free2020;@bouch2021;@pons2020].
-This is not intended as an indictment of those past efforts, but a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone.
+This is not intended as an indictment of those past efforts, but, we believe, a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone.
 Methods tested in those papers have been demonstrated to perform relatively well when given data or priors informative to specific fisheries.
@@
 For this reality to change our results though would require that the methods tested here have massively lower bias and higher accuracy for unassessed fisheries than RLSADB stocks.
-@free2020's simulation based based results of these same types of models suggests that this is unlikely to be true.
+@free2020's simulation based results of these same types of models suggests that this is unlikely to be true.
 However, while many assessed fisheries have mechanisms to actively manage catches, catch histories in unassessed and unmanaged fisheries may be more likely to reflect underlying stock status than those in heavily regulated fisheries.
@@
 Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
-Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
+Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of *F/F~MSY~* of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
+But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
 Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas of the globe results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
-
-Our value-of-information analysis also shows the high utility of having access to even a recent snapshot of F/F~MSY~ (Fig.\@ref(fig:voi-plot)).
+We found that performance of different data sources varied widely both within and among regions (Fig.\@ref(fig:cs-plot), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
+Some of this variation is likely simply due to low sample sizes of assessed stocks reported in the RLSADB in some regions.
+But, other explanations for differences in performance among stocks and regions may be an interesting area for future research.
+For example, it may be that stocks in some regions are more suitable to being represented by surplus production models than others.
+
+Our value of information analysis also shows the high utility of having access to even a recent snapshot of *F/F~MSY~* (Fig.\@ref(fig:voi-plot)).
 Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct fishery-specific priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
@@
 Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/)(https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
-However, out value of information exercise indicates that we may have to wait many years for new surveys to provide substantial improvements in status estimates (Fig.\@ref(fig:voi-plot)).
+However, our value of information exercise indicates that we may have to wait many years for new surveys to provide substantial improvements in status estimates (Fig.\@ref(fig:voi-plot)).
 
@@
 Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets.
-Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally [@costello2020].
-Catch-only methods may be a useful starting point when applied properly, but the source of their information must be clearly communicated, and achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.
+Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally [@costello2020] Achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.
 
@@
 +---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Fisheries Management Index [@melnychuk2017] | FMI              | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
+| Fisheries Management Index [@melnychuk2017] | FMI              | Priors on most recent *F/F~MSY~* values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
 +---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Swept Area Ratio [@amoroso2018]             | SAR              | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
+| Swept Area Ratio [@amoroso2018]             | SAR              | Priors on most recent *F/F~MSY~* values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
 +---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Reconstructed effort data [@rousseau2019b]  | effort           | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors.
+| Reconstructed effort data [@rousseau2019]  | effort           | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors.
 Assumed rate of technology creep reported in individual sections |
@@
 MPE = median percent error (bias), MAPE = median absolute percent error (error), Accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
-Performance is judged relative to reported values in RAM Legacy Stock Assessment Database.",
+Performance is judged relative to *B/B~MSY~* reported values in RAM Legacy Stock Assessment Database.",
               format = "pipe")
@@
 Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-Panels ordered in descending (starting from top left) mean accuracy at the FAO region level
+Panels ordered in descending (starting from top left) mean accuracy at the FAO region level.
 

### 2021-06-18 functions/fit-case-studies.R

@@ -50,7 +50,7 @@ fit_case_studies <-
     }
     
     
-    # fit heuristic
+    # fit sar_fmi
     
     sar_fmi_driors <-
       format_driors(
@@ -85,6 +85,39 @@ fit_case_studies <-
       filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy","depletion")) %>%
       mutate(year = rep(years, each = 4)) %>%
       mutate(data = "SAR & FMI")
+    
+    # fit naive sraplus
+    
+    basic_driors <-
+      format_driors(
+        taxa = scientific_name,
+        catch = catches,
+        years = seq_along(years),
+        initial_state = NA,
+        initial_state_cv = 0.2,
+        terminal_state = NA,
+        terminal_state_cv =  0.05,
+        use_heuristics = TRUE,
+        b_ref_type = "k",
+        use_b_reg = FALSE
+      )
+    
+    basic_fit <-
+      fit_sraplus(
+        driors = basic_driors,
+        include_fit = TRUE,
+        model = model,
+        engine = "sir",
+        estimate_shape = estimate_shape,
+        estimate_qslope = estimate_qslope,
+        estimate_proc_error = estimate_proc_error
+      )
+    
+    
+    basic_fit <- basic_fit$results %>%
+      filter(variable %in% c("b_div_bmsy", "u_div_umsy", "c_div_msy","depletion")) %>%
+      mutate(year = rep(years, each = 4)) %>%
+      mutate(data = "Catch Heuristic")
     
     # effort fit
     
@@ -160,7 +193,8 @@ fit_case_studies <-
     }
     
     results <- cpue_fit_results %>%
-      bind_rows(sar_fmi_fit) %>% {
+      bind_rows(sar_fmi_fit) %>% 
+      bind_rows(basic_fit) %>% {
         if (cmsy_worked){
           bind_rows(., cmsy_results)
         } else {

### 2021-06-18 make-assessing-global-fisheries.R

@@ -896,8 +896,8 @@ tmp2 <- tmp2 %>%
 
 ex_performance <- tmp2 %>% 
   group_by(variable, data) %>% 
-  summarise(r2 = round(yardstick::rsq_vec(truth, value),2)) 
-
+  summarise(r2 = round(yardstick::rsq_vec(truth, value),2),
+            rmse = round(yardstick::rmse_vec(truth, value),2)) 
 
 ex_scatter_plot <- tmp2 %>% 
   ggplot(aes(truth, value)) + 
@@ -905,7 +905,7 @@ ex_scatter_plot <- tmp2 %>%
   geom_hline(aes(yintercept = 0)) +
   geom_abline(aes(slope = 1, intercept = 0),linetype = 2) +
   geom_point(alpha = 0.5, size = 2) + 
-  ggtext::geom_richtext(data = ex_performance, aes(x = 1, y = 4, label = paste0("R<sup>2</sup> = ",r2))) +
+  ggtext::geom_richtext(data = ex_performance, size = 3,alpha = 0.75,aes(x = 1.25, y = 3.5, label = paste0("RMSE = ",rmse))) +
   facet_grid(variable ~ data, scales = "free_x")  + 
   # scale_size(trans = "sqrt", name = "Lifetime Catch") +
   scale_x_continuous(name = "RLSADB Value", expand = expansion(add = c(0, .1))) + 
@@ -1997,7 +1997,6 @@ assess_ramu_fits <- assess_ramu_fits %>%
 write_rds(assess_ramu_fits, path = file.path(results_path,"assess_ramu_fits.rds"))
 
 
-
 # assess_ram_fits %>% 
 #   ggplot(aes(pmin(5,ram_b_v_bmsy),pmin(5,mean))) + 
 #   geom_hex(binwidth = c(0.33, 0.33), color = "white") + 
@@ -2354,7 +2353,6 @@ ram_b_map_plot <- combo %>%
   theme(legend.position = "top",
         panel.background = element_rect(fill = "white"))
 
-browser()
 
 # ----save-plots----------------------------------------------------------
 

### 2021-06-18 documents/reviewer-replies.Rmd
version sent out for comments from authors

---
title: "Reviewer Replies"
author: "Dan Ovando"
date: "6/16/2021"
output: 
  bookdown::word_document2: default
bibliography: ["../references.bib"]
csl: fish-and-fisheries.csl
---

We thank both reviewers for their insightful and helpful comments. We have incorporated all of them into our revised manuscript unless otherwise stated, and we feel the suggestions have greatly improved the paper.

We include specific replies to individual comments below. Our responses to comments are noted in *italics* following each comment. We have not included individual responses to all editorial comments (e.g. missing punctuation) but have resolved each of those instances pointed out by the reviewers.

We hope the reviewers feel that our responses satisfy their concerns, but are happy to resolve any outstanding issues as required.

# Reviewer 1

## Performance metrics

The stock status of a fish stock is commonly evaluated in terms of biomass and fishing mortality. The authors present all performance metrics for the regional fishery assessments (MPE, MAPE, and accuracy) only in terms of B/BMSY . For the case study, the authors present the stock status in terms of both B/BMSY and F/FMSY . Is there a reason why the authors considered only B/BMSY for the regional fishery assessments? I would be interested in all performance metrics in terms of F/FMSY and recommend to include this quantity for the regional fishery assessments at least for the supporting information.

*Good question. We focused on B/B~MSY~ as this project was originally designed with an eye towards improving the SOFIA status estimates, which consider B/B~MSY~, not F/F~MSY~. However, F/F~MSY~ is clearly of interest as well, particularly as catch-only model F/Fmsy outputs are often used for assessment or projection, in the manner of @costello2016a.*

*We added F/F~MSY~ performance metrics summarized in the same type of table and figures used in the main document to the Supporting Information (we do not include in the main text due to limitations on numbers of available figures and tables). We also briefly mention these results in the results and discussion section. Broadly, we find similar patterns of generally poor and variable performance, with the notable exception that while in terms of B/B~MSY~ the default settings of the version of CMSY used in this paper performed in the middle of the pack in most cases, in terms of F/F~MSY~ CMSY was last in terms of performance across every metric, due to severe positive biases in F/F~MSY~ .*

## Discussion of Results

Why did the authors not apply sraplus without any additional data (SAR, FMI) for the case studies? This would constitute a fairer comparison of the comparison of CMSY and sraplus.

*We have included the requested comparison in the case study section. Under these settings, which we call "Catch Heuristic", sraplus uses the same catch heuristics as CMSY to generate priors on terminal depletion. The only real difference between the two algorithms then is the presence of a prior on midpoint depletion in CMSY, and the use the prior-predictive tuning regime in sraplus. The prior-predictive tuning regime ensures that in the absence of new data to update the priors, the "posterior" more or less match the explicit priors on terminal depletion. This is why it produces two bands of stock status corresponding to the binary values proscribed by the catch-heuristic. CMSY "updates" those priors to produce the join prior-predictive distribution of the life history priors and explicit depletion priors conditional on the model and catch history.  We did not originally include this run in the case study set as our contention is not that the sraplus algorithm is an inherently better catch-only mousetrap than CMSY (though we feel the addition of the prior-predictive tuning procedure is an important option).But, we agree that inclusion of this run in the case studies highlights an important philosophical choice of the default settings of sraplus (that we should be explicit in where our prior in terminal depletion is coming from).*

The results reveal highly variable results and no consistent patterns among regions. Did the authors find any patterns in the assessment performance regarding the life-history parameters of the stocks? This aspect is currently lacking from the discussion.

*Good question. We have not explored this idea empirically, but we now mention it explicitly in the discussion as an area for future research.*

Similarly for the case studies, I wonder if the authors found any common characteristics for the stocks with larger biases?

*One reason that we identified and more clearly articulate in the results that the effort data wasn't helpful was that construction of a CPUE index based on fishery catches and large-scale estimated effort resulted in very low CPUE values during quota or market constrained periods in a fishery's history. During these periods, according to the CPUE index abundance was very low, but RLSADB reports relatively high biomass levels, causing the model to perform badly. We address this general idea of "why" of performance in the discussion now.*

Lastly, the conclusion that catch-only methods may be useful contradicts the findings of the case study and regional fishery assessments, which indicate a poor performance (high bias, low precision and low classification accuracy) of catch-only methods if no fishery-independent data or informative and unbiased depletion priors are available. If this is one of the main conclusions, the authors should elaborate in more detail what they imply with “when applied properly

*Both reviewers pointed this discrepancy out, and we agree. We have removed that statement, as well feel it contains too many nuances to be adequately covered in the concluding paragraph of the paper. The general point that there may be some uses for these methods is made in the discussion paragraph starting with "What quality of assessment..."*

## Specific Comments

*Thank you for the edits. All specific comments were accepted unless specifically noted below.*

l. 148pp. Why do you x the shape parameter rather than using the priors recommended by Thorson et al. (2012)?

*We believe the question here is why did we fix the shape parameter the mean values reported by @thorson2012, rather than taking those values as informative priors and estimating the shape parameter? We tried this, but it slowed the models down substantially, and often resulted in very poor fits / convergence given the lack of information in the data on this parameter. It also had a tendency to shoot the shape parameter into odd spaces, given the relatively large CVs around the values reported in @thorson2012. We have clarified this in the methods.*

l. 153 Please specify whether you follow Winker et al. (2018) defining BMSY/K = 0.25 (in your case) when BMSY/K < 0:25. If another assumptions is made, this needs to be specified.

*Good catch! In the results in the paper Bmsy / K is never below 0.25, though it comes close for Clupeiformes. However, you are correct that there was not a check for this possibility in the broader sraplus package, so we are adding that in in case users try and specify a shape parameter such that Bmsy/K < 0.25.*

*We have included the relevant table in the SI now.*

l. 169 Do you mean proc instead of obs? obs not defined at this point of the manuscript. Where is defined? It is mentioned in Table S1 as the ratio of process and observation error, but not mentioned in the main text. Consider lifting Table S1 to main text and add more explanation to on estimated parameters. Be specific whether proc is estimated or fixed. Are the priors of Table S2 used? Further, consider using Bo instead of B0 and define B0 as it can also be used as virgin biomass.

*Thank you for pointing these out. We have clarified the notation here and included expanded definitions in the text at the section in question. Unfortunately we do not have space to bring in another table to the main results, but we have stated explicitly how the default priors are used and pointed the readers to that table.*

Fig. 5 I am not sure that I understood the differences between the scenarios initial state known" and initial state unfished". Thus, I do not understand, why RMSE is higher for the scenario with the initial state unfished. Please provide more information about these scenarios.

*We have added the missing explanation of these states fo the Value of Information section of the methods.*

Fig. 2-4 Keep same order of models/scenarios for all figures.

*After consultation with the group of authors, we respectfully disagree with this suggestion and as such have left the figures ordered as they are. The order of the panels reflects the performance of each data source by the metric in question, with best performing in the top left and worst in the bottom right. We feel this is helpful to the reader to parse apart the performance across the different data types. However, to clarify this we have made the caption clearer. However if you feel that this still introduces needless confusion we can make the figures in the manner suggested.*

Table S2 Where does the default prior of 1.01 for m come from? Thorson et al. (2012) report an average prior of 1.478 across all taxa.

*Good catch, that was an relic of an older set of results, we have updated the table to state that priors were drawn from @thorson2012, and provided a separate table with the specific values from @thorson2012 used.*

# Reviewer 2

General comments

Terms like biomass and notation B can often be used to refer to different things, ie., total biomass, vulnerable biomass, or spawning biomass. A non-structured model such as the SRA used here doesn’t distinguish between these three, and given the model is scaled with catches, B really represents vulnerable biomass; i.e., K = unfished vulnerable biomass, and the estimate of B/BMSY reported by the SRA corresponds with VB/VBMSY in a structured model. As the authors point out, the assessments in the RLSADB are typically more structurally complex – presumably many of them age-structured – and I assume report B/BMSY as SB/SBMSY given spawning population is generally the focus for conservation outcomes. The relationship between SB/SBMSY and VB/VBMSY would depend on the selectivity pattern relative to maturity and level of fishing mortality (eg VB/VBMSY could go to 0 while SB/SBMSY > 0). From a quick analysis with some operating models I have on hand the ratio of VB/VBMSY and SB/SBMSY ranged from 0.7 to 1.6.
If it is true that the SRA is reporting VB/VBMSY and comparing it to SB/SBMSY from the RLSADB there will be an additional level of bias and variability. This issue applies more widely any time outputs from models with fundamental structure are compared, but I rarely see it acknowledged or discussed. I don’t expect this would make a significant difference to the results of this study, and I don’t think this is a flaw in the study, but I’m curious if the authors have any thoughts on this.*

*We thank the reviewer for raising this important point. We have raised this issue broadly in the discussion section ("Our fits to the RLSADB data provide..."). We agree that this dynamic could explain the improved but still error-prone estimates of stock status when the model was fit with an index of abundance drawn directly from RLSADB. We have used this helpful suggestion to more clearly highlight the role that our results play in potentially quantifying the barrier that model misspecification plays in attempting to provide global assessments of age structured populations with surplus production models.*

It seems strange to me that having information on the most recent B/BMSY doesn’t result in lower RMSE for the predicted B/BMSY from the most recent 5 years of data (Figure 5). Presumably B/BMSY wouldn’t change a great deal between years, so it seems counter-intuitive that F/FMSY complete results in a better prediction of B/BMSY than when most recent B/BMSY is known. Perhaps the authors could add a bit of text to explain this.

*This is indeed confusing, as we did not clearly state that the value of information calculation measures performance based on B/B~MSY~ over the most recent five years of the fishery. We chose this as the metric as we noticed that catch-only models on occasion fit the last year of B/B~MSY~ well but do so by getting the prior years completely wrong. So, we chose a 5 year window to assess the ability of different data sources to get the most recent 5 years broadly right. This is why having 5 years of F/Fmsy is better, all else being equal, than 1 year of B/Bmsy (and 1 year of B/Bmsy is roughly equivalent to one year of F/Fmsy given the catch). We have clarified this in the value of information section of the Methods. We apologize for the confusion, and are happy to adjust the methods if the reviewer feels it is needed.*

The paragraph beginning on line 379 is useful, but from my perspective is a little overly generous to catch-only methods. The claim on this paper could be defended more strongly than is implied here. It’s logical that catch-only methods *need* some other data (or assumptions which are essentially assuming data look a certain way) to provide reliable estimate of stock status. E.g., even in unmanaged fisheries, the relationship between catch history and B/BMSY would have to at least depend on the time the stock has been fished, and the relative fishing effort on the stock (e.g., highly targeted, by-catch, etc). This paper makes the important point that providing estimates of stock status for unassessed fisheries requires better data (and capacity building to analyze the data); that is, essentially a stock assessment. Short-cut methods are appealing, but have the danger of over-simplifying a complex reality and could be detrimental to both sustainability and livelihoods if they don’t reliably categorize fisheries.

*Both reviewer's pointed out that some of the comments in our discussion did not seem to reflect the broadly poor performance of catch-only models shown in our results. We agree, and in the paragraph highlighted by the reviewer have sought to make clearer that any improvements have to come from new data, stronger assumptions, or improved empirical models. We then make clearer that this has largely been tried and we would argue largely failed. We have also included a paraphrased version of the reviewer's closing sentiments in the opening paragraph of our discussion.*

Minor comments

*We thank the reviewer for these comments. Unless otherwise stated all have been addressed*

56 – is it possible to say what % of the world’s fisheries this 50% of catch represents? 
*Not precisely, but it's a small percent, which we now clarify in the text. The FAO reports roughly 20000 individual "stocks", at least as defined by the intersection of taxa, country, and major FAO region. That resolution will be too coarse for some stocks (e.g. tuna) but likely too fine for others. The RLSADB contains roughly 1000 stocks at the moment, though fewer of those have consistent timeseries of reference points. So, an optimistic assessment would be 5% of fisheries?*

169 – what are sigma obs and gamma? Described later on I think.

*This was indeed unclear, we have clarified the definition of terms and added a table of prior distributions to the SI*

224 – has the SIR algorithm been described above? Hasn’t been mentioned yet – do you mean SRA?
*Our mistake, yes we mean SRA there*

305 – methods said Thorson 2012

*We have clarified that this refers to the intrinsic growth rate r, which comes from @thorson2020, as opposed to the shape parameter, which we pull from @thorson2012.*

306 – R2 values don’t match Figure 1? Fig 1 says R2 = 0.03 for CMSY

*To keep things more consistent with the rest of the paper we have replaced the R2 values in the plot with root mean squared error (RMSE) values, and double checked that the values in the text match the figures.*

313 – only SAR & FMI fit better to F/FMSY right? The others had higher R2 (though too low to consider) for B/BMSY (Figure 1)

*That is correct, much clearer now with RMSE instead of R2*

332 – did you consider sub-setting some of the RLSADB assessments to only use the earlier years where management may not have been as effective and B/BMSY in the terminal year less optimistic?

*This is a very interesting idea. We did consider this, the challenge is we only have FMI/SAR data for recent years, and so would not have been able to use those values to predict status in the past, when presumably both values would have been very different. We have another paper in the works where we have explored this exact idea though and found that catch-only models performed similarly poorly on RLSADB stocks even when restricted to pre-1990 data. We briefly mention this in the discussion now.*

441 – 449 – It may be worth pointing out that developing management methods that do not explicitly rely on an estimate of stock status could be useful as well, eg spatial closures, size regulations, empirical index-based MPs

*A very important point that we have included in the discussion in the suggested section.*

# References

### 2021-06-18 documents/ovando-etal-assessing-global-fisheries-si.Rmd
version sent out for comments from authors

@@
   rbind(c("Ratio of process to observation error", "$\\gamma$", "$LN(.5,0.25)$")) %>%
-  rbind(c("Initial State", "$B/K_0$", "Posterior probability dist. of catch-based regressions"))
+  rbind(c("Initial State", "$B_0$", "Posterior probability dist. of catch-based regressions"))
 
@@
   row.names = F,
-  caption = "Name, abbreviations, and priors distribution for parameters potentially estimated by sraplus in this manuscript.",
+  caption = "Name, abbreviations, and priors distribution for parameters potentially estimated by sraplus in this manuscript. LN refers to log normal, where the mean is reported on the unit scale.",
   format = "pipe"
@@
 
-```{r si-prior-tab}
-cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$LN(10\\times{max(catch)},4)$") %>%
-  rbind(c("Growth rate", "r", "$LN(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
-  rbind(c("Shape parameter", "m", "Thorson et al. (2012)")) %>%
-  rbind(c("Catchability", "q", "$LN(1e^{-3}, 0.3)$")) %>%
-  rbind(c("Observation Error", "$\\sigma_{obs}$", "$LN(.05,2)$")) %>%
-  rbind(c("Process Error", "$\\sigma_{proc}$", "$LN(.05,0.5)$")) %>% 
-  rbind(c("Tech Creep", "$creep$", "$LN(0.025,0.2)$"))
-
-knitr::kable(
-  cd,
-  "latex",
-  align = "l",
-  booktabs = TRUE,
-  escape = F,
-  row.names = F,
-  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
-) %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
+\newpage
 
@@
 knitr::kable(
-  sraplus::ssb0msy_to_ssb0,
-  "latex",
+  sraplus::ssb0msy_to_ssb0 %>% rename(
+    "Taxanomic Group" = tax_group,
+    "SSBMSY/SSB0" = ssbmsy_to_ssb0,
+    "SSBMSY/SSB0 SD" = ssbmsy_to_ssb0_sd
+  ),
     align = "l",
-  booktabs = TRUE,
   escape = F,
   row.names = F,
-  caption = "Ratio of SSBmsy to SSB0 by taxinomic group from Thorson et al. (2012) used as default priors in sraplus."
-) %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
+  format = "pipe",
+  caption = "SSBMSY to SSB ratios from Thorson et al.
+(2012) used in the paper.
+Taxa not within the groups assigned at the genus level by Thorson et al.
+(2012) are assigned the ratio reported for 'Other'"
+)
 

### 2021-06-18 documents/ovando-etal-assessing-global-fisheries.Rmd
version sent out for comments from authors

@@
 Marine fisheries are one of the largest anthropogenic impacts in the oceans aside from the effects of climate change.
-When managed sustainably fisheries are a critical source of economic prosperity and food security around the globe.
+Fisheries are also a critical source of economic prosperity, cultural identity, and food security around the globe.
 As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries.
-While our understanding of the state of world fisheries has improved over the last decade, the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their population size relative to reference points [@hilborn2020; @fao2020].
+While our understanding of the state of world fisheries has improved over the last decade, the majority of the world's fish populations, making up roughly 50% of marine landings (though only a few percent of the total number of fisheries in the world), lack formal statistical assessments (stock assessments) of their population size relative to reference points [@hilborn2020; @fao2020].
 This is a major impediment to ensuring the sustainable development of the world's oceans.
@@
 priors on recent stock status, or an index of abundance) in order to produce estimates of the state of a fishery over time.
-The key goal of sraplus is not substantial improvements in model fitting methods per say, but providing the ability to easily incorporate multiple kinds of fishery data potentially used in SRA style analyses in a statistically rigorous manner.
+The key goal of sraplus is not substantial improvements in model fitting methods per say, but providing the ability to easily incorporate multiple kinds of fishery data potentially used in SRA-style analyses in a statistically rigorous manner.
 We paired various types of data with fisheries in RLSADB, and then used sraplus to generate predictions of *B/B~MSY~* in the most recent year of each fishery across different combinations of data types.
@@
 Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
-We deal with this problem by following the methods described in [@winker2018] to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
+We dealt with this problem by following the methods described in @winker2018 to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
 
@@
 All stock reduction analysis runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e^6^ candidate draws.
-For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{proc}$, and *B/K~0~* (initial depletion).
+For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r* (with a prior distribution drawn from Fishlife, @thorson2020), *K*, $\sigma_{proc}$, and *B~0~* (initial depletion, $B/K$).
 *q* is also estimated when needed.
-$\sigma_{proc}$ is estimated indirectly through the parameter $\gamma$, the ratio of process (*proc*) to observation (*obs*) error ($\sigma_{obs}$).
-See Table SX for prior distributions for each of these parameters.
+$\sigma_{proc}$ (process error) is estimated indirectly through the parameter $\gamma$, the ratio of process (*proc*) to observation (*obs*) error ($\sigma_{obs}$).
+See Table.S1 for prior distributions for each of these parameters.
 
@@
 
-$$log(a_t) \sim N(f_{pt}(r,K,m,B0,\pmb{p},\pmb{c}) \times q, \sigma_{obs})$$
+$$log(a_t) \sim N(f_{pt}(r,K,m,B_0,\pmb{p},\pmb{c}) \times q, \sigma_{obs})$$
 
@@
 
-$$log(cpue_t) \sim N(f_{pt}(r,K,m,B0,\pmb{p},\pmb{c}), \sigma_{obs})$$
+$$log(cpue_t) \sim N(f_{pt}(r,K,m,B_0,\pmb{p},\pmb{c}), \sigma_{obs})$$
 
@@
 This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-Borel's Paradox poses a particular problem when there are no data to fit to though (i.e.
-when the model is simply filtering through prior distributions in the manner of a traditional SRA).
-Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse.
+However, Borel's Paradox poses a particular problem when there are no data to fit to(i.e.
+when the model is simply filtering through prior distributions in the manner of a traditional SRA) due to the fact that there are more parameter combinations that allow for a fishery to be relatively unexploited than for a fishery to be close to collapse (but never actually collapsed, i.e.
+predicted biomass less than observed catch).
 In this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020).
@@
 Our solution amounts to a two-step sampling-importance-resampling (SIR) algorithm.
-We first run the standard SIR algorithm as described in the Estimation Model section of the methods.
+We first run the standard SRA algorithm as described in the Estimation Model section of the methods.
 We then break the resulting draws into bins based on terminal stock status, and calculate the mean probability density *p* (defined by the prior distributions of estimated parameters) of each bin.
@@
 Guess assigns each stock a random *B/B~MSY~* of 1.6, 1, or 0.4, corresponding to the middle value of the FAO status bins of underfished  ($B/B~MSY~ \geq 1.2$), maximally sustainably fished ($0.8 \geq B/B~MSY~ < 1.2$), overfished  ($0 \geq B/B~MSY~ < 0.8$).
-We performed a matching analysis measuring performance in estiamting *F/F~MSY~* as well, with results presented in the SI.
+We performed a matching analysis measuring performance in estimating *F/F~MSY~* as well, with results presented in the SI.
 
@@
 We then measured the value of information as the average reduction in root mean squared error (RMSE) in *B/B~MSY~* over the most recent five years of the fishery (in order to evaluate the ability of the model to capture recent trends as well as the most recent value), resulting from use of different kinds of data.
-We considered the value of information of having data on: the most recent *B/B~MSY~*, treating initial state of the population as one of unfished ($B/K_{0}=1$), known ($B/K_{0}$ taken from stock assessment), or estimated based on initial shape of the catch history, *F/F~MSY~* values in the most recent year, over the last five years, and over the complete time series, and an abundance index spanning the complete, most recent half, or most recent quarter of the time series.
+We considered the value of information of having data on: the most recent *B/B~MSY~*, treating initial state of the population as one of unfished ($B_{0}=1$), known ($B_{0}$ taken from stock assessment), or estimated based on initial shape of the catch history, *F/F~MSY~* values in the most recent year, over the last five years, and over the complete time series, and an abundance index spanning the complete, most recent half, or most recent quarter of the time series.
 We considered the value of information of a longer time series of *F/F~MSY~* than *B/B~MSY~* to consider the potential of augmenting catch-only models with additional data that can inform *F/F~MSY~*, such as length-composition data, that might be available over the history of the fishery.
@@
 Nearly all of the fisheries used in this case study have *F/F~MSY~* values less than one, and most have *B/B~MSY~* values greater than one.
-Both the catch heuristic implemented in sraplus and CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of *B/B~MSY~*) and the estimated resilience of the target species (based on Fishlife [@thorson2020]), badly miss this trend, predicting instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing.
-Adding priors informed by stock-specific swept area ratio and Fisheries Management Index data produces some improvements (RMSE below 1 relative to above 1 for all other case studies), capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~.
-Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit in *B/B~MSY~* and *F/F~MSY~*.
-This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) but high biomass values, creating a mismatch between the CPUE trends based on the effort and catch data and actual biomass.
-However none of the models fit to the four case-study data combinations were able to explain much of the variation in *B/B~MSY~*, though the models performed better in some cases for predicting *F/F~MSY~*.
-Models informed by the SAR and FMI values performed the best in terms of *B/B~MSY~* (RMSE = 0.89).
+Both the catch heuristic implemented in sraplus and CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of *B/B~MSY~*) and the estimated resilience of the target species used to provide a prior on *r* (based on Fishlife, @thorson2020), badly miss this trend, predicting instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing (Fig.\@ref(fig:cs-plot)).
+Overall, models informed by the SAR and FMI values performed the best in terms of *B/B~MSY~* (RMSE = 0.89).
 The ranking of performance by data sources used was the same for *F/F~MSY~*, though the range of values was much higher, with the SAR and FMI based result producing an RMSE of 0.39 while the catch-only case studies produced *F/F~MSY~* RMSE values greater than 1 (Fig.\@ref(fig:cs-plot)).
 However, this improved performance of the SAR and FMI data is to be expected as the prior-generating models were trained on data from RLSADB, and therefore this is likely an optimistic assessment of the performance of these prior-generating models when applied to new fisheries.
+Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit in *B/B~MSY~* and *F/F~MSY~*.
+This is likely because a number of stocks in this case study set have very low catches (relative to historic highs) but high biomass values, creating a mismatch between the CPUE trends based on the effort and catch data and actual biomass.
+However none of the models fit to the four case-study data combinations were able to explain much of the variation in *B/B~MSY~*, though the models performed better in some cases for predicting *F/F~MSY~*.
 
@@
 The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RLSADB were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "RLSADB Index"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "RLSADB Index"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "RLSADB Index"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
+This exercise tells us that given sufficiently high quality index of abundance, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
 
@@
 While in some cases the addition of globally available data such a FMI scores, SAR values, or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RLSADB-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
+The simplicity and low data requirements of catch-only models are understandably appealing to many users, but our results highlight that without high-quality local data these methods can provide highly imprecise and biased results.
+Broad and uncritical application of these methods that frequently result in incorrect classification of stock status can be detrimental to both ecosystems and livelihoods.
+
+<!-- That use of none of the datasets collected here can achieve that, and that our tests fitting directly to abundance indices from RLSADB itself suggests that model misspecification is not the primary cause of poor performance, tells us that improvements in estimates of global stock status will likely depend on improvements in the quality and use of data themselves.
+-->
+
+Our claim is that achieving substantial progress in assessing the state of global fisheries will require improved data and capacity building.
+This claim cannot be definitively proven; it is possible that some yet-to-be discovered model will provide a reliable means of dramatically improving the accuracy of estimates produced by catch-based models using current broadly-available datasets.
+Basic logic tells us though that such a model would have to depend on strong assumptions, new data, or empirical estimation.
+Catch is a function of catchability, effort, and biomass, so given just catch we have one equation and three unknowns.
+Structurally separating out biomass (and subsequently estimating reference points etc.) requires then either data or strong assumptions on these other variables across a sufficiently representative timespan.
+Alternatively, empirical relationships could be developed linking attributes of catch histories to observed estimates of stock status [e.g.
+@costello2012b].
+However, these methods require that these empirical relationships both exist and are reliably preserved in fisheries that lack formal assessments.
+Numerous groups made up of highly skilled scientists have attempted to resolve these fundamental challenges in various ways, but external efforts to validate the performance of these methods have consistently found them to perform generally poorly when applied in bulk [@free2020;@bouch2021;@pons2020].
+This is not intended as an indictment of those past efforts, but, we believe, a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone.
+Methods tested in those papers have been demonstrated to perform relatively well when given specific data or priors informative to specific fisheries.
+We revisited this task in this paper, and found that even when augmented with other broadly available datasets, we were unable to produce consistently accurate estimates of stock status around the globe.
+For these reasons we believe that improvements to our understanding of the state of global unassessed fisheries will come from targeted use of existing but underutilized data, and expanding the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
 
@@
 
-<!-- That use of none of the datasets collected here can achieve that, and that our tests fitting directly to abundance indices from RLSADB itself suggests that model misspecification is not the primary cause of poor performance, tells us that improvements in estimates of global stock status will likely depend on improvements in the quality and use of data themselves.
--->
-
-Our claim is that achieving substantial progress in assessing the state of global fisheries will require improved data and capacity building.
-This claim cannot be definitively proven; It is possible that some yet-to-be discovered model will provide a reliable means of dramatically improving the accuracy of estimates produced by models using current broadly-available datasets, such as catch-only style methods.
-Should that occur we will be happy to have been proven wrong.
-However, efforts from numerous groups made up of highly skilled scientists tasked with solving this problem have all come up short, particularly when external efforts tried validate their performance [@free2020;@bouch2021;@pons2020].
-This is not intended as an indictment of those past efforts, but, we believe, a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone.
-Methods tested in those papers have been demonstrated to perform relatively well when given data or priors informative to specific fisheries.
-We revisited this task in this paper, and found that even when augmented with other broadly available datasets, we were unable to produce consistently accurate estimates of stock status around the globe.
-For these reasons we believe that improvements to our understanding of the state of global unassessed fisheries will come from targeted use of existing data, and expanding the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
+Our fits to the RLSADB data provide a useful diagnostic of the degree to which model mispecification might explain the poor performance of catch-only models.
+Catch-only models generally employ some form of biomass dynamics model, such as the Schaefer surplus production model [@schaefer1954].
+These models are generally much simpler than the typically age-structured population models underpinning the assessments in the RLSADB.
+While we found that fitting the Pella-Tomlinson model employed in sraplus to an index of abundance drawn directly from the RLSADB provided much improved and reasonably accurate estimates of stock status relative to catch-only approaches (Table.\@ref(tab:perf-tab)), even with the index of abundance sraplus still had an average MAPE of `r scales::percent(perf$mape[perf$'Data Used' == "RLSADB Index"])`.
+This discrepancy may provide a measure of relatively insurmountable error resulting from model misspecification that inclusion of additional data may not be able to overcome.
+For example, many assessments in the RLSADB calculate *B/B~MSY~* based on spawning stock biomass.
+However, surplus production models condition the size and state of the population on the catch, which is by definition a measure of the vulnerable biomass.
+Vulnerable biomass may not reliably track spawning biomass in highly selective fisheries, potentially resulting in model biases as the surplus production model is not capable of separately tracking spawning and vulnerable biomass.
 
@@
 @free2020's simulation based results of these same types of models suggests that this is unlikely to be true.
-However, while many assessed fisheries have mechanisms to actively manage catches, catch histories in unassessed and unmanaged fisheries may be more likely to reflect underlying stock status than those in heavily regulated fisheries.
+However, while many assessed fisheries have mechanisms to actively manage catches, catch histories in unassessed and unmanaged fisheries may be more likely to reflect underlying stock status than those in relatively well managed fisheries, though any predictive relationship will still depend on a host of factors such as changes in effort.
+This could be empirically tested by considering the performance of catch-only models on data from earlier in the history of fisheries in the RLSADB, though we did not do so here as we do not have historic estimates of FMI or SAR values.
 
@@
 Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
-Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of *F/F~MSY~* of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.
+Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of *F/F~MSY~* of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)` (see SI).
 But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
@@
 Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments.
-Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/)(https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
+Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) (https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
 However, our value of information exercise indicates that we may have to wait many years for new surveys to provide substantial improvements in status estimates (Fig.\@ref(fig:voi-plot)).
@@
 Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
+We may also need to acknowledge that in some fisheries reliable estimates of stock status relative to MSY based reference points are simply not possible (or perhaps desirable), and instead rely on more precautionary or empirical management management measures such as spatial closures, size restrictions, and indicator based harvest strategies (ideally tested through management strategy evaluation) [@prince2019;@dowling2015;@fulton2016].
 

### 2021-06-23 documents/ovando-etal-assessing-global-fisheries.Rmd
jim edits and upating equation numbering

@@
 output: 
+  bookdown::pdf_document2:
+    latex_engine: xelatex
   bookdown::word_document2:
     reference_docx: template.docx
-  bookdown::pdf_document2:
-    latex_engine: xelatex
 params:
@@
 **Authors**:
-Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^
+Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James T. Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^
 
@@
 \begin{equation}
-  p_t \sim e^{N\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
+  log(p_t) \sim {N}\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)
 (\#eq:procerror)
@@
 The model fits to a direct estimate of abundance (e.g.
-a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood is
-
-$$log(a_t) \sim N(f_{pt}(r,K,m,B_0,\pmb{p},\pmb{c}) \times q, \sigma_{obs})$$
-
-where $a_t$ is the observed abundance index and $f_{pt}$ is the Pella-Tomlinson production model (Equation.\@ref(eq:pt)).
+a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood calculated per
+
+\begin{equation}
+  B_t = f_{t}(r,K,m,B_0,\pmb{p},\pmb{c})
+(\#eq:biomass)
+\end{equation}
+
+where $B_t$ is the biomass in time *t*, given $f_{t}$ which is shorthand for the result of the Pella-Tomlinson equation in time step *t* given parameters and a vector of catches.
+The likelihood of the abundance index *a* in time *t* is then calculated as
+
+\begin{equation}
+ log(a_t) \sim N(B_t \times q, \sigma_{obs})
+(\#eq:likelihood)
+\end{equation}
+
 When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data.
@@
 When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
-This is due to the fact that in reality the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the realized fishing mortality rate must be between between 0% and 100%).
+This is due to the fact that in reality the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the realized exploitation fraction, sometimes called *u*, is between 0 and 1).
 To accommodate this, we generate a catch per effective harvest rate index of abundance, as
 
-$$cpue_t = \frac{c_t}{(1 - e^{-F_t})}$$
-
-$$F_t = q_tE_t$$
+\begin{equation}
+ cpue_t = \frac{c_t}{(1 - e^{-F_t})}
+(\#eq:cpue)
+\end{equation}
+
+\begin{equation}
+ F_t = q_tE_t
+(\#eq:f)
+\end{equation}
 
@@
 
-$$q_t = q_{t-1} \times (1  + \tau)$$
+\begin{equation}
+ q_t = q_{t-1} \times (1  + \tau)
+(\#eq:qt)
+\end{equation}
 
@@
 
-$$log(cpue_t) \sim N(f_{pt}(r,K,m,B_0,\pmb{p},\pmb{c}), \sigma_{obs})$$
+\begin{equation}
+ log(cpue_t) \sim N(f_{pt}(r,K,m,B_0,\pmb{p},\pmb{c}), \sigma_{obs})
+(\#eq:cpue)
+\end{equation}
 
@@
 
-$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$
+\begin{equation}
+ p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}
+(\#eq:pbin)
+\end{equation}
 
@@
 
-$$p(n_i) = \frac{p(bin_i)}{N_i}$$
+\begin{equation}
+ p(n_i) = \frac{p(bin_i)}{N_i}
+(\#eq:ppcor)
+\end{equation}
 
@@
 Structurally separating out biomass (and subsequently estimating reference points etc.) requires then either data or strong assumptions on these other variables across a sufficiently representative timespan.
-Alternatively, empirical relationships could be developed linking attributes of catch histories to observed estimates of stock status [e.g.
-@costello2012b].
+Alternatively, empirical relationships could be developed linking attributes of catch histories to observed estimates of stock status [@costello2012b;@thorson2012a], or through identifying a structural model for the evolution of fishing mortality in managed and unmanaged fisheries [@thorson2013].
 However, these methods require that these empirical relationships both exist and are reliably preserved in fisheries that lack formal assessments.

### 2021-06-26 documents/ovando-etal-assessing-global-fisheries.Rmd
merging final version back to main

@@
 
-<!-- data limited are not just messier versions of data rich  -->
-
-<!-- 80% in trouble  -->
-
-<!-- # rishi notes -->
-
-<!-- * more assertive on what the data needs really are -->
-
-<!--   * where to apply globally  -->
-
-<!-- * For places with less coverage, what would we actually recommend there  -->
-
-<!-- * what do we know about catchabiliy -->
-
-<!-- - e.g. 51, fleets became industrial  -->
-
-<!--   - hammer down on that's the issue with the effort  -->
-
-<!--   - how do we translate effort into f bey area -->
-
-<!--   - here's  -->
-
-<!-- https://www.nature.com/natsustain/about/content -->
-
-<!-- # check in call notes -->
-
-<!-- Ray: setting the stage: what we want is a way to assess the status of stocks and see how we're doing relative to SDG. -->
-
-<!-- We need to discuss a sampling strategy: how can we get down to a set of priority stock. A step of expert knowledge to refine what stocks you would focus in on. -->
-
-<!-- Regressions based approach, focusing in effort on stocks -->
-
-<!-- Basically a paragraph laying out that we need to do this, and discussing how we might go about doing that. Equal weight to each stock vs proportional to biomass -->
-
-<!-- add in a bit more to the discussion on guidelines -->
-
-<!-- add in error in the index for the VOI analysis -->
-
-<!-- Apples and apples comparisons between model and design developments.
-How much is the value of the effort data relative to actually improving the population model.
-Production model is ill-suited to some of these questions.
--->
-
-<!-- fishing effort uncertainty of the effort -->
-
-<!-- how credible is the information -->
-
-<!-- probably don't need to bother with the SAUP -->
-
-<!-- stock assessments -->
-
-<!-- the stock will collapse next year -->
-
-<!-- nansen data -->
-
-<!-- # Publication Requirements -->
-
-<!-- Target Journal: [Nature Sustainability](https://www.nature.com/natsustain/about/content) -->
-
-<!-- 3500 words main text -->
-
-<!-- 50 references (main text) -->
-
-<!-- 3000 words (max) methods -->
-
-<!-- # Significance Statement  {.unnumbered} -->
-
-<!-- <!-- Authors must submit a 120-word-maximum statement about the significance of their research paper written at a level understandable to an undergraduate-educated scientist outside their field of specialty.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
--->
-
-<!-- The health of ocean ecosystems and the well-being of millions of people depend on our ability to accurately estimate the state of marine fish populations.
-However, these assessments typically require extensive amounts of high-quality data that are not available for large numbers of fisheries that are critical sources of economic and nutritional benefits.
-Many studies have tried to estimate the state of these unassessed fisheries based solely on the history of catches reported for a fishery, but we show here that these catch-based estimates produce unreliable results.
-We demonstrate that improving our understanding of global fisheries will depend on redoubled efforts at data collection and capacity building.
--->
-
 # Title Page {.unnumbered}
@@
 Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added.
-Estimates of fish populations based primarily on catch histories alone on average performed 29% better than a random guess.
+Estimates of fish populations based primarily on catch histories on average performed 25% better than a random guess.
 But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time.
+Withing these broad summaries the performance of models trained on our tested data sources varied widely across regions.
 Effective improvement in estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.
@@
 
-We then fit to the index of abundance per
+We then fit to the index of abundance per Equation 3 and
 
 \begin{equation}
- log(cpue_t) \sim N(f_{pt}(r,K,m,B_0,\pmb{p},\pmb{c}), \sigma_{obs})
+ log(cpue_t) \sim N(B_t, \sigma_{obs})
 (\#eq:cpue)
@@
 Each point is a stock in the RLSADB.
-Black dashed line shows the 1:1 relationship."}
+Black dashed line shows the 1:1 relationship.
+Text displys root mean squared error (RMSE) of each panel."}
 ex_scatter_plot + 
@@
 
-ggsave(here("documents","figs","cs-plot.pdf"),ex_scatter_plot)
+ggsave(here("documents","figs","cs-plot.png"),ex_scatter_plot, height = 6, width = 6)
 
@@
 Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RLSADB index data were able to capture the overall state or geographic distribution of stock status represented in RLSADB in a consistent manner.
-Performance in estimating *F/F~MSY~ * was similarly variable and poor, with the exception that the default settings of CMSY performed much more consistently poorly in terms of *F/F~MSY~* than *B/B~MSY~* (due to systemic overestimating of *F/F~MSY~*, see SI).
+Performance in estimating *F/F~MSY~ * was similarly variable and poor, with the exception that the default settings of CMSY performed much more consistently poorly in terms of *F/F~MSY~* than *B/B~MSY~* (due to systemic overestimation of *F/F~MSY~*, see Supplementary Information [SI]).
 
@@
 ```{r mpe-map, fig.width=8, fig.height=6, fig.cap="Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-Data source panels are ordered in ascending (starting from top left) mean MPE at the FAO region level.
+Data source panels are ordered in ascending total bias at the FAO region level.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 
-ggsave(here("documents","figs","mpe-map.pdf"), ram_mpe_map_plot, width = 8, height = 6)
+ggsave(here("documents","figs","mpe-map.png"), ram_mpe_map_plot, width = 8, height = 6)
 ```
@@
 ```{r mape-map, fig.cap = "Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-Data source panels are ordered in descending (starting from top left) mean MAPE at the FAO region level.
+Data source panels are ordered in ascending (starting from top left) mean MAPE at the FAO region level.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 
-ggsave(here("documents","figs","mape-map.pdf"),ram_mape_map_plot, width = 8, height = 6)
+ggsave(here("documents","figs","mape-map.png"),ram_mape_map_plot, width = 8, height = 6)
 
@@
 ```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
-Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region level RLSADB Index refers to catch and abundance index drawn from RLSADB.
+Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region level.
+RLSADB Index refers to catch and abundance index drawn from RLSADB.
 Effective CPUE refers to an index of abundance based on reconstructed effort data.
@@
 
-ggsave(here("documents","figs","acc-map.pdf"),ram_acc_map_plot, width = 8, height = 6)
+ggsave(here("documents","figs","acc-map.png"),ram_acc_map_plot, width = 8, height = 6)
 
@@
 
-ggsave(here("documents","figs","voi-plot.pdf"),voi_plot, width = 8, height = 6)
+ggsave(here("documents","figs","voi-plot.png"),voi_plot, width = 8, height = 6)
 
@@
 However, these methods require that these empirical relationships both exist and are reliably preserved in fisheries that lack formal assessments.
-Numerous groups made up of highly skilled scientists have attempted to resolve these fundamental challenges in various ways, but external efforts to validate the performance of these methods have consistently found them to perform generally poorly when applied in bulk [@free2020;@bouch2021;@pons2020].
-This is not intended as an indictment of those past efforts, but, we believe, a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone.
+Numerous groups made up of highly skilled scientists have attempted to resolve these fundamental challenges in various ways, but external efforts to validate the performance of these methods have consistently found them to perform generally poorly when applied in bulk [@free2020;@bouch2021;@pons2020;@sharma2021].
+This summary is not intended as a critique of those past efforts, but a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone.
 Methods tested in those papers have been demonstrated to perform relatively well when given specific data or priors informative to specific fisheries.
@@
 We thank participants of the FAO "Methods for Global Assessment" workshop held in Rome, Italy February 2019 for helpful feedback on this project.
+We also thank A.
+Hordyk and one anonymous reviewer for their helpful comments.
 The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of the FAO, NOAA or the Department of Commerce.
@@
 \newpage
+
 # Tables {.unnumbered}
@@
 +---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-
-: (\#tab:data) Data sources included across model fits.
+:(\#tab:data) Data sources included across model fits.
 
@@
 
-Figure 1: RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%.
-Each point is a stock.
+Figure 1: RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], catch heuristics, priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort (Effort) trends assuming a rate of technological increase of 2.6%.
+Each point is a stock in the RLSADB.
 Black dashed line shows the 1:1 relationship.
+Text displys root mean squared error (RMSE) of each panel.
 
 Figure 2: Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+Data source panels are ordered in ascending (starting from top left) total bias at the FAO region level.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-Panels ordered in ascending (starting from top left) mean MPE at the FAO region level.
 
 Figure 3: Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
+Data source panels are ordered in ascending (starting from top left) mean MAPE at the FAO region level.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-Panels ordered in descending (starting from top left) mean MAPE at the FAO region level
 
 Figure 4: Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
+Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region level.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-Panels ordered in descending (starting from top left) mean accuracy at the FAO region level.
 
@@
 
-\newpage
-
-# Supporting Information {.unnumbered}
-
-```{r si, child = 'ovando-etal-assessing-global-fisheries-si.Rmd'}
-```
+<!-- \newpage -->
+
+<!-- # Supporting Information {.unnumbered} -->
+
+<!-- ```{r si, child = 'ovando-etal-assessing-global-fisheries-si.Rmd'} -->
+<!-- ``` -->

### 2021-06-26 README.Rmd
Accepted version (#13)
* fixed new FAO areas mapping problem

* Minor revisions

* updating README

--- 2021-06-26 documents/ovando-etal-assessing-global-fisheries.Rmd
@@
 ---
-title: Improving Estimates of the State of Global Fisheries <br> Depends on Better Data
-date: "`r Sys.Date()`"
-bibliography: ["../references.bib"]
-csl: fish-and-fisheries.csl
-output: 
-  bookdown::word_document2:
-    reference_docx: template.docx
-  bookdown::pdf_document2:
-    latex_engine: xelatex
-params:
-  results_name: ["v1.0"]
-  min_years_catch: [25]
-always_allow_html: true
-linkcolor: blue
-header-includes:
-  - \usepackage{setspace}\doublespacing
-  - \usepackage{lineno}\linenumbers
-toc: false
+output: github_document
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
-                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
-library(tidyverse)
-library(viridis)
-library(sf)
-library(hrbrthemes)
-library(extrafont)
-library(scales)
-library(patchwork)
-library(rstanarm)
-library(bayesplot)
-library(here)
-library(sraplus)
-extrafont::loadfonts()
+<!-- README.md is generated from README.Rmd. Please edit that file -->
 
-functions <- list.files(here::here("functions"))
-
-purrr::walk(functions, ~ source(here::here("functions", .x)))
-
-results_name <- params$results_name
-
-results_path <- here("results",results_name)
-
-min_years_catch <- params$min_years_catch
-
-prepare_sofia_data(lookup_fmi_names = FALSE)
-
-ram_comp_data <- ram_data %>%
-  mutate(has_things = !(is.na(catch) | catch == 0)) %>%
-  filter(has_things) %>%
-  group_by(stockid) %>%
-  mutate(delta_year = as.integer(year - lag(year))) %>%
-  mutate(delta_year = case_when(year == min(year) ~ as.integer(1),
-                                TRUE ~ delta_year)) %>%
-  mutate(missing_gaps = any(delta_year > 1)) %>%
-  filter(missing_gaps == FALSE) %>%
-  group_by(stockid) %>%
-  mutate(n = length(catch)) %>%
-  filter(n >= min_years_catch) %>%
-  ungroup()
-
-ram_comp_data <- ram_comp_data %>%
-  select(
-    year,
-    stockid,
-    country,
-    primary_FAOarea,
-    scientificname,
-    commonname,
-    year,
-    catch,
-    b_v_bmsy,
-    u_v_umsy,
-    total_biomass
-  ) %>%
-  rename(scientific_name = scientificname) %>% 
-  mutate(fao_area_code = as.numeric(primary_FAOarea)) %>% 
-  left_join(fao_species, by = "scientific_name")
-
-assess_ram_fits <- read_rds(file = file.path(results_path,"assess_ram_fits.rds"))
-
-assess_ramu_fits <- read_rds(file = file.path(results_path,"assess_ramu_fits.rds"))
-
-ram_v_sraplus_area <- read_rds(file = file.path(results_path,"ram_v_sraplus_area.rds"))
-
-ram_comp_data <- read_rds(file = file.path(results_path,"ram_comp_data.rds"))
-
-fao_sraplus_comp_data <- read_rds(file = file.path(results_path, "fao_sraplus_comp_data.rds"))
-
-load(file = file.path(results_path, "paper_plots.Rdata"))
-
-load(file = file.path(results_path, "voi_fits.RData"))
-
-  exs <- read_rds(file = file.path(results_path, "raw-case-study-fits.rds"))
-
-breaks <- c(0.8, 1.2)
-
-breaks <- c(0, breaks, Inf)
-
-labels <- c("over", "fully", "under")
-
-# rand_fits <- assess_ram_fits %>%
-#   filter(data == "ram-data") %>%
-#   mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
-#          data = "guess")
-
-perf <- assess_ram_fits %>% 
-  # bind_rows(rand_fits) %>% 
-  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
-  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
-  mutate(data = as.factor(data)) %>% 
-  filter(data != "u_umsy") %>% 
-  mutate(data = forcats::fct_recode(data, 
-                                    "RLSADB Index" = "ram-data",
-                                    "SAR" = "sar",
-                                    "FMI" = "fmi",
-                                    "Effective CPUE" = "cpue",
-                                    "Effective CPUE+" = "cpue-plus",
-                                    "Nominal CPUE" = "nominal-cpue",
-                                    "Nominal CPUE+" ="nominal-cpue-plus",
-                                    "CMSY" = "cmsy",
-                                    "Guess" = "guess",
-                                    "RLSADB F/Fmsy" = "u_umsy"
-
-                                    )) %>% 
-  group_by(data) %>% 
-  summarise(
-    mpe = median((mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
-    mape = median(abs(mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
-    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
-  ) %>% 
-  arrange((mape)) %>% 
-  rename("Data Used" = "data")
-
-uperf <- assess_ramu_fits %>% 
-  # bind_rows(rand_fits) %>% 
-  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
-  mutate(ram_bin = cut(ram_u_v_umsy, breaks = breaks, labels = labels)) %>%
-  mutate(data = as.factor(data)) %>% 
-  filter(data != "u_umsy") %>% 
-  mutate(data = forcats::fct_recode(data, 
-                                    "RLSADB Index" = "ram-data",
-                                    "SAR" = "sar",
-                                    "FMI" = "fmi",
-                                    "Effective CPUE" = "cpue",
-                                    "Effective CPUE+" = "cpue-plus",
-                                    "Nominal CPUE" = "nominal-cpue",
-                                    "Nominal CPUE+" ="nominal-cpue-plus",
-                                    "CMSY" = "cmsy",
-                                    "Guess" = "guess",
-                                    "RLSADB F/Fmsy" = "u_umsy"
-
-                                    )) %>% 
-  group_by(data) %>% 
-  summarise(
-    mpe = median((mean - ram_u_v_umsy ) / ram_u_v_umsy),
-    mape = median(abs(mean - ram_u_v_umsy ) / ram_u_v_umsy),
-    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
-  ) %>% 
-  arrange((mape)) %>% 
-  rename("Data Used" = "data")
-
-fao_area_codes <- fao %>% 
-  ungroup() %>% 
-  select(fao_area, fao_area_code) %>% 
-  unique()
-
-pub_theme <- theme_ipsum(base_size = 10,
-                         axis_text_size = 10,
-                         axis_title_size = 12) +
-    theme(panel.spacing = unit(.5,"lines"))
-
-theme_set(pub_theme)
-
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
 ```
 
-# Title Page {.unnumbered}
+#  Improving Estimates of the State of Global Fisheries Depends on Better Data Data
+## Fish and Fisheries (Link TBD)
 
-**Title 1**: Improving Estimates of the State of Global Fisheries Depends on Better Data
+  - Daniel Ovando
+  - Ray Hilborn
+  - Cole Monnahan
+  - Merrill Rudd
+  - Rishi Sharma
+  - James Thorson
+  - Yannick Rousseau
+  - Yimin Ye
 
-**Title 2**: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
-
-**Running Title**: Unassessed Fisheries
-
-**Authors**:
-Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James T. Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^
-
-**Affiliations**:
-
-^1^University of Washington, School of Aquatic and Fishery Sciences
-1122 NE Boat St, Box 355020 Seattle, WA 98195-5020
-Seattle, WA, USA 98103
-
-^2^NOAA Fisheries Alaska Fisheries Science Center, Seattle, WA, USA
-
-^3^Scaleability LLC
-
-^4^Food and Agriculture Organization of the United Nations, Fisheries and Aquaculture,Rome, IT
-
-^5^University of Tasmania, Institute for Marine and Antarctic Studies,Hobart, TAS, AUS
-
-**Authorship**: DO, RH, CM,MR, RS, JT designed model structure and ran analyses. RS, YR, and YE supplied data. All authors contributed to writing of the manuscript
-
-**Conflict of Interest**: RH receives research funding from many groups that have interests in fisheries outcomes including environmental NGOs, foundations, governments and fishing industry groups.
-
-\newpage
-
-# Abstract {.unnumbered}
+## Abstract
 
@@
 But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time.
-Withing these broad summaries the performance of models trained on our tested data sources varied widely across regions.
-Effective improvement in estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.
+Within these broad summaries the performance of models trained on our tested data sources varied widely across regions.
+Effective improvement in estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.s exploited fish populations depends on prioritizing the collection of high-priority
 
-# Introduction {.unnumbered}
-
-The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".
-Marine fisheries are one of the largest anthropogenic impacts in the oceans aside from the effects of climate change.
-Fisheries are also a critical source of economic prosperity, cultural identity, and food security around the globe.
-As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries.
-While our understanding of the state of world fisheries has improved over the last decade, the majority of the world's fish populations, making up roughly 50% of marine landings (though only a few percent of the total number of fisheries in the world), lack formal statistical assessments (stock assessments) of their population size relative to reference points [@hilborn2020; @fao2020].
-This is a major impediment to ensuring the sustainable development of the world's oceans.
-In this paper, we consider why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.
-
-What do we currently know about the state of fished marine species around the world?
-The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries.
-The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished (at or near targets for sustainable food production), 6.2% are underfished, and 34.2% are overfished [@fao2020].
-Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments summarized in the RAM Legacy Stock Assessment Database (RLSADB) [@ricard2012].
-The "assessed" fisheries in RLSADB represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the state of assessed fish populations around the globe.
-
-That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed".
-While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world.
-The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020].
-While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods.
-With the surge in data availability, such as global assessments of management strength [@melnychuk2017], trawl footprints [@amoroso2018], and fishing effort [@rousseau2019], the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.
-In addition, the SOFIA process assigns unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to the biomass that would on average result in maximized yield over the long-term.
-
-<!-- We evaluate the ability of alternative assessment models to improve our estimate of the population status of unassessed fish stocks around the world.  -->
-
-Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide numerical estimates of the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].
-Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (@free2020 and references therein).
-These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being a common measure of stock status), from characteristics of a fishery's catch history.
-
-<!-- These issues become apparent when we consider some of the macro-level predictions made by these models.
-While the estimates of *B/B~MSY~* and other reference points reported in RLSADB are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RLSADB to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model.
--->
-
-<!-- @costello2016a found similar rankings of regions in terms of stock status as RLSADB, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison.
-The superensemble method [@anderson2017a] used in @rosenberg2018 demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific.
-Simply comparing catch divided by maximum catch, in the manner of @pauly2007, finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)).
-The non-RLSADB results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status.
-However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly exploited regions as measured by RLSADB and the FAO resulting from these catch-only efforts at assessing global fisheries is concerning.
--->
-
-```{r status-tab, include=FALSE, eval=FALSE}
-
-o_by_a <- sofia_status %>% 
-  left_join(fao_area_codes, by = "fao_area_code") %>% 
-  group_by(fao_area) %>% 
-  summarise(`FAO % Overfished` = percent(mean(status == "O")))
-
-ss <-
-  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
-                    sheet = "state_space_results_08_28_2018") %>%
-  janitor::clean_names()
-
-ss_lookup <-
-  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
-                    sheet = "ss_to_fao") %>%
-  janitor::clean_names() %>% 
-  mutate(tmp = map2(region, fao_area_codes, ~data.frame(region = .x, fao_area_code = as.integer(str_split(.y, pattern = ',',simplify = TRUE))))) %>% 
-  select(tmp) %>% 
-  unnest(cols = tmp) %>% 
-    left_join(fao_area_codes, by = "fao_area_code")
-
-state_space_results <- ss_lookup %>% 
-  left_join(ss, by = "region") %>% 
-  filter(variable == "BvB") %>% 
-  group_by(fao_area) %>% 
-  filter(year == max(year)) %>% 
-  summarise(`B/BMSY RLSADB` = round(mean(dlm_geomean),2)) %>% 
-  ungroup()
-
-# ram_by_region %>% 
-#   ggplot(aes(b_v_bmsy)) +
-#   geom_histogram() + 
-#   facet_wrap(~fao_area)
-
-tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
-  select(-`B/BMSY RLSADB`) %>%  
-  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
-  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
-  arrange(`FAO % Overfished`) %>% 
-    rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
-  select(`FAO Area`,`SOFIA % Overfished`,`B/BMSY RLSADB`, everything())
-
- tmp %>% 
-  knitr::kable(caption = "Estimates of fish stock status by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
-               Rosenberg et al. 2018, Catch / max(Catch)",
-               booktabs = TRUE,
-               format = "pipe") %>% 
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-
-```
-
-However, @free2020 demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*.
-While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating the state of the individual unassessed fisheries around the world.
-The types of global assessment efforts share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure.
-Under this assumption, we can obtain better estimates of the state of unassessed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries.
-However, further evaluation of the predictions made by these models [@free2020;@bouch2021;@pons2020] as well as works such as @pauly2013 and @branch2011 call into question this very idea.
-
-In this paper we ask, how much can estimates from catch-only assessments be improved by moving beyond catch-only and augmenting catch histories with additional sources of broadly available fisheries data?
-We answer this question by aggregating a set of broadly distributed datasets that might be of use to global fishery assessment.
-We then used our new assessment method, sraplus, to evaluate the performance status estimates derived from different combinations of these broadly available data.
-We compared these results with estimates from the increasingly utilized (e.g.
-@palomares2020) default settings of the catch-only version of CMSY described in @froese2017.
-We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis, not on incremental improvements to assessment methods based solely on catch histories.
-
-# Methods {.unnumbered}
-
-The basis of this analysis is a new stock assessment software package we call [sraplus](https://github.com/DanOvando/sraplus) (https://github.com/DanOvando/sraplus).
-sraplus is an extension of stochastic stock reduction analysis (SRA) [@kimura1984;@walters2006], which allows users to combine a biomass dynamics model with a variety of data sources (e.g.
-priors on recent stock status, or an index of abundance) in order to produce estimates of the state of a fishery over time.
-The key goal of sraplus is not substantial improvements in model fitting methods per say, but providing the ability to easily incorporate multiple kinds of fishery data potentially used in SRA-style analyses in a statistically rigorous manner.
-We paired various types of data with fisheries in RLSADB, and then used sraplus to generate predictions of *B/B~MSY~* in the most recent year of each fishery across different combinations of data types.
-We then compared the predictions generated by sraplus using different types of data, along with predictions for the same fisheries generated by CMSY, to the reported values in RLSADB.
-
-All analysis were conducted in the R programming language [@rcoreteam2019].
-Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018].
-The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).
-Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.
-
-## Data Sources {.unnumbered}
-
-At its most "data-limited", sraplus can work as a catch-only model in the manner of @froese2017.
-These catch histories can then be augmented with prior information on stock status, derived from expert opinion, or using built-in prior generating models based on increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases.
-The FMI database provides self-reported scores reflecting the strength of fisheries management in many fisheries around the world, which sraplus uses to construct informative priors on stock status, e.g.
-placing a higher probability that fisheries with high FMI scores have better stock status than those with low FMI scores.
-SAR is a measure of the intensity of trawl fishing within a particular area, which sraplus uses to construct an informative prior on the magnitude of fishing mortality.
-While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries.
-As an alternative, @rousseau2019 present a global reconstruction of fishing effort for countries around the world.
-We pair these effort reconstructions (both nominal effort and effective effort assuming an annual rate of efficiency increase of 2.6%) with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status.
-In this way sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).
-
-<!-- In this paper we ask, can augmenting the types of catch-only models employed in global assessments to date with new broadly available data improve our understanding of the state of global fish populations?
-We use a flexible stock assessment software package, [sraplus](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include.
--->
-
-<!-- ## Experiment Structure {.unnumbered} -->
-
-<!-- We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database.
--->
-
-<!-- ## Data Sources {.unnumbered} -->
-
-<!-- We restricted ourselves to databases with wide coverage around the world, representing the types of information that can be leveraged in efforts to assess the status of global fisheries (Table.\@ref(tab:data)).
--->
-
-## Population Model {.unnumbered}
-
-The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of @winker2018.
-While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the data types evaluated here will support.
-
-<!-- The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status -->
-
-The population growth equation is
-
-\begin{equation}
-f(x)=\begin{cases}
-B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)p_t
-, & \text{if $B_t>0.25 \times K$}.\\
-B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)\right)p_t, & \text{otherwise}.
-\end{cases}
-(\#eq:pt)
-\end{equation}
-
-Where $B_t$ is biomass at time *t*, *K* is carrying capacity ,*r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of *B~MSY~/K* to shift.
-When *m* = 2, *B~MSY~ / K* = 0.5.
-Lower values of *m* shift the production function left, higher values right.
-The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, @thorson2012 provides estimates of the ratio of *B~MSY~* to *K* for many fish taxa.
-For each stock we fix the shape parameter based on the distributions reported in @thorson2012 for the genus of the species in question.
-We chose to fix the shape parameter at the mean stock-appropriate values from @thorson2012 rather than estimating the shape parameter with an informative prior since there is so little information regarding the shape in the data considered.
-Attempts to estimate the shape parameter with priors from @thorson2012 frequently resulted in poor model performance.
-$\pmb{c}$ is a vector of catches, and $\pmb{p}$ is vector of process errors.
-Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model.
-We dealt with this problem by following the methods described in @winker2018 to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.
-
-We allow for process error $p_t$ (in the manner of the stochastic stock reduction analysis (SRA) suggested by @walters2006). Process error $p_t$ is assumed to be log-normally distributed, such that
-
-<!-- This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average.
-Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).
--->
-
-\begin{equation}
-  log(p_t) \sim {N}\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)
-(\#eq:procerror)
-\end{equation}
-
-where *N* is the normal distribution.
-
-## Estimation Model {.unnumbered}
-
-All of our estimates are Bayesian in nature.
-sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo with the No-U-Turn sampler [@hoffman2011].
-Unless there is an abundance index to fit to, the model runs as an SRA.
-A stock reduction analysis works by specifying prior distributions on population parameters and, critically, the recent state of the fishery.
-sraplus allows users to specify the most recent status in units of depletion, *B/B~MSY~*, *F*, or *F/F~MSY~*.
-We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history.
-Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected.
-The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
-All stock reduction analysis runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e^6^ candidate draws.
-For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r* (with a prior distribution drawn from Fishlife, @thorson2020), *K*, $\sigma_{proc}$, and *B~0~* (initial depletion, $B/K$).
-*q* is also estimated when needed.
-$\sigma_{proc}$ (process error) is estimated indirectly through the parameter $\gamma$, the ratio of process (*proc*) to observation (*obs*) error ($\sigma_{obs}$).
-See Table.S1 for prior distributions for each of these parameters.
-
-<!-- .
-We can break the use of sraplus into two distinct categories: with data to fit to and without.
-By "data", we refer to measurements which are used to confront model estimates within a likelihood function.
-In our context, these include fishery-independent survey data, or a CPUE index.
-When there are no data to fit to, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters).
-Under this mode, the model is essentially a stock-reduction analysis model, in the manner of [@walters2006], in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e.
-results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches.
-This step updates the prior distribution of population parameters by eliminating combinations of priors that are impossible for a given catch history and a specified functional form.
--->
-
-<!-- The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).  -->
-
-<!-- *r* and *K* are the only two parameters that are always estimated.
-Estimation of every other parameter can be turned on or off.
-When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user.
--->
-
-<!-- sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent).
-Unless there is an abundance index to fit to the model runs as a stock reduction analysis.
-A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery.
-sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~.
-We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history.
-Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected.
-The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status.
-Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis.
-All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.
--->
-
-When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in Stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018].
-By default the model uses 2000 draws with a 1000 step warm-up and one chain.
-Any detailed fit for an individual fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria.
-The model fits to a direct estimate of abundance (e.g.
-a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood calculated per
-
-\begin{equation}
-  B_t = f_{t}(r,K,m,B_0,\pmb{p},\pmb{c})
-(\#eq:biomass)
-\end{equation}
-
-where $B_t$ is the biomass in time *t*, given $f_{t}$ which is shorthand for the result of the Pella-Tomlinson equation in time step *t* given parameters and a vector of catches.
-The likelihood of the abundance index *a* in time *t* is then calculated as
-
-\begin{equation}
- log(a_t) \sim N(B_t \times q, \sigma_{obs})
-(\#eq:likelihood)
-\end{equation}
-
-When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data.
-@rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep).
-Treating this raw effort data as the denominator in the CPUE calculation assumes that every increase in fishing effort translates to a commensurate increase in fishing mortality.
-When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population.
-This is due to the fact that in reality the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the realized exploitation fraction, sometimes called *u*, is between 0 and 1).
-To accommodate this, we generate a catch per effective harvest rate index of abundance, as
-
-\begin{equation}
- cpue_t = \frac{c_t}{(1 - e^{-F_t})}
-(\#eq:cpue)
-\end{equation}
-
-\begin{equation}
- F_t = q_tE_t
-(\#eq:f)
-\end{equation}
-
-Where $q_t$ can has a technology creep component $\tau$
-
-\begin{equation}
- q_t = q_{t-1} \times (1  + \tau)
-(\#eq:qt)
-\end{equation}
-
-We then fit to the index of abundance per Equation 3 and
-
-\begin{equation}
- log(cpue_t) \sim N(B_t, \sigma_{obs})
-(\#eq:cpue)
-\end{equation}
-
-### CMSY {.unnumbered}
-
-In addition to the results from sraplus, we include a set of results produced by the default settings of the CMSY method [@froese2017].
-For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>.
-The only modification made is to convert the underlying population model to C++ for faster computation.
-For each stock we used all the default options and priors provided and generated by CMSY, in the same manner as @palomares2020, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012].
-Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.
-
-<!-- In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation.
-The core different between CMSY and sraplus as utilized here then is that when we refer to CMSY we mean CMSY run using the default heuristics used by the algorithm to construct priors on initial, intermediate, and terminal stock status.
-In contrast, sraplus fit using for example FMI data uses the empirical models described here to generate priors on stock status.
-In addition, all sraplus runs use a prior-predictive tuning procedure to resolve issues related to Borel's paradox, described below.
--->
-
-## Priors {.unnumbered}
-
-<!-- We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus. -->
-
-### Prior Predictive Tuning {.unnumbered}
-
-<!-- We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.
-Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.
--->
-
-<!-- Suppose that the only thing we observe from a fishery is a catch history.
-Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.
--->
-
-In the absence of any data to fit to, sraplus works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history, life history priors, and model structure.
-This creates a problem for the Bayesian nature of our analysis.
-Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*.
-Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step.
-Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See @poole2000 and references therein for a discussion of Borel's Paradox in a fisheries context).
-
-This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, Borel's Paradox poses a particular problem when there are no data to fit to(i.e.
-when the model is simply filtering through prior distributions in the manner of a traditional SRA) due to the fact that there are more parameter combinations that allow for a fishery to be relatively unexploited than for a fishery to be close to collapse (but never actually collapsed, i.e.
-predicted biomass less than observed catch).
-In this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020).
-This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the posterior distributions of stock status resulting from the SRA algorithm.
-
-We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000].
-Our solution amounts to a two-step sampling-importance-resampling (SIR) algorithm.
-We first run the standard SRA algorithm as described in the Estimation Model section of the methods.
-We then break the resulting draws into bins based on terminal stock status, and calculate the mean probability density *p* (defined by the prior distributions of estimated parameters) of each bin.
-
-\begin{equation}
- p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}
-(\#eq:pbin)
-\end{equation}
-
-We then divide the mean probability density of bin *i* evenly among each of the draws within that bin *n*
-
-\begin{equation}
- p(n_i) = \frac{p(bin_i)}{N_i}
-(\#eq:ppcor)
-\end{equation}
-
-And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.
-
-The net result of this is that it allows users to place an explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior.
-While the range of possible life history values supplied still influences stock status under this approach, this prior predictive tuning process makes the resulting priors more consistent with explicit priors on recent stock status supplied by the user.
-Users can turn this functionality off and instead base priors on stock status primarily on life history.
-See Supplementary Information for a detailed explanation of this problem and our solution.
-
-### Priors Informed by Outside Data {.unnumbered}
-
-Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus.
-We paired data on catch histories, swept area ratio, and Fisheries Management Index with estimates of stock status from the RLSADB.
-We then trained a regression of the general form $log(status) \sim N(variable,\sigma)$ for each of these three data types.
-Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries.
-For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on *B/B~MSY~* or *F/F~MSY~* in the most recent year of the fishery usable by sraplus (See Supplementary Information).
-All prior regression models were tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
-The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.
-For all results presented in this paper we used these data to provide priors on *F/F~MSY~*, as we found clearer predictive relationships and subsequent model performance between catch, FMI, and SAR values and *F/F~MSY~* than we did for *B/B~MSY~*.
-
-<!-- Given the scope of this analysis, we do not claim that the regressions used here are the best possible model relating these data with the fishery status indicators of interest.
-Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.
--->
-
-## Assessing Performance {.unnumbered}
-
-Simulation testing can be preferable in many ways to comparison against model outputs.
-However it is not possible to simulation test the value of information derived from empirical relationships between variables such as fisheries management strength and fishery outcomes, which our study depends on.
-As such we assess model performance through comparison to best available estimates of stock status available in RLSADB.
-We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RLSADB, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history.
-*B/B~MSY~* values from RLSADB are themselves estimates, not data, but they are the best available information on global stock status.
-We then paired the catch histories for these RLSADB stocks with regional-level SAR, FMI, and effort data.
-Our methods approximated a regional-level assessment exercise, where data beyond catch histories are available at regional levels, but not for specific fisheries.
-We also estimated *B/B~MSY~* values of our candidate RLSADB stocks by using sraplus to fit to an abundance index drawn directly from RLSADB itself "RLSADB Index", as a measure of the ability of models like sraplus if given perfect information
-
-We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with CMSY, and a set of runs fit to the RLSADB Index (Table.\@ref(tab:dat-desc)).
-We assessed model performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy.
-Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).
-This experiment structure allows us to assess the ability of catch-only models to estimate the *B/B~MSY~* values reported in RLSADB, and evaluate how much these estimates can be improved by augmenting the catch histories with additional sources of information.
-Our benchmark model is a simple "Guess".
-Guess assigns each stock a random *B/B~MSY~* of 1.6, 1, or 0.4, corresponding to the middle value of the FAO status bins of underfished  ($B/B~MSY~ \geq 1.2$), maximally sustainably fished ($0.8 \geq B/B~MSY~ < 1.2$), overfished  ($0 \geq B/B~MSY~ < 0.8$).
-We performed a matching analysis measuring performance in estimating *F/F~MSY~* as well, with results presented in the SI.
-
-## Value of Information Calculations {.unnumbered}
-
-We performed a value of information assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries.
-The value of information analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RLSADB, and comparing the estimated values to the values reported in RLSADB.
-There are too many combinations for us to run the full expansion of possible parameter states.
-To resolve this we generated fits for 3000 combinations of a RLSADB stock and available data.
-For any one draw, we randomly sampled a RLSADB stock and a list of available data and data quality.
-For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance.
-The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.
-We then measured the value of information as the average reduction in root mean squared error (RMSE) in *B/B~MSY~* over the most recent five years of the fishery (in order to evaluate the ability of the model to capture recent trends as well as the most recent value), resulting from use of different kinds of data.
-We considered the value of information of having data on: the most recent *B/B~MSY~*, treating initial state of the population as one of unfished ($B_{0}=1$), known ($B_{0}$ taken from stock assessment), or estimated based on initial shape of the catch history, *F/F~MSY~* values in the most recent year, over the last five years, and over the complete time series, and an abundance index spanning the complete, most recent half, or most recent quarter of the time series.
-We considered the value of information of a longer time series of *F/F~MSY~* than *B/B~MSY~* to consider the potential of augmenting catch-only models with additional data that can inform *F/F~MSY~*, such as length-composition data, that might be available over the history of the fishery.
-
-## Case Study {.unnumbered}
-
-We also ran a case study demonstrating how different kinds of data led to different conclusions about stock status.
-We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores.
-We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock.
-As a benchmark, we first estimated stock status for these case study fisheries using the default settings of the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available.
-We then used sraplus to generate estimate of stock status based solely on catch heuristics, taking into account the prior-predictive tuning implemented by default in sraplus.
-We also used stock-specific data on SAR and FMI to generate priors on *F/F~MSY~* for each of the stocks, which were then passed to sraplus.
-Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.
-This case study represents a more "localized" assessment, where external data sources (FMI and SAR) are available at the stock level, rather than at the regional level.
-We summarized the performance of each model fit in the case study based on root mean squared error (RMSE).
-
-<!-- To determine what sources of data might provide the greatest value in improving our estimates of global stock status, we fit 3,000 sraplus models to randomly sampled fisheries from RLSADB, each time varying the kind and quality of data made available to the model, and what parameters the model estimated.
-We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).
--->
-
-# Results {.unnumbered}
-
-## Case Study {.unnumbered}
-
-Nearly all of the fisheries used in this case study have *F/F~MSY~* values less than one, and most have *B/B~MSY~* values greater than one.
-Both the catch heuristic implemented in sraplus and CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of *B/B~MSY~*) and the estimated resilience of the target species used to provide a prior on *r* (based on Fishlife, @thorson2020), badly miss this trend, predicting instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing (Fig.\@ref(fig:cs-plot)).
-Overall, models informed by the SAR and FMI values performed the best in terms of *B/B~MSY~* (RMSE = 0.89).
-The ranking of performance by data sources used was the same for *F/F~MSY~*, though the range of values was much higher, with the SAR and FMI based result producing an RMSE of 0.39 while the catch-only case studies produced *F/F~MSY~* RMSE values greater than 1 (Fig.\@ref(fig:cs-plot)).
-However, this improved performance of the SAR and FMI data is to be expected as the prior-generating models were trained on data from RLSADB, and therefore this is likely an optimistic assessment of the performance of these prior-generating models when applied to new fisheries.
-Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit in *B/B~MSY~* and *F/F~MSY~*.
-This is likely because a number of stocks in this case study set have very low catches (relative to historic highs) but high biomass values, creating a mismatch between the CPUE trends based on the effort and catch data and actual biomass.
-However none of the models fit to the four case-study data combinations were able to explain much of the variation in *B/B~MSY~*, though the models performed better in some cases for predicting *F/F~MSY~*.
-
-<!-- This is not in and of itself a surprising result, but demonstrates how 1) estimates of stock status based on catch histories alone can be both imprecise and biased and how 2) augmenting catch histories with additional data from the fisheries in question can produce meaningful improvements.
--->
-
-<!-- `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model.
-We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~.
-Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.
--->
-
-```{r cs-plot, fig.cap="RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], catch heuristics, priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort (Effort) trends assuming a rate of technological increase of 2.6%.
-Each point is a stock in the RLSADB.
-Black dashed line shows the 1:1 relationship.
-Text displys root mean squared error (RMSE) of each panel."}
-ex_scatter_plot + 
-  theme(strip.text = element_text(size = 8))
-
-ggsave(here("documents","figs","cs-plot.png"),ex_scatter_plot, height = 6, width = 6)
-
-```
-
-<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise.
-In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process.
-Since the SAR and FMI regressions were trained on the RLSADB stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.
--->
-
-## Performance of Regional Fishery Assessments {.unnumbered}
-
-We first considered the performance of sraplus when the model was given a perfect index of abundance from RLSADB (RLSADB Index) for each stock in RLSADB it was tasked with assessing.
-Assessment models in RLSADB are typically much more structurally complex than the simple biomass dynamics used in sraplus, and so this "perfect information" test tells us how much of a penalty in model performance we are likely to pay due to model misspecification alone.
-The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RLSADB were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "RLSADB Index"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "RLSADB Index"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "RLSADB Index"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-This exercise tells us that given sufficiently high quality index of abundance, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.
-
-<!-- Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves.  -->
-
- <!-- However, this is somewhat an artifact of the data.
-The status of most stocks in RLSADB is relatively good, with recent *B/B~MSY~* values generally near one.
-This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise.
--->
-
-We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status beyond those derived from catch-only methods.
-Many of the datasets used produced similar levels of bias as the RLSADB data (e.g.
-FMI, SAR, Nominal CPUE fits), though notable the "Guess" method actually performed near the top in terms of bias.
-However, this is somewhat an artifact of the data.
-The status of most stocks in RLSADB is relatively good, with recent *B/B~MSY~* values generally near one.
-As the mean value of the "Guess" method is one, on average the "Guess" model is an unbiased but imprecise measure of stock status in RLSADB (Table.\@ref(tab:perf-tab)).
-
-Focusing on MAPE (our measure of error rather than bias) and classification accuracy, the error of the models jumps dramatically as soon as data other than the RLSADB abundance indices are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`.
-The mean accuracy of the sraplus models across all non-RLSADB data fits was `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]))`.
-Note that there are only three bins in the FAO stock status classifications, and as such our "Guess" model has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "Guess"])`.
-This means that the accuracy of our models designed as a proxy for a global assessment process were across all non-RLSADB index data fits `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]) / perf$accuracy[perf$"Data Used" == "Guess"] - 1)` more accurate than a random guess, certainly an improvement, but on average assigned fisheries to the wrong FAO status bin `r percent(1 - mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]))` of the time.
-
-Looking geographically we found a similar pattern of a rapid decrease in performance for models besides those fit to the RLSADB Index.
-Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions.
-Models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) (Fig.\@ref(fig:mape-map)).
-We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)).
-Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RLSADB index data were able to capture the overall state or geographic distribution of stock status represented in RLSADB in a consistent manner.
-Performance in estimating *F/F~MSY~ * was similarly variable and poor, with the exception that the default settings of CMSY performed much more consistently poorly in terms of *F/F~MSY~* than *B/B~MSY~* (due to systemic overestimation of *F/F~MSY~*, see Supplementary Information [SI]).
-
-<!-- xx dig into this, which stocks are in that region? XX -->
-
-```{r mpe-map, fig.width=8, fig.height=6, fig.cap="Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-Data source panels are ordered in ascending total bias at the FAO region level.
-RLSADB Index refers to catch and abundance index drawn from RLSADB.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-"}
-ram_mpe_map_plot
-
-ggsave(here("documents","figs","mpe-map.png"), ram_mpe_map_plot, width = 8, height = 6)
-```
-
-```{r mape-map, fig.cap = "Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-Data source panels are ordered in ascending (starting from top left) mean MAPE at the FAO region level.
-RLSADB Index refers to catch and abundance index drawn from RLSADB.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6}
-ram_mape_map_plot
-
-ggsave(here("documents","figs","mape-map.png"),ram_mape_map_plot, width = 8, height = 6)
-
-```
-
-```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
+*Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
 Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region level.
@@
 2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6}
-
-ram_acc_map_plot
-
-ggsave(here("documents","figs","acc-map.png"),ram_acc_map_plot, width = 8, height = 6)
-
+Guess assigns a random recent B/B~MSY~ of 0.4,1, or 1.6.*
+```{r, echo = FALSE}
+knitr::include_graphics("documents/figs/acc-map.png")
 ```
 
-\newpage
+# Reproducing Results
 
-<!-- The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RLSADB, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.
--->
+All materials needed to reproduce our results and manuscript are contained in this repository. In order to reproduce
 
-## Value of Information {.unnumbered}
+1. Fork the repository and clone to your machine
 
-Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available.
-Interestingly though, having access to only an accurate estimate of *F/F~MSY~* in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself  (Fig.\@ref(fig:voi-plot)).
-While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average, due to the lack of historical context for recent trends in abundance.
+2. Open R and set your working directory of the cloned repository (or just use RStudio projects)
 
-<!-- We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].
--->
+3.
+This project is set up with [`renv`](https://rstudio.github.io/renv/articles/renv.html) to manage package dependencies.
+Inside R (and with your working directory set correctly) run `renv::restore()`.
+This will install the correct versions of all the packages needed to replicate our results.
+Packages are installed in a stand-alone project library for this paper, and will not affect your installed R packages anywhere else.
 
-```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error (RMSE) of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
-Distribution is full posterior probability distribution.
-Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
-Change is relative to the mean performance of a catch-only heuristic model."}
-
-voi_plot <- b_voi_plot + 
-  geom_vline(aes(xintercept = 0), alpha = 0.75)
-
-voi_plot
-
-ggsave(here("documents","figs","voi-plot.png"),voi_plot, width = 8, height = 6)
-
-```
-
-#  Discussion {.unnumbered}
-
-<!-- XX somewhere in here make clear that you're not learning anything from the catch data XX -->
-
-<!-- XX Acknolwedge the need for action the the frustration of being told "get more data" XX -->
-
-<!-- Global-level assessment are critical for guiding management agendas for the world's oceans.
-Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world.
-Numerous efforts have sought to resolve this problem through development of new modeling methods.
-These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment.
-However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes.
-While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.
--->
-
-<!-- Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RLSADB -->
-
-Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals.
-The hope of efforts using catch-only models to estimate the status of global unassessed fisheries is fundamentally that we can learn something meaningful about the state of a fished population simply by knowing something about its catch history and life history.
-While in some cases the addition of globally available data such a FMI scores, SAR values, or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RLSADB-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).
-The simplicity and low data requirements of catch-only models are understandably appealing to many users, but our results highlight that without high-quality local data these methods can provide highly imprecise and biased results.
-Broad and uncritical application of these methods that frequently result in incorrect classification of stock status can be detrimental to both ecosystems and livelihoods.
-
-<!-- That use of none of the datasets collected here can achieve that, and that our tests fitting directly to abundance indices from RLSADB itself suggests that model misspecification is not the primary cause of poor performance, tells us that improvements in estimates of global stock status will likely depend on improvements in the quality and use of data themselves.
--->
-
-Our claim is that achieving substantial progress in assessing the state of global fisheries will require improved data and capacity building.
-This claim cannot be definitively proven; it is possible that some yet-to-be discovered model will provide a reliable means of dramatically improving the accuracy of estimates produced by catch-based models using current broadly-available datasets.
-Basic logic tells us though that such a model would have to depend on strong assumptions, new data, or empirical estimation.
-Catch is a function of catchability, effort, and biomass, so given just catch we have one equation and three unknowns.
-Structurally separating out biomass (and subsequently estimating reference points etc.) requires then either data or strong assumptions on these other variables across a sufficiently representative timespan.
-Alternatively, empirical relationships could be developed linking attributes of catch histories to observed estimates of stock status [@costello2012b;@thorson2012a], or through identifying a structural model for the evolution of fishing mortality in managed and unmanaged fisheries [@thorson2013].
-However, these methods require that these empirical relationships both exist and are reliably preserved in fisheries that lack formal assessments.
-Numerous groups made up of highly skilled scientists have attempted to resolve these fundamental challenges in various ways, but external efforts to validate the performance of these methods have consistently found them to perform generally poorly when applied in bulk [@free2020;@bouch2021;@pons2020;@sharma2021].
-This summary is not intended as a critique of those past efforts, but a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone.
-Methods tested in those papers have been demonstrated to perform relatively well when given specific data or priors informative to specific fisheries.
-We revisited this task in this paper, and found that even when augmented with other broadly available datasets, we were unable to produce consistently accurate estimates of stock status around the globe.
-For these reasons we believe that improvements to our understanding of the state of global unassessed fisheries will come from targeted use of existing but underutilized data, and expanding the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
-
-What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends on the needs of those using the assessment outputs.
-It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficiently accurate.
-Catch-only based estimates of measures such as MSY are likely to be much more robust than reference points such as *B/B~MSY~* [@martell2013].
-Where catch-only methods are used for estimates of *B/B~MSY~* or *F/F~MSY~*, it is critical that users are aware of the broadly demonstrated poor performance of these methods in the absence of highly reliable fishery-specific information.
-In some instances using the data presented here did provide some improvement over use of catch-only style assessment methods; the difficulty comes in attempting to apply data types uniformly across the globe.
-While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RLSADB, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space.
-The overall poor performance of the models tested here, in terms of estimating both *B/B~MSY~* and *F/F~MSY~*, shows that improvements in estimates of global stock status depend on improvements in the quality and use of data themselves.
-
-Our fits to the RLSADB data provide a useful diagnostic of the degree to which model mispecification might explain the poor performance of catch-only models.
-Catch-only models generally employ some form of biomass dynamics model, such as the Schaefer surplus production model [@schaefer1954].
-These models are generally much simpler than the typically age-structured population models underpinning the assessments in the RLSADB.
-While we found that fitting the Pella-Tomlinson model employed in sraplus to an index of abundance drawn directly from the RLSADB provided much improved and reasonably accurate estimates of stock status relative to catch-only approaches (Table.\@ref(tab:perf-tab)), even with the index of abundance sraplus still had an average MAPE of `r scales::percent(perf$mape[perf$'Data Used' == "RLSADB Index"])`.
-This discrepancy may provide a measure of relatively insurmountable error resulting from model misspecification that inclusion of additional data may not be able to overcome.
-For example, many assessments in the RLSADB calculate *B/B~MSY~* based on spawning stock biomass.
-However, surplus production models condition the size and state of the population on the catch, which is by definition a measure of the vulnerable biomass.
-Vulnerable biomass may not reliably track spawning biomass in highly selective fisheries, potentially resulting in model biases as the surplus production model is not capable of separately tracking spawning and vulnerable biomass.
-
-We chose to test the performance of methods against estimated values reported in RLSADB.
-A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RLSADB, in addition to the errors in the RLSADB estimates themselves.
-For this reality to change our results though would require that the methods tested here have massively lower bias and higher accuracy for unassessed fisheries than RLSADB stocks.
-@free2020's simulation based results of these same types of models suggests that this is unlikely to be true.
-However, while many assessed fisheries have mechanisms to actively manage catches, catch histories in unassessed and unmanaged fisheries may be more likely to reflect underlying stock status than those in relatively well managed fisheries, though any predictive relationship will still depend on a host of factors such as changes in effort.
-This could be empirically tested by considering the performance of catch-only models on data from earlier in the history of fisheries in the RLSADB, though we did not do so here as we do not have historic estimates of FMI or SAR values.
-
-<!-- Our estimates of global stock status vary wildly depending on which source of data we use.
-Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information?
-Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)).
-Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used.
-These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability.
-The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations.
-Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks.
-None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RLSADB (XX RLSADB matches?
-need to add an analysis: match stocks from SOFIA to RLSADB stocks where possible, plot comparison).
--->
-
-Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions.
-The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e.
-those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)).
-Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet.
-Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of *F/F~MSY~* of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)` (see SI).
-But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly.
-Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas of the globe results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.
-We found that performance of different data sources varied widely both within and among regions (Fig.\@ref(fig:cs-plot), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)).
-Some of this variation is likely simply due to low sample sizes of assessed stocks reported in the RLSADB in some regions.
-But, other explanations for differences in performance among stocks and regions may be an interesting area for future research.
-For example, it may be that stocks in some regions are more suitable to being represented by surplus production models than others.
-
-Our value of information analysis also shows the high utility of having access to even a recent snapshot of *F/F~MSY~* (Fig.\@ref(fig:voi-plot)).
-Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct fishery-specific priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution.
-Another avenue would be to prioritize the development of a global repository for length and age composition data.
-Given appropriate conditions, these length measurements can be used to estimate local fishing mortality rates [@hordyk2016; @prince2019; @rudd2017].
-While length-based assessments come with a host of assumptions and potential pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone.
-Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f; @rudd2017].
-
-<!-- Using catch data alone does not reveal about recent stock status beyond what is encoded in our priors and functional forms.
-So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest).
-Given the demonstrated value of F/F~MSY~ values in improving estimates of stock status, and the increased feasibility in estimating F/F~MSY~ at a local scale relative to B/B~MSY~, how might we go about obtaining estimates of fishing mortality rates at a global scale?
-We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/F~MSY~ for individual fisheries (see Supplemental Information).
-While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful.
--->
-
-We must also prioritize collection and curation of fish population survey data worldwide.
-Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://www.FishStats.org) ( www.FishStats.org).
-Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments.
-Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) (https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes.
-However, our value of information exercise indicates that we may have to wait many years for new surveys to provide substantial improvements in status estimates (Fig.\@ref(fig:voi-plot)).
-
-Expanded training of fisheries scientists around the globe is another critical need.
-Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise.
-This is why stock assessments even in data-rich fisheries are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used.
-Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results.
-We may also need to acknowledge that in some fisheries reliable estimates of stock status relative to MSY based reference points are simply not possible (or perhaps desirable), and instead rely on more precautionary or empirical management management measures such as spatial closures, size restrictions, and indicator based harvest strategies (ideally tested through management strategy evaluation) [@prince2019;@dowling2015;@fulton2016].
-
-The coming decades are a critical time for the future of fisheries and ocean health.
-Achieving the United Nations Sustainable Development Goal 14 for the conservation and sustainable use of the world's oceans depends on our ability to effectively assess the status of fish stocks around the world.
-The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
-However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed.
-Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a solution to this problem which has proven to be unbiased and sufficiently precise at a global or regional level.
-
-<!-- Our results show that improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists. -->
-
-The lack of strong information on stock status within catch histories alone means that differences in models and assumptions between catch-based assessment efforts can produce starkly contrasting conclusions on global stock status, leading to debates that are inconclusive as they are inherently driven by assumptions.
-The FAO is leading efforts to increase technical capacity and monitoring and evaluation infrastructure to improve fisheries management in places with limited data.
-Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets.
-Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally [@costello2020] Achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.
-
-<!-- The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries. -->
-
-# Data Availability Statement {.unnumbered}
-
-All data and materials needed to reproduce our results are publicly available or queried by code available at https://github.com/DanOvando/assessing-global-fisheries.
-
-# Acknowledgements {.unnumbered}
-
-Funding for this work was provided by the Food and Agriculture Organization of the United Nations (FAO).
-We thank participants of the FAO "Methods for Global Assessment" workshop held in Rome, Italy February 2019 for helpful feedback on this project.
-We also thank A.
-Hordyk and one anonymous reviewer for their helpful comments.
-The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of the FAO, NOAA or the Department of Commerce.
-
-# References {.unnumbered}
-
-::: {#refs}
-:::
-
-\newpage
-
-# Tables {.unnumbered}
-
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Data Source                                 | Short Name       | Data Use                                                                 | Caveats                                                                                                         |
-+=============================================+==================+==========================================================================+=================================================================================================================+
-| Catch data [@fao2020]                       | catches          | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Fisheries Management Index [@melnychuk2017] | FMI              | Priors on most recent *F/F~MSY~* values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Swept Area Ratio [@amoroso2018]             | SAR              | Priors on most recent *F/F~MSY~* values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Reconstructed effort data [@rousseau2019]  | effort           | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors.
-Assumed rate of technology creep reported in individual sections |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-:(\#tab:data) Data sources included across model fits.
-
-```{r dat-desc}
-dat_description <- tribble(~`Data Name`, ~Description,
-        "RLSADB Index", "Fit to abundance index from RLSADB",
-        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
-        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
-        "Efective CPUE", "Fit to CPUE index created from RLSADB catch and regional effort index. 2.6% technology creep",
-        "Effective CPUE+", "Fit to CPUE index created from RLSADB catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
-        "Nominal CPUE", "Fit to CPUE index created from RLSADB catch and regional effort index. 0% tech. creep",
-         "Nominal CPUE+", "Fit to CPUE index created from RLSADB catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
-        "Guess", "Priors on terminal B/Bmsy randomly sampled from 0.4,1.0,1.6")
-
-knitr::kable(dat_description,
-             caption = "Data sources used for terminal stock status estimate",
-             format = "pipe") %>%
-  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
-  kableExtra::row_spec(0, bold = T)
-```
-
-```{r perf-tab}
-
-knitr::kable(perf %>% rename(MPE = mpe, MAPE = mape, Accuracy = accuracy), digits = 2,
-             caption = "Global performance statistics in the most recent year available of models using different sources of data.
-MPE = median percent error (bias), MAPE = median absolute percent error (error), Accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished).
-Performance is judged relative to *B/B~MSY~* reported values in RAM Legacy Stock Assessment Database.",
-              format = "pipe")
-
-```
-
-\newpage
-# Figure Legends  {.unnumbered}
-
-Figure 1: RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], catch heuristics, priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort (Effort) trends assuming a rate of technological increase of 2.6%.
-Each point is a stock in the RLSADB.
-Black dashed line shows the 1:1 relationship.
-Text displys root mean squared error (RMSE) of each panel.
-
-Figure 2: Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-Data source panels are ordered in ascending (starting from top left) total bias at the FAO region level.
-RLSADB Index refers to catch and abundance index drawn from RLSADB.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-
-Figure 3: Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources.
-Data source panels are ordered in ascending (starting from top left) mean MAPE at the FAO region level.
-RLSADB Index refers to catch and abundance index drawn from RLSADB.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-
-Figure 4: Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
-Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region level.
-RLSADB Index refers to catch and abundance index drawn from RLSADB.
-Effective CPUE refers to an index of abundance based on reconstructed effort data.
-Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data.
-For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed.
-FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio.
-CMSY uses the methods from Froese et al.
-2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-
-Figure 5: Posterior probability distributions of estimated effect of different data types on root mean squared error (RMSE) of *B/B~MSY~* in the most recent 5 years of data available for each model fit.
-Distribution is full posterior probability distribution.
-Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
-Change is relative to the mean performance of a catch-only heuristic model.
-
-<!-- \newpage -->
-
-<!-- # Supporting Information {.unnumbered} -->
-
-<!-- ```{r si, child = 'ovando-etal-assessing-global-fisheries-si.Rmd'} -->
-<!-- ``` -->
+4.
+Run make-assessing-global-fisheries.R, setting all run_ options to TRUE (this will likely take over 48 hours to run).
+This will knit the manuscript for this paper automatically, generating ovando-etal-assessing-global-fisheries.docx in the documents folder

### 2021-06-26 documents/ovando-etal-assessing-global-fisheries-si.Rmd
Accepted version (#13)
* fixed new FAO areas mapping problem

* Minor revisions

* updating README

@@
+---
+title: Improving Estimates of the State of Global Fisheries <br> Depends on Better Data
+subtitle: Supplementary Information
+bibliography: ["../references.bib"]
+csl: fish-and-fisheries.csl
+output: 
+  bookdown::pdf_document2:
+    latex_engine: xelatex
+  bookdown::word_document2:
+    reference_docx: template.docx
+params:
+  results_name: ["v1.0"]
+  min_years_catch: [25]
+always_allow_html: true
+linkcolor: blue
+header-includes:
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}\linenumbers
+toc: false
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
+                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
+library(tidyverse)
+library(viridis)
+library(sf)
+library(hrbrthemes)
+library(extrafont)
+library(scales)
+library(patchwork)
+library(rstanarm)
+library(bayesplot)
+library(here)
+library(sraplus)
+extrafont::loadfonts()
+
+functions <- list.files(here::here("functions"))
+
+purrr::walk(functions, ~ source(here::here("functions", .x)))
+
+results_name <- params$results_name
+
+results_path <- here("results",results_name)
+
+min_years_catch <- params$min_years_catch
+
+prepare_sofia_data(lookup_fmi_names = FALSE)
+
+ram_comp_data <- ram_data %>%
+  mutate(has_things = !(is.na(catch) | catch == 0)) %>%
+  filter(has_things) %>%
+  group_by(stockid) %>%
+  mutate(delta_year = as.integer(year - lag(year))) %>%
+  mutate(delta_year = case_when(year == min(year) ~ as.integer(1),
+                                TRUE ~ delta_year)) %>%
+  mutate(missing_gaps = any(delta_year > 1)) %>%
+  filter(missing_gaps == FALSE) %>%
+  group_by(stockid) %>%
+  mutate(n = length(catch)) %>%
+  filter(n >= min_years_catch) %>%
+  ungroup()
+
+ram_comp_data <- ram_comp_data %>%
+  select(
+    year,
+    stockid,
+    country,
+    primary_FAOarea,
+    scientificname,
+    commonname,
+    year,
+    catch,
+    b_v_bmsy,
+    u_v_umsy,
+    total_biomass
+  ) %>%
+  rename(scientific_name = scientificname) %>% 
+  mutate(fao_area_code = as.numeric(primary_FAOarea)) %>% 
+  left_join(fao_species, by = "scientific_name")
+
+assess_ram_fits <- read_rds(file = file.path(results_path,"assess_ram_fits.rds"))
+
+assess_ramu_fits <- read_rds(file = file.path(results_path,"assess_ramu_fits.rds"))
+
+ram_v_sraplus_area <- read_rds(file = file.path(results_path,"ram_v_sraplus_area.rds"))
+
+ram_comp_data <- read_rds(file = file.path(results_path,"ram_comp_data.rds"))
+
+fao_sraplus_comp_data <- read_rds(file = file.path(results_path, "fao_sraplus_comp_data.rds"))
+
+load(file = file.path(results_path, "paper_plots.Rdata"))
+
+load(file = file.path(results_path, "voi_fits.RData"))
+
+  exs <- read_rds(file = file.path(results_path, "raw-case-study-fits.rds"))
+
+breaks <- c(0.8, 1.2)
+
+breaks <- c(0, breaks, Inf)
+
+labels <- c("over", "fully", "under")
+
+# rand_fits <- assess_ram_fits %>%
+#   filter(data == "ram-data") %>%
+#   mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
+#          data = "guess")
+
+perf <- assess_ram_fits %>% 
+  # bind_rows(rand_fits) %>% 
+  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
+  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
+  mutate(data = as.factor(data)) %>% 
+  filter(data != "u_umsy") %>% 
+  mutate(data = forcats::fct_recode(data, 
+                                    "RLSADB Index" = "ram-data",
+                                    "SAR" = "sar",
+                                    "FMI" = "fmi",
+                                    "Effective CPUE" = "cpue",
+                                    "Effective CPUE+" = "cpue-plus",
+                                    "Nominal CPUE" = "nominal-cpue",
+                                    "Nominal CPUE+" ="nominal-cpue-plus",
+                                    "CMSY" = "cmsy",
+                                    "Guess" = "guess",
+                                    "RLSADB F/Fmsy" = "u_umsy"
+
+                                    )) %>% 
+  group_by(data) %>% 
+  summarise(
+    mpe = median((mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
+    mape = median(abs(mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
+    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
+  ) %>% 
+  arrange((mape)) %>% 
+  rename("Data Used" = "data")
+
+uperf <- assess_ramu_fits %>% 
+  # bind_rows(rand_fits) %>% 
+  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
+  mutate(ram_bin = cut(ram_u_v_umsy, breaks = breaks, labels = labels)) %>%
+  mutate(data = as.factor(data)) %>% 
+  filter(data != "u_umsy") %>% 
+  mutate(data = forcats::fct_recode(data, 
+                                    "RLSADB Index" = "ram-data",
+                                    "SAR" = "sar",
+                                    "FMI" = "fmi",
+                                    "Effective CPUE" = "cpue",
+                                    "Effective CPUE+" = "cpue-plus",
+                                    "Nominal CPUE" = "nominal-cpue",
+                                    "Nominal CPUE+" ="nominal-cpue-plus",
+                                    "CMSY" = "cmsy",
+                                    "Guess" = "guess",
+                                    "RLSADB F/Fmsy" = "u_umsy"
+
+                                    )) %>% 
+  group_by(data) %>% 
+  summarise(
+    mpe = median((mean - ram_u_v_umsy ) / ram_u_v_umsy),
+    mape = median(abs(mean - ram_u_v_umsy ) / ram_u_v_umsy),
+    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
+  ) %>% 
+  arrange((mape)) %>% 
+  rename("Data Used" = "data")
+
+fao_area_codes <- fao %>% 
+  ungroup() %>% 
+  select(fao_area, fao_area_code) %>% 
+  unique()
+
+pub_theme <- theme_ipsum(base_size = 10,
+                         axis_text_size = 10,
+                         axis_title_size = 12) +
+    theme(panel.spacing = unit(.5,"lines"))
+
+theme_set(pub_theme)
+
+```
+
+# Title Page {.unnumbered}
+
+**Title 1**: Improving Estimates of the State of Global Fisheries Depends on Better Data
+
+**Title 2**: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
+
+**Running Title**: Unassessed Fisheries
+
+**Authors**:
+Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James T. Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^
+
+**Affiliations**:
+
+^1^University of Washington, School of Aquatic and Fishery Sciences
+1122 NE Boat St, Box 355020 Seattle, WA 98195-5020
+Seattle, WA, USA 98103
+
+^2^NOAA Fisheries Alaska Fisheries Science Center, Seattle, WA, USA
+
+^3^Scaleability LLC
+
+^4^Food and Agriculture Organization of the United Nations, Fisheries and Aquaculture,Rome, IT
+
+^5^University of Tasmania, Institute for Marine and Antarctic Studies,Hobart, TAS, AUS
+
+**Authorship**: DO, RH, CM,MR, RS, JT designed model structure and ran analyses. RS, YR, and YE supplied data. All authors contributed to writing of the manuscript
+
+**Conflict of Interest**: RH receives research funding from many groups that have interests in fisheries outcomes including environmental NGOs, foundations, governments and fishing industry groups.
+
+\newpage
+
 \renewcommand{\thefigure}{S\arabic{figure}}
@@
 cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "$K$", `Default Prior` = "Prior predictive tuning") %>%
-  rbind(c("Growth rate", "$r$", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
-  rbind(c("Shape parameter", "$m$", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
+  rbind(c("Growth rate", "$r$", "@thorson2020 updated by prior predictive tuning")) %>%
+  rbind(c("Shape parameter", "$m$", "Drawn from @thorson2012")) %>%
   rbind(c("Catchability", "$q$", "$LN(1e^{-3}, 1)$")) %>%
@@
   format = "pipe",
-  caption = "SSBMSY to SSB ratios from Thorson et al.
-(2012) used in the paper.
-Taxa not within the groups assigned at the genus level by Thorson et al.
-(2012) are assigned the ratio reported for 'Other'"
+  caption = "SSBMSY to SSB ratios from @thorson2012 used in the paper. Taxa not within the groups assigned at the genus level by @thorson2012 are assigned the ratio reported for 'Other'"
 )
@@
 ```{r umpe-map, fig.width=8, fig.height=6, fig.cap="Median percent error (MPE, predicted relative to observed) in most recent *F/F~MSY~* by FAO statistical area from different data sources.
+Data source panels are ordered in ascending (starting from top left) mean MPE at the FAO region level.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 ```{r umape-map, fig.cap = "Median absolute percent error (MAPE) in most recent *F/F~MSY~* by FAO statistical area from different data sources.
+Data source panels are ordered in ascending (starting from top left) total bias at the FAO region level.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 ```{r uacc-map, fig.cap="Mean classification accuracy (assignment to general bin of overfishing, fishing near F~MSY~, and underfishing) by FAO statistical area arising from different data sources.
+Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region leve.
 RLSADB Index refers to catch and abundance index drawn from RLSADB.
@@
 ```
+
+\newpage
+
+# References {.unnumbered}
+
+::: {#refs}
+:::

### 2021-06-26 make-assessing-global-fisheries.R

@@ -25,6 +25,7 @@ library(tidybayes)
 library(rstan)
 library(rstanarm)
 library(sf)
+library(rgdal)
 library(portedcmsy)
 extrafont::loadfonts()
 
@@ -807,11 +808,11 @@ if (run_case_studies){
     mutate(fit = map(data,
                      safely(fit_case_studies)))
   
-  write_rds(exs, path = file.path(results_path, "raw-case-study-fits.rds"))
+  write_rds(exs, file = file.path(results_path, "raw-case-study-fits.rds"))
   
 } else {
   
-  exs <- read_rds(path = file.path(results_path, "raw-case-study-fits.rds"))
+  exs <- read_rds(file = file.path(results_path, "raw-case-study-fits.rds"))
   
 }
 
@@ -895,8 +896,8 @@ tmp2 <- tmp2 %>%
 
 ex_performance <- tmp2 %>% 
   group_by(variable, data) %>% 
-  summarise(r2 = round(yardstick::rsq_vec(truth, value),2)) 
-
+  summarise(r2 = round(yardstick::rsq_vec(truth, value),2),
+            rmse = round(yardstick::rmse_vec(truth, value),2)) 
 
 ex_scatter_plot <- tmp2 %>% 
   ggplot(aes(truth, value)) + 
@@ -904,7 +905,7 @@ ex_scatter_plot <- tmp2 %>%
   geom_hline(aes(yintercept = 0)) +
   geom_abline(aes(slope = 1, intercept = 0),linetype = 2) +
   geom_point(alpha = 0.5, size = 2) + 
-  ggtext::geom_richtext(data = ex_performance, aes(x = 1, y = 4, label = paste0("R<sup>2</sup> = ",r2))) +
+  ggtext::geom_richtext(data = ex_performance, size = 3,alpha = 0.75,aes(x = 1.25, y = 3.5, label = paste0("RMSE = ",rmse))) +
   facet_grid(variable ~ data, scales = "free_x")  + 
   # scale_size(trans = "sqrt", name = "Lifetime Catch") +
   scale_x_continuous(name = "RLSADB Value", expand = expansion(add = c(0, .1))) + 
@@ -1381,7 +1382,7 @@ if (run_sofia_comparison == TRUE) {
   
 } else {
   fao_status_fits <-
-    read_rds(path = file.path(results_path, "fao_status_fits.rds"))
[…]

### 2021-06-26 tmp.R

@@ -1,145 +1,0 @@
-run_case_studies <- TRUE
-
-exs <- ram_data %>%
-  filter(
-    stockid %in% unique(sraplus::fmi$stockid) &
-      stockid %in% unique(sraplus::sar$stockid)
-  ) %>%
-  mutate(fao_area_code = as.numeric(primary_FAOarea),
-         country = as.character(primary_country)) %>%
-  mutate(country =
-           countrycode::countrycode(country, "country.name", "un.name.en"))
-
-comp_stocks <- exs %>%
-  select(scientificname) %>%
-  unique() %>%
-  mutate(fishbase_vuln = map_dbl(
-    scientific_name,
-    ~ rfishbase::species(.x, fields = "Vulnerability")$Vulnerability
-  )) %>%
-  mutate(
-    resilience = dplyr::case_when(
-      fishbase_vuln < 33 ~ "Low",
-      fishbase_vuln >= 66 ~ "High",
-      TRUE ~ "Medium"
-    )
-  ) %>%
-  select(scientificname, resilience)
-
-
-exs %>%
-  ggplot(aes(year, b_v_bmsy, color = stockid)) +
-  geom_line(show.legend = FALSE)
-
-total_nominal_effort <- rous_data %>%
-  left_join(effort_region_to_country, by = "country") %>%
-  rename(fao_area_code = area) %>%
-  group_by(fao_area_code, year, country) %>%
-  summarise(total_effort = sum(effort_cell_reported_nom, na.rm = TRUE)) %>%
-  ungroup()
-
-matched_fmi <- sraplus::fmi %>%
-  filter(stockid %in% unique(exs$stockid)) %>%
-  group_by(stockid) %>%
-  nest() %>%
-  ungroup() %>%
-  mutate(tmp = map(data, ~ map_df(.x[,c("research", "management", "enforcement","socioeconomics")], mean))) %>%
-  select(-data) %>%
-  unnest(cols = tmp) %>%
-  select(stockid, research, management, enforcement, socioeconomics)
-
-
-matched_sar <- sraplus::sar %>%
-  filter(stockid %in% unique(exs$stockid)) %>%
-  group_by(stockid) %>%
-  nest() %>%
-  ungroup() %>%
-  mutate(tmp = map(data, ~ purrrlyr::dmap_at(.x, c("sar"), mean))) %>%
-  select(-data) %>%
-  unnest(cols = tmp) %>%
[…]

### 2021-07-16 documents/ovando-etal-assessing-global-fisheries.Rmd
updating README
back to main and stashing as published version

@@
 ---
-title: Improving Estimates of the State of Global Fisheries <br> Depends on Better Data
-date: "`r Sys.Date()`"
 bibliography: ["../references.bib"]
@@
   - \usepackage{lineno}\linenumbers
-toc: false
 ---
@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
-                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
+                      cache = FALSE, fig.width = 6, fig.asp = .75)
 library(tidyverse)
@@
 
-**Title 1**: Improving Estimates of the State of Global Fisheries Depends on Better Data
-
-**Title 2**: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
-
-**Running Title**: Unassessed Fisheries
+**Title**: Improving Estimates of the State of Global Fisheries Depends on Better Data
 
 **Authors**:
-Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James T. Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^
+Daniel Ovando^1^*, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James T. Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^
 
@@
 
+**Corresponding Author** *Daniel Ovando, danovan@uw.edu, University of Washington, School of Aquatic and Fishery Sciences
+1122 NE Boat St, Box 355020 Seattle, WA 98195-5020
+Seattle, WA, USA 98103
+
 **Authorship**: DO, RH, CM,MR, RS, JT designed model structure and ran analyses. RS, YR, and YE supplied data. All authors contributed to writing of the manuscript
@@
 
+**Running Title**: Unassessed Fisheries
+
 \newpage
 
-# Abstract {.unnumbered}
+# Abstract {.unnumbered .unlisted}
 
@@
 But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time.
-Withing these broad summaries the performance of models trained on our tested data sources varied widely across regions.
-Effective improvement in estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.
+Within these broad summaries the performance of models trained on our tested data sources varied widely across regions.
+Substantial improvements to estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.
+
+# Keywords  {.unnumbered}
+- Stock assessment
+- Fisheries management
+- Data-limited assessment
+- Catch-only
+- Global fisheries
+- United Nations Sustainable Development Goals
+
+# Table of Contents {.unnumbered}
+
+  - Introduction
+  - Methods
+      - Data Sources
+      - Population Model
+      - Estimation Model
+      - CMSY
+      - Priors
+      - Assessing Performance
+      - Value of Information
+      - Case Study
+  - Results
+      - Case Study
+      - Performance of Regional Fishery Assessments
+      - Value of Information
+  - Discussion
+  - Acknowledgments
+  - Data Availability Statement
+  - References
 
@@
 
-We allow for process error $p_t$ (in the manner of the stochastic stock reduction analysis (SRA) suggested by @walters2006). Process error $p_t$ is assumed to be log-normally distributed, such that
+We allow for process error $p_t$ (in the manner of the stochastic stock reduction analysis suggested by @walters2006). Process error $p_t$ is assumed to be log-normally distributed, such that
 
@@
 
-When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in Stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018].
+When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in Stan [@standevelopmentteam2020] accessed through the `tmbstan` interface [@monnahan2018].
 By default the model uses 2000 draws with a 1000 step warm-up and one chain.
@@
 
-### CMSY {.unnumbered}
+## CMSY {.unnumbered}
 
@@
 This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.
-However, Borel's Paradox poses a particular problem when there are no data to fit to(i.e.
+However, Borel's Paradox poses a particular problem when there are no data to fit to (i.e.
 when the model is simply filtering through prior distributions in the manner of a traditional SRA) due to the fact that there are more parameter combinations that allow for a fishery to be relatively unexploited than for a fishery to be close to collapse (but never actually collapsed, i.e.
 predicted biomass less than observed catch).
-In this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020).
+This causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020).
 This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the posterior distributions of stock status resulting from the SRA algorithm.
@@
 Users can turn this functionality off and instead base priors on stock status primarily on life history.
-See Supplementary Information for a detailed explanation of this problem and our solution.
+See Supplementary Information (SI) for a detailed explanation of this problem and our solution.
 
@@
 Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries.
-For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on *B/B~MSY~* or *F/F~MSY~* in the most recent year of the fishery usable by sraplus (See Supplementary Information).
+For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on *B/B~MSY~* or *F/F~MSY~* in the most recent year of the fishery usable by sraplus (See SI).
 All prior regression models were tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017].
@@
 Black dashed line shows the 1:1 relationship.
-Text displys root mean squared error (RMSE) of each panel."}
+Text displys root mean squared error (RMSE) of each panel.", include=TRUE}
 ex_scatter_plot + 
@@
 
-ggsave(here("documents","figs","cs-plot.png"),ex_scatter_plot, height = 6, width = 6)
+ggsave(here("documents","figs","figure_1.eps"),ex_scatter_plot, height = 6, width = 6, device = cairo_ps, dpi = 600)
 
@@
 Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RLSADB index data were able to capture the overall state or geographic distribution of stock status represented in RLSADB in a consistent manner.
-Performance in estimating *F/F~MSY~ * was similarly variable and poor, with the exception that the default settings of CMSY performed much more consistently poorly in terms of *F/F~MSY~* than *B/B~MSY~* (due to systemic overestimation of *F/F~MSY~*, see Supplementary Information [SI]).
+Performance in estimating *F/F~MSY~ * was similarly variable and poor, with the exception that the default settings of CMSY performed much more consistently poorly in terms of *F/F~MSY~* than *B/B~MSY~* (due to systemic overestimation of *F/F~MSY~*, see SI).
 
@@
 2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.
-"}
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.", include=TRUE}
 ram_mpe_map_plot
 
-ggsave(here("documents","figs","mpe-map.png"), ram_mpe_map_plot, width = 8, height = 6)
+ggsave(here("documents","figs","figure_2.eps"), ram_mpe_map_plot, width = 8, height = 6, device = cairo_ps,dpi = 600)
 ```
@@
 2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6}
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6, include=TRUE}
 ram_mape_map_plot
 
-ggsave(here("documents","figs","mape-map.png"),ram_mape_map_plot, width = 8, height = 6)
+ggsave(here("documents","figs","figure_3.eps"),ram_mape_map_plot, width = 8, height = 6,device = cairo_ps,dpi = 600)
 
@@
 2017 [@froese2017].
-Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6}
+Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6,include=TRUE}
 
@@
 
-ggsave(here("documents","figs","acc-map.png"),ram_acc_map_plot, width = 8, height = 6)
+ggsave(here("documents","figs","figure_4.eps"),ram_acc_map_plot, width = 8, height = 6,device = cairo_ps,dpi = 600)
 
@@
 Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th.
-Change is relative to the mean performance of a catch-only heuristic model."}
+Change is relative to the mean performance of a catch-only heuristic model.", include = TRUE}
 
@@
 
-ggsave(here("documents","figs","voi-plot.png"),voi_plot, width = 8, height = 6)
+ggsave(here("documents","figs","figure_5.eps"),voi_plot, width = 8, height = 6,device = cairo_ps,dpi = 600)
 
@@
 The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years.
-However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed.
+However, this process still leaves a substantial number of fisheries and proportion of global catch lacking specific assessments of stock status relative to reference points.
 Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a solution to this problem which has proven to be unbiased and sufficiently precise at a global or regional level.
@@
 <!-- The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries. -->
-
-# Data Availability Statement {.unnumbered}
-
-All data and materials needed to reproduce our results are publicly available or queried by code available at https://github.com/DanOvando/assessing-global-fisheries.
 
@@
 
+# Data Availability Statement {.unnumbered}
+
+All data and materials needed to reproduce our results are publicly available or queried by code available at https://github.com/DanOvando/assessing-global-fisheries.
+The data used at time of submission have also been archived at https://figshare.com/articles/preprint/assessing-global-fisheries/13070627.
+
 # References {.unnumbered}
@@
 
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
 | Data Source            | Short Name | Data Use | Caveats |
-+=============================================+==================+==========================================================================+=================================================================================================================+
-| Catch data [@fao2020]                       | catches          | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-| Fisheries Management Index [@melnychuk2017] | FMI              | Priors on most recent *F/F~MSY~* values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
+|------------------------|------------|----------|---------|
+| Catch data [@fao2020] |  Catch    | Priors on stock status, scaling of population size, exploitation history |  Heuristics or regressions used to translate shape of catch history into priors on stock status |
+| Fisheries Management Index [@melnychuk2017] | FRI  | Priors on most recent *F/F~MSY~* values   |  Priors produced by regression trained on data from RAM Legacy Stock Assessment Database |
 | Swept Area Ratio [@amoroso2018]   |  SAR | Priors on most recent *F/F~MSY~* values | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database |  
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
 | Reconstructed effort data [@rousseau2019] | effort |  Combined with catch data to create an index of abundance | Total reconstructed effort across all sectors.
 Assumed rate of technology creep reported in individual sections |
-+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
-:(\#tab:data) Data sources included across model fits.
+Table: (\#tab:data) Candidate data types included across sraplus model fits.
 

### 2021-07-16 make-assessing-global-fisheries.R
merging back to main

@@ -60,29 +60,31 @@ results_name <- "v1.0"
 results_description <-
   "publication version of results"
 
-run_voi_models <- TRUE
+message("First time running results, set all the following options to TRUE")
+
+run_voi_models <- TRUE # run value of information calculations
 # sub options for run_voi_models
-fit_models <- FALSE
+fit_models <- FALSE # actually fit the VOI models
 
-write_results <- FALSE
+write_results <- TRUE # leave as TRUE generally
 
-process_fits <- FALSE
+process_fits <- FALSE # process fits to RAM data
 
-run_case_studies <- FALSE
+run_case_studies <- FALSE # run case studies
 
-run_sofia_comparison <- FALSE
+run_sofia_comparison <- FALSE # run comparison to SOFIA estimates
 
-run_ram_tests <- FALSE
+run_ram_comparison <- FALSE # run comparison to RAM estimates, the main results
 
-run_ram_comparison <- FALSE
+run_ram_tests <- FALSE # mostly diagnostic tests, no included in published results
 
-knit_paper <- FALSE
+knit_paper <- FALSE # set to TRUE to knit paper based on results
 
-warning("Running full analysis takes upwards of 24 hours on 2 cores. Recommend starting on a Friday night and then having a nice weekend. Given memory constraints of models, more than 2 cores is not recommended (memory routinely runs out on a 16 core 36GB machine when cores > 2)")
+message("Running full analysis takes upwards of 24 hours on 2 cores. Recommend starting on a Friday night and then having a nice weekend. Given memory constraints of models, more than 2 cores is not recommended (memory routinely runs out on a 16 core 36GB machine when cores > 2)")
 
-engine <-  "stan"
+engine <-  "stan" # leave as stan
 
-catchability = 1e-2
+catchability = 1e-2 # defauly catchability coefficient
 
 pub_theme <- theme_ipsum(base_size = 10,
                          axis_text_size = 10,
@@ -1681,16 +1683,16 @@ compare_to_ram <- function(data, fit){
 
 ram_test_comparison <- map2_df(ram_fit_tests$data, ram_fit_tests$fit, compare_to_ram,.id = "stock")
 
-# ram_test_comparison %>% 
-#   ggplot(aes(observed, mean, color = stock)) + 
[…]

### 2021-07-16 README.Rmd
cleaning up README

@@
 Within these broad summaries the performance of models trained on our tested data sources varied widely across regions.
-Effective improvement in estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.s exploited fish populations depends on prioritizing the collection of high-priority
+Substantial improvements to estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.
 
-*Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
+<!-- *Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources.
 Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region level.
@@
 2017 [@froese2017].
-Guess assigns a random recent B/B~MSY~ of 0.4,1, or 1.6.*
-```{r, echo = FALSE}
-knitr::include_graphics("documents/figs/acc-map.png")
+Guess assigns a random recent B/B~MSY~ of 0.4,1, or 1.6.* -->
+
+```{r, echo = FALSE, include=FALSE}
+knitr::include_graphics("documents/figs/figure_4.eps")
 ```

### 2022-01-25 make-assessing-global-fisheries.R
cleaning up .Rmd

@@ -221,25 +221,11 @@ if (run_voi_models == TRUE) {
     saveRDS(ram_fits,
             file = here::here("results", results_name, "ram_experiments.rds"))
     
-    # sra_model <-
-    #   rstan::stan_model(file = here::here("src", "biosra.stan"))
-    
+ 
     future::plan(multiprocess, workers = n_cores)
     
     
     ram_fits <- ram_fits %>%
-      # filter(initial_state_type == "unfished",
-      #        u_window == "snapshot",
-      #        use_index == TRUE,
-      #        use_u_priors == TRUE,
-      #        estimate_shape == FALSE) %>%
-      # filter(use_u_priors == FALSE,
-      #        use_index == TRUE,
-      #        index_window == 1) %>%
-      # mutate(sciname = map_chr(data,~unique(.x$"scientificname"))) %>%
-      # filter(sciname == "Ophiodon elongatus") %>%
-      # sample_n(4) %>%
-    # slice(6) %>%
     mutate(
       fits = future_pmap(
         list(
@@ -273,10 +259,6 @@ if (run_voi_models == TRUE) {
   }
   
   
-  # wtf <- map(ram_fits$fits,'error') %>% map_lgl(is.null)
-  #
-  # a <- ram_fits %>%
-  #   filter(!wtf)
   
   # process fits ------------------------------------------------------------
   
@@ -321,8 +303,7 @@ if (run_voi_models == TRUE) {
         return(out)
       }
     
-    # future::plan(multiprocess, workers = n_cores)
-    
+
     ram_fits <- ram_fits %>%
       mutate(
         fits = map2(
@@ -372,20 +353,7 @@ if (run_voi_models == TRUE) {
       summary_plot = map(fit, "summary_plot"),
       summary = map(fit, "fit_summary")
     )
-  # test <- test %>%
-  #   mutate(
-  #     summary_plot = map_plot(fit_performance, "summary_plot"),
-  #     summary = map(fit_performance, "fit_summary")
-  #   )
-  
-  
-  
-  # trelliscopejs::trelliscope(
-  #   ram_fits %>%
-  #     filter(use_index == TRUE),
-  #   name = "blah",
-  #   panel_col = "summary_plot"
-  # )
+
   
   
   fits <- ram_fits %>%
[…]
